### 2016-10-12 constrained_MLE/StandardErrors.tex

\documentclass[11pt]{article}
\usepackage{url,graphicx,tabularx,array,amsmath,amssymb,amsthm,booktabs,float}
\usepackage[margin=.65in]{geometry}
\setlength{\parskip}{1ex} %--skip lines between paragraphs
\setlength{\parindent}{0pt} %--don't indent paragraphs

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\newcommand{\vcentergraphics}[2]{\ensuremath{\vcenter{\hbox{\includegraphics[#1]{#2}}}}}

%-- Commands for header
\renewcommand{\title}[1]{\textbf{#1}\\}
\renewcommand{\line}{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}\hline\\\end{tabularx}\\[-0.5cm]}
\newcommand{\leftright}[2]{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}#1%
& #2\\\end{tabularx}\\[-0.5cm]}

\begin{document}

\line
\leftright{10.12.16}{Colin Lewis-Beck} %-- left and right positions in the header

\begin{description}
\item \textit{Getting Proper Standard Errors From R Stan:}\\
The Hessian Stan's optimization program returns, results in a covariance matrix in terms of:
\begin{align*}
[\log(t_{p_1}),\log(t_{p_2}),\log(\sigma_1), \log(\sigma_2), p]
\end{align}

Let's use the $\Delta$ method to get a covariance matrix for: [$ \mu_1,\mu_2,\sigma_1,\sigma_2,p$].  Not the following two relationships between $\mu$ and the quantiles.

\begin{equation*}
\begin{align}
\Phi^{-1}_{SEV}(p) &=\log(-\log(1-p)) \\
\log(t_p)&=\mu +\log(\sigma)\Phi^{-1}_{SEV}(p)\\
\end{align}
\end{equation}

Solving for $\mu$ we get:
\begin{equation*}
\mu&=\log(t_p)-\exp({\log(\sigma)})\Phi^{-1}_{SEV}(p)
\end{equation}
Taking partials with respect to  $\frac{d\mu}{d\log(t_p)}$ and $\frac{d\mu}{d\log(\sigma)}$ we get:
\begin{equation*}
\begin{align}
\frac{d\mu}{d\log(t_p)} &= 1 \\
\frac{d\mu}{d\log(\sigma)} &= -\exp(\log(\sigma))\Phi^{-1}_{SEV}(p)
\end{align}
\end{equation}

It is easy to transform $\log(\sigma)$ and as far as I can tell, $p$ doesn't need to be transformed, but we might want to check if the standard errors should be changed.  Putting these pieces together we get a $\Delta$ matrix:
\begin{equation*}
\begin{align}
\Delta = \begin{bmatrix}
1 & 0 & -\exp(\log(\sigma_1)\Phi^{-1}_{SEV}(p_1) & 0 & 0\\
0 & 1 & 0 & -\exp(\log(\sigma_2)\Phi^{-1}_{SEV}(p_2) & 0\\
0 & 0 & \exp(\log(\sigma_1)) & 0 & 0\\
0 & 0 & 0 & \exp(\log(\sigma_2)) & 0\\
0 & 0 & 0 & 0 & 1\\
\end{bmatrix}
\end{align}
\end{equation}
Now, following Meeker and Escobar (Section 8.4.3), we need to get the standard errors for the GFLP CDF to make Wald bands.  Recall, the CDF: $g(\mu_1,\mu_2,\sigma_1,\sigma_2,p)=1-(1-pF_1)(1-F_2)$, which we can re-write as:
\begin{equation*}
\begin{align}
\Pr(t<T)&=1-(1-p\Phi_{SEV}(\mu_1,\sigma_1)(z_1))(1-\Phi_{SEV}(\mu_2,\sigma_2)(z_2))\\
&=\Phi_{SEV}(\mu_2,\sigma_2)(z_2)+p\Phi_{SEV}(\mu_1,\sigma_1)(z_1)-p\Phi_{SEV}(\mu_1,\sigma_1)(z_1)\Phi_{SEV}(\mu_2,\sigma_2)(z_2)
\end{align}
\end{equation}
Note that the SEV distribution is standardized in terms of $ z =(\log(t)-\mu)/\sigma$.  Let's again take partials of the GFLP CDF with respect to the 5 parameters.  Note: $z_1$ or $z_2$ should be plugged into the proper SEV function for each term.
\begin{equation*}
\begin{align}
\frac{dg}{d\mu_1}&=-p\phi_{SEV}(\mu_1,\sigma_1)\frac{1}{\sigma_1}+p\phi_{SEV}(\mu_1,\sigma_1)\Phi_{SEV}(\mu_2,\sigma_2)\frac{1}{\sigma_1}\\
\frac{dg}{d\mu_2}&=-\phi_{SEV}(\mu_2,\sigma_2)\frac{1}{\sigma_2}+p\Phi_{SEV}(\mu_1,\sigma_1)\phi_{SEV}(\mu_2,\sigma_2)\frac{1}{\sigma_2}\\
\frac{dg}{d\sigma_1}&=p\phi_{SEV}(\mu_1,\sigma_1)\left(\frac{-log(t)+\mu_1}{\sigma^{2}_1}\right)-p\Phi_{SEV}(\mu_2,\sigma_2)\phi_{SEV}(\mu_1,\sigma_1)\left(\frac{-log(t)+\mu_1}{\sigma^{2}_1}\right)\\
\frac{dg}{d\sigma_1}&=\phi_{SEV}(\mu_2,\sigma_2)\left(\frac{-log(t)+\mu_2}{\sigma^{2}_2}\right)-p\Phi_{SEV}(\mu_1,\sigma_1)\phi_{SEV}(\mu_2,\sigma_2)\left(\frac{-log(t)+\mu_2}{\sigma^{2}_2}\right)\\
\frac{dg}{dp}&=\Phi_{SEV}(\mu_1,\sigma_1)-\Phi_{SEV}(\mu_1,\sigma_1)\Phi_{SEV}(\mu_2,\sigma_2)\\
\end{align}
\end{equation}
Putting these into a 5 x 5 matrix we can once again use the $\Delta$ method for the standard error of g.  Then with this standard error we can use a normal approximation and get $\hat{g} \pm z_{1-\alpha/2} \hat{g_{se}}$\\\\

Taking this a step further (Meeker and Hong, p. 169) we can use a tangent method to get a CI that respects the bounds of a CDF.

Let $y_e=log(t)$.  The lower and upper bounds are calculated as:
\begin{equation*}
\begin{align*}
y_l&=y_e-\frac{z_{1-\alpha/2}\hat{g_{se}}}{\frac{dF(t)}{log(t)}}\\
y_u&=y_e+\frac{z_{1-\alpha/2}\hat{g_{se}}}{\frac{dF(t)}{log(t)}}
\end{align}
\end{equation}
Taking the derivative of the GFLP in terms of $log(t)$, which it is already parameterized in terms of is:
\begin{equation*}
\begin{align}
\frac{dg(t)}{dlog(t)}&=p\phi_{SEV}(\mu_1,\sigma_1)(z_1)(1-\Phi_{SEV}(\mu_2,\sigma_2)(z_2))(1/\sigma_1)+\phi_{SEV}(\mu_2,\sigma_2)(z_2)(1-p\Phi_{SEV}(\mu_1,\sigma_1)(z_1))(1/\sigma_2)\\
\end{align}
\end{equation}
Finally, we plug these end points back into the GFLP CDF to get our final confidence intervals.
\begin{align*}
[\hat{F}(\exp(y_l)),\hat{F}(\exp(y_u))]
\end{align}

Phew.  I have checked this a few times, but it's kind of tedious so there very well be some errors and or typos.

\end{document}

### 2016-10-21 constrained_MLE/StandardErrors.tex
Checked analytical derivatives with numerical derivates to ensure calculations were done correctly.  Transformed the SE's for p, too.

@@
 \setlength{\parindent}{0pt} %--don't indent paragraphs
+\graphicspath{{C:/Users/Colin/Documents/GitHub/BB_data_analysis/constrained_MLE}
 
@@
 \line
-\leftright{10.12.16}{Colin Lewis-Beck} %-- left and right positions in the header
+\leftright{10.21.16}{Colin Lewis-Beck} %-- left and right positions in the header
 
@@
 
-Let's use the $\Delta$ method to get a covariance matrix for: [$ \mu_1,\mu_2,\sigma_1,\sigma_2,p$].  Not the following two relationships between $\mu$ and the quantiles.
+Let's use the $\Delta$ method to get a covariance matrix for: [$ \mu_1,\mu_2,\sigma_1,\sigma_2,p$].  Not the following three relationships between $\mu$ and the quantiles.
 
@@
 \Phi^{-1}_{SEV}(p) &=\log(-\log(1-p)) \\
-\log(t_p)&=\mu +\log(\sigma)\Phi^{-1}_{SEV}(p)\\
+\log(t_{pi})&=\mu_i +\log(\sigma_i)\Phi^{-1}_{SEV}(p_i)\\
+F(t;\mu;\sigma)&=\Phi_{SEV}\left[\frac{\log(t)-\mu}{\sigma}\right]
 \end{align}
@@
 \begin{equation*}
-\mu&=\log(t_p)-\exp({\log(\sigma)})\Phi^{-1}_{SEV}(p)
+\mu_i&=\log(t_{pi})-\exp({\log(\sigma_i)})\Phi^{-1}_{SEV}(p_i)
 \end{equation}
-Taking partials with respect to  $\frac{d\mu}{d\log(t_p)}$ and $\frac{d\mu}{d\log(\sigma)}$ we get:
+Taking partials with respect to  $\frac{d\mu_i}{d\log(t_{pi})}$ and $\frac{d\mu_i}{d\log(\sigma_i)}$ we get:
 \begin{equation*}
 \begin{align}
-\frac{d\mu}{d\log(t_p)} &= 1 \\
-\frac{d\mu}{d\log(\sigma)} &= -\exp(\log(\sigma))\Phi^{-1}_{SEV}(p)
+\frac{d\mu_i}{d\log(t_{pi})} &= 1 \\
+\frac{d\mu_i}{d\log(\sigma_i)} &= -\exp(\log(\sigma_i))\Phi^{-1}_{SEV}(p_i)
 \end{align}
@@
 
-It is easy to transform $\log(\sigma)$ and as far as I can tell, $p$ doesn't need to be transformed, but we might want to check if the standard errors should be changed.
+It is easy to transform $\log(\sigma_i)$.
+Stan estimates the SE's for p on the unconstrained space, so we use the inverse logit transformation ($\text{logit}^{-1}(p)=\frac{1}{1+e^{-p}}$) to get the SE's on the constrained space.
 Putting these pieces together we get a $\Delta$ matrix:
@@
 0 & 0 & 0 & \exp(\log(\sigma_2)) & 0\\
-0 & 0 & 0 & 0 & 1\\
+0 & 0 & 0 & 0 & (\frac{1}{1+e^{-p}})(\frac{e^{-p}}{1+e^{-p}})\\
 \end{bmatrix}
@@
 Now, following Meeker and Escobar (Section 8.4.3), we need to get the standard errors for the GFLP CDF to make Wald bands.
-Recall, the CDF: $g(\mu_1,\mu_2,\sigma_1,\sigma_2,p)=1-(1-pF_1)(1-F_2)$, which we can re-write as:
+Recall, the CDF: $g(\mu_1,\mu_2,\sigma_1,\sigma_2,p)=1-(1-pF(\mu_1;\sigma_1)(1-F(\mu_2;\sigma_2))$, which we can re-write as:
 \begin{equation*}
@@
 \frac{dg}{d\sigma_1}&=p\phi_{SEV}(\mu_1,\sigma_1)\left(\frac{-log(t)+\mu_1}{\sigma^{2}_1}\right)-p\Phi_{SEV}(\mu_2,\sigma_2)\phi_{SEV}(\mu_1,\sigma_1)\left(\frac{-log(t)+\mu_1}{\sigma^{2}_1}\right)\\
-\frac{dg}{d\sigma_1}&=\phi_{SEV}(\mu_2,\sigma_2)\left(\frac{-log(t)+\mu_2}{\sigma^{2}_2}\right)-p\Phi_{SEV}(\mu_1,\sigma_1)\phi_{SEV}(\mu_2,\sigma_2)\left(\frac{-log(t)+\mu_2}{\sigma^{2}_2}\right)\\
+\frac{dg}{d\sigma_2}&=\phi_{SEV}(\mu_2,\sigma_2)\left(\frac{-log(t)+\mu_2}{\sigma^{2}_2}\right)-p\Phi_{SEV}(\mu_1,\sigma_1)\phi_{SEV}(\mu_2,\sigma_2)\left(\frac{-log(t)+\mu_2}{\sigma^{2}_2}\right)\\
 \frac{dg}{dp}&=\Phi_{SEV}(\mu_1,\sigma_1)-\Phi_{SEV}(\mu_1,\sigma_1)\Phi_{SEV}(\mu_2,\sigma_2)\\
@@
 
-Phew.  I have checked this a few times, but it's kind of tedious so there very well be some errors and or typos.
+\begin{figure}[H]
+\caption{Wald Band 1 for Drive Model 3}
+\centering
+\includegraphics[height=8cm]{bandsmod3.pdf}
+\end{figure}
+
+\begin{figure}[H]
+\caption{Wald Band for Drive Model 3 Hong and Meeker}
+\centering
+\includegraphics[height=8cm]{hongmeek.pdf}
+\end{figure}
 

### 2017-02-07 paper/paper.tex
If accepted, this will serve as an outline for the GFLP paper.

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf,float}
\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/plots}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

\newcommand{\myequation}{\begin{equation}}
\newcommand{\myendequation}{\end{equation}}
\let\[\myequation
\let\]\myendequation

\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes}
  \author{Eric Mittman 1\thanks{
    The authors gratefully acknowledge Bill Meeker for his comments and suggestions}\hspace{.2cm}\\
    Department of Statistics, Iowa State University\\
    and \\
    Colin Lewis-Beck \\
    Department of Statistics, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The text of your abstract.  100 or fewer words.
\end{abstract}

\noindent%
{\it Keywords:}  3 to 6 keywords, (don't reuse words appearing in title)
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
\subsection{Background}
This section will follow last paper.

\subsection{Motivation}
This section will follow last paper.

\subsection{Overview}
This section will follow last paper.

\section{Lifetime Model}
\subsection{GFLP Model}
I think we introduce the general GFLP model much like in Meeker and Chan.   
\subsection{Weibull Lifetime Distribution}
We introduce the specific Weibull lifetime distribution and write out the pdf, cdf, and interpretation of the parameters. 
\subsection{Censoring and Truncation}
Similar to last paper, review the issues due to left truncation and write truncation.
\subsection{Data}
Introduce the data and show some summary statistics, but NOT box plots.  Perhaps show an adjusted Kaplan Meier plot to show a Weibull lifetime distribution could be reasonable.

\section{Weibull Model}
\subsection{Likelihood}
Write out the likelihood for the single Weibull with censoring and right truncation.  This can come from the previous paper.  Also mention why MLE model alone by brand has its limitations.
\subsection{Hierarchical Model}
Write out the model, priors, etc for the single Weibull Distribution model.
\subsection{Comparison of Approaches}
Compare MLE estimates to Weibull and show improved precision due to pooling of information.  This can follow the last paper in terms of the plots.  I think this section should be shorter, however, since we are just warming up for the GFLP Model.
\section{GFLP Model}
\subsection{Motivation}
Point out limitation of single failure mode model (perhaps show a drive model with a kink in the distribution) and then present the GFLP Model.  Could also discuss the bathtub hazard.  Mention Wayne Nelson.
\subsection{Model}
Write out the full Bayes hierarchical model for the GFLP model
\subsection{Goodness of Fit}
Show some of the fits of the GFLP Model.  Mention some convergence statistics for the MCMC.  
\subsection{Brand Comparisons}
How can we use this model to compare brands?  Present comparisons of Quantiles, other parameters.  Highlight how this model could be used in an applied sense.  
\section{Concluding Remarks and Extensions}
Review the advantages of fitting the GFLP model and offer future ideas.

\bigskip
\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

\begin{description}

\item Put R Stan code here

\end{description}

\bibliographystyle{plain}
\bibliography{sample}

\end{document}

### 2017-02-10 paper/gflp_model.tex
Go back to non-centered parameterization for sigma1

\documentclass{article}
\usepackage{amsmath}
\newcommand{\ind}{\stackrel{ind.}{\sim}}
\newcommand{\op}{\operatorname}

\begin{document}
\section{Hierarchical GLFP model}
Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
\begin{equation*}
P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)),
\end{equation*}
where $\pi_d$ is in the interval $(0,1)$.

Our interpretation of this model is that, while all units are prone to a mode of failure which can be approximated by a distribution with cdf $F_{d2}$, a proportion $\pi$ are {\em also} defective, and thus subject to another failure mode ($F_{d1}$).

To borrow strength across models, we further model the scales, $\sigma_j$, and log-locations, $\mu_j$, of the component distributions as follows:

\begin{align*}
\sigma_{d,j} &\ind \op{Lognormal} \left( \eta_{\sigma,j}, \tau^2_{\sigma,j} \right) \mbox{ for } j=1,2\; d=1,\ldots,D\\
t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j) & \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D\\
\pi_d &\ind \op{Beta}(\alpha, \beta) \mbox{ for } d=1,\ldots,D.
\end{align*}

Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution. The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = \Phi^{-1}(.5)$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data. For this data we selected $p_1=0.5,\, p_2 = 0.2$.

\subsection{Priors}
To complete the full probability model, we need to select priors such that the posterior distribution is proper. We chose informative priors; we argue that doing so helps us identify the model parameters while still allowing for substantial Bayesian learning about the hierarchical distributions.

\begin{align*}
 \eta_{\sigma ,1} & \sim \op{Normal}(0, 0.5)\\
 \eta_{\sigma ,2} & \sim \op{Normal}(-0.75, .5)\\
 \tau_{\sigma ,1} & \sim \op{Gamma}(1, 2)\\
 \tau_{\sigma ,2} & \sim \op{Gamma}(1, 2)\\
 \eta_{t_{p_1},1} & \sim \op{Normal}(7, 1)\\
 \eta_{t_{p_2},2} & \sim \op{Normal}(9, 2)\\
 \tau_{t_{p_1},1} & \sim \op{Gamma}(1, 2)\\
 \tau_{t_{p_2},2} & \sim \op{Gamma}(1, 2)
 \end{align*}

\end{document}

### 2017-02-10 workflow/functions.R

@@ -33,7 +33,7 @@ run_mcmc <- function(dataset, chains = 4, iter = 2000, warmup = iter/2, p = c(.5
                          p = p)
   )
   
-  m <- stan_model(file = "../stan_models/glfp.stan")
+  m <- stan_model(file = "../stan_models/glfp_logodds_sigma_centered.stan")
   s <- sampling(obj = m, data = stan_data, chains = chains, iter = iter, warmup = warmup, control = list(adapt_delta = .99), init=0)
   return(s)
   # return(stan_data)

### 2017-02-10 workflow/workflow.R

@@ -6,7 +6,7 @@ datasets = commands(
 
 analyses = analyses(
   commands = commands(
-    samples_logodds_2_3 = run_mcmc_logodds(..dataset.., chains = 16, iter = 3000, warmup = 1500)
+    samples_logodds_2_10 = run_mcmc_logodds(..dataset.., chains = 16, iter = 3000, warmup = 1500)
   ),
   datasets = datasets
 )

### 2017-02-10 paper/paper.tex
Update bfcn, add to outliine

@@
 
-\section{Lifetime Model}
-\subsection{GFLP Model}
-I think we introduce the general GFLP model much like in Meeker and Chan.   
-\subsection{Weibull Lifetime Distribution}
-We introduce the specific Weibull lifetime distribution and write out the pdf, cdf, and interpretation of the parameters. 
-\subsection{Censoring and Truncation}
-Similar to last paper, review the issues due to left truncation and write truncation.
-\subsection{Data}
-Introduce the data and show some summary statistics, but NOT box plots.  Perhaps show an adjusted Kaplan Meier plot to show a Weibull lifetime distribution could be reasonable.
+\section{Data}
+Introduce the data and show some summary statistics, but NOT box plots.  Perhaps show an adjusted Kaplan Meier plots to show a Weibull lifetime distribution could be reasonable.
+
+Possible plots (by drive model):
+avg. time on test vs. proportion failed
+total time on test (histogram)
+total records (by drive model)
 
@@
 I think this section should be shorter, however, since we are just warming up for the GFLP Model.
+
 \section{GFLP Model}
@@
 Write out the full Bayes hierarchical model for the GFLP model
-\subsection{Goodness of Fit}
-Show some of the fits of the GFLP Model.  Mention some convergence statistics for the MCMC.  
+
+\subsection{Computation}
+
+\section{Data analysis}
+\subsection{Results}
+Show some of the fits of the GFLP Model.  Show parameter estimates with uncertainty. Mention some convergence statistics for the MCMC. 
+\subsection{collapsing \mu_1, \sigma_1}
+Based on posterior plots, we might fit a simpler model...
 \subsection{Brand Comparisons}
-How can we use this model to compare brands?  Present comparisons of Quantiles, other parameters.  Highlight how this model could be used in an applied sense.  
+How can we use this model to compare brands?  Present comparisons of Quantiles, other parameters.  Highlight how this model could be used in an applied sense. Quantiles by brand, etc.
 \section{Concluding Remarks and Extensions}

### 2017-02-10 plotting_fns/quantile.R

@@ -1,6 +1,6 @@
 ## Load In Samples ##
 
-s <- readRDS("workflow/samples_2_1.rds")
+s <- readRDS("samples_2_1.rds")
 samp <- extract(s)
 
 ## Get Quantiles from GFLP Model ##
@@ -16,20 +16,22 @@ quant <- function(loc1, loc2, scl1, scl2, pi, b, t) {
   }
 
 #Function to Get Quantiles of Interest for All Posterior Draws.  Takes Model Argument and Quantile.
-bfcn <- function(model, b){
+bfcn <- function(model, b, quantiles){
     j <- model 
     Qp <- sapply(1:24000, function(i){
           uniroot(function(t) quant(samp$mu1[i,j],samp$mu2[i,j],samp$sigma1[i,j],samp$sigma2[i,j], exp(samp$log_pi[i,j]), b, t),c(0,300000))$root
   })
-  q <- quantile(Qp, c(.025,.5,.975))
-  data.frame(model=j,lower=q[1],med=q[2],upper=q[3])
+  q <- quantile(Qp, quantiles)
+  df <- data.frame(model=j,t(q))
+  rownames(df) <- NULL
+  names(df)[-1] <- sapply(quantiles, function(q) as.character(round(q,3)))
+  df
 }
 
 #Run this Function for All Models.  Note:  Should be a quick way to combine this with first function.
 df <- NULL
 for (j in 1:21){
   df <- rbind(df,bfcn(j,.1))
-  rownames(df) <- NULL
 }
 
 

### 2017-02-10 workflow/functions.R
centered model

@@ -56,7 +56,7 @@ run_mcmc_logodds <- function(dataset, chains = 4, iter = 10, warmup = iter/2, p 
                          p = p)
   )
   
-  m <- stan_model(file = "../stan_models/glfp_logodds_sigma.stan")
+  m <- stan_model(file = "../stan_models/glfp_logodds_sigma_centered.stan")
   s <- sampling(obj = m, data = stan_data, chains = chains, iter = iter, warmup = warmup, control = list(adapt_delta = .9999), init=0)
   return(s)
   # return(stan_data)

### 2017-03-10 plotting_fns/KM_plot.R
Push .gitignore changes

@@ -76,7 +76,8 @@ get_tr_adj <- function(starttime, pi, loc1, scl1, loc2, scl2){
 }
 
 ## ggplot object containing Weibull probability plot with point estimate and optional credible band
-KM_plot <- function(data, model, tr_adj = 0, title = NULL, linear_axes = FALSE, fixed=FALSE){
+KM_plot <- function(data, model, tr_adj = 0, title = NULL,
+                    linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000), ylimits=c(.0001, .9999)){
   require(ggplot2)
   require(plyr)
   require(dplyr)
@@ -111,9 +112,11 @@ KM_plot <- function(data, model, tr_adj = 0, title = NULL, linear_axes = FALSE, 
       
     } else{
       p <- p +
-        scale_x_continuous(trans = "log", breaks = xbrks(1000, 50000, prec=0), limits=c(1000, 50000)) +
-        scale_y_continuous(trans = ytrans, breaks = ybrks(.0001, .9999, model="weibull"), limits=c(.0001, .9999))
+        scale_x_continuous(trans = "log", breaks = xbrks(xlimits[1], xlimits[2], prec=0), limits=xlimits) +
+        scale_y_continuous(trans = ytrans, breaks = ybrks(ylimits[1], ylimits[2], model="weibull"), limits=ylimits)
     }
+  } else{
+    p <- p + scale_x_continuous(limits=xlimits) + scale_y_continuous(limits=ylimits)
   }
   p <- p +
     theme_bw(base_size = 14) + ggtitle(title)+xlab("Hours")+ylab("Fraction Failing")

### 2017-03-10 plotting_fns/tmp.R

@@ -14,7 +14,7 @@ id <- with(overview, which(overview$early >= 1 & overview$late_f >= 1 & f >=5))
 overview$stan_id <- NA
 overview[id,]$stan_id <- 1:length(id)
 
-s <- readRDS("../workflow/samples_2_1.rds")
+s <- readRDS("../workflow/reduced_relaxed.rds")
 samp <- extract(s)
 
 plot(s, pars="log_pi")
@@ -47,8 +47,8 @@ summary(s)$summary[c("mu1[14]","mu2[14]","log_sigma1[14]", "log_sigma2[14]", "lo
 
 source("../plotting_fns/KM_plot.R")
 
-filter(dat, model==id[8]) %>%
-  KM_plot("weibull")
+filter(dat, model==id[1]) %>%
+  KM_plot("weibull") + xlim(c(200,50000))
 
 filter(dat, model==id[8]) %>%
   KM_plot("weibull") + theme(axis.text=element_blank(),
@@ -61,29 +61,30 @@ plot_list <-NULL
 for(j in 1:length(id)){
   orig_id <- id[j]
   pi_j = exp(samp$log_pi[max_id,j])
-  loc1_j = samp$mu1[max_id,j]
+  loc1_j = samp$mu1[max_id]
   loc2_j = samp$mu2[max_id,j]
-  scl1_j = samp$sigma1[max_id,j]
+  scl1_j = samp$sigma1[max_id]
   scl2_j = samp$sigma2[max_id,j]
   adj <- with(subset(mods, model == orig_id),get_tr_adj(min(starttime), pi_j, loc1_j, scl1_j, loc2_j, scl2_j))
   
   kmp <- subset(mods, model == orig_id) %>%
-    KM_plot(model="weibull", tr_adj = adj, fixed=T)
-  band_df <- data.frame(x = exp(seq.int(log(1000),
+    filter(model != 6 | endtime>1200) %>%
+    KM_plot(model="weibull", tr_adj = adj, fixed=T, linear_axes = T, xlimits=c(200, 50000), ylimits=c(.001,.8))
+  band_df <- data.frame(x = exp(seq.int(log(200),
                                              log(50000),
                                              length.out=25))) %>%
     ddply(.(x), function(g){
       Fp <- sapply(1:1000, function(i) {
[…]

### 2017-03-10 workflow/workflow.R
Relax filter for drive models

@@ -1,7 +1,8 @@
 library(remakeGenerator)
 
 datasets = commands(
-  data_2_21 = prepare_data(lb_fails = 5, lb_late_fails = 1, lb_early_fails = 1)
+  data_2_21 = prepare_data(lb_fails = 5, lb_late_fails = 1, lb_early_fails = 1),
+  data_3_10 = prepare_data(lb_fails = 3, lb_late_fails = 0, lb_early_fails = 0)
 )
 
 analyses = analyses(

### 2017-03-10 constrained_MLE/MLE_stan.R
Updating MLE files.

@@ -37,7 +37,7 @@ dat <- prepare_data(5,1,1,8760)  # 5 total failures, 1 early, 1 late, Less than 
 
 mle_stan <- function(brand,start){
   
-  df<-subset(dat,model==brand)
+  df <- subset(dat,model==brand)
   #Get Data Ready for Stan
   stan_dats <- with(df,
                   list(N_obs = sum(!(censored) & !(mode1)),
@@ -75,7 +75,9 @@ get_mle_stan <- function(model,initial){
   out <- list(est=est, cov=cov, cov2=cov2) #note: point estimates are still in original scale
   attr(out,"model") <- model
   return(out)
-  }
+}
+
+mod2 <- get_mle_stan(2,inits)
   
 #Return Parameters and CI for MLE, i diag of cov2 (mu1,mu2, sigma1,sigma2,pi).
 bandsmle <- function(out,alpha, pm, i){

### 2017-03-10 constrained_MLE/tp_mle.R

@@ -79,14 +79,21 @@ tp_bounds <- function(mle, tp, se, alpha){
   return(out)
 }
 
-## Test the Series of Functions for Model 2 ##
+## Run for All Models ##
+
+tp_all <- matrix(NA, ncol=4, nrow=21)
+
+for (i in 1:21){
+  mod <- i
+  mlefit <- get_mle_stan(mod, inits)  #get MLE estimate
+  tp_mle <- mletp(mlefit,.10)     #use MLE point estimates for tp
+  idev <- impli_deriv(mlefit, tp_mle) #get implicit derivative of GFLP at log(tp)
+  se_ylog <- se_tp(idev, mlefit) #se for log(tp)
+  tp.bounds <- tp_bounds(mlefit, tp_mle, se_ylog, .05) #get 95% tp CI
+  tp_all[i,2:4] <- tp.bounds
+  tp_all[i, 1] <- mod
+}
 
 
-mod2 <- get_mle_stan(2,inits)  #get MLE estimate
-tp_mod2 <- mletp(mod2,.10)     #use MLE point estimates for tp
-idev_mod2 <- impli_deriv(mod2, tp_mod2) #get implicit derivative of GFLP at log(tp)
-se_mod2 <- se_tp(idev_mod2,mod2) #se for log(tp)
-mod2_bounds <- tp_bounds(mod2, tp_mod2, se_mod2, .05) #get 95% tp CI
-mod2_bounds
 
 

### 2017-03-13 stan_models/workflow2.R
Add two Stan models: one with no pooling (expecting problems with sampling for several drive models), one with increasingly relaxed priors, including no upper bound on sigma2[m]

@@ -1,0 +1,16 @@
+library(remakeGenerator)
+
+datasets = commands(
+  data_2_21 = prepare_data(lb_fails = 5, lb_late_fails = 1, lb_early_fails = 1)
+)
+
+analyses = analyses(
+  commands = commands(
+    samples_logodds_nopool_3_13 = run_mcmc_logodds_nopool(..dataset.., chains = 4, iter = 3000, warmup = 1500)
+  ),
+  datasets = datasets
+)
+
+targets = targets(datasets = datasets, analyses = analyses)
+
+workflow(targets, sources = "functions.R", packages = "rstan")

### 2017-03-13 workflow/functions.R

@@ -63,6 +63,30 @@ run_mcmc_logodds <- function(dataset, chains = 4, iter = 10, warmup = iter/2, p 
   # return(stan_data)
 }
 
+run_mcmc_logodds_nopool <- function(dataset, chains = 4, iter = 10, warmup = iter/2, p = c(.5,.2)){
+  require(rstan)
+  rstan_options(auto_write = TRUE)
+  options(mc.cores = parallel::detectCores())#format for Stan
+  stan_data <- with(dataset,
+                    list(M = length(unique(model)),
+                         N_obs = sum(!censored),
+                         N_cens = sum(censored),
+                         starttime_obs = log(starttime[!censored]+1),
+                         starttime_cens = log(starttime[censored]+1),
+                         endtime_obs = log(endtime[!censored]+1),
+                         endtime_cens = log(endtime[censored]+1),
+                         dm_obs = model[!censored],
+                         dm_cens = model[censored],
+                         p = p)
+  )
+  
+  
+  m <- stan_model(file = "../stan_models/glfp_no_pooling.stan")
+  s <- sampling(obj = m, data = stan_data, chains = chains, iter = iter, warmup = warmup, control = list(adapt_delta = .999), init=0)
+  return(s)
+  # return(stan_data)
+}
+
 count_divergences <- function(fit, chain, inc_warmup) {
   sampler_params <- get_sampler_params(fit, inc_warmup=inc_warmup)
   sum(sapply(sampler_params, function(x) c(x[,'divergent__']))[,chain])

### 2017-03-13 workflow2/job2.sh
Set up a 2nd directory for another concurrent job

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --nodes=1 #request one node
+#SBATCH --cpus-per-task=16
+#SBATCH --time=200:00:00 #ask that the job be allowed to run for 150 hours.
+#SBATCH --error=job.%J.err # tell it to store the output console text to a file
+#SBATCH --output=job.%J.out #tell it to store the error messages to a file
+module load R
+srun R --vanilla CMD BATCH make2.R
+make

### 2017-03-13 workflow2/make2.R

@@ -1,0 +1,2 @@
+source("workflow2.R")
+system("make")

### 2017-03-13 plotting_fns/KM_plot.R
Improve code used to create Kaplan-Meier plots

@@ -77,7 +77,7 @@ get_tr_adj <- function(starttime, pi, loc1, scl1, loc2, scl2){
 
 ## ggplot object containing Weibull probability plot with point estimate and optional credible band
 KM_plot <- function(data, model, tr_adj = 0, title = NULL,
-                    linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000), ylimits=c(.0001, .9999)){
+                    linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000), ylimits=c(.0001, .9999), verbose=F){
   require(ggplot2)
   require(plyr)
   require(dplyr)
@@ -101,6 +101,10 @@ KM_plot <- function(data, model, tr_adj = 0, title = NULL,
   df$Ft <- (1 - df$St) * (1 - tr_adj) + tr_adj
   df <- df[which(df$Ft < 1),]
   
+  if(verbose){
+    print(df)
+  }
+  
   p <- df %>%
     ggplot(aes(t, Ft)) + geom_point()
   
@@ -121,6 +125,35 @@ KM_plot <- function(data, model, tr_adj = 0, title = NULL,
   p <- p +
     theme_bw(base_size = 14) + ggtitle(title)+xlab("Hours")+ylab("Fraction Failing")
   p
+}
+
+KM_band <- function(id, samp, n_iter=NULL, xlim, ylim, quantiles=c(.05, .5, .95), x_logscale=T, n=30, verbose=F){
+  total_iter <- dim(samp[[1]])[1]
+  iter <- floor(total_iter/n_iter) * 1:n_iter
+  
+  if(x_logscale){
+    x <- exp(seq(log(xlim[1]), log(xlim[2]), length.out=n))
+  } else{
+    x <- seq(xlim[1], xlim[2], length.out=n)
+  }
+  
+  band <- data.frame(x=x) %>%
+    ddply(.(x), function(g){
+      Fp <- sapply(iter, function(i) {
+        1 -  (1 - exp(samp$log_pi[i,id]) * my_pweibull(g$x, samp$mu1[id], samp$sigma1[id])) *
+          (1 - my_pweibull(g$x, samp$mu2[i,j], samp$sigma2[i,j]))
+      })
+      q <- quantile(Fp, quantiles)
+      data.frame(y = q[2], lower = max(ylim[1],q[1]), upper = min(ylim[2],q[3]))
+    })
+  if(verbose) print(band)
[…]

### 2017-03-24 paper/gflp_model.tex
Update model.tex

@@
 
-Our interpretation of this model is that, while all units are prone to a mode of failure which can be approximated by a distribution with cdf $F_{d2}$, a proportion $\pi$ are {\em also} defective, and thus subject to another failure mode ($F_{d1}$).
+Our interpretation of this model is that,
+\begin{enumerate}
+\item All units of a given drive-model are exchangable
+\item All units of drive-model $d$ have a primary failure mode
+  which can be approximated by a distribution with cdf $F_{d2}$
+\item A fraction of units of drive-model $d$, $\pi_d$, are defective,
+  and thus subject to another failure mode. This failure mode can be
+  approximated by a distribution, $F_{d1}$.
+\end{enumerate}
 
@@
 \begin{align*}
- \eta_{\sigma ,1} & \sim \op{Normal}(0, 0.5)\\
- \eta_{\sigma ,2} & \sim \op{Normal}(-0.75, .5)\\
- \tau_{\sigma ,1} & \sim \op{Gamma}(1, 2)\\
- \tau_{\sigma ,2} & \sim \op{Gamma}(1, 2)\\
- \eta_{t_{p_1},1} & \sim \op{Normal}(7, 1)\\
+  \eta_{\pi} & \sim \op{Normal}(-3, 1)\\
+  \tau_{\pi} & \sim \op{Cauchy}^+(0, 1)\\
+  \eta_{\sigma ,2} & \sim \op{Normal}(0, 2)\\
+  \tau_{\sigma ,2} & \sim \op{Cauchy}^+(0, 1)\\
   \eta_{t_{p_2},2} & \sim \op{Normal}(9, 2)\\
- \tau_{t_{p_1},1} & \sim \op{Gamma}(1, 2)\\
- \tau_{t_{p_2},2} & \sim \op{Gamma}(1, 2)
+  \tau_{t_{p_2},2} & \sim \op{Cauchy}^+(0, 1)
  \end{align*}

### 2017-03-25 paper/paper.tex
Working on intro to paper and added bib file.

@@
 \usepackage{graphicx,psfrag,epsf,float}
-\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/plots}
+\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper}
 \usepackage{enumerate}
@@
 \subsection{Background}
-This section will follow last paper.
+Hard drive failure is the primary cause of data loss and or damage \cite{harris}.
+In addition to the cost of repairing a broken hard drive, lost or corrupted data can have major economic consequences for businesses, researchers, and consumers \cite{smith}.
+To protect against hard drive failure multiple companies now offer backup storage.
+One of the major online backup companies is Backblaze.
+While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.
+The purpose is to provide consumers and businesses with reliability information on different hard drive brands and models.
+The hard drives continuously spin in controlled storage pods where they run until failure.
+When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.
+In addition, the number of storage pods is increasing as Backblaze adds new hard drive brands to the sample.
+Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
+In addition, Backblaze publishes summary statistics of the different hard drive models currently operating.
+No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.
 
 \subsection{Motivation}
-This section will follow last paper.
+The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using parametric lifetime models.
+There are numerous features of the data, however, that make standard estimation approaches problematic.
+The first issue is many hard drives have been running for quite some time before entering the sample.
+Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.
+Also, although we have four years of data, hard drives rarely fail so many units are still in service.
+Thus, we have both left truncation and right censoring.
+The other challenge is each hard drive model varies in number units in operation and total time on test.
+Some models have hundreds of units in service; others have as few as two.
+Combined with the truncation this makes stable maximum likelihood (ML) estimates difficult to obtain for many models.
+\\
+
+Another issue is many product populations contain a mixture of defective and nondefective units.
+The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
+Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.
+A model that generalizes fitting a single failure mode distribution is the general limited failure population model (GFLP) that combines multiple failure time distributions; for example, one failure mode for defective units (infant mortality) and a second mode for wearout \cite{chan}.
+Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model as discussed in Chan and Meeker (1999).
+Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.
+With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.
+This is especially helpful for the Backblaze data as there are many drive models with fewer than 5 failures.
 
 \subsection{Overview}
-This section will follow last paper.
+The structure of the paper is as follows.
+Section 1 presents the Backblaze data and the unique modeling challenges it presents, such as multiple failure modes, right censoring and left truncation.
+Section 2 introduces the GFLP model as a mixture of two Weibull distributions.
+Using the Backblaze data, in Section 3 we fit a 5 parameter GFLP model that allows for separate Weibull terms for each failure mode and estimates $p$, the proportion of units subject to infant mortality.
+In Section 4 we discuss computational and modeling issues that results in us selecting a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
+In Section 5 we compare hard drive brand models based on our final model.
+In Section 6  we examine the posterior distributions of the model parameters to assess goodness of fit for the hierarchical model.
+Finally, in Section 7 we review the final model and propose extensions of the GFLP model.
 
@@
 
-\section{Weibull Model}
+\section{GFLP Model}
 \subsection{Likelihood}

### 2017-03-28 paper/paper.tex
Paper edits.  Intro and Data Section.

@@
 Some models have hundreds of units in service; others have as few as two.
-Combined with the truncation this makes stable maximum likelihood (ML) estimates difficult to obtain for many models.
+Combined with the truncation this makes stable maximum likelihood (ML) estimates difficult to obtain for many drive models.
 \\
@@
 Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.
-A model that generalizes fitting a single failure mode distribution is the general limited failure population model (GFLP) that combines multiple failure time distributions; for example, one failure mode for defective units (infant mortality) and a second mode for wearout \cite{chan}.
-Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model as discussed in Chan and Meeker (1999).
+A model that accounts for multiple failure modes is the general limited failure population model (GFLP) that combines different failure time distributions; for example, one failure mode for defective units (infant mortality) and a second mode for wearout \cite{chan}.
+Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).
 Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.
 With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.
-This is especially helpful for the Backblaze data as there are many drive models with fewer than 5 failures.
+This is especially helpful for the Backblaze data as there are many drive models with as few as one or two failures.
 
@@
 The structure of the paper is as follows.
-Section 1 presents the Backblaze data and the unique modeling challenges it presents, such as multiple failure modes, right censoring and left truncation.
+Section 1 presents the Backblaze data and the unique modeling challenges it presents, such as multiple failure modes, right censoring, and left truncation.
 Section 2 introduces the GFLP model as a mixture of two Weibull distributions.
 Using the Backblaze data, in Section 3 we fit a 5 parameter GFLP model that allows for separate Weibull terms for each failure mode and estimates $p$, the proportion of units subject to infant mortality.
-In Section 4 we discuss computational and modeling issues that results in us selecting a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
+In Section 4 we discuss computational and modeling issues that sugguests a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
 In Section 5 we compare hard drive brand models based on our final model.
 In Section 6  we examine the posterior distributions of the model parameters to assess goodness of fit for the hierarchical model.
-Finally, in Section 7 we review the final model and propose extensions of the GFLP model.
+Finally, in Section 7 we review the GFLP model and propose future extensions.
 
 \section{Data}
-Introduce the data and show some summary statistics, but NOT box plots.  Perhaps show an adjusted Kaplan Meier plots to show a Weibull lifetime distribution could be reasonable.
+As of the first quarter of 2016, Backblaze was collecting data on 63 different hard drive models.
+Some drive models have been running since 2013, while others have been added at a later date.
+There are a total of 75,297 hard drives running.
+The distribution of drives by model varies: some models have only 1 drive in testing; the maximum number of drives running for a specific drive model is 35,860.
+We can also look at the distribution of total number of failures and total time on test across all drive models.
+As seen below, some drive models contain a lot of information, while others have few failures or a little time in operation.
+[INSERT FIGURE 1].
+\\
 
-Possible plots (by drive model):
-avg. time on test vs. proportion failed
-total time on test (histogram)
-total records (by drive model)
+Probability plotting is an effective method to check the adequacy of various parametric models to the data.
+Identifying whether failure data follows a specific distribution is difficult to do by eye: especially when looking at pdfs or other non-linear plots.
+However, if we can linearize the cumulative distribution function it is easier to visually assess a distributional goodness of fit; if a model is appropriate, a non parametric estimate of $\hat{F(t)}$ graphed on linarized probability scales should approximately follow a straight line.
+\\
 
-\section{GFLP Model}
-\subsection{Likelihood}
-Write out the likelihood for the single Weibull with censoring and right truncation.  This can come from the previous paper.  Also mention why MLE model alone by brand has its limitations.
-\subsection{Hierarchical Model}
-Write out the model, priors, etc for the single Weibull Distribution model.
-\subsection{Comparison of Approaches}
-Compare MLE estimates to Weibull and show improved precision due to pooling of information.
-This can follow the last paper in terms of the plots.
-I think this section should be shorter, however, since we are just warming up for the GFLP Model.
+For the hard drive data we first estimate the empirical cdf for each model using the Kaplan-Meier estimator \cite{kaplan}.
+With left truncation, however, the Kaplan-Meier estimate, $\widehat{F(t)_{KM}}$, is conditional based on hard drive $i$ surviving up to time $\tau_i^L$ where $L$ is the amount of time the hard drive was running before Backblaze starting monitoring the drive.
+To get the unconditional distribution we correct the non-parametric estimates using a parametric adjustment outlined by Turnbull, and given in more detail by Meeker and Escobar \cite{turnbull,meeker}.
+For each hard drive model we select $\tau_{\text{min}}^L$, the smallest left truncated time in the sample.
+Then, using the parametric estimates of the parameters, we calculate $Pr(T>\tau_\text{min}^L)$ the probability a hard drive has survived up to $\tau_{\text{min}}^L$.
+We then apply this correction to the non parametric estimates, which results in the unconditional distribution of time to failure.\\
+
+In Figure 2 we plot the Kaplan-Meier adjusted cdf for hard drive model X on Weibull paper.
+Each point on the plot corresponds to a hard drive failure.
+Censored drives are not plotted.
+As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.
+One mode is a result of manufacturing defects, which cause early failures, known as infant mortality.
+The second mode is non-defective hard drives that eventually fail due to wearout.
+Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring around hour Z.
+Therefore, fitting a single Weibull model would not be flexible enough to model the failure distribution.
+[INSERT FIGURE, Do we Want to Put Greenwood Bands on Too?]
 
@@
 \subsection{Motivation}
-Point out limitation of single failure mode model (perhaps show a drive model with a kink in the distribution) and then present the GFLP Model.
-Could also discuss the bathtub hazard.
-Mention Wayne Nelson.
+We suggest the motivation for the GFLP model in the Data Section.  Perhaps expand on this a little more?
+
 \subsection{Model}
@@
 \subsection{Computation}
+Discuss the MCMC approach, RStan, etc.
 
@@
 \subsection{Results}
-Show some of the fits of the GFLP Model.  Show parameter estimates with uncertainty. Mention some convergence statistics for the MCMC. 
+Show some of the fits of the GFLP Model.  Also present the unpooled output for a few models.  Show parameter estimates with uncertainty. Mention some convergence statistics for the MCMC. 
 \subsection{collapsing \mu_1, \sigma_1}

### 2017-03-29 paper/paper.tex
Fix KM_plot_NP so that user can set xlim, ylim

@@
 \section{Introduction}
+We present a analysis of real hard drive failure data exhibiting heavy multiple right censoring and left truncation.
+Our analysis includes estimation of a "bathtub" curve for each drive model in the data set which is made possible by the use of hierarchical model to borrow information across drive models.
+We discuss a principled approach to modeling the data, using Stan.
+
 \subsection{Background}
+
 Hard drive failure is the primary cause of data loss and or damage \cite{harris}.
@@
 Show some of the fits of the GFLP Model.  Also present the unpooled output for a few models.  Show parameter estimates with uncertainty. Mention some convergence statistics for the MCMC. 
-\subsection{collapsing \mu_1, \sigma_1}
+\subsection{collapsing $\mu_1, \sigma_1$}
 Based on posterior plots, we might fit a simpler model...

### 2017-03-29 plotting_fns/greenwood_errors.R

@@ -18,8 +18,8 @@ KM_plot_NP <- function(data, model, tr_adj = 0, title = NULL,
   # df <- with(data, data.frame(time = endtime[!censored]))
   df <- data.frame(t = sort(unique(data$endtime[!(data$censored)]))) %>%
     ddply(.(t), summarise,
-          n = sum(data$starttime < t & data$endtime >=t),
-          d = sum(data$starttime < t & data$endtime == t & data$censored == 0)) %>%
+          n = as.numeric(sum(data$starttime < t & data$endtime >=t)),
+          d = as.numeric(sum(data$starttime < t & data$endtime == t & data$censored == 0))) %>%
     mutate(p = (n-d)/n)
   df$St <- cumprod(df$p)
   df$Vt <- with(df, St * cumsum(d/(n*(n-d))))
@@ -45,8 +45,8 @@ KM_plot_NP <- function(data, model, tr_adj = 0, title = NULL,
   if(!linear_axes){
     if(!fixed){
       p <- p +
-        scale_x_continuous(trans = "log", breaks = xbrks(min(df$t), max(df$t))) +
-        scale_y_continuous(trans = ytrans, breaks = ybrks(min(df$Ft), max(df$Ft), model="weibull"))
+        scale_x_continuous(trans = "log", breaks = xbrks(min(df$t), max(df$t)), limits=xlimits) +
+        scale_y_continuous(trans = ytrans, breaks = ybrks(min(df$Ft), max(df$Ft), model="weibull"), limits=ylimits)
       
     } else{
       p <- p +

### 2017-03-29 workflow/functions.R
Run 'null' model

@@ -43,6 +43,26 @@ prepare_data <- function(dm = NULL, lb_fails = 0, lb_late_fails = 0, lb_early_fa
 #   return(s)
 #   # return(stan_data)
 # }
+run_mcmc_logodds_null <- function(dataset, chains = 4, iter = 10, warmup = iter/2, p = c(.5,.2)){
+  require(rstan)
+  rstan_options(auto_write = TRUE)
+  options(mc.cores = parallel::detectCores())#format for Stan
+  stan_data <- with(dataset,
+                    list(N_obs = sum(!censored),
+                         N_cens = sum(censored),
+                         starttime_obs = log(starttime[!censored]+1),
+                         starttime_cens = log(starttime[censored]+1),
+                         endtime_obs = log(endtime[!censored]+1),
+                         endtime_cens = log(endtime[censored]+1),
+                         p = p)
+  )
+  
+  
+  m <- stan_model(file = "../stan_models/glfp_lo_null_model.stan")
+  s <- sampling(obj = m, data = stan_data, chains = chains, iter = iter, warmup = warmup, control = list(adapt_delta = .99), init=0)
+  return(s)
+  # return(stan_data)
+}
 
 run_mcmc_logodds <- function(dataset, chains = 4, iter = 10, warmup = iter/2, p = c(.5,.2)){
   require(rstan)

### 2017-03-29 workflow_null/workflow_null.R

@@ -1,0 +1,16 @@
+library(remakeGenerator)
+
+datasets = commands(
+  data_all = prepare_data(lb_fails = 0, lb_late_fails = 0, lb_early_fails = 0)
+)
+
+analyses = analyses(
+  commands = commands(
+    samples_lo_null_3_28 = run_mcmc_logodds_null(..dataset.., chains = 4, iter = 3000, warmup = 1500)
+  ),
+  datasets = datasets
+)
+
+targets = targets(datasets = datasets, analyses = analyses)
+
+workflow(targets, sources = "../workflow/functions.R", packages = "rstan")

### 2017-03-29 workflow_pis_vary/workflow_pis_vary.R
Functions for varying pis

@@ -1,0 +1,16 @@
+library(remakeGenerator)
+
+datasets = commands(
+  data_all = prepare_data(lb_fails = 3, lb_late_fails = 0, lb_early_fails = 0)
+)
+
+analyses = analyses(
+  commands = commands(
+    samples_lo_pis_vary_3_28 = run_mcmc_logodds_pis(..dataset.., chains = 4, iter = 3000, warmup = 1500)
+  ),
+  datasets = datasets
+)
+
+targets = targets(datasets = datasets, analyses = analyses)
+
+workflow(targets, sources = "../workflow/functions.R", packages = "rstan")

### 2017-03-29 workflow/functions.R
Add function

@@ -64,6 +64,30 @@ run_mcmc_logodds_null <- function(dataset, chains = 4, iter = 10, warmup = iter/
   # return(stan_data)
 }
 
+run_mcmc_logodds_pis <- function(dataset, chains = 4, iter = 10, warmup = iter/2, p = c(.5,.2)){
+  require(rstan)
+  rstan_options(auto_write = TRUE)
+  options(mc.cores = parallel::detectCores())#format for Stan
+  stan_data <- with(dataset,
+                    list(M = length(unique(model)),
+                         N_obs = sum(!censored),
+                         N_cens = sum(censored),
+                         starttime_obs = log(starttime[!censored]+1),
+                         starttime_cens = log(starttime[censored]+1),
+                         endtime_obs = log(endtime[!censored]+1),
+                         endtime_cens = log(endtime[censored]+1),
+                         dm_obs = model[!censored],
+                         dm_cens = model[censored],
+                         p = p)
+  )
+  
+  
+  m <- stan_model(file = "../stan_models/glfp_lo_pis_vary.stan")
+  s <- sampling(obj = m, data = stan_data, chains = chains, iter = iter, warmup = warmup, control = list(adapt_delta = .99), init=0)
+  return(s)
+  # return(stan_data)
+}
+
 run_mcmc_logodds <- function(dataset, chains = 4, iter = 10, warmup = iter/2, p = c(.5,.2)){
   require(rstan)
   rstan_options(auto_write = TRUE)

### 2017-03-29 paper/paper.tex
Add to introduction

@@
 \section{Introduction}
-We present a analysis of real hard drive failure data exhibiting heavy multiple right censoring and left truncation.
-Our analysis includes estimation of a "bathtub" curve for each drive model in the data set which is made possible by the use of hierarchical model to borrow information across drive models.
-We discuss a principled approach to modeling the data, using Stan.
+We present an approach to modeling failure data with heavy censoring, multiple censoring and multiple truncation.
+We present the generalized limited failure population model of Meeker and Hong which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation).
+This model may provide an adequate fit to failures of complex systems by differentiating between early and late types of failures as well as the propensity toward early failure.
+We discuss the case where the data are grouped, suggesting that we might expand the model, accounting for differences among groups.
+We show that by modeling the data hierarchically, we can deal with groups of various sizes by borrowing information across groups.
+We illustrate our approach on a large set of failure data for hard disk drives made available by Backblaze (citation).
 

### 2017-03-31 bs_hazard/test.R
Tighten up elaborate model slightly

@@ -1,0 +1,72 @@
+library(plyr)
+library(dplyr)
+library(bshazard)
+
+dat <- readRDS("../BB_data/clean_unit_summaries.rds")
+
+overview <- ddply(dat, .(model), summarise,
+                  n=length(model),
+                  f=sum(failed>0),
+                  early_f = sum(failed>0 & endtime<365*24*1),
+                  late_f = sum(failed>0 & endtime>365*24*2))
+
+dat$model <- as.integer(dat$model)
+id <- with(overview, which(overview$early >= 5 & overview$late_f >= 5 & f >=5))
+overview$stan_id <- NA
+overview[id,]$stan_id <- 1:length(id)
+
+fits <- list()
+
+for(i in 1:length(id)){
+Survtmp <- with(filter(dat, model==id[i]), Surv(starttime, endtime, pmin(failed,1)))
+
+fits[[i]] <- bshazard(Survtmp~1, data = dat)
+}
+length(fits)
+pdf("hazards.pdf")
+opar <- par(mfrow=c(3,3))
+  for(i in 1:length(id)){
+    plot(fits[[i]], main=paste(c("Drive model %d", id[i]), collapse=""))
+  }
+dev.off()
+
+#all data
+Survall <- with(filter(dat, model == 7), Surv(starttime, endtime, pmin(failed, 1)))
+fitall <- bshazard(Survall~1, data=dat, phi=7)
+plot(fitall)
+
+#dummy data
+
+my_rweibull <- function(n, mu, sigma){
+  rweibull(n, shape=1/sigma, scale=exp(mu))
+}
+
+rgflp <- function(n, p, mu1, sigma1, mu2, sigma2, censortime=NULL){
+  def <- rbinom(n, size=1, prob=p)
+  out <- ifelse(!def, my_rweibull(n, mu2, sigma2),
+                pmin(my_rweibull(n, mu1, sigma1), my_rweibull(n, mu2, sigma2)))
+  cens <- 0
+  if(!is.null(censortime)){
+    cens <- out>censortime
+    out <- pmin(censortime, out)
+  }
+  round(out)
+  return(data.frame(endtime=out, cens=cens))
+}
+
+par(mfrow=c(3,3))
+for(i in 1:9){
+  dummy <- rgflp(5000, .05, 8, 2, 11, .8, 60000)
+  dummy$died <- dummy$cens != 1
+  dummy$start <- 0
+  
[…]

### 2017-04-02 paper/paper.tex
Additions from our meeting
Paper edits.

@@
 We present an approach to modeling failure data with heavy censoring, multiple censoring and multiple truncation.
-We present the generalized limited failure population model of Meeker and Hong which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation).
+We present the generalized limited failure population model (GFLP) of Chan and Meeker which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation).
 This model may provide an adequate fit to failures of complex systems by differentiating between early and late types of failures as well as the propensity toward early failure.
@@
 \subsection{Background}
-
-Hard drive failure is the primary cause of data loss and or damage \cite{harris}.
-In addition to the cost of repairing a broken hard drive, lost or corrupted data can have major economic consequences for businesses, researchers, and consumers \cite{smith}.
-To protect against hard drive failure multiple companies now offer backup storage.
-One of the major online backup companies is Backblaze.
-While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.
-The purpose is to provide consumers and businesses with reliability information on different hard drive brands and models.
-The hard drives continuously spin in controlled storage pods where they run until failure.
-When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.
-In addition, the number of storage pods is increasing as Backblaze adds new hard drive brands to the sample.
-Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
-In addition, Backblaze publishes summary statistics of the different hard drive models currently operating.
-No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.
+Products can frequently fail due to more than one cause.
+For example, there are many parts in a washing machine that can break and cause the whole machine to fail.
+Decorative tree lighting can fail if one bulb in the chain burns out.
+The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.
+A common assumption is the times to failure for each risk, $s$, are statistically independent.
+Thus, the overall reliability of a unit can be modeled using the product rule across all $s$ risks.
+The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.
+Sometimes the cause of the failure is know; but more often the observations do not contain any information about the failure mode \cite{chan}.
+The GFLP model was first used to describe the early and late failure modes of CB radios.
+No failure information was available for this data, which made fitting the model a challenge using maximum likelihood estimation.
+Other authors, for example, Basu et al., used a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.
+More recently, Ranjan et al.
+considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.
 
@@
 \section{Data}
+Backblaze is a company that offers backup storage to protect against hard drive failure.
+While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.
+The purpose is to provide consumers and businesses with reliability information on different hard drive brands and models.
+The hard drives continuously spin in controlled storage pods where they run until failure.
+When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.
+In addition, the number of storage pods is increasing as Backblaze adds new hard drive brands to the sample.
+Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
+In addition, Backblaze publishes summary statistics of the different hard drive models currently operating.
+No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.
+
 As of the first quarter of 2016, Backblaze was collecting data on 63 different hard drive models.
@@
 \section{GFLP Model}
-\subsection{Motivation}
-We suggest the motivation for the GFLP model in the Data Section.  Perhaps expand on this a little more?
-
 \subsection{Model}
 Write out the full Bayes hierarchical model for the GFLP model
+\subsubsection{Varying proportion defective}
+\subsubsection{Varying wearout mode}
+\subsubsection{Varying defective failure model}
 

### 2017-04-14 paper/glfp_model.tex
Merge model.tex into paper.tex, add options to KM_plot

\documentclass{article}
\usepackage{amsmath}
\usepackage{natbib}
\newcommand{\ind}{\stackrel{ind.}{\sim}}
\newcommand{\op}{\operatorname}

\begin{document}
\section{Hierarchical GLFP model}
For modeling the lifetime of hard-drives, we select the Generalized Limited Failure Population model of \citet{chan}.
Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
$$P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), \mbox{ for }t>0 \mbox{, and where } \pi_d \mbox{ is in }(0,1).$$

As in \cite{chan}, we assume $F_{dj}$ is a member of the Weibull family of cdfs and parameterize in terms of a log-location parameter $\mu_{dj}$ and log-scale parameter $\sigma_{dj}$ so that

$$F_{dj}(t) = 1 - \exp \left\{ -\exp \left\{ \frac{ \log (t) - \mu_{dj}}{\sigma_{dj}} \right\} \right\},\; t>0, j=1,2$$

For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{d1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The interpretation of the parameter $\pi_d$ is to represent the proportion of units susceptible to early failure, hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.

To borrow strength across models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters. For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j} = \Phi^{-1}(\mu_j)$, and proportions, $\pi_d$, of the component distributions as follows:

$$\sigma_{d,j} \ind \op{Lognormal} \left( \eta_{\sigma,j}, \tau^2_{\sigma,j} \right) \mbox{ for } j=1,2\; d=1,\ldots,D$$

$$t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j)  \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D$$

$$\op{logit} \pi_d \ind \op{N}(\eta_pi, \tau_pi) \mbox{ for } d=1,\ldots,D.$$

Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution. The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = \Phi^{-1}(.5)$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data. For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.

We consider the following set of restrictions:

\begin{enumerate}
\item[Model 1:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
\item[Model 2:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
\item[Model 3:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
\item[Model 4:] $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
\end{enumerate}

\subsection{Priors}
To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior. For the scale hyper-parameters we follow the recommendations of \citet{gelman2014bayesian} and use half-Cauchy priors. As for the location hyper-parameters, we selected mildly informative priors consistent with our prior understanding of hard-drives. We believe that using such priors improves the identifiability of the model parameters while still allowing for substantial Bayesian learning about the hierarchical distributions.

\begin{align*}
  \eta_{\pi} & \sim \op{Normal}(-3, 1)\\
  \tau_{\pi} & \sim \op{Cauchy}^+(0, 1)\\
  \eta_{\sigma ,2} & \sim \op{Normal}(0, 2)\\
  \tau_{\sigma ,2} & \sim \op{Cauchy}^+(0, 1)\\
  \eta_{t_{p_2},2} & \sim \op{Normal}(9, 2)\\
  \tau_{t_{p_2},2} & \sim \op{Cauchy}^+(0, 1)
 \end{align*}

 \bibliographystyle{plainnat}
 \bibliography{sample}
\end{document}

### 2017-04-14 paper/paper.tex
Merge model.tex into paper.tex, add options to KM_plot

@@
 \usepackage{url} % not crucial - just used below for the URL
+
+% math macros 
+\newcommand{\ind}{\stackrel{ind.}{\sim}}
+\newcommand{\op}{\operatorname}
 
@@
 
-\section{GFLP Model}
-\subsection{Model}
-Write out the full Bayes hierarchical model for the GFLP model
-\subsubsection{Varying proportion defective}
-\subsubsection{Varying wearout mode}
-\subsubsection{Varying defective failure model}
+\section{Hierarchical GLFP model}
+For modeling the lifetime of hard-drives, we select the Generalized Limited Failure Population model of \citet{chan}.
+Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
+We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
+$$P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), \mbox{ for }t>0 \mbox{, and where } \pi_d \mbox{ is in }(0,1).$$
+
+As in \cite{chan}, we assume $F_{dj}$ is a member of the Weibull family of cdfs and parameterize in terms of a log-location parameter $\mu_{dj}$ and log-scale parameter $\sigma_{dj}$ so that
+
+$$F_{dj}(t) = 1 - \exp \left\{ -\exp \left\{ \frac{ \log (t) - \mu_{dj}}{\sigma_{dj}} \right\} \right\},\; t>0, j=1,2$$
+
+For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{d1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The interpretation of the parameter $\pi_d$ is to represent the proportion of units susceptible to early failure, hence susceptible to both failure modes.
+Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.
+
+To borrow strength across models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters.
+For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j} = \Phi^{-1}(\mu_j)$, and proportions, $\pi_d$, of the component distributions as follows:
+
+$$\sigma_{d,j} \ind \op{Lognormal} \left( \eta_{\sigma,j}, \tau^2_{\sigma,j} \right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
+
+$$t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j)  \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
+
+$$\op{logit} \pi_d \ind \op{N}(\eta_pi, \tau_pi) \mbox{ for } d=1,\ldots,D.$$
+
+Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution.
+The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = \Phi^{-1}(.5)$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
+For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.
+
+We consider the following set of restrictions:
+
+\begin{enumerate}
+\item[Model 1:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
+\item[Model 2:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
+\item[Model 3:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
+\item[Model 4:] $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
+\end{enumerate}
+
+\subsection{Priors}
+To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model.
+We select proper priors to ensure a proper posterior.
+For the scale hyper-parameters we follow the recommendations of \citet{gelman2014bayesian} and use half-Cauchy priors.
+As for the location hyper-parameters, we selected mildly informative priors consistent with our prior understanding of hard-drives.
+We believe that using such priors improves the identifiability of the model parameters while still allowing for substantial Bayesian learning about the hierarchical distributions.
+
+\begin{align*}
+  \eta_{\pi} & \sim \op{Normal}(-3, 1)\\
+  \tau_{\pi} & \sim \op{Cauchy}^+(0, 1)\\
+  \eta_{\sigma ,2} & \sim \op{Normal}(0, 2)\\
+  \tau_{\sigma ,2} & \sim \op{Cauchy}^+(0, 1)\\
+  \eta_{t_{p_2},2} & \sim \op{Normal}(9, 2)\\
+  \tau_{t_{p_2},2} & \sim \op{Cauchy}^+(0, 1)
+ \end{align*}
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{sample}

### 2017-04-14 bs_hazard/test.R

@@ -11,7 +11,7 @@ overview <- ddply(dat, .(model), summarise,
                   late_f = sum(failed>0 & endtime>365*24*2))
 
 dat$model <- as.integer(dat$model)
-id <- with(overview, which(overview$early >= 5 & overview$late_f >= 5 & f >=5))
+id <- with(overview, which(overview$early >= 5 & overview$late_f >= 10))
 overview$stan_id <- NA
 overview[id,]$stan_id <- 1:length(id)
 

### 2017-04-22 paper/paper.tex
If accepted, paper edits.

@@
 We present an approach to modeling failure data with heavy censoring, multiple censoring and multiple truncation.
-We present the generalized limited failure population model (GFLP) of Chan and Meeker which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation).
-This model may provide an adequate fit to failures of complex systems by differentiating between early and late types of failures as well as the propensity toward early failure.
+We extend the generalized limited failure population model (GFLP) of Chan and Meeker which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation).
+This model provides an adequate fit to failures of complex systems by differentiating between early and late types of failures as well as the propensity toward early failure.
 We discuss the case where the data are grouped, suggesting that we might expand the model, accounting for differences among groups.
-We show that by modeling the data hierarchically, we can deal with groups of various sizes by borrowing information across groups.
-We illustrate our approach on a large set of failure data for hard disk drives made available by Backblaze (citation).
+We show that by modeling the data hierarchically, we can deal with groups of various sample sizes by borrowing information across groups.
+We illustrate our approach on a large set of failure data for hard disk drives made available by the hard drive backup company, Backblaze (citation).
 
 \subsection{Background}
-Products can frequently fail due to more than one cause.
-For example, there are many parts in a washing machine that can break and cause the whole machine to fail.
+Products can frequently fail due to more than one reason.
+For example, there are many parts in a washing machine that can break causing the whole machine to fail.
 Decorative tree lighting can fail if one bulb in the chain burns out.
 The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.
-A common assumption is the times to failure for each risk, $s$, are statistically independent.
+A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.
 Thus, the overall reliability of a unit can be modeled using the product rule across all $s$ risks.
 The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.
-Sometimes the cause of the failure is know; but more often the observations do not contain any information about the failure mode \cite{chan}.
-The GFLP model was first used to describe the early and late failure modes of CB radios.
-No failure information was available for this data, which made fitting the model a challenge using maximum likelihood estimation.
+Sometimes the cause of the failure is know; but more often the observations do not contain information about the failure mode.
+The GFLP model was first introduced to describe the early and late failure modes of CB radios \cite{chan}.
+No failure information was available for this data, which made estimation a challenge using maximum likelihood estimation.
 Other authors, for example, Basu et al., used a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.
@@
 \subsection{Motivation}
-The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using parametric lifetime models.
-There are numerous features of the data, however, that make standard estimation approaches problematic.
+The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.
+There are numerous features of the data, however, that make standard estimation approaches for this parametric model problematic.
 The first issue is many hard drives have been running for quite some time before entering the sample.
 Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.
+When an observation is at risk before entering the study, this is known as left truncation.
 Also, although we have four years of data, hard drives rarely fail so many units are still in service.
@@
 Some models have hundreds of units in service; others have as few as two.
-Combined with the truncation this makes stable maximum likelihood (ML) estimates difficult to obtain for many drive models.
+Combined with the truncation and censoring this makes stable maximum likelihood (ML) estimates difficult to obtain for many drive models.
 \\
 
-Another issue is many product populations contain a mixture of defective and nondefective units.
+Another issue is product populations often contain a mixture of defective and nondefective units.
 The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
 Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.
-A model that accounts for multiple failure modes is the general limited failure population model (GFLP) that combines different failure time distributions; for example, one failure mode for defective units (infant mortality) and a second mode for wearout \cite{chan}.
+The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
 Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).
@@
 The structure of the paper is as follows.
-Section 1 presents the Backblaze data and the unique modeling challenges it presents, such as multiple failure modes, right censoring, and left truncation.
+Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.
 Section 2 introduces the GFLP model as a mixture of two Weibull distributions.
-Using the Backblaze data, in Section 3 we fit a 5 parameter GFLP model that allows for separate Weibull terms for each failure mode and estimates $p$, the proportion of units subject to infant mortality.
-In Section 4 we discuss computational and modeling issues that sugguests a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
-In Section 5 we compare hard drive brand models based on our final model.
-In Section 6  we examine the posterior distributions of the model parameters to assess goodness of fit for the hierarchical model.
-Finally, in Section 7 we review the GFLP model and propose future extensions.
+In Section 3 we fit multiple specifications of the GFLP model to best model the Backblaze data.
+We also discuss computational and modeling issues that suggests a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
+In Section 4 we compare hard drive brand models based on our final GFLP model.
+Finally, in Section 5we review the GFLP model and propose future extensions.
 
@@
 Therefore, fitting a single Weibull model would not be flexible enough to model the failure distribution.
-[INSERT FIGURE, Do we Want to Put Greenwood Bands on Too?]
 
@@
 \end{enumerate}
+
+The set of models to compare was chosen based on the data, interpretation of the model, as well as estimation considerations.
+For all model specifications the location and scale parameter of the first failure mode was fixed across drive brand models.
+Partly this was due to a lack of early failures in the data.
+Moreover, though, from a consumer or business standpoint, comparing the lifetime distribution of infant mortality across hard drive brand models is not particularly informative.
+The more practical questions are: is this hard drive likely to suffer from infant morality?
+and how long should it expect to run before it fails from wearout?
+Assuming a constant failure distribution for infant mortality allows the first question to be answered since $\pi$'s are comparable across brands, which would not be the case if each drive brand model had a unique Weibull distribution fit to the early failure mode.
+Flexibility in the parameters for the 2nd failure mode is necessary as initial plots of the data indicate heterogeneity across brands in the right tails of the failure distribution.
+Going from a fixed model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4, which has the most parameters.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by brand drive model.
 

### 2017-04-26 plots/BB_kaplan_meier_plots.R
Paper updates

@@ -15,7 +15,8 @@ id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
 overview$stan_id <- NA
 overview[id,]$stan_id <- 1:length(id)
 
-s <- readRDS("../workflow/samples_lor_only3fails.rds")
+# s <- readRDS("../workflow/samples_lor_only3fails.rds")
+s <- readRDS("../workflow_tp2_vary/samples_tp2_vary.rds")
 samp <- extract(s)
 
 source("../plotting_fns/KM_plot.R")
@@ -25,17 +26,18 @@ max_id <- which.max(samp$lp__)
 plot_list <-NULL
 for(j in 1:length(id)){
   orig_id <- id[j]
-  pi_j = exp(samp$log_pi[max_id,j])
+  pi_j = exp(samp$log_pi[max_id])
   loc1_j = samp$mu1[max_id]
   loc2_j = samp$mu2[max_id,j]
   scl1_j = samp$sigma1[max_id]
-  scl2_j = samp$sigma2[max_id,j]
+  scl2_j = samp$sigma2[max_id]
   adj <- with(subset(mods, model == orig_id),get_tr_adj(min(starttime), pi_j, loc1_j, scl1_j, loc2_j, scl2_j))
   
   dat_tmp <- subset(mods, model == orig_id)
 
   kmb <- KM_with_band(data = dat_tmp, id =j, samp = samp, n_iter = 200, n = 50, quantiles = c(.05,.5,.95), tr_adj = adj,
-                      xlimits = c(0,50000), ylimits = c(0,.9), fixed = T, linear_axes = T, verbose = F, model = "weibull")
+                      xlimits = c(0,50000), ylimits = c(0,.9), fixed = T, linear_axes = T, verbose = F, model = "weibull",
+                      pi.free=F, sigma2.free=F)
   
   plot_list[[j]] <- kmb
 }

### 2017-04-26 paper/paper.tex

@@
 \subsection{Background}
-Products can frequently fail due to more than one reason.
+Often products can fail due to more than one failure mode.
 For example, there are many parts in a washing machine that can break causing the whole machine to fail.
@@
 A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.
-Thus, the overall reliability of a unit can be modeled using the product rule across all $s$ risks.
+Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.
 The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.
-Sometimes the cause of the failure is know; but more often the observations do not contain information about the failure mode.
 The GFLP model was first introduced to describe the early and late failure modes of CB radios \cite{chan}.
-No failure information was available for this data, which made estimation a challenge using maximum likelihood estimation.
-Other authors, for example, Basu et al., used a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.
+Chan and Meeker used maximum likelihood (ML) estimation to estimate the model parameters, and developed likelihood based approximation methods to construct confidence intervals for functions of model parameters.
+Other authors, for example, Basu et al., have employed a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.
 More recently, Ranjan et al.
 considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.
+There has not been, however, any extensions of the GFLP model to a hierarchical framework.
 
@@
 Some models have hundreds of units in service; others have as few as two.
-Combined with the truncation and censoring this makes stable maximum likelihood (ML) estimates difficult to obtain for many drive models.
+Combined with the truncation and censoring this makes stable ML estimates difficult to obtain for many drive models.
 \\
@@
 With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.
-This is especially helpful for the Backblaze data as there are many drive models with as few as one or two failures.
+This is especially helpful for the Backblaze data as there are many drive models with one or two failures.
 
@@
 Section 2 introduces the GFLP model as a mixture of two Weibull distributions.
-In Section 3 we fit multiple specifications of the GFLP model to best model the Backblaze data.
-We also discuss computational and modeling issues that suggests a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
+In Section 3 we fit multiple specifications of the GFLP model to best describe the Backblaze data.
+We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
 In Section 4 we compare hard drive brand models based on our final GFLP model.
-Finally, in Section 5we review the GFLP model and propose future extensions.
+Finally, in Section 5, we review the GFLP model and propose future extensions.
 

### 2017-05-03 paper/paper.tex
Paper edits and plots.

@@
 
-The set of models to compare was chosen based on the data, interpretation of the model, as well as estimation considerations.
-For all model specifications the location and scale parameter of the first failure mode was fixed across drive brand models.
+The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations.
+For all 4 models the location and scale parameter of the first failure mode was fixed across drive brand models.
 Partly this was due to a lack of early failures in the data.
@@
 The more practical questions are: is this hard drive likely to suffer from infant morality?
-and how long should it expect to run before it fails from wearout?
-Assuming a constant failure distribution for infant mortality allows the first question to be answered since $\pi$'s are comparable across brands, which would not be the case if each drive brand model had a unique Weibull distribution fit to the early failure mode.
-Flexibility in the parameters for the 2nd failure mode is necessary as initial plots of the data indicate heterogeneity across brands in the right tails of the failure distribution.
-Going from a fixed model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4, which has the most parameters.
+and how long should it expect to run before failing from wearout?
+Fixed failure parameters for infant mortality allows the first question to be answered since $\pi$'s are comparable across brands.
+Drive brand specific parameters for the 2nd failure mode seemed reasonable as plots of the data indicate heterogeneity across brands in the right tails of the failure distribution.
+Going from a fixed model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
 Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by brand drive model.
@@
 \subsection{Results}
-Show some of the fits of the GFLP Model.  Also present the unpooled output for a few models.  Show parameter estimates with uncertainty. Mention some convergence statistics for the MCMC. 
-\subsection{collapsing $\mu_1, \sigma_1$}
-Based on posterior plots, we might fit a simpler model...
+Show some of the fits of the GFLP Model.  Also present the unpooled output for a few models.  Show parameter estimates with uncertainty.
+
+\subsection{Model Comparisons}
+First look at the overlaid posterior credible bands. \\
+
+In addition to graphically examining model fits via posterior credible bands, we compared the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
+The predictive accuracy from a fitted Bayesian model can be estimated simply with posterior simulations of the model parameters \cite{vehtari}.
+For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior parameters with the $i$th data point removed.
+The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\logp(y_i|y_{-i}$)}).
+To obtain a more stable estimate of elpd, Vehtari et.
+al suggest using Pareto smoothed importance weights (PSIS) since the tail of the importance weights used to estimate elpd is often long.
+We computed the elpd for all 4 models using the R package loo \cite{loopackage}.
+When $n$ is large, the distribution of the elpd is approximately normal and models can be statistically compared.
+We calculated the difference in expected predictive accuracy for all 4 models as well as the standard error of the difference.
+As the model complexity increased, the predictive accuracy improved (Table 1).
+Of all the models, Model 4 had the best predictive accuracy and was significantly better than models 1-3.
+
 \subsection{Brand Comparisons}
-How can we use this model to compare brands?  Present comparisons of Quantiles, other parameters.  Highlight how this model could be used in an applied sense. Quantiles by brand, etc.
+Ideas for plots: A graph comparing the pi's for all drive models.
+
+I think a graph looking at the B10 using the full model could be good, too.  Something like the graph below.  Perhaps we could sort drive model based on sample size?
+
+\begin{figure}[H]
+    \centering
+   \includegraphics[width=6.0in]{B10_full.pdf}
+		\caption{B10 in Years for All Drive Models \label{fig:first}} 
+\end{figure}
+
 \section{Concluding Remarks and Extensions}

### 2017-05-03 plotting_fns/quantile.R

@@ -1,7 +1,8 @@
 ## Load In Samples ##
+library(rstan)
 
 # s <- readRDS("reduce_relaxed.rds")
-s <- readRDS("samples_2_1.rds")
+s <- readRDS("MCMC_draws/samples_lor_only3fails.rds")
 samp <- extract(s)
 
 ## Get Quantiles from GFLP Model ##
@@ -34,7 +35,7 @@ bfcn2 <- function(model, b, samp, quantiles = c(.025, .5, .975)){
   
   j <- model 
   Qp <- sapply(1:dim(samp$lp__), function(i){
-    uniroot(function(t) quant(samp$mu1[i],samp$mu2[i,j],samp$sigma1[i],samp$sigma2[i,j], exp(samp$log_pi[i,j]), b, t),c(0,300000))$root
+    uniroot(function(t) quant(samp$mu1[i],samp$mu2[i,j],samp$sigma1[i],samp$sigma2[i,j], exp(samp$log_pi[i,j]), b, t),c(1,3000000))$root
   })
   q <- quantile(Qp, quantiles)
   df <- data.frame(model=j,t(q))
@@ -50,8 +51,34 @@ for (j in 1:21){
 }
 
 df <- NULL
-for (j in 1:21){
-  df <- rbind(df,bfcn(j,.1, samp))
+for (j in 1:44){
+  df <- rbind(df,bfcn2(j,.1, samp))
 }
+
+
+
+#B10_Model4.rds is in paper folder for future use
+b10 <- readRDS("B10_Model4.rds")
+colnames(b10) <- c("model","lb","med","ub")
+b10$lb <- b10$lb/(24*365)
+b10$med <- b10$med/(24*365)
+b10$ub <- b10$ub/(24*365)
+
+
+#Make Catepillar Plot for B10; Perhaps Sort by Sample Size?
+p <- ggplot(b10, aes(x=as.factor(model), y=med, ymin=lb, ymax=ub)) +
+  geom_pointrange() +
+  coord_flip() +
+  xlab('Drive Model') + theme_bw() + 
+  ylab("Years")   +
+  scale_x_discrete(breaks=seq(1,44,1)) + 
+p
+
+
+
+
+
+
+
 
 xtable::xtable(cbind(df, df_reduced), digits = 0)

### 2017-05-08 paper/paper.tex
Prepend to intro, fix typo in model compare

@@
 \section{Introduction}
-We present an approach to modeling failure data with heavy censoring, multiple censoring and multiple truncation.
-We extend the generalized limited failure population model (GFLP) of Chan and Meeker which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation).
-This model provides an adequate fit to failures of complex systems by differentiating between early and late types of failures as well as the propensity toward early failure.
-We discuss the case where the data are grouped, suggesting that we might expand the model, accounting for differences among groups.
-We show that by modeling the data hierarchically, we can deal with groups of various sample sizes by borrowing information across groups.
-We illustrate our approach on a large set of failure data for hard disk drives made available by the hard drive backup company, Backblaze (citation).
+Failure data for series systems may be collected at the system level by end-users interested only in the lifetime of the system.
+In the case of complex, high-reliability systems where cause of failure is not available, a model of failure which
+\begin{enumerate}[a]
+\item can be estimated from the observed data
+\item is flexible enough to fit the observed data
+\item relies on an approximation of the understood underlying process
+\end{enumerate}
+is desireable.
+We present a Bayesian, hierarchical modeling approach for grouped failure data.
+We demonstrate this approach with a large, heavily censored real data set with multiple right-censoring and left truncation times.
+We extend the generalized limited failure population model (GLFP) by \citet{chan}, which provides a generative model for failure times in a population with infant mortality.
+This model is an example of a bathtub distribution, exhibiting distinct early and late failure modes.
+The GLFP model assumes there are two causes of failure, one of which affects only a fraction of the population, and that the cause of failure is unobserved.
+We extend this model to grouped data and show that by borrowing information across groups we can acheieve good model fits for almost all groups.
+By working with posterior samples, we can easily make inference on a wide range of quantities of interest while accounting for uncertainty.
 
@@
 For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior parameters with the $i$th data point removed.
-The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\logp(y_i|y_{-i}$)}).
+The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
 To obtain a more stable estimate of elpd, Vehtari et.

### 2017-05-10 paper/paper.tex
Small edit

@@
 
-$$\sigma_{d,j} \ind \op{Lognormal} \left( \eta_{\sigma,j}, \tau^2_{\sigma,j} \right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
-
+$$\sigma_{d,1} \ind \op{Lognormal} \left( \eta_{\sigma,1}, \tau^2_{\sigma,1} \right) \mbox{ for } d=1,\ldots,D$$
+$$\sigma_{d,2} \ind \op{Truncated-Lognormal} \left( \eta_{\sigma,2}, \tau^2_{\sigma,2} , 0, 1, \right) \mbox{ for } d=1,\ldots,D$$
 $$t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j)  \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
@@
 Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution.
+We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that wearout implies an increasing hazard function.
 The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = \Phi^{-1}(.5)$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
@@
 
-We consider the following set of restrictions:
+We consider the models with the following set of restrictions (from most to least restrictive):
 
 \begin{enumerate}
-\item[Model 1:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
-\item[Model 2:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
-\item[Model 3:] $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
-\item[Model 4:] $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
+\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
+\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
+\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
+\item $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
 \end{enumerate}
@@
 The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations.
-For all 4 models the location and scale parameter of the first failure mode was fixed across drive brand models.
-Partly this was due to a lack of early failures in the data.
-Moreover, though, from a consumer or business standpoint, comparing the lifetime distribution of infant mortality across hard drive brand models is not particularly informative.
-The more practical questions are: is this hard drive likely to suffer from infant morality?
-and how long should it expect to run before failing from wearout?
-Fixed failure parameters for infant mortality allows the first question to be answered since $\pi$'s are comparable across brands.
-Drive brand specific parameters for the 2nd failure mode seemed reasonable as plots of the data indicate heterogeneity across brands in the right tails of the failure distribution.
-Going from a fixed model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
+Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
+Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
 Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by brand drive model.
+
+For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive brand models.
+We found that there was insufficient information in the data to model these parameters hierarchically.
+Moreover, assuming a common distribution for infant mortality allows for straightforward interpretation of the quantities like $\frac{\pi_{d}}{\pi_{d'}}$.
 
@@
 For the scale hyper-parameters we follow the recommendations of \citet{gelman2014bayesian} and use half-Cauchy priors.
-As for the location hyper-parameters, we selected mildly informative priors consistent with our prior understanding of hard-drives.
-We believe that using such priors improves the identifiability of the model parameters while still allowing for substantial Bayesian learning about the hierarchical distributions.
+As for the location hyper-parameters, we select weakly informative priors consistent with our prior information on hard-drives.
+The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification?
+Should maybe be lower.) For $\eta_{\sigma, 2}$, 95\% of prior mass is on values greater than $0.037$.
+Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this correponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$.
+The prior for $\eta_{t_{p_2},2}$ implies that the median 20th percentile for non-defective units is less than greater than 3 days and less than 24 years.
 
@@
   \tau_{\sigma ,2} & \sim \op{Cauchy}^+(0, 1)\\
-  \eta_{t_{p_2},2} & \sim \op{Normal}(9, 2)\\
-  \tau_{t_{p_2},2} & \sim \op{Cauchy}^+(0, 1)
+  \eta_{t_{.2_2},2} & \sim \op{Normal}(9, 2)\\
+  \tau_{t_{.2_2},2} & \sim \op{Cauchy}^+(0, 1)
  \end{align*}

### 2017-05-10 paper/paper.tex
Revert "Small edit"
This reverts commit a3c23a10574b2b4c124b0bb16e9f3f4ade2d96e3.

@@
 The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations.
-Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
+Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered to account for heterogeneity across brands in the right tails of the failure distribution.
 Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
@@
 The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification?
-Should maybe be lower.) For $\eta_{\sigma, 2}$, 95\% of prior mass is on values greater than $0.037$.
+Should maybe be lower.) For $\eta__{\sigma, 2}$, 95\% of prior mass is on values greater than $0.037$.
 Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this correponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$.

### 2017-05-10 paper/paper.tex
Revert "Modified model and priors sections"
This reverts commit 04f131ee55044338130f0925a8c7d74ee45d273f.

--- 2017-05-08 paper/paper.tex
@@
 \end{enumerate}
-is desireable.
+is desirable.
 We present a Bayesian, hierarchical modeling approach for grouped failure data.
@@
 The GLFP model assumes there are two causes of failure, one of which affects only a fraction of the population, and that the cause of failure is unobserved.
-We extend this model to grouped data and show that by borrowing information across groups we can acheieve good model fits for almost all groups.
+We extend this model to grouped data and show that by borrowing information across groups we can achieve good model fits for almost all groups.
 By working with posterior samples, we can easily make inference on a wide range of quantities of interest while accounting for uncertainty.
@@
 \subsection{Background}
-Often products can fail due to more than one failure mode.
-For example, there are many parts in a washing machine that can break causing the whole machine to fail.
-Decorative tree lighting can fail if one bulb in the chain burns out.
-The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.
+In engineering applications, a product can often fail due to one out of a set of possible malfunctioning components.
+For example, a computer system can fail if the mother board, disc drive or power supply stop working.
+Circuit boards (CB) can fail due to a manufacturing defect or later as a result of normal wearout.
+The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \citet{nelson}.
 A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.
 Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.
-The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.
-The GFLP model was first introduced to describe the early and late failure modes of CB radios \cite{chan}.
-Chan and Meeker used maximum likelihood (ML) estimation to estimate the model parameters, and developed likelihood based approximation methods to construct confidence intervals for functions of model parameters.
-Other authors, for example, Basu et al., have employed a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.
+Parameter estimation is straightforward if the cause of failure is known for each observation.
+With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the researcher.
+\\
+
+Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.
+When modeling computer system failures Reiser et al.
+assumed the cause of failure for each observation came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.
+Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.
+This helped identify parameters when using maximum likelihood (ML) estimation.
+Extending Chan and Meeker's analysis, Basu et.
+al used a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.
+Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.
 More recently, Ranjan et al.
 considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.
-There has not been, however, any extensions of the GFLP model to a hierarchical framework.
+Treating the unknown failure modes as incomplete data, an expectation maximization algorithm with ML was used, in addition to Bayesian estimation.
 
@@
 The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.
-There are numerous features of the data, however, that make standard estimation approaches for this parametric model problematic.
-The first issue is many hard drives have been running for quite some time before entering the sample.
-Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.
-When an observation is at risk before entering the study, this is known as left truncation.
-Also, although we have four years of data, hard drives rarely fail so many units are still in service.
-Thus, we have both left truncation and right censoring.
-The other challenge is each hard drive model varies in number units in operation and total time on test.
-Some models have hundreds of units in service; others have as few as two.
-Combined with the truncation and censoring this makes stable ML estimates difficult to obtain for many drive models.
-\\
-
-Another issue is product populations often contain a mixture of defective and nondefective units.
+Product populations often contain a mixture of defective and nondefective units.
 The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
@@
 The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
-Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).
+Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are masked, which presents challenges for parameter identification in the GFLP model.
 Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.
 With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.
-This is especially helpful for the Backblaze data as there are many drive models with one or two failures.
+In cases where drive models with only one or two failures the hierarchical model shrinks unstable parameter estimates back towards the overall group average.
+This is advantageous since for many groups it was not possible to estimate the GFLP model individually by drive model.
 

### 2017-05-10 paper/paper.tex
More edits to the Motivation Section.

@@
 \subsection{Motivation}
-The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.
-Product populations often contain a mixture of defective and nondefective units.
+The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GFLP lifetime model.
+Product populations often contain a mixture of defective and non defective units.
 The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
-Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.
+Ignoring this hazard structure, which exists for series systems and computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.
 The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
-Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are masked, which presents challenges for parameter identification in the GFLP model.
+Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are masked, which presents challenges for parameter identification: especially with sample sample sizes.
+One option is to combine all drive models and estiamte an overall GFLP model.
+While this model can be precisely estimated, the data indicate a common model for all drive models fits well for only a few drive models.
+At the other extreme, fitting individual models to each hard drive brand excludes drive brand models with one or two failures.
+The GFLP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.
 Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.
-With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.
-In cases where drive models with only one or two failures the hierarchical model shrinks unstable parameter estimates back towards the overall group average.
-This is advantageous since for many groups it was not possible to estimate the GFLP model individually by drive model.
+With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of individual drive model parameters.
+Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures can be estimated precisely; and in cases where little data is available the hierarchical model borrows information from other models to obtain stable estimates.
+This allows us to include all of the hard drive data and more importantly, make reliability inference for all the drive models in the data set.
 

### 2017-05-11 paper/paper.tex
More paper edits (Model Comparison) and updated references bibtek.

@@
 \subsection{Background}
-In engineering applications, a product can often fail due to one out of a set of possible malfunctioning components.
+In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.
 For example, a computer system can fail if the mother board, disc drive or power supply stop working.
-Circuit boards (CB) can fail due to a manufacturing defect or later as a result of normal wearout.
+Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.
 The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \citet{nelson}.
@@
 When modeling computer system failures Reiser et al.
-assumed the cause of failure for each observation came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.
+assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.
 Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.
@@
 Extending Chan and Meeker's analysis, Basu et.
-al used a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.
+al performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.
 Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.
@@
 The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
-Ignoring this hazard structure, which exists for series systems and computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.
+Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{sujata}.
 The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
@@
 With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of individual drive model parameters.
-Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures can be estimated precisely; and in cases where little data is available the hierarchical model borrows information from other models to obtain stable estimates.
+Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive brand models to obtain stable estimates.
 This allows us to include all of the hard drive data and more importantly, make reliability inference for all the drive models in the data set.
@@
 In addition to graphically examining model fits via posterior credible bands, we compared the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
-The predictive accuracy from a fitted Bayesian model can be estimated simply with posterior simulations of the model parameters \cite{vehtari}.
-For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior parameters with the $i$th data point removed.
+As outlined in Vehtari et.
+al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters \citet{vehtari}.
+For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with the $i$th data point removed.
 The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
@@
 al suggest using Pareto smoothed importance weights (PSIS) since the tail of the importance weights used to estimate elpd is often long.
-We computed the elpd for all 4 models using the R package loo \cite{loopackage}.
+We computed the elpd for all 4 models using the R package loo \citet{loo}.
 When $n$ is large, the distribution of the elpd is approximately normal and models can be statistically compared.
-We calculated the difference in expected predictive accuracy for all 4 models as well as the standard error of the difference.
-As the model complexity increased, the predictive accuracy improved (Table 1).
-Of all the models, Model 4 had the best predictive accuracy and was significantly better than models 1-3.
+We calculated the difference in expected predictive accuracy for Model 4 vs 3, Model 3 vs 2, and Model 2 vs 1, as well as the standard error of the difference (Table 1).
+As the model complexity increased, the expected predictive accuracy improved.
+Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 models.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & elpd\ & Diff(elpd)\ & SE(Diff) \\ 
+  \hline
+Model 4 & -13309.5 & 40.7 & 11.3 \\ 
+Model 3 & -13350.2 & 458.8 & 31.0  \\ 
+Model 2 & -13809.0 & 3674.6 & 96.2 \\ 
+Model 1 & -17483.6  \\ 
+   \hline
+\end{tabular}
+\caption{Model Comparison Based on LOO}
+\label{table:1}
+\end{table}
 

### 2017-05-12 paper/paper.tex
Work on intro, remove redundant bibtex entry

@@
 \section{Introduction}
-Failure data for series systems may be collected at the system level by end-users interested only in the lifetime of the system.
-In the case of complex, high-reliability systems where cause of failure is not available, a model of failure which
-\begin{enumerate}[a]
-\item can be estimated from the observed data
+Failure distributions of high-reliability products, can be difficult
+to assess. Accelerated testing tries to address this problem, but
+requires strong assumptions to extrapolate to use conditions in the
+field (citation). When field data are available, information may be
+very limited. In this paper, we propose the use of Bayesian hierarchical modeling to borrow
+information to improve inferences on lifetime
+distributions where information is limited by right-censoring and
+left-truncation.
+
+Failure data for series systems may be collected at the system level
+by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
+model (GLFP), which provides a generative model for
+failure times in a population where some units are susceptible to infant mortality. This model
+covers a subset of the class of distributions known as bathtub distributions
+\cite{sujata}. Bathtub distributions are characterized by a U-shaped
+hazard function, where early failures are due to infant mortality and
+late failure are due to wearout. While complex systems actually
+exhibit more than two modes of failure, suitable parametric models
+which can suitably approximate the overall lifetime distribution are
+useful.
+
+In the case of complex, high-reliability systems where cause of failure is
+not available, a model which
+\begin{enumerate}[a.]
+\item has parameters that can be estimated from the observed data
 \item is flexible enough to fit the observed data
-\item relies on an approximation of the understood underlying process
+\item in interpretable with respect to available theory of the underlying process
 \end{enumerate}
-is desirable.
-We present a Bayesian, hierarchical modeling approach for grouped failure data.
-We demonstrate this approach with a large, heavily censored real data set with multiple right-censoring and left truncation times.
-We extend the generalized limited failure population model (GLFP) by \citet{chan}, which provides a generative model for failure times in a population with infant mortality.
-This model is an example of a bathtub distribution, exhibiting distinct early and late failure modes.
-The GLFP model assumes there are two causes of failure, one of which affects only a fraction of the population, and that the cause of failure is unobserved.
-We extend this model to grouped data and show that by borrowing information across groups we can achieve good model fits for almost all groups.
-By working with posterior samples, we can easily make inference on a wide range of quantities of interest while accounting for uncertainty.
+is desired. We present a Bayesian, hierarchical modeling approach for
+grouped failure data. As a benefit of working with Monte Carlo
+samples, we can easily make inference on a wide range of quantities of
+interest while accounting for uncertainty. We demonstrate this approach on real data hard
+drive failure data from a
+cloud-based storage company.
 

### 2017-05-12 paper/paper.tex
Begin computation subsection, add references

@@
 \subsection{Computation}
-Discuss the MCMC approach, RStan, etc.
+Each model was fit in {\t R} \nocite{r} using Hamiltonian Monte Carlo
+\cite{betancourt} as implemented in {\tt
+  rstan}\cite{rstan}. Multiple chain were run with _000 iterations of
+warmup. Convergence was assessed by examining posterior
+plots and checking that Gelman-Rubin diagnostics were close to 1.
 

### 2017-05-16 paper/paper.tex
Made some edits to Motivation.  Even using \cite{} I am still getting the whole name of the author put into the citation when I just want the year of publication.  Not sure what's going on.  Also, added a Figure folder with contents, so it now should appear in the repo.

@@
 \usepackage{graphicx,psfrag,epsf,float}
-\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper}
+\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
 \usepackage{enumerate}
@@
 Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.
-The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \citet{nelson}.
+The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.
 A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.
@@
 \subsection{Motivation}
-The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GFLP lifetime model.
+The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GLFP lifetime model.
 Product populations often contain a mixture of defective and non defective units.
@@
 Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{sujata}.
-The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
-Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are masked, which presents challenges for parameter identification: especially with sample sample sizes.
-One option is to combine all drive models and estiamte an overall GFLP model.
-While this model can be precisely estimated, the data indicate a common model for all drive models fits well for only a few drive models.
-At the other extreme, fitting individual models to each hard drive brand excludes drive brand models with one or two failures.
-The GFLP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.
-Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.
-With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of individual drive model parameters.
+The GLFP combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
+\\
+
+While the GLFP model allows for multiple hazard functions, it does not account for the group structure of the hard drive data.
+A naive approach would combine all drive models and fit an overall GLFP model.
+This model can be easily estimated, but a common GLFP model for all drive brand groups fits poorly: the model captures drive models with lots of observed failures, while the predicted time to failure for drive brand models with less information is over or under estimated.
+At the other extreme, fitting individual models to each hard drive brand excludes many drive brand models with fewer than 5 failures.
+The GLFP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.
+\\
+
+We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
+With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of the GLFP parameters for each drive model group.
 Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive brand models to obtain stable estimates.
-This allows us to include all of the hard drive data and more importantly, make reliability inference for all the drive models in the data set.
+This allows us to include all of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
 
@@
 Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.
-Section 2 introduces the GFLP model as a mixture of two Weibull distributions.
-In Section 3 we fit multiple specifications of the GFLP model to best describe the Backblaze data.
+Section 2 introduces the GLFP model as a mixture of two Weibull distributions.
+In Section 3 we fit multiple specifications of the GLFP model to best describe the Backblaze data.
 We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
-In Section 4 we compare hard drive brand models based on our final GFLP model.
-Finally, in Section 5, we review the GFLP model and propose future extensions.
+In Section 4 we compare hard drive brand models based on our final GLFP model.
+Finally, in Section 5, we review the GLFP model and propose future extensions.
 
@@
 As outlined in Vehtari et.
-al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters \citet{vehtari}.
+al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters \cite{vehtari}.
 For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with the $i$th data point removed.
 The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
-To obtain a more stable estimate of elpd, Vehtari et.
-al suggest using Pareto smoothed importance weights (PSIS) since the tail of the importance weights used to estimate elpd is often long.
-We computed the elpd for all 4 models using the R package loo \citet{loo}.
+We computed the elpd for all 4 models using the R package loo \cite{loo}.
 When $n$ is large, the distribution of the elpd is approximately normal and models can be statistically compared.
@@
     \centering
-   \includegraphics[width=6.0in]{B10_full.pdf}
+   \includegraphics[width=5.0in]{B10_full.pdf}
 		\caption{B10 in Years for All Drive Models \label{fig:first}} 

### 2017-05-17 paper/paper.tex
Adjusted KM_plot_multi to be a step function, added figure, added to Results

@@
 \section{Data}
-Backblaze is a company that offers backup storage to protect against hard drive failure.
+Backblaze is a company that offers backup storage to protect against data loss.
 While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.
@@
 There are a total of 75,297 hard drives running.
-The distribution of drives by model varies: some models have only 1 drive in testing; the maximum number of drives running for a specific drive model is 35,860.
+The distribution of drives by model varies: some models have only have a service record for a single drive; the maximum number of service records for a single drive model is 35,860.
 We can also look at the distribution of total number of failures and total time on test across all drive models.
@@
 
-In Figure 2 we plot the Kaplan-Meier adjusted cdf for hard drive model X on Weibull paper.
+\begin{figure}
+\includegraphics[width=\textwidth]{fig/heterogeneity_compare_preliminary.pdf}
+\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
+Right: Pointwise posterior median time to failure curves for Models 1-4, orderered left to right, top to bottom.}
+\label{fig2}
+\end{figure}
+
+In Figure \ref{fig2} we plot the Kaplan-Meier adjusted cdf for hard drive model X on Weibull paper.
 Each point on the plot corresponds to a hard drive failure.
@@
 \cite{betancourt} as implemented in {\tt
-  rstan}\cite{rstan}. Multiple chain were run with _000 iterations of
-warmup. Convergence was assessed by examining posterior
+  rstan}\cite{rstan}. Multiple chain were run with warmup iterations. Convergence was assessed by examining posterior
 plots and checking that Gelman-Rubin diagnostics were close to 1.
@@
 \subsection{Results}
-Show some of the fits of the GFLP Model.  Also present the unpooled output for a few models.  Show parameter estimates with uncertainty.
+The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit.
+The upper left panel shows the estimated failure curve fit to all data ignoring drive model as a factor (Model 1).
+The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.
+
+The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which mean that the ranking of models with respect to fraction failing differs as a function of time.
+
+Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.
 
@@
 
+This should probably be truncated at something like eight years, since the right tails seem unlikely. Sort by left endpoint?
+
 \begin{figure}[H]
     \centering
-   \includegraphics[width=5.0in]{B10_full.pdf}
+   \includegraphics[width=5.0in]{fig/B10_full.pdf}
 		\caption{B10 in Years for All Drive Models \label{fig:first}} 

### 2017-05-17 plots/km_overplot.R

@@ -44,7 +44,7 @@ y.scale <- scale_y_continuous(name="Fraction failing", breaks=y.breaks, limits=y
 no_legend <- theme(legend.position="none")
 x_log <- scale_x_continuous(trans="log")
 
-p1 <- KM_plot_multi(df1, linear_axes = T, xlimits = xlimits, ylimits=ylimits)
+p1 <- KM_plot_multi(df1, linear_axes = T, xlimits = xlimits, ylimits=ylimits, grayscale = TRUE)
 
 null_s <- readRDS("../workflow_null/samples_null_model_3_29.rds")
 null_samp <- extract(null_s)

### 2017-05-17 plotting_fns/KM_plot.R

@@ -161,7 +161,8 @@ KM_with_band <- function(title = NULL, data, id, samp, n_iter, n, quantiles, tr_
 }
 
 KM_plot_multi <- function(data, prob_model="weibull", tr_adj = 0, title = NULL,
-                    linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000), ylimits=c(.0001, .9999), verbose=F){
+                    linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000),
+                    ylimits=c(.0001, .9999), verbose=F, grayscale=F){
   require(ggplot2)
   require(plyr)
   require(dplyr)
@@ -193,9 +194,9 @@ KM_plot_multi <- function(data, prob_model="weibull", tr_adj = 0, title = NULL,
   if(verbose){
     print(df)
   }
-  
-  p <- df %>%
-    ggplot(aes(t, Ft, group=model, color=model)) + geom_line()
+
+  p <- df %>% 
+    ggplot(aes(t, Ft, group=model, color=model)) + geom_step()
   
   if(!linear_axes){
     if(!fixed){
@@ -210,6 +211,9 @@ KM_plot_multi <- function(data, prob_model="weibull", tr_adj = 0, title = NULL,
     }
   } else{
     p <- p + scale_x_continuous(limits=xlimits) + scale_y_continuous(limits=ylimits)
+  }
+  if(grayscale){
+    p <- p + scale_colour_grey(start = 0, end = .9) 
   }
   p <- p +
     theme_bw(base_size = 14) + ggtitle(title)+xlab("Hours")+ylab("Fraction Failing")

### 2017-05-19 paper/paper.tex
Added some plots and small paper edits.

@@
 We found that there was insufficient information in the data to model these parameters hierarchically.
-Moreover, assuming a common distribution for infant mortality allows for straightforward interpretation of the quantities like $\frac{\pi_{d}}{\pi_{d'}}$.
+Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of the various ${\pi_{d}$, the probability of infant mortality for drive model $d$.
 
@@
 \subsection{Model Comparisons}
-First look at the overlaid posterior credible bands. \\
+First look at the overlaid posterior credible bands.
+Perhaps pick 4 different drive models and present a plot like the figure below.
+Bands are for full model, and the solid line is the full model fit.
+\\
+
+\begin{figure}[H]
+    \centering
+   \includegraphics[width=5.0in]{fig/mod40_bands.pdf}
+		\caption{Drive Model 40 \label{fig:first}} 
+\end{figure}
 
@@
 \subsection{Brand Comparisons}
-Ideas for plots: A graph comparing the pi's for all drive models.
-
-I think a graph looking at the B10 using the full model could be good, too.  Something like the graph below.  Perhaps we could sort drive model based on sample size?
-
-This should probably be truncated at something like eight years, since the right tails seem unlikely. Sort by left endpoint?
 
@@
 		\caption{B10 in Years for All Drive Models \label{fig:first}} 
+\end{figure}
+
+Am wondering if we should truncate this graph at a reasonable value, perhaps .3 or .2?
+
+\begin{figure}[H]
+    \centering
+   \includegraphics[width=5.0in]{fig/pi_bymod.pdf}
+		\caption{$\pi$ for All Drive Models \label{fig:first}} 
 \end{figure}

### 2017-05-19 plots/compare_fits.R

@@ -20,6 +20,13 @@ ssigtp <- readRDS("../workflow_sig2_and_tp2_vary/vary_s2_and_tp2_4_17.rds")
 stp2 <- readRDS("../workflow_tp2_vary/samples_tp2_vary.rds")
 snull <- readRDS("../workflow_null/samples_null_model_3_29.rds")
 sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
+#
+ssigtp <- readRDS("MCMC_draws/vary_s2_and_tp2_4_17.rds")
+stp2 <- readRDS("MCMC_draws/samples_tp2_vary_new.rds")
+snull <- readRDS("MCMC_draws/samples_null_model_3_29.rds")
+sfull <- readRDS("MCMC_draws/samples_lor_only3fails.rds")
+
+
 samps2tp <- extract(ssigtp)
 samptp2 <- extract(stp2)
 sampnull <- extract(snull)
@@ -36,7 +43,7 @@ xlimits <- c(100,50000)
 ylimits <- c(.0001, .9)
 null_band <- KM_band(id=1, n_iter= 100, samp=sampnull, xlim=xlimits, ylim=ylimits, quantiles=c(.05,.5,.95),
                      x_logscale=T, verbose=F, n = 100, pi.free=F, mu1.free=F, sigma1.free = F, mu2.free = F, sigma2.free = F,
-                     colband="green", colline="darkblue")
+                     linetp = "longdash", colband="green", colline="darkblue")
 for(j in 1:length(id)){
   orig_id <- id[j]
   pi_j = exp(sampfull$log_pi[max_id,j])
@@ -53,22 +60,31 @@ for(j in 1:length(id)){
                              xlimits=xlimits, ylimits=ylimits, verbose=F, conf=.90)
   tp2_list[[j]] <- KM_band(id=j, samp=samptp2, xlim=xlimits, ylim=ylimits, n_iter=100, quantiles=c(.05,.5,.95),
                            x_logscale=T, verbose=F, n = 100, pi.free=F, mu1.free=F, sigma1.free = F, mu2.free = T, sigma2.free = F,
-                           colband="red", colline="black")
+                           colband="red", linetp="dotted", colline="black")
   s2tp_list[[j]] <- KM_band(id=j, samp=samps2tp, xlim=xlimits, ylim=ylimits, n_iter=100, quantiles=c(.05,.5,.95),
[…]

### 2017-05-19 plotting_fns/KM_plot.R

@@ -128,7 +128,7 @@ KM_plot <- function(data, model, tr_adj = 0, title = NULL,
 }
 
 KM_band <- function(id, samp, n_iter=NULL, xlim, ylim, quantiles=c(.05, .5, .95), x_logscale=T, n=30, verbose=F,
-                    pi.free=T, mu1.free=F, sigma1.free=F, mu2.free=T, sigma2.free=T, colline="black", colband="red"){
+                    pi.free=T, mu1.free=F, sigma1.free=F, mu2.free=T, sigma2.free=T, linetp="dashed", colline="black", colband="red"){
   total_iter <- length(samp$lp__)
   iter <- floor(total_iter/n_iter) * 1:n_iter
   
@@ -150,7 +150,7 @@ KM_band <- function(id, samp, n_iter=NULL, xlim, ylim, quantiles=c(.05, .5, .95)
     })
   if(verbose) print(band)
   list(ptws_ci = geom_ribbon(data = band, inherit.aes = FALSE, aes(x=x, ymin=lower, ymax=upper), fill=colband, alpha=.2),
-       ptws_est = geom_line(data= band, inherit.aes = FALSE, aes(x=x, y=y), color=colline))
+       ptws_est = geom_line(data= band, inherit.aes = FALSE, aes(x=x, y=y), linetype=linetp, color=colline))
 }
 
 KM_with_band <- function(title = NULL, data, id, samp, n_iter, n, quantiles, tr_adj, xlimits, ylimits, fixed=T,

### 2017-05-19 plotting_fns/quantile.R

@@ -58,20 +58,23 @@ for (j in 1:44){
 
 
 #B10_Model4.rds is in paper folder for future use
-b10 <- readRDS("B10_Model4.rds")
+b10 <- readRDS("paper/B10_Model4.rds")
 colnames(b10) <- c("model","lb","med","ub")
 b10$lb <- b10$lb/(24*365)
 b10$med <- b10$med/(24*365)
 b10$ub <- b10$ub/(24*365)
 
+#Sort by Lower End Point Time
+b10$model <- factor(b10$model,levels=b10$model[order(b10$lb)])  #Sort By Lower End Point
+
 
 #Make Catepillar Plot for B10; Perhaps Sort by Sample Size?
 p <- ggplot(b10, aes(x=as.factor(model), y=med, ymin=lb, ymax=ub)) +
-  geom_pointrange() +
-  coord_flip() +
+  geom_pointrange() +  
+  coord_flip(ylim = c(0,8)) + 
   xlab('Drive Model') + theme_bw() + 
   ylab("Years")   +
-  scale_x_discrete(breaks=seq(1,44,1)) + 
+  scale_x_discrete(breaks=seq(1,44,1)) 
 p
 
 

### 2017-05-23 paper/paper.tex
Updates to Priors, updated pi_compare

@@
 \usepackage{enumerate}
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 \usepackage{url} % not crucial - just used below for the URL
@@
 Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution.
-We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that wearout implies an increasing hazard function.
-The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = \Phi^{-1}(.5)$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
+We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that,by nature, wearout produces an increasing hazard function.
+The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = t_{0.632}$, is that lifetime data often, as is true of the data presented here, features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
 For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.
@@
 \begin{enumerate}
-\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
+\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
 \item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
@@
 We found that there was insufficient information in the data to model these parameters hierarchically.
-Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of the various ${\pi_{d}$, the probability of infant mortality for drive model $d$.
+Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).
 
 \subsection{Priors}
-To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model.
-We select proper priors to ensure a proper posterior.
+To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior.
+
+For models 1-4, different sets of restrictions required different prior specifications, which were assigned as follows:
+
+\begin{enumerate}
+\item This model constrains all drive models to the same failure distribution. For this ``reduced" model we assume the priors :
+
+$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad t_{0.5,1} \sim \op{Normal}(8,4)$$
+$$\sigma_2 \sim \op{Lognormal}(0, 2.5) \quad t_{0.2,2} \sim \op{N}(10,4)$$
+
+The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
+The prior on $\sigma_1$ and $\sigma_2$ put it on the interval $(0.02, 61.1)$ with $90\%$ probability.
+(Note that, due to symmetry, the Weibull shape parameter has the same prior interval and that a value of 1, corresponding to a constant hazard, is the median of this prior distribuion.) $t_{0.5,1}$ has a prior $90\%$ interval of $(1.4,14.6)$.
+For $t_{0.2,2}$ it is $(3.4,16.6)$.
+Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.
+
+\item We allow $t_{pd}$ to vary by drive-model. To help with model identifiability, we tighten the priors on the defective mode:
+$$ \mbox{logit}^{-1}\pi \sim \op{Normal}(-3,1),\quad \sigma_1 \sim \op{Lognormal}(0, 1), \quad t_{p1} \sim \op{Normal}(7,2),$$
+implying now a prior $90\%$ probability that $\sigma_1$ is between $0.19$ and $5.18$ and $t_{0.5,1}$ between $3.7$ and $10.3$, which, translating from the log-scale, is equivalent to an interval from 1.5 days to 3.4 years.
+
+\item $\sigma_{d2}$ is allowed to vary. Priors for constrained parameters are the same as for model 2.
+
+\item $\pi_d$ varies. Priors for constrained parameters are the same as for model 2.
+
+\end{enumerate}
+
 For the scale hyper-parameters we follow the recommendations of \citet{gelman2014bayesian} and use half-Cauchy priors.
@@
 The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification?
-Should maybe be lower.) For $\eta_{\sigma, 2}$, 95\% of prior mass is on values greater than $0.037$.
+Should maybe be lower.) For $\eta_{\sigma, 2}$, $95\%$ of prior mass is on values greater than $0.037$.
 Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this correponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$.
@@
 \subsection{Computation}
-Each model was fit in {\t R} \nocite{r} using Hamiltonian Monte Carlo
-\cite{betancourt} as implemented in {\tt
-  rstan}\cite{rstan}. Multiple chain were run with warmup iterations. Convergence was assessed by examining posterior
-plots and checking that Gelman-Rubin diagnostics were close to 1.
+Each model was fit using the {\tt
+  rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
+\cite{betancourt}. Multiple chain were run for 1500 iterations after 1500 warmup iterations. Convergence was assessed by examining posterior
+plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$. Four chains were run for models 1,2 and 3 and 16 chains were run for model 4.
 
 \section{Data analysis}
-\subsection{Results}
+
+\subsection{Model Comparisons}
 The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit.
@@
 Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.
-
-\subsection{Model Comparisons}
 First look at the overlaid posterior credible bands.
@@
     \centering
-   \includegraphics[width=5.0in]{fig/pi_bymod.pdf}
+   \includegraphics[width=5.0in]{fig/pi_compare.pdf}
 		\caption{$\pi$ for All Drive Models \label{fig:first}} 
@@
 
-\section{Concluding Remarks and Extensions}
+\section{Discussion}
 Review the advantages of fitting the GFLP model and offer future ideas.
+
+Advances in computer technology have made computationally intensive Bayesian methods for fitting hierarchical models feasible in practice.
+Analysis using HMC is accessible to practitioners by Stan.
+HMC can be very efficient in terms of effective samples because of the way that it uses joint updates for all parameters using Hamiltonian dynamics to rapidly explore the target distribution.
+Hierarchical modeling can be a useful way to pool information to estimate large models.
+By using Bayesian methods like the one presented here we can avoid using potentially problematic asymptotic approximations, since all posterior uncertainty is automatically integrated into averages computed from functions of the samples.
 

### 2017-05-23 plots/compare_fits.R

@@ -21,10 +21,10 @@ stp2 <- readRDS("../workflow_tp2_vary/samples_tp2_vary.rds")
 snull <- readRDS("../workflow_null/samples_null_model_3_29.rds")
 sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
 #
-ssigtp <- readRDS("MCMC_draws/vary_s2_and_tp2_4_17.rds")
-stp2 <- readRDS("MCMC_draws/samples_tp2_vary_new.rds")
-snull <- readRDS("MCMC_draws/samples_null_model_3_29.rds")
-sfull <- readRDS("MCMC_draws/samples_lor_only3fails.rds")
+# ssigtp <- readRDS("MCMC_draws/vary_s2_and_tp2_4_17.rds")
+# stp2 <- readRDS("MCMC_draws/samples_tp2_vary_new.rds")
+# snull <- readRDS("MCMC_draws/samples_null_model_3_29.rds")
+# sfull <- readRDS("MCMC_draws/samples_lor_only3fails.rds")
 
 
 samps2tp <- extract(ssigtp)
@@ -104,9 +104,9 @@ out.pi <- matrix(ncol=4, nrow=44)
 for (i in 1:44){
   num=i
   seta <- paste0("log_pi[",num,"]",collapse="")
-  etal <- summary(sfull)$summary[seta,"2.5%"]
+  etal <- summary(sfull)$summary[seta,"25%"]
   etam <- summary(sfull)$summary[seta,"50%"]
-  etah <- summary(sfull)$summary[seta,"97.5%"]
+  etah <- summary(sfull)$summary[seta,"75%"]
   out.pi[i,1] <- exp(etal)
   out.pi[i,2] <- exp(etam)
   out.pi[i,3]<- exp(etah)
@@ -127,7 +127,8 @@ p <- ggplot(pi.dat, aes(x=as.factor(model), y=med, ymin=lower, ymax=upper)) +
   coord_flip() + 
   xlab('Drive Model') + theme_bw() + 
   ylab(expression(pi))   +
-  scale_x_discrete(breaks=seq(1,44,1)) 
+  scale_x_discrete(breaks=seq(1,44,1))+
+  scale_y_continuous(trans="logit", breaks=c(.01, .02, .05, .1, .2, .4, .6, .8, .9))
 p
 
 

### 2017-05-25 paper/paper.tex
Updated plots

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{fig/heterogeneity_compare_preliminary.pdf}
-\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
-Right: Pointwise posterior median time to failure curves for Models 1-4, orderered left to right, top to bottom.}
-\label{fig2}
+\includegraphics[width=.8\textwidth]{fig/data_overview}
+\caption{Left: Kaplan-Meier plots showing nonparametric estimates of fraction failing for individual drive-models with truncation adjustments based on fitted model, Right: Total observed failures (top) and the sum of lengths of all observational intervals by drive-model.}
 \end{figure}
@@
 \subsection{Model Comparisons}
+\begin{figure}
+\includegraphics[width=\textwidth]{fig/heterogeneity_compare_preliminary.pdf}
+\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
+Right: Pointwise posterior median time to failure curves for Models 1-4, orderered left to right, top to bottom.}
+\label{fig2}
+\end{figure}
+
 The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit.

### 2017-05-25 plotting_fns/KM_plot.R

@@ -160,9 +160,9 @@ KM_with_band <- function(title = NULL, data, id, samp, n_iter, n, quantiles, tr_
   p + q[[1]] + q[[2]]
 }
 
-KM_plot_multi <- function(data, prob_model="weibull", tr_adj = 0, title = NULL,
+KM_plot_multi <- function(data, prob_model="weibull", tr_adj = NULL, title = NULL,
                     linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000),
-                    ylimits=c(.0001, .9999), verbose=F, grayscale=F){
+                    ylimits=c(.0001, .9999), verbose=F, grayscale=F, size=.2){
   require(ggplot2)
   require(plyr)
   require(dplyr)
@@ -176,6 +176,11 @@ KM_plot_multi <- function(data, prob_model="weibull", tr_adj = 0, title = NULL,
          ytrans <- "qgev"},
          {stop("`model' must be one of lnorm, weibull, frechet")})
   
+  if(is.null(tr_adj)){
+    message("No truncation adjustment specified: setting to 0.")
+    tr_adj <- rep(0, length(unique(data$model)))
+  }
+  
   # df <- with(data, data.frame(time = endtime[!censored]))
   df <- ddply(data, .(model), function(m){
               t = sort(unique(m$endtime[!(m$censored)]))
@@ -186,7 +191,7 @@ KM_plot_multi <- function(data, prob_model="weibull", tr_adj = 0, title = NULL,
           d = sum(data$starttime[data$model==model] < t & data$endtime[data$model==model] == t & data$censored[data$model==model] == 0)) %>%
     mutate(p = (n-d)/n,
            St = cumprod(p))
-  df$Ft <- (1 - df$St) * (1 - tr_adj) + tr_adj
+  df$Ft <- (1 - df$St) * (1 - tr_adj[df$model]) + tr_adj[df$model]
   df <- df[which(df$Ft < 1),]
   
   df$model <- factor(df$model)
@@ -196,7 +201,7 @@ KM_plot_multi <- function(data, prob_model="weibull", tr_adj = 0, title = NULL,
   }
 
   p <- df %>% 
-    ggplot(aes(t, Ft, group=model, color=model)) + geom_step()
+    ggplot(aes(t, Ft, group=model, color=model)) + geom_step(size=size)
   
   if(!linear_axes){
     if(!fixed){

### 2017-05-25 paper/paper.tex
Data Section Edits.

--- 2017-05-23 paper/paper.tex
@@
 While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.
-The purpose is to provide consumers and businesses with reliability information on different hard drive brands and models.
+The purpose is to provide consumers and businesses with reliability information on different hard drive brand models.
 The hard drives continuously spin in controlled storage pods where they run until failure.
 When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.
-In addition, the number of storage pods is increasing as Backblaze adds new hard drive brands to the sample.
+In addition, the number of storage pods is increasing as Backblaze adds new hard drive models to the sample.
 Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
@@
 As of the first quarter of 2016, Backblaze was collecting data on 63 different hard drive models.
-Some drive models have been running since 2013, while others have been added at a later date.
-There are a total of 75,297 hard drives running.
+Some drive models have been running since 2013, while others were added at a later date.
+There are a total of 75,297 hard drives in operation.
 The distribution of drives by model varies: some models have only have a service record for a single drive; the maximum number of service records for a single drive model is 35,860.
-We can also look at the distribution of total number of failures and total time on test across all drive models.
-As seen below, some drive models contain a lot of information, while others have few failures or a little time in operation.
-[INSERT FIGURE 1].
-\\
+We can also look at the distribution of total number of failures and total time on test across all drive brand models.
+As seen in Figure \ref{fig1}, some drive models contain a lot of information, while others have few failures or little time in operation.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{fig/data_sum.pdf}
+\caption{Left: Total number of failures (log scale) for each of the drive-models in the Backblaze data.
+Right: Total time in operation (log scale) for each of the drive-models in the Backblaze data.
+Both plots are sorted based on the total number of failures.}
+\label{fig2}
+\end{figure}
 
@@
 As outlined in Vehtari et.
-al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters \cite{vehtari}.
+al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.
 For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with the $i$th data point removed.
-The final expected point-wise predictive density is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
+The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
 We computed the elpd for all 4 models using the R package loo \cite{loo}.
-When $n$ is large, the distribution of the elpd is approximately normal and models can be statistically compared.
+When $n$ is large, the distribution of the elpd is approximately normal and different models can be statistically compared.
 We calculated the difference in expected predictive accuracy for Model 4 vs 3, Model 3 vs 2, and Model 2 vs 1, as well as the standard error of the difference (Table 1).
 As the model complexity increased, the expected predictive accuracy improved.
-Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 models.
+Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 model specifications.
 
@@
   \hline
- & elpd\ & Diff(elpd)\ & SE(Diff) \\ 
+ & ELPD \ & Difference in ELPD \ & SE of the Difference \\ 
   \hline
@@
 \end{tabular}
-\caption{Model Comparison Based on LOO}
+\caption{Expected Log Pointwise Predictive Density (ELPD) for each model specification.
+Each model is compared to the more parsimonious model below.
+The estimated difference of expected leave-one-out prediction errors between the two models, as well as the standard error of the difference, is also presented.}
 \label{table:1}

### 2017-05-25 plots/plots_forpaper.R

@@ -1,30 +1,54 @@
-library(backblaze)
 library(plyr)
 library(dplyr)
 library(rstan)
+library(cowplot)
 
-source("../workflow/functions.R")
-source("../plotting_fns/KM_plot.R")
-source("../plotting_fns/greenwood_errors.R")
-
-dat <- backblaze
-
-
+dat <- readRDS("BB_data/clean_unit_summaries.rds")
 dat$model <- as.integer(dat$model)
 dat$failed[dat$failed==2] <- 1
 
+overview <- ddply(dat, .(model), summarise,
+                  n=length(model),
+                  f=sum(failed>0),
+                  early_f = sum(failed>0 & endtime<365*24*1),
+                  late_f = sum(failed>0 & endtime>365*24*2))
+# id <- unique(dat$model)
+id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
+overview$stan_id <- NA
+overview[id,]$stan_id <- 1:length(id)
+
+
 #Look at Distribution of Failures and Total Time on Test
 #Total Filures By Model
-tot<- dat %>% select(model, censored) %>% group_by(model) %>% summarise(sum(censored==FALSE))
+tot <- dat %>% select(model, censored) %>% group_by(model) %>% summarise(sum(censored==FALSE))
 names(tot)[2]<-"count"
+
+#Sort by Failures
+tot <- as.data.frame(tot)
+tot$model <- factor(tot$model,levels=tot$model[order(tot$count)])  #Sort by Total Failures
+
+
 theme_set(theme_bw(base_size=18))
-ggplot(tot, aes(x = count)) + geom_histogram() + scale_x_log10()
-ggplot(tot, aes(x=factor(model),y=count))+geom_bar(stat="identity")+xlab("Model")+ylab("Total Failures")+scale_x_discrete(breaks=seq(1, 62, by=2))+
-  ggtitle("Total Failures by Model") + scale_y_log10()
+#ggplot(tot, aes(x = count)) + geom_histogram(bins = 15) + scale_x_log10()
+fail <- ggplot(tot, aes(x=model, y=count)) + geom_bar(stat="identity") + xlab("Drive Brand Model") + ylab("Total Failures") + 
+   ggtitle("Total Failures by Drive Brand Model") + scale_y_log10() + coord_flip() 
 
 #Total Time on Test
[…]

### 2017-05-25 paper/paper.tex
Reworking KM section

--- 2017-05-25 paper/paper.tex
@@
 
-For the hard drive data we first estimate the empirical cdf for each model using the Kaplan-Meier estimator \cite{kaplan}.
-With left truncation, however, the Kaplan-Meier estimate, $\widehat{F(t)_{KM}}$, is conditional based on hard drive $i$ surviving up to time $\tau_i^L$ where $L$ is the amount of time the hard drive was running before Backblaze starting monitoring the drive.
-To get the unconditional distribution we correct the non-parametric estimates using a parametric adjustment outlined by Turnbull, and given in more detail by Meeker and Escobar \cite{turnbull,meeker}.
-For each hard drive model we select $\tau_{\text{min}}^L$, the smallest left truncated time in the sample.
-Then, using the parametric estimates of the parameters, we calculate $Pr(T>\tau_\text{min}^L)$ the probability a hard drive has survived up to $\tau_{\text{min}}^L$.
-We then apply this correction to the non parametric estimates, which results in the unconditional distribution of time to failure.\\
+For the hard drive data we first estimate the empirical cdf for each model using the Kaplan-Meier estimator (KME)\cite{kaplan}.
+With left truncation, however, the KME, which we denote by $\widehat{F(t)_{KM}}$, is conditional based on hard drive $i$ surviving up to time $\tau_i^L$ where $L$ is the amount of time the hard drive was running before Backblaze starting monitoring the drive.
+To correct the non-parametric estimates, we adapt the adjustment method outlined by Turnbull, and given in more detail by Meeker and Escobar \cite{turnbull,meeker}.
+For each hard drive model we select $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
+Since $Pr(T>t_{d,\text{min}}^L)$, the probability a hard drive has survived up to $t_{d,\text{min}}^L$, is a function of the model parameters, by sampling from the full posterior distribution we can easily compute its posterior median, $\widehat{Pr_(T>t_{d,\text{min}}^L)}$.
+We use this to shift and rescale the KM thus  the unconditional distribution of time to failure, as in
+
+$$\widehat{F(t)_{KMadj}} = .$$
 

### 2017-05-25 paper/paper.tex
Edits to Plots.  Working on Model Comparison plot for 4 drive models.

--- 2017-05-25 paper/paper.tex
@@
 Both plots are sorted based on the total number of failures.}
-\label{fig2}
+\label{fig1}
 \end{figure}
@@
 Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.
-First look at the overlaid posterior credible bands.
-Perhaps pick 4 different drive models and present a plot like the figure below.
-Bands are for full model, and the solid line is the full model fit.
 \\
+Comment: I think I prefer to use points for Kaplan Meier.  Using the connected dots almost makes it look like another model fit, rather than the actual observed failure data.  What do you think? \\
 
@@
     \centering
-   \includegraphics[width=5.0in]{fig/mod40_bands.pdf}
-		\caption{Drive Model 40 \label{fig:first}} 
+   \includegraphics[width=5.0in]{fig/brand_compare_nobands.pdf}
+		\caption{Estimated GFLP for all 4 Model Specifications.  Truncated Adjusted Kaplan Meier points are also presented.    \label{fig:first}} 
 \end{figure}

### 2017-05-25 paper/paper.tex
Data section edits.  Will wait to add more edits till you push.

@@
 
-In Figure \ref{fig2} we plot the Kaplan-Meier adjusted cdf for hard drive model X on Weibull paper.
+In Figure \ref{fig2} we plot the Kaplan-Meier adjusted cdf for hard drive model 20 on Weibull paper.
 Each point on the plot corresponds to a hard drive failure.
@@
 The second mode is non-defective hard drives that eventually fail due to wearout.
-Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring around hour Z.
+Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring around 15,000 hours.
 Therefore, fitting a single Weibull model would not be flexible enough to model the failure distribution.

### 2017-05-25 paper/paper.tex
CHanged data section

--- 2017-05-25 paper/paper.tex
@@
 Both plots are sorted based on the total number of failures.}
-\label{fig1}
+\label{fig2}
 \end{figure}
 
-Probability plotting is an effective method to check the adequacy of various parametric models to the data.
-Identifying whether failure data follows a specific distribution is difficult to do by eye: especially when looking at pdfs or other non-linear plots.
-However, if we can linearize the cumulative distribution function it is easier to visually assess a distributional goodness of fit; if a model is appropriate, a non parametric estimate of $\hat{F(t)}$ graphed on linarized probability scales should approximately follow a straight line.
-\\
+Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
+Identifying whether failure data are consistent with a specific family of distributions is difficult to do by examining a histogram or other density estimate.
+By transform the y-axis of the estimated cumulative distribution function (cdf), denoted by $\widehat{F(t)}$, by the inverse cdf of for the standardized distribution, a visual comparison of the plot to a straight line can be an effective informal test for distributional goodness of fit.
 
-For the hard drive data we first estimate the empirical cdf for each model using the Kaplan-Meier estimator (KME)\cite{kaplan}.
-With left truncation, however, the KME, which we denote by $\widehat{F(t)_{KM}}$, is conditional based on hard drive $i$ surviving up to time $\tau_i^L$ where $L$ is the amount of time the hard drive was running before Backblaze starting monitoring the drive.
-To correct the non-parametric estimates, we adapt the adjustment method outlined by Turnbull, and given in more detail by Meeker and Escobar \cite{turnbull,meeker}.
+Applying this method to the hard drive data, we first must estimate the empirical cdf for each model using the Kaplan-Meier estimator \cite{kaplan}.
+With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by $\widehat{F_d(t)}_{KM}$, is conditional on survival up to $t_{d,\text{min}}^L$, the shortest reported running time of all units of drive-model $d$ for which records are available.
+To produce unconditional estimates, we adapt the adjustment method outlined by Turnbull, and given in more detail by Meeker and Escobar \cite{turnbull,meeker}.
 For each hard drive model we select $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
-Since $Pr(T>t_{d,\text{min}}^L)$, the probability a hard drive has survived up to $t_{d,\text{min}}^L$, is a function of the model parameters, by sampling from the full posterior distribution we can easily compute its posterior median, $\widehat{Pr_(T>t_{d,\text{min}}^L)}$.
-We use this to shift and rescale the KM thus  the unconditional distribution of time to failure, as in
+By sampling from the full posterior distribution, since $Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has survived up to $t_{d,\text{min}}^L$) is a function of the model parameters, we can easily compute its posterior median, $\widehat{A}_{\text{med}} = \widehat{Pr}_(T>t_{d,\text{min}}^L|\theta_d)$.
+We compute the adjusted estimate by
 
-$$\widehat{F(t)_{KMadj}} = .$$
+$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_d(t)}_{KM},\; t>t_{d,\text{min}}^L.$$
+
+While this adjustment is negligible for the majority of drive-models, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
+2.3 years.
 
@@
 Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.
+First look at the overlaid posterior credible bands.
+Perhaps pick 4 different drive models and present a plot like the figure below.
+Bands are for full model, and the solid line is the full model fit.
 \\
-Comment: I think I prefer to use points for Kaplan Meier.  Using the connected dots almost makes it look like another model fit, rather than the actual observed failure data.  What do you think? \\
 
@@
     \centering
-   \includegraphics[width=5.0in]{fig/brand_compare_nobands.pdf}
-		\caption{Estimated GFLP for all 4 Model Specifications.  Truncated Adjusted Kaplan Meier points are also presented.    \label{fig:first}} 
+   \includegraphics[width=5.0in]{fig/mod40_bands.pdf}
+		\caption{Drive Model 40 \label{fig:first}} 
 \end{figure}

### 2017-05-25 workflow/tr_adj.R
Compute truncation adjustments

@@ -1,0 +1,36 @@
+dat <- readRDS("../BB_data/clean_unit_summaries.rds")
+dat$model <- as.integer(dat$model)
+
+library(plyr)
+library(dplyr)
+library(rstan)
+overview <- ddply(dat, .(model), summarise,
+                  n=length(model),
+                  f=sum(failed>0),
+                  early_f = sum(failed>0 & endtime<365*24*1),
+                  late_f = sum(failed>0 & endtime>365*24*2))
+# id <- unique(dat$model)
+id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
+overview$stan_id <- NA
+overview[id,]$stan_id <- 1:length(id)
+
+samp <- extract(readRDS("../workflow/samples_lor_only3fails.rds"))
+S <- length(samp$lp__)
+
+source("../plotting_fns/KM_plot.R")
+
+dat2 <- filter(dat, model %in% id)
+tr_adj_df <- ddply(dat2, .(model), function(x){
+  start = min(x$starttime)
+  modelid = overview$stan_id[which(overview$model==x$model[1])]
+  samp_tr = get_tr_adj(start, exp(as.numeric(samp$log_pi[,modelid])),
+                       as.numeric(samp$mu1),
+                       as.numeric(samp$sigma1),
+                       as.numeric(samp$mu2[,modelid]), 
+                       as.numeric(samp$sigma2[modelid]))
+  adj = quantile(samp_tr, c(.25,.5,.75))
+  data.frame(modelid=modelid, start=start, lower=adj[1], median=adj[2], upper=adj[3])
+})
+
+saveRDS(tr_adj_df, "../BB_data/tr_adj_tp2s2pi.rds")
+

### 2017-05-26 paper/paper.tex
Added new plots to data section.

@@
 \includegraphics[width=\textwidth]{fig/data_sum.pdf}
-\caption{Left: Total number of failures (log scale) for each of the drive-models in the Backblaze data.
+\caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data.
 Right: Total time in operation (log scale) for each of the drive-models in the Backblaze data.
@@
 
-\begin{figure}
-\includegraphics[width=.8\textwidth]{fig/data_overview}
-\caption{Left: Kaplan-Meier plots showing nonparametric estimates of fraction failing for individual drive-models with truncation adjustments based on fitted model, Right: Total observed failures (top) and the sum of lengths of all observational intervals by drive-model.}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=.8\textwidth]{fig/data_overview}
+%\caption{Left: Kaplan-Meier plots showing nonparametric estimates of fraction failing for individual drive-models with truncation adjustments based on fitted model, Right: Total observed failures (top) and the sum of lengths of all observational intervals by drive-model.}
+%\end{figure}
 
-In Figure \ref{fig2} we plot the Kaplan-Meier adjusted cdf for hard drive model 20 on Weibull paper.
+In Figure 2 we plot the Kaplan-Meier adjusted cdf for hard drive model 14 on Weibull paper.
+Drive model 14 had the most observed failures of all the brands.
 Each point on the plot corresponds to a hard drive failure.
 Censored drives are not plotted.
+Standard error bands are calculated using Greenwood's method (reference).
 As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.
@@
 Therefore, fitting a single Weibull model would not be flexible enough to model the failure distribution.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=.5\textwidth]{fig/mod14_KM_adj}
+\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive model 14.  Dashed lines are Greenwood standard errors.}
+\end{figure}
 

### 2017-05-26 plots/compare_fits.R

@@ -20,11 +20,14 @@ ssigtp <- readRDS("../workflow_sig2_and_tp2_vary/vary_s2_and_tp2_4_17.rds")
 stp2 <- readRDS("../workflow_tp2_vary/samples_tp2_vary.rds")
 snull <- readRDS("../workflow_null/samples_null_model_3_29.rds")
 sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
-#
+tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")$median
+
+
 #ssigtp <- readRDS("MCMC_draws/vary_s2_and_tp2_4_17.rds")
-# stp2 <- readRDS("MCMC_draws/samples_tp2_vary_new.rds")
-# snull <- readRDS("MCMC_draws/samples_null_model_3_29.rds")
-# sfull <- readRDS("MCMC_draws/samples_lor_only3fails.rds")
+#stp2 <- readRDS("MCMC_draws/samples_tp2_vary_new.rds")
+#snull <- readRDS("MCMC_draws/samples_null_model_3_29.rds")
+#sfull <- readRDS("MCMC_draws/samples_lor_only3fails.rds")
+
 
 
 samps2tp <- extract(ssigtp)
@@ -39,12 +42,13 @@ full_list <- list()
 s2tp_list <- list()
 tp2_list <-list()
 KM_list <- list()
-null_list <- list()
 xlimits <- c(100,50000)
 ylimits <- c(.0001, .9)
-#null_band <- KM_band(id=1, n_iter= 100, samp=sampnull, xlim=xlimits, ylim=ylimits, quantiles=c(.05,.5,.95),
+
+null_band <- KM_band(id=1, n_iter= 100, samp=sampnull, xlim=xlimits, ylim=ylimits, quantiles=c(.05,.5,.95),
                      x_logscale=T, verbose=F, n = 100, pi.free=F, mu1.free=F, sigma1.free = F, mu2.free = F, sigma2.free = F,
                      linetp = "longdash", colband="green", colline="darkblue")
+
 for(j in 1:length(id)){
   orig_id <- id[j]
   pi_j = exp(sampfull$log_pi[max_id,j])
@@ -52,17 +56,14 @@ for(j in 1:length(id)){
   loc2_j = sampfull$mu2[max_id,j]
   scl1_j = sampfull$sigma1[max_id]
   scl2_j = sampfull$sigma2[max_id,j]
[…]

### 2017-05-26 plots/compare_fits.R
Adjusted truncation piece in compare fits.

@@ -5,6 +5,7 @@ dat$model <- as.integer(dat$model)
 library(plyr)
 library(dplyr)
 library(rstan)
+library(ggplot2)
 overview <- ddply(dat, .(model), summarise,
                   n=length(model),
                   f=sum(failed>0),

### 2017-05-26 plots/plots_forpaper.R

@@ -17,14 +17,22 @@ id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
 overview$stan_id <- NA
 overview[id,]$stan_id <- 1:length(id)
 
+#new id
+
+
 
 #Look at Distribution of Failures and Total Time on Test
 #Total Filures By Model
-tot <- dat %>% select(model, censored) %>% group_by(model) %>% summarise(sum(censored==FALSE))
+
+#Subset based on Model with >=3 failures
+datsub = subset(dat, model %in% id)
+
+tot <- datsub %>% select(model, censored) %>% group_by(model) %>% summarise(sum(censored==FALSE))
 names(tot)[2]<-"count"
 
 #Sort by Failures
 tot <- as.data.frame(tot)
+tot$model <- 1:length(id)  #Make New ID based on Subset
 tot$model <- factor(tot$model,levels=tot$model[order(tot$count)])  #Sort by Total Failures
 
 
@@ -34,12 +42,13 @@ fail <- ggplot(tot, aes(x=model, y=count)) + geom_bar(stat="identity") + xlab("D
    ggtitle("Total Failures by Drive Brand Model") + scale_y_log10() + coord_flip() 
 
 #Total Time on Test
-tt <- dat %>% group_by(model) %>% mutate(tt=(endtime-starttime)/(24*365))
+tt <- datsub %>% group_by(model) %>% mutate(tt=(endtime-starttime)/(24*365))
 tt2 <- tt %>% group_by(model) %>% summarize(sum(tt))
 names(tt2)[2]<-"total"
 
 tt2 <- as.data.frame(tt2)
 tt2$count <- tot$count
+tt2$model <- 1:length(id)
 tt2$model <- factor(tt2$model,levels=tt2$model[order(tt2$count)])  #Sort by Total Failures
 
 theme_set(theme_bw(base_size=18))

### 2017-05-30 paper/paper.tex
Paper, BibTek, and a few plot files edited.

@@
 There are a total of 75,297 hard drives in operation.
-The distribution of drives by model varies: some models have only have a service record for a single drive; the maximum number of service records for a single drive model is 35,860.
-We can also look at the distribution of total number of failures and total time on test across all drive brand models.
-As seen in Figure \ref{fig1}, some drive models contain a lot of information, while others have few failures or little time in operation.
+The distribution of drives by model varies: some models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.
+Figure \ref{fig1} shows the distribution of total number of failures and total time on test for drive brand models with at least 3 failures.
+As discussed later, for model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.
 
@@
 Drive model 14 had the most observed failures of all the brands.
-Each point on the plot corresponds to a hard drive failure.
+Each point on the plot corresponds to an observed failure.
 Censored drives are not plotted.
-Standard error bands are calculated using Greenwood's method (reference).
+Standard error bands are calculated using Greenwood's method \cite{green}.
 As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.
@@
 Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring around 15,000 hours.
-Therefore, fitting a single Weibull model would not be flexible enough to model the failure distribution.
+Therefore, fitting a single Weibull model is not flexible enough to model the failure distribution.
 
@@
 \includegraphics[width=.5\textwidth]{fig/mod14_KM_adj}
-\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive model 14.  Dashed lines are Greenwood standard errors.}
+\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive model 14.  Dashed lines are Greenwood standard errors.  Plotted on Weibull paper.}
 \end{figure}
@@
 Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.
-\\
-Comment: I think I prefer to use points for Kaplan Meier.  Using the connected dots almost makes it look like another model fit, rather than the actual observed failure data.  What do you think? \\
 
@@
 \centering
-   \includegraphics[width=5.0in]{fig/B10_full.pdf}
-		\caption{B10 in Years for All Drive Models \label{fig:first}} 
-\end{figure}
-
-Am wondering if we should truncate this graph at a reasonable value, perhaps .3 or .2?
-
-\begin{figure}[H]
-    \centering
-   \includegraphics[width=5.0in]{fig/pi_compare.pdf}
-		\caption{$\pi$ for All Drive Models \label{fig:first}} 
+ \includegraphics[width=4.0in]{fig/b10andpi.pdf}
+\caption{Left: 50\% credible intervals for B10 (time in years till 10\% of the drives fail).
+Right: 50\% credible intervals for $\pi$ plotted on the logit scale.
+Both plots are sorted based on the median value of B10.}
+\label{fig4}
 \end{figure}

### 2017-05-30 plots/compare_fits.R

@@ -117,7 +117,7 @@ ggplot(mean_df, aes(x=means)) + geom_histogram(aes(y=..density..), bins=10) + fa
 
 #Plot B10
 #Note: 50% CI for B10
-b10 <- readRDS("../paper/b10_fullmod_50CI.rds")
+b10 <- readRDS("b10_50ci.rds")
 colnames(b10) <- c("model","lb","med","ub")
 b10$lb <- b10$lb/(24*365)
 b10$med <- b10$med/(24*365)

### 2017-05-31 paper/paper.tex
Morning changes

@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{fig/data_sum.pdf}
+\includegraphics[width=\textwidth]{fig/data-sum.pdf}
 \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data.
@@
 
-Applying this method to the hard drive data, we first must estimate the empirical cdf for each model using the Kaplan-Meier estimator \cite{kaplan}.
-With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by $\widehat{F_d(t)}_{KM}$, is conditional on survival up to $t_{d,\text{min}}^L$, the shortest reported running time of all units of drive-model $d$ for which records are available.
-To produce unconditional estimates, we adapt the adjustment method outlined by Turnbull, and given in more detail by Meeker and Escobar \cite{turnbull,meeker}.
-For each hard drive model we select $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
-By sampling from the full posterior distribution, since $Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has survived up to $t_{d,\text{min}}^L$) is a function of the model parameters, we can easily compute its posterior median, $\widehat{A}_{\text{med}} = \widehat{Pr}_(T>t_{d,\text{min}}^L|\theta_d)$.
-We compute the adjusted estimate by
+Applying this method to the hard drive data, we first must estimate
+the empirical cdf for each model using the Kaplan-Meier estimator
+\cite{kaplan}.  With left truncation, however, the standard
+Kaplan-Meier estimator for drive-model $d$, denoted by
+$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
+$t_{d,\text{min}}^L$, the shortest reported running time of all units
+of drive-model $d$ for which records are available. To produce
+unconditional estimates, we adapt the adjustment method outlined by
+Turnbull, and given in more detail by Meeker and Escobar
+\cite{turnbull,meeker}.  For each hard drive model we select
+$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
+By sampling from the full posterior distribution, since
+$Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
+survived up to $t_{d,\text{min}}^L$) is a function of the model
+parameters, we can easily compute its posterior median,
+$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by
 
@@
 %\begin{figure}
-%\includegraphics[width=.8\textwidth]{fig/data_overview}
+%\includegraphics[width=.8\textwidth]{fig/data-overview}
 %\caption{Left: Kaplan-Meier plots showing nonparametric estimates of fraction failing for individual drive-models with truncation adjustments based on fitted model, Right: Total observed failures (top) and the sum of lengths of all observational intervals by drive-model.}
@@
 \centering
-\includegraphics[width=.5\textwidth]{fig/mod14_KM_adj}
+\includegraphics[width=.5\textwidth]{fig/mod14-KM-adj}
 \caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive model 14.  Dashed lines are Greenwood standard errors.  Plotted on Weibull paper.}
@@
 To borrow strength across models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters.
-For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j} = \Phi^{-1}(\mu_j)$, and proportions, $\pi_d$, of the component distributions as follows:
+For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
 
 $$\sigma_{d,1} \ind \op{Lognormal} \left( \eta_{\sigma,1}, \tau^2_{\sigma,1} \right) \mbox{ for } d=1,\ldots,D$$
-$$\sigma_{d,2} \ind \op{Truncated-Lognormal} \left( \eta_{\sigma,2}, \tau^2_{\sigma,2} , 0, 1, \right) \mbox{ for } d=1,\ldots,D$$
+$$\sigma_{d,2} \ind \op{Truncated-Lognormal} \left( \eta_{\sigma,2}, \tau^2_{\sigma,2} , 0, 1 \right) \mbox{ for } d=1,\ldots,D$$
 $$t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j)  \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
 
-$$\op{logit} \pi_d \ind \op{N}(\eta_pi, \tau_pi) \mbox{ for } d=1,\ldots,D.$$
+$$\op{logit} \pi_d \ind \op{N}(\eta_\pi, \tau_\pi) \mbox{ for } d=1,\ldots,D.$$
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{fig/heterogeneity_compare_preliminary.pdf}
+\includegraphics[width=\textwidth]{fig/heterogeneity-compare-preliminary}
 \caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
@@
     \centering
-   \includegraphics[width=5.0in]{fig/brand_compare_nobands.pdf}
+   \includegraphics[width=5.0in]{fig/brand-compare-nobands}
 		\caption{Estimated GFLP for all 4 Model Specifications.  Truncated Adjusted Kaplan Meier points are also presented.    \label{fig:first}} 
@@
 
-\subsection{Brand Comparisons}
+\subsection{Drive-model Comparisons}
+We consider the problem of selecting between drive-models. From the
+perspective of return on investment, is clear
+that we should prefer the drive-models which will provide more years
+of service. There are two sources of variation at play; the
+posterior uncertainty in the parameters, which largely depends on the
+sample size, and the uncertainty in future
+observations conditional on the parameters. The {\em posterior
+  predictive} distribution incorporates both sources.
+\begin{equation}
+
+\end{equation}
+
+Figure \ref{fig5} shows the predi
 
@@
   \centering
- \includegraphics[width=4.0in]{fig/b10andpi.pdf}
+  \includegraphics[width=.8\textwidth]{fig/b10andpi}
   \caption{Left: 50\% credible intervals for B10 (time in years till 10\% of the drives fail).
@@
   \label{fig4}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=.8\textwidth]{fig/dm-eval}
+  \caption{Left: Posterior predictive TTF (years) for new
+    unit, by drive-model. Right: TTF (top) and loss based on exponential
+    devaluation, $\lambda=5$ (bottom), for best 11 drive-models.}
+  \label{fig5}
 \end{figure}

### 2017-05-31 workflow/cost_analysis.R
Renames pdfs, replace "_" with "-"

@@ -49,6 +49,7 @@ p3 <- res %>%
   ggplot(aes(x=model, y=annualized_loss, ymin=lower2, ymax=upper2)) +
   geom_pointrange()+
   ylab("") +
+  scale_y_continuous(trans="sqrt", breaks=c(.01,.05,.10,.15,.25))+
   ggtitle("Annualized adjusted loss")
 
 c1 <- cowplot::plot_grid(p1,p3,ncol=1)

### 2017-05-31 paper/paper.tex
Start writing cost analysis section

@@
 \subsection{Drive-model Comparisons}
-We consider the problem of selecting between drive-models. From the
-perspective of return on investment, is clear
+We consider the problem of selecting between drive-models. From a
+business perspective, it is clear
 that we should prefer the drive-models which will provide more years
-of service. There are two sources of variation at play; the
+of service. For ease of exposition, we will assume that the purchase
+price of a hard-drive is the same across drive-models.
+
+There are two sources of variation at play; the
 posterior uncertainty in the parameters, which largely depends on the
@@
   predictive} distribution incorporates both sources.
-\begin{equation}
+\begin{align*}
+  p(t_{d,new}|t) = \int p(\t_{d,new}|\theta_d) p(\theta_d|t) \, d\theta_d\\
+\end{align*}
+We can sample from this distribution by drawing $t_{d,new}^{(s)}$ from
+$\operatorname{GLFP}(\pi_d^{(s)},
+\mu_1^{(s)},\sigma_1^{(s)},\mu_{2,d}^{(s)},\sigma_{2,d}^{(s)})$, for
+$s=1,...,S$, using the saved posterior draws for the model parameters.
 
-\end{equation}
+Predicted time to failure does not account for depreciation. Assuming that
+antiquated computer hardware tends to have less practical value, we
+may want to use a different measure which will reward more the drive-models
+that are very reliable in the first 0-5 years compared with those with a
+larger number of early failures, but which may last 20 years.
 

### 2017-06-01 paper/paper.tex
Edited figures and added legend functionality to model comparison plots.

--- 2017-05-31 paper/paper.tex
@@
 The distribution of drives by model varies: some models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.
-Figure \ref{fig1} shows the distribution of total number of failures and total time on test for drive brand models with at least 3 failures.
-As discussed later, for model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.
+Figure \ref{fig1} shows the distribution of total number of failures and total observed time on test for drive brand models with at least 3 failures.
+For model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.
 
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{fig/data-sum.pdf}
+\includegraphics[width=.9\textwidth]{fig/data-sum.pdf}
+\centering
 \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data.
-Right: Total time in operation (log scale) for each of the drive-models in the Backblaze data.
+Right: Total observed time in operation (log scale) for each of the drive-models in the Backblaze data.
 Both plots are sorted based on the total number of failures.}
@@
 The second mode is non-defective hard drives that eventually fail due to wearout.
-Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring around 15,000 hours.
+Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring between 8000 and 20,000 hours.
 Therefore, fitting a single Weibull model is not flexible enough to model the failure distribution.
@@
     \centering
-   \includegraphics[width=5.0in]{fig/brand-compare-nobands}
-		\caption{Estimated GFLP for all 4 Model Specifications.  Truncated Adjusted Kaplan Meier points are also presented.    \label{fig:first}} 
+   \includegraphics[width=.8\textwidth]{fig/mod_compare_legend}
+		\caption{Estimated GLFP for 4 Drive Brand Models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan Meier points.  \label{fig:first}} 
 \end{figure}
@@
   \centering
-  \includegraphics[width=.8\textwidth]{fig/b10andpi}
+  \includegraphics[width=\textwidth]{fig/b10andpi}
   \caption{Left: 50\% credible intervals for B10 (time in years till 10\% of the drives fail).

### 2017-06-01 plots/compare_fits.R

@@ -46,11 +46,12 @@ KM_list <- list()
 xlimits <- c(100,50000)
 ylimits <- c(.0001, .9)
 
-null_band <- KM_band(id=1, n_iter= 100, samp=sampnull, xlim=xlimits, ylim=ylimits, quantiles=c(.05,.5,.95),
+null_band <- KM_band("Model 1", id=1, n_iter= 100, samp=sampnull, xlim=xlimits, ylim=ylimits, quantiles=c(.05,.5,.95),
                      x_logscale=T, verbose=F, n = 100, pi.free=F, mu1.free=F, sigma1.free = F, mu2.free = F, sigma2.free = F,
-                     linetp = "longdash", colband="green", colline="darkblue")
+                     linetp = "longdash", colband="green")
 
-for(j in 1:length(id)){
+#for(j in 1:length(id)){
+for(j in 1:44){
   orig_id <- id[j]
   pi_j = exp(sampfull$log_pi[max_id,j])
   loc1_j = sampfull$mu1[max_id]
@@ -61,20 +62,23 @@ for(j in 1:length(id)){
   
   dat_tmp <- subset(mods, model == orig_id)
   
-  #KM_list[[j]] <- KM_plot(data = dat_tmp, model = "weibull", tr_adj = adj, title="", linear_axes = FALSE, fixed= TRUE, xlimits=xlimits, ylimits = ylimits,verbose = F)
+  KM_list[[j]] <- KM_plot(data = dat_tmp, model = "weibull", tr_adj = adj, title="", linear_axes = FALSE, fixed= TRUE, xlimits=xlimits, 
+                          size=.2, ylimits = ylimits,verbose = F)
   
-  KM_list[[j]] <- KM_plot_NP(data=dat_tmp, model = "weibull", tr_adj=adj, title = NULL, linear_axes = FALSE, fixed=TRUE,
-                             xlimits=xlimits, ylimits=ylimits, verbose=F, conf=.90)
-  tp2_list[[j]] <- KM_band(id=j, samp=samptp2, xlim=xlimits, ylim=ylimits, n_iter=100, quantiles=c(.05,.5,.95),
+  #KM_list[[j]] <- KM_plot_NP(data=dat_tmp, model = "weibull", tr_adj=adj, title = NULL, linear_axes = FALSE, fixed=TRUE,
+                            # xlimits=xlimits, ylimits=ylimits, verbose=F, conf=.90)
+  
[…]

### 2017-06-01 plots/plots_forpaper.R

@@ -3,7 +3,7 @@ library(dplyr)
 library(rstan)
 library(cowplot)
 
-dat <- readRDS("BB_data/clean_unit_summaries.rds")
+dat <- readRDS("../BB_data/clean_unit_summaries.rds")
 dat$model <- as.integer(dat$model)
 dat$failed[dat$failed==2] <- 1
 
@@ -39,7 +39,7 @@ tot$model <- factor(tot$model,levels=tot$model[order(tot$count)])  #Sort by Tota
 theme_set(theme_bw(base_size=18))
 #ggplot(tot, aes(x = count)) + geom_histogram(bins = 15) + scale_x_log10()
 fail <- ggplot(tot, aes(x=model, y=count)) + geom_bar(stat="identity") + xlab("Drive Brand Model") + ylab("Total Failures") + 
-   ggtitle("Total Failures by Drive Brand Model") + scale_y_log10() + coord_flip() 
+   ggtitle("Total Failures") + scale_y_log10(breaks=c(10,100,1000)) + coord_flip() 
 
 #Total Time on Test
 tt <- datsub %>% group_by(model) %>% mutate(tt=(endtime-starttime)/(24*365))
@@ -53,8 +53,8 @@ tt2$model <- factor(tt2$model,levels=tt2$model[order(tt2$count)])  #Sort by Tota
 
 theme_set(theme_bw(base_size=18))
 testtime <- ggplot(tt2, aes(x=factor(model),y=total)) + geom_bar(stat="identity") + xlab("Drive Brand Model") + 
-  ylab("Total Time (Years)") + ggtitle("Total Running Time by Drive Brand Model") + scale_y_log10() +
-  coord_flip()
+  ylab("Total Time (Years)") + ggtitle("Total Observed Running Time") + scale_y_log10(breaks=c(10,100,1000, 10000)) +
+  coord_flip() + theme(axis.ticks.y = element_blank(), axis.text.y = element_blank())
 
 #Make Cow Plot
 plot_grid(fail, testtime, ncol = 2, nrow = 1)

### 2017-06-01 plotting_fns/KM_plot.R

@@ -77,7 +77,7 @@ get_tr_adj <- function(starttime, pi, loc1, scl1, loc2, scl2){
 
 ## ggplot object containing Weibull probability plot with point estimate and optional credible band
 KM_plot <- function(data, model, tr_adj = 0, title = NULL,
-                    linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000), ylimits=c(.0001, .9999), verbose=F){
+                    linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000), ylimits=c(.0001, .9999), verbose=F, size=.3){
   require(ggplot2)
   require(plyr)
   require(dplyr)
@@ -106,7 +106,7 @@ KM_plot <- function(data, model, tr_adj = 0, title = NULL,
   }
   
   p <- df %>%
-    ggplot(aes(t, Ft)) + geom_point()
+    ggplot(aes(t, Ft)) + geom_step(size=size)
   
   if(!linear_axes){
     if(!fixed){
@@ -120,15 +120,15 @@ KM_plot <- function(data, model, tr_adj = 0, title = NULL,
         scale_y_continuous(trans = ytrans, breaks = ybrks(ylimits[1], ylimits[2], model="weibull"), limits=ylimits)
     }
   } else{
-    p <- p + scale_x_continuous(limits=xlimits) + scale_y_continuous(limits=ylimits)
+    p <- p + scale_x_continuous(limits=xlimits) + scale_y_continuous(limits=ylimits) 
   }
   p <- p +
-    theme_bw(base_size = 14) + ggtitle(title)+xlab("Hours")+ylab("Fraction Failing")
+    theme_bw(base_size = 14) + ggtitle(title) + xlab("Hours") + ylab("Fraction Failing") 
   p
 }
 
-KM_band <- function(id, samp, n_iter=NULL, xlim, ylim, quantiles=c(.05, .5, .95), x_logscale=T, n=30, verbose=F,
-                    pi.free=T, mu1.free=F, sigma1.free=F, mu2.free=T, sigma2.free=T, linetp="dashed", colline="black", colband="red"){
+KM_band <- function(num, id, samp, n_iter=NULL, xlim, ylim, quantiles=c(.05, .5, .95), x_logscale=T, n=30, verbose=F,
[…]

### 2017-06-01 plotting_fns/greenwood_errors.R

@@ -1,6 +1,6 @@
 KM_plot_NP <- function(data, model, tr_adj = 0, title = NULL,
                     linear_axes = FALSE, fixed=FALSE, xlimits=c(1000, 50000),
-                    ylimits=c(.0001, .9999), conf=.95, verbose=F){
+                    ylimits=c(.0001, .9999), conf=.95, verbose=F, size=.2){
   require(ggplot2)
   require(plyr)
   require(dplyr)
@@ -37,7 +37,7 @@ KM_plot_NP <- function(data, model, tr_adj = 0, title = NULL,
   }
   
   p <- df %>%
-    ggplot(aes(t, Ft)) + geom_point() +
+    ggplot(aes(t, Ft)) + geom_step(size=size) +
     geom_line(aes(y=lowFt), color="green", lty=2)+
     geom_line(aes(y=uprFt), color="green", lty=2)
     

### 2017-06-01 plotting_fns/quantile.R

@@ -2,7 +2,7 @@
 library(rstan)
 
 # s <- readRDS("reduce_relaxed.rds")
-s <- readRDS("MCMC_draws/samples_lor_only3fails.rds")
+s <- readRDS("../MCMC_draws/samples_lor_only3fails.rds")
 samp <- extract(s)
 
 ## Get Quantiles from GFLP Model ##
@@ -31,7 +31,7 @@ bfcn <- function(model, b, samp, quantiles = c(.025, .5, .975)){
 }
 
 #reduced version
-bfcn2 <- function(model, b, samp, quantiles = c(.025, .5, .975)){
+bfcn2 <- function(model, b, samp, quantiles = c(.25, .5, .75)){
   
   j <- model 
   Qp <- sapply(1:dim(samp$lp__), function(i){
@@ -45,10 +45,10 @@ bfcn2 <- function(model, b, samp, quantiles = c(.025, .5, .975)){
 }
 
 #Run this Function for All Models.  Note:  Should be a quick way to combine this with first function.
-df_reduced <- NULL
-for (j in 1:21){
-  df_reduced <- rbind(df_reduced,bfcn2(j,.1, samp))
-}
+#df_reduced <- NULL
+#for (j in 1:21){
+#  df_reduced <- rbind(df_reduced,bfcn2(j,.1, samp))
+#}
 
 df <- NULL
 for (j in 1:44){
@@ -58,14 +58,14 @@ for (j in 1:44){
 
 
 #B10_Model4.rds is in paper folder for future use
-b10 <- readRDS("paper/B10_Model4.rds")
+b10 <- readRDS("../paper/B10_Model4.rds")
 colnames(b10) <- c("model","lb","med","ub")
 b10$lb <- b10$lb/(24*365)
 b10$med <- b10$med/(24*365)
 b10$ub <- b10$ub/(24*365)
 
 #Sort by Lower End Point Time
-b10$model <- factor(b10$model,levels=b10$model[order(b10$lb)])  #Sort By Lower End Point
+b10$model <- factor(b10$model,levels=b10$model[order(b10$med)])  #Sort By Lower End Point
 
 
 #Make Catepillar Plot for B10; Perhaps Sort by Sample Size?

### 2017-06-01 plots/compare_fits.R
Updated mod_compare_legend (shared legend)

@@ -24,10 +24,10 @@ sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
 tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")$median
 
 
-ssigtp <- readRDS("../MCMC_draws/vary_s2_and_tp2_4_17.rds")
-stp2 <- readRDS("../MCMC_draws/samples_tp2_vary_new.rds")
-snull <- readRDS("../MCMC_draws/samples_null_model_3_29.rds")
-sfull <- readRDS("../MCMC_draws/samples_lor_only3fails.rds")
+# ssigtp <- readRDS("../MCMC_draws/vary_s2_and_tp2_4_17.rds")
+# stp2 <- readRDS("../MCMC_draws/samples_tp2_vary_new.rds")
+# snull <- readRDS("../MCMC_draws/samples_null_model_3_29.rds")
+# sfull <- readRDS("../MCMC_draws/samples_lor_only3fails.rds")
 
 
 
@@ -46,7 +46,7 @@ KM_list <- list()
 xlimits <- c(100,50000)
 ylimits <- c(.0001, .9)
 
-null_band <- KM_band("Model 1", id=1, n_iter= 100, samp=sampnull, xlim=xlimits, ylim=ylimits, quantiles=c(.05,.5,.95),
+null_band <- KM_band("1", id=1, n_iter= 100, samp=sampnull, xlim=xlimits, ylim=ylimits, quantiles=c(.05,.5,.95),
                      x_logscale=T, verbose=F, n = 100, pi.free=F, mu1.free=F, sigma1.free = F, mu2.free = F, sigma2.free = F,
                      linetp = "longdash", colband="green")
 
@@ -68,15 +68,15 @@ for(j in 1:44){
   #KM_list[[j]] <- KM_plot_NP(data=dat_tmp, model = "weibull", tr_adj=adj, title = NULL, linear_axes = FALSE, fixed=TRUE,
                             # xlimits=xlimits, ylimits=ylimits, verbose=F, conf=.90)
   
-  tp2_list[[j]] <- KM_band("Model 2", id=j, samp=samptp2, xlim=xlimits, ylim=ylimits, n_iter=100, quantiles=c(.05,.5,.95),
+  tp2_list[[j]] <- KM_band("2", id=j, samp=samptp2, xlim=xlimits, ylim=ylimits, n_iter=100, quantiles=c(.05,.5,.95),
[…]

### 2017-06-01 plots/indepedent_fits.R
update/add new figures

@@ -1,0 +1,47 @@
+setwd("workflow2/")
+#3 8  14
+s <- readRDS("samples_nopool_data_dm14.rds")
+print(s)
+library(plyr)
+library(rstan)
+library(dplyr)
+library(ggplot2)
+dat <- readRDS("../BB_data/clean_unit_summaries.rds")
+ov <- readRDS("../BB_data/overview.rds")
+mod_id <- filter(ov, f>=5, early_f>=1, late_f>=1) %>%
+  select(model) %>%
+  unlist
+xlim = c(2000,50000)
+ylim = c(.005,.8)
+todo <- c(3,8,14)
+
+#these corresponds to full_model ids: 
+# match(mod_id[todo], filter(ov, f>=3)$model) 
+#[1]  6 14 23
+
+plots <- list()
+bands <- list()
+
+for(id in 1:3){
+  dm = todo[id]
+  subdat <- filter(dat, model == mod_id[dm])
+  samp <- extract(readRDS(paste(c("samples_nopool_data_dm",dm,".rds"), collapse="")))
+  plots[[id]] <- KM_plot_NP(subdat, "weibull", conf=.9, xlimits=xlim, ylimits=ylim)+
+    theme_bw()+
+    theme(legend.position="none",
+          axis.title = element_blank(),
+          axis.ticks = element_blank(),
+          axis.text = element_blank(),
+          panel.grid = element_blank()
+          )
+  bands[[id]] <- KM_band("1", 1, samp, 100, xlim=xlim, ylim=ylim, pi.free=F, mu1.free=F, sigma1.free=F,
+               mu2.free=F, sigma2.free=F)
+}
+
+
+pdf("../plots/ind_estimates.pdf")
+library(cowplot)
+plot_grid(plots[[1]]+bands[[1]],
+          plots[[2]]+bands[[2]],
+          plots[[3]]+bands[[3]], ncol=3)
+dev.off()

### 2017-06-01 plots/indepedent_fits.R
shuffle placement of figures, update plot

@@ -11,8 +11,8 @@ ov <- readRDS("../BB_data/overview.rds")
 mod_id <- filter(ov, f>=5, early_f>=1, late_f>=1) %>%
   select(model) %>%
   unlist
-xlim = c(2000,50000)
-ylim = c(.005,.8)
+xlim = c(1000,50000)
+ylim = c(.008,.8)
 todo <- c(3,8,14)
 
 #these corresponds to full_model ids: 
@@ -29,19 +29,20 @@ for(id in 1:3){
   plots[[id]] <- KM_plot_NP(subdat, "weibull", conf=.9, xlimits=xlim, ylimits=ylim)+
     theme_bw()+
     theme(legend.position="none",
-          axis.title = element_blank(),
-          axis.ticks = element_blank(),
-          axis.text = element_blank(),
+          axis.title.y = element_blank(),
           panel.grid = element_blank()
-          )
-  bands[[id]] <- KM_band("1", 1, samp, 100, xlim=xlim, ylim=ylim, pi.free=F, mu1.free=F, sigma1.free=F,
+          )+
+    scale_x_continuous(limits=xlim, breaks=c(1000, 2000, 5000, 10000, 20000, 40000), trans="log")+
+    scale_y_continuous(limits=ylim, breaks=c(.01, .05, .1, .2, .5, .7, .9, .95, .99), trans="qsev")
+  bands[[id]] <- KM_band("1", 1, samp, 2000, xlim=xlim, ylim=ylim, pi.free=F, mu1.free=F, sigma1.free=F,
                mu2.free=F, sigma2.free=F)
 }
 
+no_yaxis_nums <- theme(axis.text.y=element_blank())
+library(cowplot)
 
 pdf("../plots/ind_estimates.pdf")
-library(cowplot)
-plot_grid(plots[[1]]+bands[[1]],
-          plots[[2]]+bands[[2]],
-          plots[[3]]+bands[[3]], ncol=3)
+plot_grid(plots[[1]]+bands[[1]]+ggtitle("6"),
+          plots[[2]]+bands[[2]]+no_yaxis_nums+ggtitle("14"),
+          plots[[3]]+bands[[3]]+no_yaxis_nums+ggtitle("23"), ncol=3)
 dev.off()

### 2017-06-01 paper/paper.tex
shuffle placement of figures, update plot
Add to paper, rethinking analysis of loss

--- 2017-06-01 paper/paper.tex
@@
 \subsection{Model Comparisons}
-\begin{figure}
-\includegraphics[width=\textwidth]{fig/heterogeneity-compare-preliminary}
-\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
-Right: Pointwise posterior median time to failure curves for Models 1-4, orderered left to right, top to bottom.}
-\label{fig2}
-\end{figure}
 
@@
 \begin{figure}[H]
+\includegraphics[width=\textwidth]{fig/heterogeneity-compare-preliminary}
+\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
+Right: Pointwise posterior median time to failure curves for Models 1-4, orderered left to right, top to bottom.}
+\label{fig2}
+\end{figure}
+
+\begin{figure}[H]
     \centering
-   \includegraphics[width=.8\textwidth]{fig/mod_compare_legend}
+   \includegraphics[width=\textwidth]{fig/mod_compare_legend}
 		\caption{Estimated GLFP for 4 Drive Brand Models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan Meier points.  \label{fig:first}} 
@@
 \subsection{Drive-model Comparisons}
-We consider the problem of selecting between drive-models. From the
-perspective of return on investment, is clear
+We consider the problem of ranking drive-models. From a
+business perspective, it is clear
 that we should prefer the drive-models which will provide more years
-of service. There are two sources of variation at play; the
+of service. For ease of exposition, we will assume that the purchase
+price of a hard-drive is the same across drive-models.
+
+There are two sources of variation at play; the
 posterior uncertainty in the parameters, which largely depends on the
@@
   predictive} distribution incorporates both sources.
-\begin{equation}
+\begin{equation*}
+  p(t_{d,new}|t) = \int p(t_{d,new}|\theta_d) p(\theta_d|t) \, d\theta_d
+\end{equation*}
 
-\end{equation}
+We can sample from this distribution by drawing $t_{d,new}^{(s)}$ from
+$\operatorname{GLFP}(\pi_d^{(s)},
+\mu_1^{(s)},\sigma_1^{(s)},\mu_{2,d}^{(s)},\sigma_{2,d}^{(s)})$, for
+$s=1,...,S$, using the saved posterior draws for the model
+parameters. The expected time-to-failure (TTF) for drive-model $d$ can be estimated by
+$$\frac{1}{S} \sum_{s=1}^{S} t_{d,new}^{(s)}.$$
+
+While TTF is clearly important, due to the
+anticipation of advancements in technology, we can expect that the
+relative value of computer hardware will depreciate rather quickly. To account
+for this, we can posit a model for this depreciation and instead of
+TTF, we can consider the value at replacement. The IRS considered
+computer hardware to be ``5-year'' equipment.(form 4562) Using``the declining
+balance method'' the rate of depreciation is 40\% per year.
+
+Let $L(t) = e^{-.4t}$ represent the value at the time of failure
+relative to its original value. We can then define the predicted loss
+of a unit as its value at time of failure.
 

### 2017-06-01 plots/indepedent_fits.R

@@ -11,7 +11,7 @@ ov <- readRDS("../BB_data/overview.rds")
 mod_id <- filter(ov, f>=5, early_f>=1, late_f>=1) %>%
   select(model) %>%
   unlist
-xlim = c(1000,50000)
+xlim = c(1000,55000)
 ylim = c(.008,.8)
 todo <- c(3,8,14)
 
@@ -41,7 +41,7 @@ for(id in 1:3){
 no_yaxis_nums <- theme(axis.text.y=element_blank())
 library(cowplot)
 
-pdf("../plots/ind_estimates.pdf")
+pdf("../plots/ind_estimates.pdf", width=10, height=5)
 plot_grid(plots[[1]]+bands[[1]]+ggtitle("6"),
           plots[[2]]+bands[[2]]+no_yaxis_nums+ggtitle("14"),
           plots[[3]]+bands[[3]]+no_yaxis_nums+ggtitle("23"), ncol=3)

### 2017-06-01 workflow/cost_analysis.R

@@ -5,7 +5,7 @@ source("../predictive/sample_from_predictive.R")
 # tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")
 V <- function(x){
   #devaluation curve
-  5*dexp(x,1/5)
+  exp(-.4*x)
 }
 
 library(plyr)
@@ -19,10 +19,10 @@ res <- ldply(1:44, function(dm){
   s2 <- samp$sigma2[,dm]
   newtime <- sample_glfp(n=length(samp$lp__), p, m1, s1, m2, s2)/24/365
   meantime <- mean(newtime)
-  annualized_loss <- mean(V(newtime)) * 1/mean(newtime)
+  loss <- V(newtime)
   q <- quantile(newtime, c(.25,.5,.75))
-  q2 <- quantile(V(newtime)/newtime, c(.25,.5,.75))
-  data.frame(model = dm, mean_time = meantime, annualized_loss = annualized_loss,
+  q2 <- quantile(loss, c(.25,.5,.75))
+  data.frame(model = dm, mean_time = meantime, mean_loss = mean(loss),
              lower = q[1], median = q[2], upper = q[3],
              lower2 = q2[1], median2 = q2[2], upper2 = q2[3])
 })
@@ -55,8 +55,8 @@ p3 <- res %>%
 c1 <- cowplot::plot_grid(p1,p3,ncol=1)
 
 p4 <- res %>%
-  ggplot(aes(x=model, y=median, ymin=lower, ymax=upper)) + geom_pointrange(linetype=2) +
-  scale_y_continuous()+geom_hline(yintercept=9, color="red") +
+  ggplot(aes(x=model, y=median2, ymin=lower2, ymax=upper2)) + geom_pointrange(linetype=2) +
+  scale_y_continuous()+#geom_hline(yintercept=9, color="red") +
   theme(axis.title.x = element_blank(),
         axis.ticks.x = element_blank(),
         axis.text.x = element_blank()) +

### 2017-06-03 paper/paper.tex
Added discussion piece and made small edits to first part.

@@
 This model can be easily estimated, but a common GLFP model for all drive brand groups fits poorly: the model captures drive models with lots of observed failures, while the predicted time to failure for drive brand models with less information is over or under estimated.
-At the other extreme, fitting individual models to each hard drive brand excludes many drive brand models with fewer than 5 failures.
+At the other extreme, fitting individual models to each hard drive brand excludes many drive brand models with few failures.
 The GLFP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.
@@
 Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive brand models to obtain stable estimates.
-This allows us to include all of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
+This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
 
@@
 We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
-In Section 4 we compare hard drive brand models based on our final GLFP model.
+In Section 4 we compare the different model specifications, in addition to the hard drive brand models based on our final GLFP model.
 Finally, in Section 5, we review the GLFP model and propose future extensions.
@@
 The distribution of drives by model varies: some models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.
-Figure \ref{fig1} shows the distribution of total number of failures and total observed time on test for drive brand models with at least 3 failures.
+Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive brand models with at least 3 failures.
 For model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.
@@
 \cite{betancourt}. Multiple chain were run for 1500 iterations after 1500 warmup iterations. Convergence was assessed by examining posterior
-plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$. Four chains were run for models 1,2 and 3 and 16 chains were run for model 4.
+plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$. Four chains were run for Models 1,2 and 3 and 16 chains were run for Model 4.
 
@@
 \caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
-Right: Pointwise posterior median time to failure curves for Models 1-4, orderered left to right, top to bottom.}
+Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
 \label{fig2}
@@
 
+We can also compare the model fits for each drive brand model individually.
+Using the set of parameters that maximize the posterior log probability, we examine the GLFP cdf for each of the 4 model specifications.
+The GLFP fit for drive brand models 2, 9, 14, and 40 are presented in Figure 4.
+The black step function again corresponds to the adjusted Kaplan-Meier estimates.
+As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.
+All the GLFP models overestimate the proportion of failed hard drives for drive model 2, but Model 4 is the least biased.
+For drive model 14, Models 1-3 underestimate the proportion of drives failing, and for drive model 40, they slightly overestimate the data.
+Conversely, Model 4 follows the observed failures quite closely--accurately capturing failures at the lower and upper ends of the distribution.
+The difference in accuracy between Models 3 and 4 are and Models 1 and 2 are visually quite apparent.
+Differentiating between Model 3 and 4 is less clear; for instance, the fits from Model 3 and 4 line almost exactly on top of each other for drive model 40.
 \begin{figure}[H]
@@
    \includegraphics[width=\textwidth]{fig/mod_compare_legend}
-		\caption{Estimated GLFP for 4 Drive Brand Models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan Meier points.  \label{fig:first}} 
+		\caption{Estimated GLFP for 4 Drive Brand Models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
 \end{figure}
 
-In addition to graphically examining model fits via posterior credible bands, we compared the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
+To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
 As outlined in Vehtari et.

### 2017-06-05 paper/paper.tex
Edits to Comparison Section.

@@
 
-Figure \ref{fig5} shows the predi
+We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a central measure of tendency.
+On the left panel of Figure 5 we plot B10, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive brand models.
+The best six drive models ranked by TTF are also the drive models with the longest time to B10.
+Drive brand models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.
+Another interesting feature to compare is $\pi$, the proportion failing due to infant mortality.
+One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each hard drive model.
+The plot is again sorted according to B10 so it can be compared to the left hand side.
+While for many of the drive models the ordinal ranking is the same as on the left, there are some drive model comparisons, for example 23 and 18, that differ in ranking if compared using B10 or $\pi$.
+Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.
 

### 2017-06-13 paper/paper.tex
Add citation for depreciation rate

@@
   \includegraphics[width=.9\textwidth]{fig/data-sum.pdf}
-\centering
   \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data.
@@
 relative value of computer hardware will depreciate rather quickly. To account
-for this, we can posit a model for this depreciation and instead of
-TTF, we can consider the value at replacement. The IRS considered
-computer hardware to be ``5-year'' equipment.(form 4562) Using``the declining
+for this depreciation, rather than using TTF as the metric for drive-model comparison, we use the value at replacement. The IRS considers
+computer hardware to be ``5-year'' equipment\cite{f4562}. Using``the declining
 balance method'' the rate of depreciation is 40\% per year.
@@
 Let $L(t) = e^{-.4t}$ represent the value at the time of failure
-relative to its original value. We can then define the predicted loss
-of a unit as its value at time of failure.
+relative to a new unit. We can rank the drive-models by 
+$$\op{E}(L(t_{d,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{d,new}^{(s)}).$$
 
@@
   \includegraphics[width=.8\textwidth]{fig/dm-eval}
-  \caption{Left: Posterior predictive TTF (years) for new
-    unit, by drive-model. Right: TTF (top) and loss based on exponential
-    devaluation, $\lambda=5$ (bottom), for best 11 drive-models.}
+  \caption{Left: Posterior predictive relative value-at-loss for new
+    unit, by drive-model. Right: TTF (top) and VAF based on 40\% annual depreciation (bottom), for best 11 drive-models.}
   \label{fig5}

### 2017-06-16 paper/paper.tex
Adding to Conclusion section

@@
 
-\section{Discussion}
-Review the advantages of fitting the GFLP model and offer future ideas.
+\section{Conclusions}
+We should point out some assumptions made in this analysis.
+First, we assume exchangability of units within drive models.
+This assumption is due to our ignorance about the causes of failure and of potentially important covariates.
+If we had information about distinguishing characteristics of the two identified failure modes, then we should use this information.
+Also, there could be many environmental or human-related factors that impact lifetime, such as operating conditions associated with location in the data center, which are likely confounded with drive-model.
 
-Advances in computer technology have made computationally intensive Bayesian methods for fitting hierarchical models feasible in practice.
+Second, it is probably true that many modes of failure occur and that the heterogeneity that we observe among drive models is due to different combinations of these.
+Our ignorance of the causes of failure leads us to approximate the main failure mode with a single drive-model specific Weibull distribution.
+Because of this, we suspect that our forecasts are too confident; especially for ``well-estimated" models, our ``model certainty" makes the uncertainty about these parameters too small.
+On the other hand, our model borrows strength across drive-models, including those with information about late failures.
+In addition, we suggest that using metrics for comparing drive-models that reduce the importance of the right tail of the distribution are appropriate for these data.
+With respect to decisions based on such metrics, we believe the assumptions in the previous paragraph to be the most critical.
+Those issues are typical to observational data such as these.
+
+Consumers of high-tech manufactured goods usually lack the expertise to perform a failure analysis to determine the cause of failure.
+The ability to fit marginal models, such as GLFP, can allow investigators to answer many pertinent questions while still appropriately accounting for uncertainty and avoiding substantial model bias.
+
+Advances in technology have made computationally intensive Bayesian methods for fitting hierarchical models feasible in practice.
+Hierarchical modeling can be an effective way to regularize a large number of models fits by borrowing information.
+Because all the data contribute some information about the global parameters, it is often sufficient to use weakly informative priors on the global parameters, making the resulting shrinkage data dependent.
+Using samples from the full joint posterior avoids the need for potential problematic asymptotic approximations.
+
 Analysis using HMC is accessible to practitioners by Stan.
 HMC can be very efficient in terms of effective samples because of the way that it uses joint updates for all parameters using Hamiltonian dynamics to rapidly explore the target distribution.
-Hierarchical modeling can be a useful way to pool information to estimate large models.
-By using Bayesian methods like the one presented here we can avoid using potentially problematic asymptotic approximations, since all posterior uncertainty is automatically integrated into averages computed from functions of the samples.
 

### 2017-06-19 paper/paper.tex
Consistent usage of "drive-model" to distinguish from many other uses of the word model

@@
 by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
-model (GLFP), which provides a generative model for
+model (GLFP), for
 failure times in a population where some units are susceptible to infant mortality. This model
@@
 The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
-Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{sujata}.
+Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive-models \cite{sujata}.
 The GLFP combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
@@
 While the GLFP model allows for multiple hazard functions, it does not account for the group structure of the hard drive data.
-A naive approach would combine all drive models and fit an overall GLFP model.
-This model can be easily estimated, but a common GLFP model for all drive brand groups fits poorly: the model captures drive models with lots of observed failures, while the predicted time to failure for drive brand models with less information is over or under estimated.
-At the other extreme, fitting individual models to each hard drive brand excludes many drive brand models with few failures.
-The GLFP model applied to these data produces very imprecise parameter estimates or can only be fit if highly informative priors are imposed.
+A naive approach would combine all drive-models and fit an overall GLFP model.
+This model can be easily estimated, but a common GLFP model for all drive-models fits poorly: the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
+At the other extreme, fitting individual models to each drive-mode is difficult for drive-models with few failures.
+Due to multimodality, in these cases the model fit is unstable unless strongly informative priors are chosen.
 \\
@@
 We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
-With over 60 brand drive models in testing a hierarchical model is advantageous as it pools information from across brands to get more precise estimates of the GLFP parameters for each drive model group.
-Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive brand models to obtain stable estimates.
+With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.
+Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.
 This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
@@
 We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
-In Section 4 we compare the different model specifications, in addition to the hard drive brand models based on our final GLFP model.
+In Section 4 we compare the different model specifications, in addition to the hard drive-models based on our final GLFP model.
 Finally, in Section 5, we review the GLFP model and propose future extensions.
@@
 While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.
-The purpose is to provide consumers and businesses with reliability information on different hard drive brand models.
+The purpose is to provide consumers and businesses with reliability information on different hard drive-models.
 The hard drives continuously spin in controlled storage pods where they run until failure.
 When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.
-In addition, the number of storage pods is increasing as Backblaze adds new hard drive models to the sample.
+In addition, the number of storage pods is increasing as Backblaze adds new drive-models to the sample.
 Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
-In addition, Backblaze publishes summary statistics of the different hard drive models currently operating.
+In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.
 No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.
 
-As of the first quarter of 2016, Backblaze was collecting data on 63 different hard drive models.
-Some drive models have been running since 2013, while others were added at a later date.
+As of the first quarter of 2016, Backblaze was collecting data on 63 different drive-models.
+Some drive-models have been running since 2013, while others were added at a later date.
 There are a total of 75,297 hard drives in operation.
-The distribution of drives by model varies: some models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.
-Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive brand models with at least 3 failures.
+The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.
+Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive-models with at least 3 failures.
 For model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.
@@
 Applying this method to the hard drive data, we first must estimate
-the empirical cdf for each model using the Kaplan-Meier estimator
+the empirical cdf for each drive-model using the Kaplan-Meier estimator
 \cite{kaplan}.  With left truncation, however, the standard
@@
 Turnbull, and given in more detail by Meeker and Escobar
-\cite{turnbull,meeker}.  For each hard drive model we select
+\cite{turnbull,meeker}.  For each drive-model we select
 $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
@@
 
-In Figure 2 we plot the Kaplan-Meier adjusted cdf for hard drive model 14 on Weibull paper.
-Drive model 14 had the most observed failures of all the brands.
+In Figure 2 we plot the Kaplan-Meier adjusted cdf for drive-model 14 on Weibull paper.
+Drive-model 14 had the most observed failures of all the brands.
 Each point on the plot corresponds to an observed failure.
@@
 \includegraphics[width=.5\textwidth]{fig/mod14-KM-adj}
-\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive model 14.  Dashed lines are Greenwood standard errors.  Plotted on Weibull paper.}
+\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive-model 14.  Dashed lines are Greenwood standard errors.  Plotted on Weibull paper.}
 \end{figure}
@@
 \section{Hierarchical GLFP model}
-For modeling the lifetime of hard-drives, we select the Generalized Limited Failure Population model of \citet{chan}.
+For modeling the lifetime of hard drives, we select the Generalized Limited Failure Population model of \citet{chan}.
 Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
@@
 
-To borrow strength across models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters.
+To borrow strength across drive-models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters.
 For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
@@
 Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by brand drive model.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by drive-model.
 
-For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive brand models.
+For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.
 We found that there was insufficient information in the data to model these parameters hierarchically.
@@
 \begin{enumerate}
-\item This model constrains all drive models to the same failure distribution. For this ``reduced" model we assume the priors :
+\item This model constrains all drive-models to the same failure distribution. For this ``reduced" model we assume the priors :
 
@@
 The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit.
-The upper left panel shows the estimated failure curve fit to all data ignoring drive model as a factor (Model 1).
-The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.
+The upper left panel shows the estimated failure curve fit to all data ignoring drive-model as a factor (Model 1).
+The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive-models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.
 
-The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which mean that the ranking of models with respect to fraction failing differs as a function of time.
+The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which mean that the ranking of drive-models with respect to fraction failing differs over time.
 
-Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive model, there is greater variability in the early rate of failure and, for some drive models, the forecasted fraction failing is adjusted downward.
+Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive-model, there is greater variability in the early rate of failure and, for some drive-models, the forecasted fraction failing is adjusted downward.
 
@@
 
-We can also compare the model fits for each drive brand model individually.
+We can also compare the model fits for each drive-model individually.
 Using the set of parameters that maximize the posterior log probability, we examine the GLFP cdf for each of the 4 model specifications.
-The GLFP fit for drive brand models 2, 9, 14, and 40 are presented in Figure 4.
+The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure 4.
 The black step function again corresponds to the adjusted Kaplan-Meier estimates.
 As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.
-All the GLFP models overestimate the proportion of failed hard drives for drive model 2, but Model 4 is the least biased.
-For drive model 14, Models 1-3 underestimate the proportion of drives failing, and for drive model 40, they slightly overestimate the data.
+All the GLFP models appear to overestimate the proportion of failed hard drives for drive-model 2, but Model 4 agrees the most with the nonparametric estimate.
+For drive-model 14, Models 1-3 underestimate the proportion of drives failing, and for drive-model 40, they slightly overestimate the data.
 Conversely, Model 4 follows the observed failures quite closely--accurately capturing failures at the lower and upper ends of the distribution.
 The difference in accuracy between Models 3 and 4 are and Models 1 and 2 are visually quite apparent.
-Differentiating between Model 3 and 4 is less clear; for instance, the fits from Model 3 and 4 line almost exactly on top of each other for drive model 40.
+Differentiating between Model 3 and 4 is less clear; for instance, the fits from Model 3 and 4 line almost exactly on top of each other for drive-model 40.
 \begin{figure}[H]
@@
    \includegraphics[width=\textwidth]{fig/mod_compare_legend}
-		\caption{Estimated GLFP for 4 Drive Brand Models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
+		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
 \end{figure}
@@
 We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a central measure of tendency.
-On the left panel of Figure 5 we plot B10, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive brand models.
-The best six drive models ranked by TTF are also the drive models with the longest time to B10.
-Drive brand models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.
+On the left panel of Figure 5 we plot B10, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
+The best six drive-models ranked by TTF are also the drive-models with the longest time to B10.
+Drive-models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.
 Another interesting feature to compare is $\pi$, the proportion failing due to infant mortality.
-One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each hard drive model.
+One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each drive-model.
 The plot is again sorted according to B10 so it can be compared to the left hand side.
-While for many of the drive models the ordinal ranking is the same as on the left, there are some drive model comparisons, for example 23 and 18, that differ in ranking if compared using B10 or $\pi$.
+While for many of the drive-models the ordinal ranking is the same as on the left, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using B10 or $\pi$.
 Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.
@@
 We should point out some assumptions made in this analysis.
-First, we assume exchangability of units within drive models.
+First, we assume exchangability of units within drive-models.
 This assumption is due to our ignorance about the causes of failure and of potentially important covariates.
@@
 Our ignorance of the causes of failure leads us to approximate the main failure mode with a single drive-model specific Weibull distribution.
-Because of this, we suspect that our forecasts are too confident; especially for ``well-estimated" models, our ``model certainty" makes the uncertainty about these parameters too small.
+Because of this, we suspect that our forecasts are too confident; especially for ``well-estimated" models, our uncertainty in the right tail of the distribution is probably too small.
 On the other hand, our model borrows strength across drive-models, including those with information about late failures.
@@
 
-Analysis using HMC is accessible to practitioners by Stan.
-HMC can be very efficient in terms of effective samples because of the way that it uses joint updates for all parameters using Hamiltonian dynamics to rapidly explore the target distribution.
+Analysis using HMC is accessible to practitioners by Stan. HMC has been shown to be very efficient in effective samples per iteration because its joint updating scheme.
 

### 2017-06-23 paper/paper.tex
New outline

@@
 
-%\bibliographystyle{natbib}
+\bibliographystyle{natbib}
 
@@
 \newpage
+\tableofcontents
+\newpage
 \spacingset{1.45} % DON'T change the spacing!
 \section{Introduction}
+You are a consumer with reliability data for products from multiple sources and wish to estimate lifetime, possibly for purposes of comparison.
+
+We present a method for joint estimation of reliability data.
+This model is appropriate for products that have both early failures due to defect as well as due to wearout such as electronics.
+We illustrate this method in a setting where the information in the data is unevenly distributed across the populations of interest.
+
+\subsection{Motivation}
+
 Failure distributions of high-reliability products, can be difficult
@@
 left-truncation.
+
+The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GLFP lifetime model.
+Product populations often contain a mixture of defective and non defective units.
+The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
+Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive-models \cite{sujata}.
+The GLFP combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
+\\
+
+While the GLFP model allows for multiple hazard functions, it does not account for the group structure of the hard drive data.
+A naive approach would combine all drive-models and fit an overall GLFP model.
+This model can be easily estimated, but a common GLFP model for all drive-models fits poorly: the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
+At the other extreme, fitting individual models to each drive-mode is difficult for drive-models with few failures.
+Due to multimodality, in these cases the model fit is unstable unless strongly informative priors are chosen.
+\\
+
+We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
+With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.
+Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.
+This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
 
@@
 
-\subsection{Background}
+\subsection{Overview}
+The structure of the paper is as follows.
+Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.
+Section 2 introduces the GLFP model as a mixture of two Weibull distributions.
+In Section 3 we fit multiple specifications of the GLFP model to best describe the Backblaze data.
+We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
+In Section 4 we compare the different model specifications, in addition to the hard drive-models based on our final GLFP model.
+Finally, in Section 5, we review the GLFP model and propose future extensions.
+
+\section{Background}
 In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.
@@
 
-\subsection{Motivation}
-The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GLFP lifetime model.
-Product populations often contain a mixture of defective and non defective units.
-The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
-Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive-models \cite{sujata}.
-The GLFP combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
-\\
+\section{GLFP model}
+\subsection{Likelihood}
+\subsubsection{With Truncation}
+\subsection{Example 1}
+Analysis of single drive model
 
-While the GLFP model allows for multiple hazard functions, it does not account for the group structure of the hard drive data.
-A naive approach would combine all drive-models and fit an overall GLFP model.
-This model can be easily estimated, but a common GLFP model for all drive-models fits poorly: the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
-At the other extreme, fitting individual models to each drive-mode is difficult for drive-models with few failures.
-Due to multimodality, in these cases the model fit is unstable unless strongly informative priors are chosen.
-\\
+\section{Hierarchical GLFP model}
+For modeling the lifetime of hard drives, we select the Generalized Limited Failure Population model of \citet{chan}.
+Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
+We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
+$$P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), \mbox{ for }t>0 \mbox{, and where } \pi_d \mbox{ is in }(0,1).$$
 
-We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
-With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.
-Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.
-This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
+As in \cite{chan}, we assume $F_{dj}$ is a member of the Weibull family of cdfs and parameterize in terms of a log-location parameter $\mu_{dj}$ and log-scale parameter $\sigma_{dj}$ so that
 
-\subsection{Overview}
-The structure of the paper is as follows.
-Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.
-Section 2 introduces the GLFP model as a mixture of two Weibull distributions.
-In Section 3 we fit multiple specifications of the GLFP model to best describe the Backblaze data.
-We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
-In Section 4 we compare the different model specifications, in addition to the hard drive-models based on our final GLFP model.
-Finally, in Section 5, we review the GLFP model and propose future extensions.
+$$F_{dj}(t) = 1 - \exp \left\{ -\exp \left\{ \frac{ \log (t) - \mu_{dj}}{\sigma_{dj}} \right\} \right\},\; t>0, j=1,2$$
 
-\section{Data}
+For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{d1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The interpretation of the parameter $\pi_d$ is to represent the proportion of units susceptible to early failure, hence susceptible to both failure modes.
+Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.
+
+To borrow strength across drive-models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters.
+For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
+
+$$\sigma_{d,1} \ind \op{Lognormal} \left( \eta_{\sigma,1}, \tau^2_{\sigma,1} \right) \mbox{ for } d=1,\ldots,D$$
+$$\sigma_{d,2} \ind \op{Truncated-Lognormal} \left( \eta_{\sigma,2}, \tau^2_{\sigma,2} , 0, 1 \right) \mbox{ for } d=1,\ldots,D$$
+$$t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j)  \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
+
+$$\op{logit} \pi_d \ind \op{N}(\eta_\pi, \tau_\pi) \mbox{ for } d=1,\ldots,D.$$
+
+Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution.
+We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that,by nature, wearout produces an increasing hazard function.
+The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = t_{0.632}$, is that lifetime data often, as is true of the data presented here, features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
+For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.
+
+We consider the models with the following set of restrictions (from most to least restrictive):
+
+\begin{enumerate}
+\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
+\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
+\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
+\item $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
+\end{enumerate}
+
+The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations.
+Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
+Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by drive-model.
+
+For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.
+We found that there was insufficient information in the data to model these parameters hierarchically.
+Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).
+
+\section{Data analysis}
+\subsection{Data}
 Backblaze is a company that offers backup storage to protect against data loss.
@@
 
-\section{Hierarchical GLFP model}
-For modeling the lifetime of hard drives, we select the Generalized Limited Failure Population model of \citet{chan}.
-Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
-We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
-$$P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), \mbox{ for }t>0 \mbox{, and where } \pi_d \mbox{ is in }(0,1).$$
-
-As in \cite{chan}, we assume $F_{dj}$ is a member of the Weibull family of cdfs and parameterize in terms of a log-location parameter $\mu_{dj}$ and log-scale parameter $\sigma_{dj}$ so that
-
-$$F_{dj}(t) = 1 - \exp \left\{ -\exp \left\{ \frac{ \log (t) - \mu_{dj}}{\sigma_{dj}} \right\} \right\},\; t>0, j=1,2$$
-
-For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{d1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The interpretation of the parameter $\pi_d$ is to represent the proportion of units susceptible to early failure, hence susceptible to both failure modes.
-Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.
-
-To borrow strength across drive-models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters.
-For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
-
-$$\sigma_{d,1} \ind \op{Lognormal} \left( \eta_{\sigma,1}, \tau^2_{\sigma,1} \right) \mbox{ for } d=1,\ldots,D$$
-$$\sigma_{d,2} \ind \op{Truncated-Lognormal} \left( \eta_{\sigma,2}, \tau^2_{\sigma,2} , 0, 1 \right) \mbox{ for } d=1,\ldots,D$$
-$$t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j)  \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
-
-$$\op{logit} \pi_d \ind \op{N}(\eta_\pi, \tau_\pi) \mbox{ for } d=1,\ldots,D.$$
-
-Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution.
-We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that,by nature, wearout produces an increasing hazard function.
-The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = t_{0.632}$, is that lifetime data often, as is true of the data presented here, features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
-For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.
-
-We consider the models with the following set of restrictions (from most to least restrictive):
-
-\begin{enumerate}
-\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
-\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
-\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
-\item $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
-\end{enumerate}
-
-The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations.
-Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
-Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by drive-model.
-
-For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.
-We found that there was insufficient information in the data to model these parameters hierarchically.
-Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).
-
 \subsection{Priors}
@@
 plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$. Four chains were run for Models 1,2 and 3 and 16 chains were run for Model 4.
-
-\section{Data analysis}
 
@@
 
-\bibliographystyle{plainnat}
 \bibliography{sample}

### 2017-06-23 plots/compare_fits.R

@@ -206,23 +206,75 @@ pi.plot
 #Make Cow Plot
 plot_grid(b10.plot, pi.plot, ncol = 2, nrow = 1)
 
-
+#needs xlimits, ylimits in global env. as well as KM_list, null_band, full_list
 full_pls_km <- function(j){
   datKM <- KM_list[[j]]$data
   datKM$t <- pmax(xlimits[1], datKM$t)
   datNULL <- null_band[[2]]$data
   datBAND <- full_list[[j]][[1]]$data
+  xlabels <- c(.01,.05,.25,1,2,5)
   ggplot(datBAND) + geom_line(aes(x,y), color="red")+
     geom_ribbon(aes(x,y,ymin=lower,ymax=upper), fill="red",alpha=.3)+
     geom_step(data=datKM, inherit.aes =FALSE, aes(x=t,y=Ft), color="black")+
     geom_line(data=datNULL, aes(x,y), color="green")+
-    scale_x_continuous(name="time(hours)",trans="log", limits = xlimits,
-                       breaks=xbrks(xlimits[1], xlimits[2], prec=0, N = 5))+
+    scale_x_continuous(name="time(years)",trans="log", limits = xlimits,
+                       breaks=xlabels*365*24,
+                       labels=xlabels)+
     scale_y_continuous(name="proportion failing", trans="qsev", limits = ylimits,
                        breaks=ybrks(ylimits[1], ylimits[2], model = "weibull", len=5))+
-    theme(legend.position="none") + ggtitle(paste(j))+
-    theme_bw()
+    theme_bw()+
+    theme(legend.position="none"#,
+#          axis.text.x=element_text(angle=45, hjust=1, vjust=1)
+          ) + ggtitle(paste(j))
 }
 cowplot::plot_grid(full_pls_km(11), full_pls_km(9), 
                    full_pls_km(10), full_pls_km(15))
-p
+
+cowplot::plot_grid(full_pls_km(4), full_pls_km(5), full_pls_km(2), full_pls_km(3),
+                   full_pls_km(6), full_pls_km(8), full_pls_km(26), full_pls_km(23),
+                   nrow=2)
+
+#lifetime plots
+j = 11
+
+lifetime_plot <- function(mod, to_show=NULL){
+  xlabels <- c(.01,.05,.25,1,2,5)
+  subdat <- filter(dat, model==id[mod])
+  if(!is.null(to_show)){
[…]

### 2017-07-01 paper/paper.tex
Edits on GLFP Section for Single Drive Model.

@@
 \usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf,float}
+\usepackage{graphicx,psfrag,epsf,float, bm}
 \graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
@@
 \section{GLFP model}
-\subsection{Likelihood}
-\subsubsection{With Truncation}
+For modeling the lifetime of hard drives, we use the Generalized Limited Failure Population model of \citet{chan}.
+Let $T_{d,i}$ be the time of failure for drive $i$ of drive-model $d$.
+We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
+$$F_d(T_{d,i}) = P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), t>0 \mbox{, where } 0 < \pi_d < 1.$$
+
+As in \cite{chan}, we assume $F_{dj}$ has a Weibull distribtion. The standard parametrization of the Weibull distribution is
+\begin{equation*}
+F(T;\mu,\sigma)= 1-\text{exp}\left[-\left(\frac{T}{\eta}\right)^{\beta}\right] = \Phi_{sev} \left(\frac{\text{log}(T)-\mu}{\sigma}\right) 
+\end{equation*}
+where $\sigma=\frac{1}{\beta}$, $\mu=\text{log}(\eta)$ and $\Phi_{sev}(z)=1-\text{exp}[-\text{exp}(z)]$.
+The flexibility of the Weibull distribution makes it quite useful empirically; it can model failure distributions with positive or negative hazard functions \cite{meeker}.
+The parameter $\beta$ captures this flexibility: if $\beta>1$ the Weibull distribution has an increasing hazard function, $\beta=1$ corresponds to constant hazard, and if $0<\beta<1$ the hazard is decreasing over time.
+The GLFP model is a mixture of two Weibull distributions.
+We will define each of the Weibull distributions in terms of a log-location parameter $\mu_{dj}$ and log-scale parameter $\sigma_{dj}$ so that
+$$F_{dj}(t) = 1 - \exp \left\{ -\exp \left\{ \frac{ \log (t) - \mu_{dj}}{\sigma_{dj}} \right\} \right\},\; t>0, j=1,2$$
+For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{d1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
+Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.
+
+\subsection{Censoring and Truncation}
+A common feature of lifetime data is right censoring.
+Due to time or budget constraints it is rare all units are observed until failure.
+Therefore, any units still in service when the data is analyzed are considered right censored.
+Despite over three years of data the hard drive times are heavily right censored.
+Hard drives are built to last for decades under normal use conditions.
+Even when running continuously fewer than two percent of all the drives Backblaze operates fail per year.
+For our analysis we code any hard drive still in operation as of the first quarter of 2016 as right censored.\\
+
+While right censoring imposes an upper bound on an observation, truncation means the existence of an observation is unknown if it falls within a certain interval \cite{meeker}.
+Other than the previous amount of running time, Backblaze does not provide information about why hard drives are added the sample at different times.
+Some hard drives enter the storage pod with zero hours of service.
+Others are added after thousands of hours of usage.
+Ignoring the left truncation and treating the data as identically distributed would lead to biased estimates.
+Dropping left truncated observations would substantially reduce our sample size.
+Therefore, both the right censoring and left truncation are incorporated into the likelihood function.
+
+\subsubsection{Likelihood}
+We will now define the general form for the likelihood function that takes into account left truncation, as well as right censoring.
+We will use the same likelihood form for each hard drive brand, and thus drop the drive-model, $d$, subscript for simplicity.
+Let $T_{i}$ denote the lifetime of hard drive $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.
+As mentioned earlier, the sample size, $n$, varies by hard drive brand.
+Let $t_i^L$ define the left truncation time in hours.
+Specifically, for how many hours has a hard drive been running for before it is added to the storage pod for observation.
+For hard drives that are right censored, $c_i$ denotes the number of hours a hard drive has survived without failing.
+We also specify the following indicator functions.
+Let $\nu_i=0$ if a hard drive is left truncated and $\nu_i=1$ otherwise.
+Let $\delta_i=0$ if a hard drive is right censored and $\delta_i=1$ otherwise.
+Given the two parametric forms for $F_1$ and $F_2$ the joint likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
+Assuming hard drives are independent, the complete likelihood for one hard drive brand is
+\begin{equation*}
+L(\bm{\theta})= \prod_{i=1}^{n} f(T_i; \bm{\theta})^{\nu_i\delta_i} \left[\frac{f(T_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{(1-\nu_i) \delta_i}
+[1-F(c_i;\bm{\theta})]^{\nu_i (1-\delta_i)} \left[ \frac{1-F(c_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{(1-\nu_i)(1-\delta_i)}
+\end{equation*}
+
 \subsection{Example 1}
-Analysis of single drive model
+We first fit the GLFP model to individual hard drive brand models.
+For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.
+If $\text{log}(Y) \sim \text{SEV}(\mu, \sigma)$ then $Y \sim \text{Weibull}(\beta = 1/\sigma, \eta = \exp(\mu))$.
+This decision was chosen after encountering numerical issues when performing estimation using the standard Weibull parametrization.
+In addition, we transformed the parameters to the lower log quantile, $\log(t_{p_j})$ (where $p_{1} = 0.5$, and $p_2 = 0.2$), and log scale, $\log(\sigma_j)$.
+This parameterization was selected as the majority of failures occur at the lower tail of the distribution.
+\\
+
+Given the instability of maximum likelihood estimates for many of the hard drive brand models, we used a Bayeisan approach and specified priors for following parameters where $\Phi^{-1}$ is the quantile function for the SEV distribution.
+
+$$t_{p_j} \equiv \mu_{j} + \sigma_{j}\,\Phi^{-1}(p_j) \quad j=1,2$$
+\[\log(t_{p_1})  \sim \operatorname{N}(\mu = 7, \sigma = 1 )\]
+\[\log(t_{p_2}) \sim \operatorname{N}(\mu = 9, \sigma = 2 )\]
+\[\sigma_j \sim \operatorname{Lognormal}(\mu = 0, \sigma =1 ) \quad j = 1,2\]
 
 \section{Hierarchical GLFP model}
-For modeling the lifetime of hard drives, we select the Generalized Limited Failure Population model of \citet{chan}.
-Let $T_{d,i}$ be the time of failure for the $i^{th}$ drive of drive-model $d$.
-We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
-$$P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), \mbox{ for }t>0 \mbox{, and where } \pi_d \mbox{ is in }(0,1).$$
-
-As in \cite{chan}, we assume $F_{dj}$ is a member of the Weibull family of cdfs and parameterize in terms of a log-location parameter $\mu_{dj}$ and log-scale parameter $\sigma_{dj}$ so that
-
-$$F_{dj}(t) = 1 - \exp \left\{ -\exp \left\{ \frac{ \log (t) - \mu_{dj}}{\sigma_{dj}} \right\} \right\},\; t>0, j=1,2$$
-
-For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{d1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The interpretation of the parameter $\pi_d$ is to represent the proportion of units susceptible to early failure, hence susceptible to both failure modes.
-Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.
 

### 2017-07-03 paper/paper.tex
Updates to top of document

@@
 \begin{document}
-
-\bibliographystyle{natbib}
 
@@
 \begin{abstract}
-The text of your abstract.  100 or fewer words.
+Consumers have an interest in accurate assessment of product reliability.
+Toward this end, there is a need for models that can fit the data together with methods that account for uncertainty.
+While computationally convenient, simple parametric models can fail to capture key characteristics found in real data.
+
+In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}.
+This 5-parameter model for lifetime data accomodates lifetime distributions displaying early failures due to "infant mortality" and wearout.
+By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
+By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.
+
+We illustrate this method on real data with left-truncation and right-censoring where the information in the data is unevenly distributed across the populations of interest.
 \end{abstract}
@@
 \section{Introduction}
-You are a consumer with reliability data for products from multiple sources and wish to estimate lifetime, possibly for purposes of comparison.
+Consumers have an interest in accurate assessment of product reliability.
+Toward this end, there is a need for models that can fit the data together with methods that account for uncertainty.
+While computationally convenient, simple parametric models can fail to capture key characteristics found in real data.
 
-We present a method for joint estimation of reliability data.
-This model is appropriate for products that have both early failures due to defect as well as due to wearout such as electronics.
-We illustrate this method in a setting where the information in the data is unevenly distributed across the populations of interest.
+In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}.
+This 5-parameter model for lifetime data accomodates lifetime distributions displaying early failures due to "infant mortality" and wearout.
+By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
+By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.
+
+We illustrate this method on real data with left-truncation and right-censoring where the information in the data is unevenly distributed across the populations of interest.
 
@@
 
-Failure distributions of high-reliability products, can be difficult
-to assess. Accelerated testing tries to address this problem, but
-requires strong assumptions to extrapolate to use conditions in the
-field (citation). When field data are available, information may be
-very limited. In this paper, we propose the use of Bayesian hierarchical modeling to borrow
-information to improve inferences on lifetime
+Failure distributions of high-reliability consumer products can be difficult
+to assess; when field data are available, information may be limited. In this paper, we propose the use of Bayesian hierarchical modeling to borrow
+information, improving inferences on lifetime
 distributions where information is limited by right-censoring and
@@
 
-The goal of this paper is to model, and compare, the lifetime distribution of different hard drive brands using the GLFP lifetime model.
-Product populations often contain a mixture of defective and non defective units.
-The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.
-Ignoring this hazard structure, which exists for series systems, could lead to spurious inference when comparing the reliability of hard drive-models \cite{sujata}.
-The GLFP combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.
-\\
+The reliability of many engineered products follow a similar pattern.
+Relatively high rates of failure occur in early life ("infant mortality").
+This can often be explained by defects from the manufacturing process.
+Following a period of "burn-in", after which the majority of the defective units have failed, failure rates flatten out during a period.
+Finally, after prolonged use, rates of failure increase due to wearout.
+Ignoring this pattern in failure rates can lead to suboptimal decisions.
+While simple nonparametric methods are available, these are not suitable for forecasting.
 
-While the GLFP model allows for multiple hazard functions, it does not account for the group structure of the hard drive data.
-A naive approach would combine all drive-models and fit an overall GLFP model.
-This model can be easily estimated, but a common GLFP model for all drive-models fits poorly: the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
-At the other extreme, fitting individual models to each drive-mode is difficult for drive-models with few failures.
-Due to multimodality, in these cases the model fit is unstable unless strongly informative priors are chosen.
+\citet{chan} proposed the Generalized Limited Failure Population (GLFP) model.
+This model includes the Weibull as a special case.
+Additionally, it theorizes an unknown proportion of defective units which are susceptible to competing risks.\\
+
+THIS NEEDS TO BE SEPARATED OUT
+While the GLFP model provides flexibility, it can be difficult to fit with limited data.
+When multiple populations are of interest, comparisons based on GLFP model fits will be restricted to those with sufficient data.
+A naive approach would be to combine the data from all drive-models and fit a single GLFP model.
+This model can be well-estimated, but this common GLFP model fits poorly.% the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
+At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.
+Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are chosen.
 \\
@@
 
+\bibliographystyle{plainnat}
 \bibliography{sample}

### 2017-07-05 paper/paper.tex
Edits to the Motivation section.

@@
   \title{\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes}
-  \author{Eric Mittman 1\thanks{
+  \author{Eric Mittman\thanks{
     The authors gratefully acknowledge Bill Meeker for his comments and suggestions}\hspace{.2cm}\\
@@
 In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}.
-This 5-parameter model for lifetime data accomodates lifetime distributions displaying early failures due to "infant mortality" and wearout.
+This 5-parameter model for lifetime data accommodates lifetime distributions displaying early failures due to "infant mortality" and wearout.
 By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
@@
 \noindent%
-{\it Keywords:}  3 to 6 keywords, (don't reuse words appearing in title)
+{\it Keywords:} GLFP, Bathtub hazard, Censored data, Hierarchical models, Bayesian estimation
 \vfill
@@
 
-Failure distributions of high-reliability consumer products can be difficult
-to assess; when field data are available, information may be limited. In this paper, we propose the use of Bayesian hierarchical modeling to borrow
-information, improving inferences on lifetime
-distributions where information is limited by right-censoring and
-left-truncation.
+Failure distributions of high-reliability consumer products can be difficult to assess.
+Even when field data are available, information may be limited for a variety of reasons.
+Certain brands may have few products in operation resulting in small sample sizes.
+Other products may have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.
+In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.
 
 The reliability of many engineered products follow a similar pattern.
-Relatively high rates of failure occur in early life ("infant mortality").
-This can often be explained by defects from the manufacturing process.
+Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process.
 Following a period of "burn-in", after which the majority of the defective units have failed, failure rates flatten out during a period.
 Finally, after prolonged use, rates of failure increase due to wearout.
-Ignoring this pattern in failure rates can lead to suboptimal decisions.
-While simple nonparametric methods are available, these are not suitable for forecasting.
+Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability.
+While simple nonparametric methods are available, such as the Kaplan-Meier estimator, these are not suitable for forecasting.
+They also do not allow for a straight forward statistical comparisons of different brands in order to select the most reliable product.
 
-\citet{chan} proposed the Generalized Limited Failure Population (GLFP) model.
-This model includes the Weibull as a special case.
-Additionally, it theorizes an unknown proportion of defective units which are susceptible to competing risks.\\
-
-THIS NEEDS TO BE SEPARATED OUT
-While the GLFP model provides flexibility, it can be difficult to fit with limited data.
-When multiple populations are of interest, comparisons based on GLFP model fits will be restricted to those with sufficient data.
-A naive approach would be to combine the data from all drive-models and fit a single GLFP model.
-This model can be well-estimated, but this common GLFP model fits poorly.% the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
-At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.
-Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are chosen.
+To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.
+In addition to describing the failure time distribution for products that are susceptible to infant mortality and wearout failure modes, it theorizes an unknown proportion of defective units which are susceptible to defects.
+While the GLFP model provides flexibility over a simple or piece-wise parametric model, it can be difficult to fit with limited data  As Chan and Meeker discuss in their original paper, few observed failures in addition to unknown causes of failure cause maximum likelihood estimates to be unstable or impossible to obtain.
+Therefore, when multiple populations are of interest, comparisons based on the standard GLFP model are restricted to those with sufficient data and observed failure from each failure mode.
+A naive approach to this limitation is to combine the data from all products and fit a single GLFP model.
+This model can be well-estimated, but the pooled GLFP model fits poorly.
+At the other extreme, fitting a separate GLFP model to each product is problematic for products where the data are weakly informative.
+Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are placed on the model parameters.
 \\
@@
 We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
-With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.
-Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.
-This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
+A hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for a whole class of products.
+Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: product brands with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other products to obtain stable estimates.
+This allows us to include all of the data and more importantly, make valid inferences for each individual product.
 

### 2017-07-05 paper/paper.tex
Eric changes to paper 7/5

--- 2017-07-03 paper/paper.tex
@@
 \usepackage{url} % not crucial - just used below for the URL
+
+\usepackage[x11names]{xcolor}
+\usepackage{framed} % provides framed env., should be removed in final draft
+\definecolor{shadecolor}{rgb}{1,.5,.5}
 
@@
 
+For real reliability data, often no failure analysis is available allowing objective discernment about when a failure is due to infant mortality.
+This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution.
+However, without any knowledge of cause of failure, traditional competing risk models are not identifiable.
 \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model.
-This model includes the Weibull as a special case.
-Additionally, it theorizes an unknown proportion of defective units which are susceptible to competing risks.\\
+This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life.
+It avoids non-identifiability by introducing a parameter representing the population proportion defective.
+Where this parameter is zero, the GLFP model reduces to a unimodal distribution.
+\\
 
-THIS NEEDS TO BE SEPARATED OUT
-While the GLFP model provides flexibility, it can be difficult to fit with limited data.
-When multiple populations are of interest, comparisons based on GLFP model fits will be restricted to those with sufficient data.
-A naive approach would be to combine the data from all drive-models and fit a single GLFP model.
-This model can be well-estimated, but this common GLFP model fits poorly.% the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
+While the GLFP model provides flexibility and accords with a macro-level theory of product failure, it requires a lot of data to fit because of its complexity.
+Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information.
+When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those with sufficient data.
+As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations, imposing "soft" constraints on model parameters.
+This can enables estimation for all populations; the posterior distribution of time-to-failure for populations with limited data will contain greater uncertainty, while being shrunk toward a "pooled" model.
+For model fitting, samples obtained from the posterior distribution via Markov Chain Monte Carlo for all model parameters allow easy numerical integration for computing various quantities of interest.\\
+
+% the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
+
+\begin{shaded}
+  THIS SHOULD BE SEPARATED OUT
   At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.
@@
 Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.
-This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.
+This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.\\
 
@@
   cloud-based storage company.
+\end{shaded}
 
@@
 The structure of the paper is as follows.
-Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.
-Section 2 introduces the GLFP model as a mixture of two Weibull distributions.
-In Section 3 we fit multiple specifications of the GLFP model to best describe the Backblaze data.
-We also discuss computational and modeling issues that suggest a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.
-In Section 4 we compare the different model specifications, in addition to the hard drive-models based on our final GLFP model.
-Finally, in Section 5, we review the GLFP model and propose future extensions.
+Section~\ref{sec:Background} describes some of the potential applications of our method and summarizes previous related work found in the literature.
+Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates an application in an example.
+Section~\ref{sec:Hierarchical GLFP model} introduces a hierarchical model for extending the GLFP model to multiple populations.
+In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set.
+Subsection~\ref{sec:Data} describes and summarizes the Backblaze Hard Drive data set.
+Subsection~\ref{sec:Model Comparisons} shows the use of approximate leave-one-out cross-validation as a way to do model selection between different hierarchical model specifications.
+We also address computational issues and modeling decisions that lead to a final model.
+In subsection~\ref{sec:Drive-model Comparisons} we present an analysis comparing drive-model lifetime distributions.
+Finally, in section~\ref{sec:Conclusions}, we discuss the strengths and limitations of our method and consider extensions and alternatives.
 
 \section{Background}
+\label{sec:Background}
 In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.
@@
 
-\section{GLFP model}
-For modeling the lifetime of hard drives, we use the Generalized Limited Failure Population model of \citet{chan}.
-Let $T_{d,i}$ be the time of failure for drive $i$ of drive-model $d$.
-We assume that $T_{d,1},\ldots T_{d,n_d}$ are independent and have a probability distribution with cdf given by
-$$F_d(T_{d,i}) = P(T_{d,i}\le t) = 1 - (1-\pi_d\, F_{d1}(t))(1 - F_{d2}(t)), t>0 \mbox{, where } 0 < \pi_d < 1.$$
+\section{Generalized Limited Failure Population model}
+\label{sec:GLFP model}
 
-As in \cite{chan}, we assume $F_{dj}$ has a Weibull distribtion. The standard parametrization of the Weibull distribution is
-\begin{equation*}
-F(T;\mu,\sigma)= 1-\text{exp}\left[-\left(\frac{T}{\eta}\right)^{\beta}\right] = \Phi_{sev} \left(\frac{\text{log}(T)-\mu}{\sigma}\right) 
-\end{equation*}
-where $\sigma=\frac{1}{\beta}$, $\mu=\text{log}(\eta)$ and $\Phi_{sev}(z)=1-\text{exp}[-\text{exp}(z)]$.
-The flexibility of the Weibull distribution makes it quite useful empirically; it can model failure distributions with positive or negative hazard functions \cite{meeker}.
+\subsection{Weibull parameterization}
+\label{sec:Weibull parameterization}
+For a random variable $Y$ following a Weibull distribution, the standard parameterization of the cumulative distribution function, in terms of scale parameter $\alpha$ and shape parameter $\beta$, is
+
+$$ P(Y \le y) = 1 - \exp \left\{ \left( \frac{y}{\alpha} \right)^\beta \right\}. $$
+
+An alternative parameterization, in terms of log-location parameter $\mu = \log \alpha$ and log-scale parameter $\sigma = 1/\beta$ is
+
+$$ P(Y \le y) = 1 - \exp \left\{ -\exp \left( \frac{ \log y - \mu }{\sigma} \right) \right\}. $$
+
+This make clear that $(\log Y - \mu)/\sigma$ follows the smallest extreme value distribution (SEV).
+The Weibull distribution is widely used in practice as it can model failure distributions with both increasing and decreasing hazard functions \cite{meeker}.
 The parameter $\beta$ captures this flexibility: if $\beta>1$ the Weibull distribution has an increasing hazard function, $\beta=1$ corresponds to constant hazard, and if $0<\beta<1$ the hazard is decreasing over time.
-The GLFP model is a mixture of two Weibull distributions.
-We will define each of the Weibull distributions in terms of a log-location parameter $\mu_{dj}$ and log-scale parameter $\sigma_{dj}$ so that
-$$F_{dj}(t) = 1 - \exp \left\{ -\exp \left\{ \frac{ \log (t) - \mu_{dj}}{\sigma_{dj}} \right\} \right\},\; t>0, j=1,2$$
-For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{d1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
+
+\subsection{GLFP model}
+\label{subsec:GLFP model}
+Following \cite{chan}, let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$ respectively.
+The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $Y_1,\ldots Y_{n} \ind \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
+$$P(Y_i \le y) = 1 - (1-\pi\, F_{1}(y))(1 - F_{2}(y)),\, t>0,\, 0 < \pi < 1.$$
+
+The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $D_i\ind \op Bern(\pi)$.
+$D_i$ is an indicator for whether or not an observational unit is defective.
+Expressed conditional on $D_i$,
+\begin{align*}
+P(Y_i\le y | D_i=1) =& 1 -(1-F_1(y))(1-F_2(y))\\
+P(Y_i\le y | D_i=0) =& F_2(y).
+\end{align*}
+
+For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
 Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.
@@
 \section{Hierarchical GLFP model}
+\label{sec:Hierarchical GLFP model}
 
@@
 \section{Data analysis}
+\label{sec:Data analysis}
 \subsection{Data}
+\label{sec:Data}
 Backblaze is a company that offers backup storage to protect against data loss.
@@
 
-\subsection{Priors}
+\subsection{Prior Distributions}
+\label{sec:Prior Distributions}
 To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior.
@@
 \subsection{Computation}
+\label{sec:Computation}
 Each model was fit using the {\tt
@@
 \subsection{Model Comparisons}
-
+\label{sec:Model Comparisons}
 The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit.
@@
 \subsection{Drive-model Comparisons}
+\label{sec:Drive-model Comparisons}
 We consider the problem of ranking drive-models. From a
@@
 \section{Conclusions}
+\label{sec:Conclusions}
 We should point out some assumptions made in this analysis.
@@
 
-\bigskip
+\pagebreak
+
 \begin{center}
@@
 
+\begin{appendix}
+\scriptsize
+\begin{verbatim}
+functions {
+  real sev_logpdf(real y, real mu, real sigma){
+    real z;
+    z = (y - mu) / sigma;
+    return -log(sigma) + z - exp(z);
+  }
+
+  real sev_logccdf(real y, real mu, real sigma){
+    return -exp((y - mu) / sigma);
+  }
+
+  real sev_logcdf(real y, real mu, real sigma){
+    return log1m_exp(-exp((y - mu) / sigma));
+  }
+}
+
+data {
+  int M;
+  int N_obs;
+  int N_cens;
+  real endtime_obs[N_obs];
+  real endtime_cens[N_cens];
+  real starttime_obs[N_obs];
+  real starttime_cens[N_cens];
+  int<lower=1, upper=M> dm_obs[N_obs];
+  int<lower=1, upper=M> dm_cens[N_cens];
+  vector<lower=0, upper=1>[2] p; # quantiles to model
+}
+transformed data{
+  vector[2] z_corr;
+  for(i in 1:2)
+    z_corr[i] = log(-1.0 * log1m(p[i])); # used to get location(mu) from quantile(t_p)
+}
+parameters{
+  real log_tp1;
+  real<lower=0> sigma1;
+  real eta_tp2;
+  real eta_s2;
+  real eta_pi;
+  real<lower=0> tau_tp2;
+  real<lower=0> tau_s2;
+  real<lower=0> tau_pi;
+  vector[M] log_tp2_raw;
+  real<lower=0, upper=1> sigma2[M];
+  vector[M] logit_pi_raw;
+}
+
+transformed parameters{
+  real mu1;
+  vector[M] mu2;
+  vector[M] log_pi;
+  mu1 = log_tp1 - sigma1 * z_corr[1];
+  for(m in 1:M){
+    mu2[m] = (eta_tp2 + tau_tp2 * log_tp2_raw[m]) - (sigma2[m] * z_corr[2]);
+  }
+  for(m in 1:M)
+    log_pi[m] = log_inv_logit(eta_pi + tau_pi * logit_pi_raw[m]);
+}
+
+model{
+  real tmp[2];
+  int m;
+  real logpi;
+  real mu_1;
+  real mu_2;
+  real sig_1;
+  real sig_2;
+
+  for(i in 1:N_obs){
+    m = dm_obs[i];
+    logpi = log_pi[m];
+    mu_2 = mu2[m];
+    sig_2 = sigma2[m];
+    // numerator:   log( p * f1 * (1 - F2) + f2 * (1 - p * F1) )
+    //            = log( exp(log(p) + log(f1) + log(1 - F2)) + exp(log(f2) + log(1 - exp(log(p) + log(F1)))) )
+    tmp[1] = log_sum_exp(logpi + sev_logpdf(endtime_obs[i], mu1, sigma1) +
+               sev_logccdf(endtime_obs[i], mu_2, sig_2),
+               sev_logpdf( endtime_obs[i], mu_2, sig_2) + 
+               log1m_exp(logpi + sev_logcdf(endtime_obs[i], mu1, sigma1)
+               )
+             );
+    // denominator:  log((1 - p * F1) * (1 - F2))
+    //            =  log(1 - p * F1) + log(1 - F2)
+    tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_obs[i], mu1, sigma1)) + 
+             sev_logccdf(starttime_obs[i], mu_2, sig_2);
+
+    target += tmp[1] - tmp[2];
+  }
+
+  for(i in 1:N_cens){
+    m = dm_cens[i];
+    logpi = log_pi[m];
+    mu_2 = mu2[m];
+    sig_2 = sigma2[m];
+
+    // numerator:   log((1 - p * F1) * (1 - F2))
+    //            =  log(1 - p * F1) + log(1 - F2)
+    tmp[1] = log1m_exp(logpi + sev_logcdf(endtime_cens[i], mu1, sigma1)) + 
+             sev_logccdf(endtime_cens[i], mu_2, sig_2);
+    // denominator:  log((1 - p * F1) * (1 - F2))
+    //            =  log(1 - p * F1) + log(1 - F2)
+    tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_cens[i], mu1, sigma1)) + 
+             sev_logccdf(starttime_cens[i], mu_2, sig_2);
+
+    target += tmp[1] - tmp[2];
+  }
+
+  log_tp1        ~ normal(7, 2);
+  log_tp2_raw    ~ normal(0, 1);
+  sigma1         ~ lognormal(0, 1);
+  sigma2         ~ lognormal(eta_s2, tau_s2);
+  logit_pi_raw   ~ normal(0, 1);
+  eta_tp2       ~ normal(9, 2);
+  eta_s2         ~ normal(0, 2);
+  eta_pi         ~ normal(-3, 1);
+  tau_tp2       ~ cauchy(0,1);
+  tau_s2         ~ cauchy(0,1);
+  tau_pi         ~ cauchy(0,1);
+}
+
+\end{verbatim}
+\end{appendix}
+
+\pagebreak
+
 \bibliographystyle{plainnat}

### 2017-07-05 paper/paper.tex

@@
 % math macros
+
 \newcommand{\ind}{\stackrel{ind.}{\sim}}
@@
 
-Failure distributions of high-reliability consumer products can be difficult
-to assess; when field data are available, information may be limited. In this paper, we propose the use of Bayesian hierarchical modeling to borrow
-information, improving inferences on lifetime
-distributions where information is limited by right-censoring and
-left-truncation.
+Failure distributions of high-reliability consumer products can be difficult to assess.
+Even when field data are available, information may be limited for a variety of reasons.
+Certain brands may have few products in operation resulting in small sample sizes.
+Other products may have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.
+In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.
 
 The reliability of many engineered products follow a similar pattern.
-Relatively high rates of failure occur in early life ("infant mortality").
-This can often be explained by defects from the manufacturing process.
+Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process.
 Following a period of "burn-in", after which the majority of the defective units have failed, failure rates flatten out during a period.
 Finally, after prolonged use, rates of failure increase due to wearout.
-Ignoring this pattern in failure rates can lead to suboptimal decisions.
-While simple nonparametric methods are available, these are not suitable for forecasting.
+Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability.
+While simple nonparametric methods are available, such as the Kaplan-Meier estimator, these are not suitable for forecasting.
+They also do not allow for a straight forward statistical comparisons of different brands in order to select the most reliable product.
 
@@
 \end{shaded}
+
+To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.
+In addition to describing the failure time distribution for products that are susceptible to infant mortality and wearout failure modes, it theorizes an unknown proportion of defective units which are susceptible to defects.
+While the GLFP model provides flexibility over a simple or piece-wise parametric model, it can be difficult to fit with limited data  As Chan and Meeker discuss in their original paper, few observed failures in addition to unknown causes of failure cause maximum likelihood estimates to be unstable or impossible to obtain.
+Therefore, when multiple populations are of interest, comparisons based on the standard GLFP model are restricted to those with sufficient data and observed failure from each failure mode.
+A naive approach to this limitation is to combine the data from all products and fit a single GLFP model.
+This model can be well-estimated, but the pooled GLFP model fits poorly.
+At the other extreme, fitting a separate GLFP model to each product is problematic for products where the data are weakly informative.
+Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are placed on the model parameters.
+\\
+
+We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
+A hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for a whole class of products.
+Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: product brands with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other products to obtain stable estimates.
+This allows us to include all of the data and more importantly, make valid inferences for each individual product.
+
+Failure data for series systems may be collected at the system level
+by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
+model (GLFP), for
+failure times in a population where some units are susceptible to infant mortality. This model
+covers a subset of the class of distributions known as bathtub distributions
+\cite{sujata}. Bathtub distributions are characterized by a U-shaped
+hazard function, where early failures are due to infant mortality and
+late failure are due to wearout. While complex systems actually
+exhibit more than two modes of failure, suitable parametric models
+which can suitably approximate the overall lifetime distribution are
+useful.
+
+In the case of complex, high-reliability systems where cause of failure is
+not available, a model which
+\begin{enumerate}[a.]
+\item has parameters that can be estimated from the observed data
+\item is flexible enough to fit the observed data
+\item in interpretable with respect to available theory of the underlying process
+\end{enumerate}
+is desired. We present a Bayesian, hierarchical modeling approach for
+grouped failure data. As a benefit of working with Monte Carlo
+samples, we can easily make inference on a wide range of quantities of
+interest while accounting for uncertainty. We demonstrate this approach on real data hard
+drive failure data from a
+cloud-based storage company.
 

### 2017-07-06 paper/paper.tex
Edits to Abstract, Motivation, and Model Section.

@@
 Consumers have an interest in accurate assessment of product reliability.
-Toward this end, there is a need for models that can fit the data together with methods that account for uncertainty.
-While computationally convenient, simple parametric models can fail to capture key characteristics found in real data.
+However, available reliability data often contains little information.
+Products with long lifetimes produce few observed failures, and when products do fail the cause of failure is often unknown.
+While computationally convenient, simple parametric models are often limited in modeling this type of data.
+Small sample sizes make parameter estimates unstable or difficult to obtain.
+Multiple failure modes cause bends in the lifetime distributions that single parametric models cannot capture.
 
-In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}.
-This 5-parameter model for lifetime data accommodates lifetime distributions displaying early failures due to "infant mortality" and wearout.
+In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}.
+This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  an early failure due to "infant mortality" and a second failure mode due to wearout.
+We fit the GLFP model using a hierarchical Bayesian approach.
 By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.
+Moreover, by using a Bayeisan approach, comparison of different product brands is straightforward since the estimation of arbitrary functionals are easily computable from the full posterior distribution.
 
-We illustrate this method on real data with left-truncation and right-censoring where the information in the data is unevenly distributed across the populations of interest.
+We illustrate this method on hard drive reliability data from a cloud-based computing company.
+The data contains both left-truncation and right-censoring.
+Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous.
 \end{abstract}
@@
 Certain brands may have few products in operation resulting in small sample sizes.
-Other products may have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.
+Other products could have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.
 In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.
@@
 Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process.
-Following a period of "burn-in", after which the majority of the defective units have failed, failure rates flatten out during a period.
+After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.
 Finally, after prolonged use, rates of failure increase due to wearout.
@@
 Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information.
-When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those with sufficient data.
+When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
 As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations, imposing "soft" constraints on model parameters.
-This can enables estimation for all populations; the posterior distribution of time-to-failure for populations with limited data will contain greater uncertainty, while being shrunk toward a "pooled" model.
-For model fitting, samples obtained from the posterior distribution via Markov Chain Monte Carlo for all model parameters allow easy numerical integration for computing various quantities of interest.\\
+This enables estimation for all product groups.
+Populations with lots of data will be precisely estimated.
+For products with limited data the posterior distribution of time-to-failure will contain greater uncertainty, while being shrunk toward a "pooled" model.
+Finally, in order to compare the reliability across brands, the Bayesian approach is appealing as samples obtained from the posterior distribution via Markov Chain Monte Carlo allow for easy numerical integration for computing  various quantities of interest.
+\\
 
@@
 
-\begin{shaded}
-  THIS SHOULD BE SEPARATED OUT
-  At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.
+%\begin{shaded}
+  %THIS SHOULD BE SEPARATED OUT
+  %At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.
 Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are chosen.
 \\
-
-  We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
+  %
+  %We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
 With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.
@@
 This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.\\
-
-  Failure data for series systems may be collected at the system level
-  by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
-  model (GLFP), for
-  failure times in a population where some units are susceptible to infant mortality. This model
-  covers a subset of the class of distributions known as bathtub distributions
-  \cite{sujata}. Bathtub distributions are characterized by a U-shaped
-  hazard function, where early failures are due to infant mortality and
-  late failure are due to wearout. While complex systems actually
-  exhibit more than two modes of failure, suitable parametric models
-  which can suitably approximate the overall lifetime distribution are
-  useful.
-
-  In the case of complex, high-reliability systems where cause of failure is
-  not available, a model which
-  \begin{enumerate}[a.]
-  \item has parameters that can be estimated from the observed data
-  \item is flexible enough to fit the observed data
-  \item in interpretable with respect to available theory of the underlying process
-  \end{enumerate}
-  is desired. We present a Bayesian, hierarchical modeling approach for
-  grouped failure data. As a benefit of working with Monte Carlo
-  samples, we can easily make inference on a wide range of quantities of
-  interest while accounting for uncertainty. We demonstrate this approach on real data hard
-  drive failure data from a
-  cloud-based storage company.
-\end{shaded}
-
-To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.
+  %
+  %
+  %Failure data for series systems may be collected at the system level
+  %by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
+  %model (GLFP), for
+  %failure times in a population where some units are susceptible to infant mortality. This model
+  %covers a subset of the class of distributions known as bathtub distributions
+  %\cite{sujata}. Bathtub distributions are characterized by a U-shaped
+  %hazard function, where early failures are due to infant mortality and
+  %late failure are due to wearout. While complex systems actually
+  %exhibit more than two modes of failure, suitable parametric models
+  %which can suitably approximate the overall lifetime distribution are
+  %useful. 
+  %
+  %In the case of complex, high-reliability systems where cause of failure is
+  %not available, a model which
+  %\begin{enumerate}[a.]
+  %\item has parameters that can be estimated from the observed data
+  %\item is flexible enough to fit the observed data
+  %\item in interpretable with respect to available theory of the underlying process
+  %\end{enumerate}
+  %is desired. We present a Bayesian, hierarchical modeling approach for
+  %grouped failure data. As a benefit of working with Monte Carlo
+  %samples, we can easily make inference on a wide range of quantities of
+  %interest while accounting for uncertainty. We demonstrate this approach on real data hard
+  %drive failure data from a
+  %cloud-based storage company.
+%\end{shaded}
+%
+%To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.
 In addition to describing the failure time distribution for products that are susceptible to infant mortality and wearout failure modes, it theorizes an unknown proportion of defective units which are susceptible to defects.
@@
 \\
-
-We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
+%
+%We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
 A hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for a whole class of products.
@@
 This allows us to include all of the data and more importantly, make valid inferences for each individual product.
-
-Failure data for series systems may be collected at the system level
-by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
-model (GLFP), for
-failure times in a population where some units are susceptible to infant mortality. This model
-covers a subset of the class of distributions known as bathtub distributions
-\cite{sujata}. Bathtub distributions are characterized by a U-shaped
-hazard function, where early failures are due to infant mortality and
-late failure are due to wearout. While complex systems actually
-exhibit more than two modes of failure, suitable parametric models
-which can suitably approximate the overall lifetime distribution are
-useful.
-
-In the case of complex, high-reliability systems where cause of failure is
-not available, a model which
-\begin{enumerate}[a.]
-\item has parameters that can be estimated from the observed data
-\item is flexible enough to fit the observed data
-\item in interpretable with respect to available theory of the underlying process
-\end{enumerate}
-is desired. We present a Bayesian, hierarchical modeling approach for
-grouped failure data. As a benefit of working with Monte Carlo
-samples, we can easily make inference on a wide range of quantities of
-interest while accounting for uncertainty. We demonstrate this approach on real data hard
-drive failure data from a
-cloud-based storage company.
+%
+%
+%Failure data for series systems may be collected at the system level
+%by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
+%model (GLFP), for
+%failure times in a population where some units are susceptible to infant mortality. This model
+%covers a subset of the class of distributions known as bathtub distributions
+%\cite{sujata}. Bathtub distributions are characterized by a U-shaped
+%hazard function, where early failures are due to infant mortality and
+%late failure are due to wearout. While complex systems actually
+%exhibit more than two modes of failure, suitable parametric models
+%which can suitably approximate the overall lifetime distribution are
+%useful. 
+%
+%In the case of complex, high-reliability systems where cause of failure is
+%not available, a model which
+%\begin{enumerate}[a.]
+%\item has parameters that can be estimated from the observed data
+%\item is flexible enough to fit the observed data
+%\item in interpretable with respect to available theory of the underlying process
+%\end{enumerate}
+%is desired. We present a Bayesian, hierarchical modeling approach for
+%grouped failure data. As a benefit of working with Monte Carlo
+%samples, we can easily make inference on a wide range of quantities of
+%interest while accounting for uncertainty. We demonstrate this approach on real data hard
+%drive failure data from a
+%cloud-based storage company.
 
@@
 Parameter estimation is straightforward if the cause of failure is known for each observation.
-With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the researcher.
+With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the analyst.
 \\
@@
 This helped identify parameters when using maximum likelihood (ML) estimation.
-Extending Chan and Meeker's analysis, Basu et.
-al performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.
+Extending Chan and Meeker's analysis, Basu et.al performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.
 Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.
@@
 considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.
-Treating the unknown failure modes as incomplete data, an expectation maximization algorithm with ML was used, in addition to Bayesian estimation.
+Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.
 
@@
 \subsection{Example 1}
-We first fit the GLFP model to individual hard drive brand models.
+We first demonstrate the GLFP model on real hard drive failure data from a cloud-based storage company.
+The data set has over 60 hard drive models in operation, but for now we focus on one model.
 For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.
@@
 $$t_{p_j} \equiv \mu_{j} + \sigma_{j}\,\Phi^{-1}(p_j) \quad j=1,2$$
-\[\log(t_{p_1})  \sim \operatorname{N}(\mu = 7, \sigma = 1 )\]
-\[\log(t_{p_2}) \sim \operatorname{N}(\mu = 9, \sigma = 2 )\]
-\[\sigma_j \sim \operatorname{Lognormal}(\mu = 0, \sigma =1 ) \quad j = 1,2\]
+$$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
+$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad \sigma_2 \sim \op{Lognormal}(0, 2.5) $$
+
+Question: Do we Want to explain these priors here or wait till the hierarchical section where we lay out all the priors?
 
@@
 \label{sec:Data}
-Backblaze is a company that offers backup storage to protect against data loss.
-While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.
+Backblaze is a company that offers cloud backup storage to protect against data loss.
+Since 2013 it has been collecting data on hard drives operating at its facility.
 The purpose is to provide consumers and businesses with reliability information on different hard drive-models.
-The hard drives continuously spin in controlled storage pods where they run until failure.
-When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.
-In addition, the number of storage pods is increasing as Backblaze adds new drive-models to the sample.
+The hard drives continuously spin in controlled storage pods.
+Drives are run until failure.
+When a hard drive fails it is permanently removed and replaced.
+In addition, the number of storage pods is increasing as Backblaze adds to its storage capacity.
 Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
 In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.
-No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.
+No other analysis or modeling of the failure data is provided.
+Backblaze does, however, encourage others to further analyze its data.
 
-As of the first quarter of 2016, Backblaze was collecting data on 63 different drive-models.
-Some drive-models have been running since 2013, while others were added at a later date.
-There are a total of 75,297 hard drives in operation.
-The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.
+As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.
+Some drive-models have been running since 2013 or before, while others were added at a later date.
+Data have been reported on 75,297 different hard drives that are or where in operation.
+The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
 Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive-models with at least 3 failures.
-For model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.
+For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analyses.
 
@@
   \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data.
-Right: Total observed time in operation (log scale) for each of the drive-models in the Backblaze data.
-Both plots are sorted based on the total number of failures.}
+Right: Total accumulated time in operation (log scale) for each of the drive-models in the Backblaze data.
+The plots are sorted based on the total number of failures on the left.}
   \label{fig1}

### 2017-07-08 paper/paper.tex
Edits, edits, edits.  Mostly adding to Examle 1 and making paper more broad.

@@
 Consumers have an interest in accurate assessment of product reliability.
-However, available reliability data often contains little information.
-Products with long lifetimes produce few observed failures, and when products do fail the cause of failure is often unknown.
+However, lifetime data is often sparse and has limited information.
+Products with long lifetimes produce few observed failures; and when products do fail the cause of failure is often unknown.
 While computationally convenient, simple parametric models are often limited in modeling this type of data.
 Small sample sizes make parameter estimates unstable or difficult to obtain.
-Multiple failure modes cause bends in the lifetime distributions that single parametric models cannot capture.
+Multiple failure modes cause bends in the lifetime distribution single parametric models cannot capture.
 
@@
 We fit the GLFP model using a hierarchical Bayesian approach.
-By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-Moreover, by using a Bayeisan approach, comparison of different product brands is straightforward since the estimation of arbitrary functionals are easily computable from the full posterior distribution.
-
-We illustrate this method on hard drive reliability data from a cloud-based computing company.
-The data contains both left-truncation and right-censoring.
-Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous.
+Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
+Moreover, using a Bayesian approach, comparison of different product brands is straightforward as estimation of arbitrary functionals are easily computable from the full posterior distribution.
 \end{abstract}
@@
 
-We illustrate this method on real data with left-truncation and right-censoring where the information in the data is unevenly distributed across the populations of interest.
+We illustrate this method on hard drive reliability data from a cloud-based computing company.
+The data contains both left-truncation and right-censoring.
+Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous.
 
@@
 In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.
-For example, a computer system can fail if the mother board, disc drive or power supply stop working.
+For example, a computer system can break if the mother board, disc drive or power supply stop working.
 Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.
@@
 For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
-Here the cause of failure is not assumed to be known, thus units of the same drive-model are exchangable.
+Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.
 
@@
 Therefore, any units still in service when the data is analyzed are considered right censored.
-Despite over three years of data the hard drive times are heavily right censored.
+Products with long expected lifetimes, such as hard drives, are heavily right censored.
 Hard drives are built to last for decades under normal use conditions.
-Even when running continuously fewer than two percent of all the drives Backblaze operates fail per year.
-For our analysis we code any hard drive still in operation as of the first quarter of 2016 as right censored.\\
+Even when running continuously, hard drives can exhibit failure rates of less than two percent per year \cite{backblaze}.
+\\
 
 While right censoring imposes an upper bound on an observation, truncation means the existence of an observation is unknown if it falls within a certain interval \cite{meeker}.
-Other than the previous amount of running time, Backblaze does not provide information about why hard drives are added the sample at different times.
-Some hard drives enter the storage pod with zero hours of service.
-Others are added after thousands of hours of usage.
-Ignoring the left truncation and treating the data as identically distributed would lead to biased estimates.
-Dropping left truncated observations would substantially reduce our sample size.
+For example, a product may appear in a sample after operating for thousands of hours.
+Ignoring the left truncation and treating the data as identically distributed leads to biased estimates.
+Dropping left truncated observations substantially reduces the sample size.
 Therefore, both the right censoring and left truncation are incorporated into the likelihood function.
@@
 \subsubsection{Likelihood}
-We will now define the general form for the likelihood function that takes into account left truncation, as well as right censoring.
-We will use the same likelihood form for each hard drive brand, and thus drop the drive-model, $d$, subscript for simplicity.
-Let $T_{i}$ denote the lifetime of hard drive $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.
-As mentioned earlier, the sample size, $n$, varies by hard drive brand.
+We now define the general form for the likelihood function that takes into account left truncation, as well as right censoring.
+Let $T_{i}$ denote the lifetime of product $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.
 Let $t_i^L$ define the left truncation time in hours.
-Specifically, for how many hours has a hard drive been running for before it is added to the storage pod for observation.
-For hard drives that are right censored, $c_i$ denotes the number of hours a hard drive has survived without failing.
+Specifically, for how many hours has a unit been running for before it is added to the sample for observation.
+For units that are right censored, $c_i$ denotes the number of hours a unit has survived without failing.
 We also specify the following indicator functions.
-Let $\nu_i=0$ if a hard drive is left truncated and $\nu_i=1$ otherwise.
-Let $\delta_i=0$ if a hard drive is right censored and $\delta_i=1$ otherwise.
+Let $\nu_i=0$ if a product is left truncated and $\nu_i=1$ otherwise.
+Let $\delta_i=0$ if a product is right censored and $\delta_i=1$ otherwise.
 Given the two parametric forms for $F_1$ and $F_2$ the joint likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
-Assuming hard drives are independent, the complete likelihood for one hard drive brand is
+Assuming all products are independent, the complete likelihood for one product brand is
 \begin{equation*}
@@
 \subsection{Example 1}
-We first demonstrate the GLFP model on real hard drive failure data from a cloud-based storage company.
-The data set has over 60 hard drive models in operation, but for now we focus on one model.
+We fit the GLFP model on real hard drive failure data from the cloud-based storage company, Backblaze.
+Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model \cite{chan}.
+Backblaze provides the number of hours the hard drive is in operation, and whether the drive failed or was right censored.
+No information is available about the cause of failure.
+We first focus on the drive brand model with the most observed failures to demonstrate the GLFP model.
+\\
+
 For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.
@@
 
-Given the instability of maximum likelihood estimates for many of the hard drive brand models, we used a Bayeisan approach and specified priors for following parameters where $\Phi^{-1}$ is the quantile function for the SEV distribution.
+Given the instability of maximum likelihood estimates for many of the hard drive brand models, we used a Bayesian approach and specified priors for following parameters where $\Phi^{-1}$ is the quantile function for the SEV distribution.
 
@@
 
-Question: Do we Want to explain these priors here or wait till the hierarchical section where we lay out all the priors?
+The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
+The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.
+(Note that, due to symmetry, the Weibull shape parameter has the same prior distribution and that a value of 1, corresponding to a constant hazard, is the median of this prior distribution.) The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$.
+For $t_{0.2,2}$ it is $(3.4,16.6)$.
+Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.
+They do, however, help identify the two failure modes.
+\\
+
+The model model was fit using the {\tt rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC) \cite{betancourt}.
+Multiple chains were run for 1500 iterations after 1500 burnin iterations.
+Convergence was assessed by examining posterior plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$.
+
+Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for drive model 14.
+The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.
+As expected, this fraction is small with a median value of 0.05.
+The parameter estimates for the two Weibull distributions also match intuition.
+The wearout failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.
+Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, an increasing hazard.
+The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than for the wearout mode.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & 2.5\% & 50\% & 97.5\% \\ 
+  \hline
+\pi & 0.03 & 0.05 & 0.10 \\ 
+ \beta_1 & 0.47 & 1.13 & 1.72 \\ 
+  \beta_2 & 4.47 & 4.70 & 4.95 \\ 
+  \mu_1 & 7.46 & 8.06 & 8.76 \\ 
+  \mu_2 & 10.12 & 10.13 & 10.14 \\ 
+   \hline
+\end{tabular}
+\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive Model 14.}
+\label{table:1}
+\end{table}
+
+To graphically examine the model fit, we plot the GLFP cdf using the median values of the posterior distribution along with 90\% credible interval bands.
+For reference, the solid line corresponds to a truncated adjusted non-parametric  estimator of $F(t)$.
+The posterior median follows the non-parametric estimate quite closely.
+Evidence eof at least two failure modes is seen in the probability plot having a kink occurring between 1 and 1.5 years.
+Moreover, the fitted GLFP curves bends with the non-parametric estimate, accurately following the curvature of the lifetime distribution.
+
+\begin{figure}[H]
+\centering
+  \includegraphics[width=.9\textwidth]{fig/drive14cred.pdf}
+  \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.
+Dashed line corresponds to the median of the posterior draws.
+Solid blue line is a non-parametric Kaplan-Meier estimator.
+Credible 90\% bands are in red.
+Greenwood non-parametric error bands are in blue.}
+  \label{fig1}
+\end{figure}
 
@@
 
-To borrow strength across drive-models, we can either share parameters across drive-models, or model the drive-model specific parameters hierarchically, allowing the data to inform the hyperparameters.
+To borrow strength across different product populations, we can either share parameters across products, or model the product specific parameters hierarchically, allowing the data to inform the hyperparameters.
 For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
@@
 Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution.
-We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that,by nature, wearout produces an increasing hazard function.
-The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = t_{0.632}$, is that lifetime data often, as is true of the data presented here, features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
-For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.
+We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that, by nature, wearout produces an increasing hazard function.
+The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = t_{0.632}$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
+%For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.
 
@@
 
-The set of model specifications were chosen based on the data, interpretation of the model, as well as estimation considerations.
-Drive brand specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
-Going from a common model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by drive-model.
+The set of model specifications were chosen based data, interpretation of the model, as well as estimation considerations.
+Product specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
+Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by product.
 
-For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.
-We found that there was insufficient information in the data to model these parameters hierarchically.
+For all of the models we consider, the parameters for the infant mortality failure mode are held in common across products.
+We found that there was often insufficient information to model these parameters hierarchically.
 Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).
@@
 Identifying whether failure data are consistent with a specific family of distributions is difficult to do by examining a histogram or other density estimate.
-By transform the y-axis of the estimated cumulative distribution function (cdf), denoted by $\widehat{F(t)}$, by the inverse cdf of for the standardized distribution, a visual comparison of the plot to a straight line can be an effective informal test for distributional goodness of fit.
+By properly transforming the axes of a plot of an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a graphical test for distributional goodness of fit.
 
-Applying this method to the hard drive data, we first must estimate
-the empirical cdf for each drive-model using the Kaplan-Meier estimator
-\cite{kaplan}.  With left truncation, however, the standard
-Kaplan-Meier estimator for drive-model $d$, denoted by
+Applying this method to the hard drive data, we start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
+\cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
 $\widehat{F_d(t)}_{KM}$, is conditional on survival up to
@@
 By sampling from the full posterior distribution, since
-$Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
+$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
 survived up to $t_{d,\text{min}}^L$) is a function of the model
@@
 2.3 years.
-
-%\begin{figure}
-%\includegraphics[width=.8\textwidth]{fig/data-overview}
-%\caption{Left: Kaplan-Meier plots showing nonparametric estimates of fraction failing for individual drive-models with truncation adjustments based on fitted model, Right: Total observed failures (top) and the sum of lengths of all observational intervals by drive-model.}
-%\end{figure}
 
@@
 \begin{enumerate}
-\item This model constrains all drive-models to the same failure distribution. For this ``reduced" model we assume the priors :
+\item This model constrains all drive-models to the same failure distribution. For this ``reduced" model we assume the same priors as in Example 1.
 
-$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad t_{0.5,1} \sim \op{Normal}(8,4)$$
-$$\sigma_2 \sim \op{Lognormal}(0, 2.5) \quad t_{0.2,2} \sim \op{N}(10,4)$$
-
-The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
-The prior on $\sigma_1$ and $\sigma_2$ put it on the interval $(0.02, 61.1)$ with $90\%$ probability.
-(Note that, due to symmetry, the Weibull shape parameter has the same prior interval and that a value of 1, corresponding to a constant hazard, is the median of this prior distribuion.) $t_{0.5,1}$ has a prior $90\%$ interval of $(1.4,14.6)$.
-For $t_{0.2,2}$ it is $(3.4,16.6)$.
-Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.
-
-\item We allow $t_{pd}$ to vary by drive-model. To help with model identifiability, we tighten the priors on the defective mode:
+\item We allow $t_{pd}$ to vary by drive-model. To help with model identifiably, we tighten the priors on the defective mode:
 $$ \mbox{logit}^{-1}\pi \sim \op{Normal}(-3,1),\quad \sigma_1 \sim \op{Lognormal}(0, 1), \quad t_{p1} \sim \op{Normal}(7,2),$$
@@
 Should maybe be lower.) For $\eta_{\sigma, 2}$, $95\%$ of prior mass is on values greater than $0.037$.
-Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this correponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$.
+Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this corresponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$.
 The prior for $\eta_{t_{p_2},2}$ implies that the median 20th percentile for non-defective units is less than greater than 3 days and less than 24 years.
@@
 The upper left panel shows the estimated failure curve fit to all data ignoring drive-model as a factor (Model 1).
-The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive-models; note that, since the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.
+The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive-models; note that, because the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.
 
-The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which mean that the ranking of drive-models with respect to fraction failing differs over time.
+The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which means that the ranking of drive-models with respect to fraction failing differs over time.
 
@@
 The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure 4.
-The black step function again corresponds to the adjusted Kaplan-Meier estimates.
+The black step functions again corresponds to the adjusted non-parametric estimates.
 As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.
-All the GLFP models appear to overestimate the proportion of failed hard drives for drive-model 2, but Model 4 agrees the most with the nonparametric estimate.
+All of the GLFP models appear to overestimate the proportion of failed hard drives for drive-model 2, but Model 4 agrees best with the nonparametric estimate.
 For drive-model 14, Models 1-3 underestimate the proportion of drives failing, and for drive-model 40, they slightly overestimate the data.
@@
    \includegraphics[width=\textwidth]{fig/mod_compare_legend}
-		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  Black step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
+		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  The solid step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
 \end{figure}
@@
 al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.
-For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with the $i$th data point removed.
+For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.
 The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
 We computed the elpd for all 4 models using the R package loo \cite{loo}.
-When $n$ is large, the distribution of the elpd is approximately normal and different models can be statistically compared.
+When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically.
 We calculated the difference in expected predictive accuracy for Model 4 vs 3, Model 3 vs 2, and Model 2 vs 1, as well as the standard error of the difference (Table 1).
 As the model complexity increased, the expected predictive accuracy improved.
-Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 model specifications.
+Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 models.
 
@@
 We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a central measure of tendency.
-On the left panel of Figure 5 we plot B10, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
-The best six drive-models ranked by TTF are also the drive-models with the longest time to B10.
+On the left panel of Figure 5 we plot $t_{0.10}$, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
+The best six drive-models ranked by TTF are also the drive-models with the longest time to $t_{0.10}$.
 Drive-models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.
@@
 One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each drive-model.
-The plot is again sorted according to B10 so it can be compared to the left hand side.
-While for many of the drive-models the ordinal ranking is the same as on the left, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using B10 or $\pi$.
+The plot is again sorted according to $t_{0.10}$ so it can be compared to the left hand side.
+While for many of the drive-models the ordinal ranking is the same as on the left, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
 Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.
@@
   \includegraphics[width=\textwidth]{fig/b10andpi}
-  \caption{Left: 50\% credible intervals for B10 (time in years till 10\% of the drives fail).
+  \caption{Left: 50\% credible intervals for  $t_{0.10}$ (time in years till 10\% of the drives fail).
 Right: 50\% credible intervals for $\pi$ plotted on the logit scale.
-Both plots are sorted based on the median value of B10.}
+Both plots are sorted based on the median value of $t_{0.10}$.}
   \label{fig4}

### 2017-07-13 paper/paper.tex
Modified abstract

@@
 \begin{abstract}
-Consumers have an interest in accurate assessment of product reliability.
-However, lifetime data is often sparse and has limited information.
-Products with long lifetimes produce few observed failures; and when products do fail the cause of failure is often unknown.
-While computationally convenient, simple parametric models are often limited in modeling this type of data.
-Small sample sizes make parameter estimates unstable or difficult to obtain.
-Multiple failure modes cause bends in the lifetime distribution single parametric models cannot capture.
+% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
+While reliability is an important aspect of product quality, by its nature, the lifetime data available to consumers tends to be limited; high-reliability consumer products fail infrequently.
+Due in part to multiple causes of failure, simple parametric models have limited application, being unable to capture relevant lifetime characteristics.
+On the other hand, more suitable models are difficult to fit because of small sample sizes.
+If it were available, failure analysis mitigate identifiability issues; however, consumers do not have this information, nor is it have practical value for them.
 
 In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}.
-This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  an early failure due to "infant mortality" and a second failure mode due to wearout.
+This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout.
 We fit the GLFP model using a hierarchical Bayesian approach.

### 2017-07-13 paper/paper.tex
typo

@@
 On the other hand, more suitable models are difficult to fit because of small sample sizes.
-If it were available, failure analysis mitigate identifiability issues; however, consumers do not have this information, nor is it have practical value for them.
+If it were available, failure analysis could mitigate identifiability issues; however, consumers do not have this information, nor is it have practical value for them.
 

### 2017-07-13 paper/paper.tex
Add missing $ to table
Streamline intro/motivation section

@@
 
-In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}.
+% In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}.
 This 5-parameter model for lifetime data accommodates lifetime distributions displaying early failures due to "infant mortality" and wearout.
@@
 By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.
-
-We illustrate this method on hard drive reliability data from a cloud-based computing company.
+% 
+% We illustrate this method on hard drive reliability data from a cloud-based computing company.
 The data contains both left-truncation and right-censoring.
 Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous.
-
-\subsection{Motivation}
-
-Failure distributions of high-reliability consumer products can be difficult to assess.
+% 
+% \subsection{Motivation}
+% 
+% Failure distributions of high-reliability consumer products can be difficult to assess.
 Even when field data are available, information may be limited for a variety of reasons.
@@
 Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability.
-While simple nonparametric methods are available, such as the Kaplan-Meier estimator, these are not suitable for forecasting.
-They also do not allow for a straight forward statistical comparisons of different brands in order to select the most reliable product.
+This highlights the importance of choosing a sufficiently flexible model.
 
-For real reliability data, often no failure analysis is available allowing objective discernment about when a failure is due to infant mortality.
+Nonparametric methods, which provide such flexibility are well-known, Kaplan-Meier for example.
+However, truncation, censoring and finite samples limit the scope of inference for these methods.
+Parametric models do not have these limitations.
+Truncated and/or censored data can be accomodated simply by adjusting the likelihood.
+Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength and incorporate prior information.
+
+Often no failure analysis is available allowing objective discernment about when a failure is due to infant mortality.
 This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution.
@@
 
-While the GLFP model provides flexibility and accords with a macro-level theory of product failure, it requires a lot of data to fit because of its complexity.
+The GLFP model has a meaningful parameterization while accomodating lifetime data with infant mortality and wearout but it can require a lot of data to fit because of this complexity.
 Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information.
 When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
-As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations, imposing "soft" constraints on model parameters.
-This enables estimation for all product groups.
-Populations with lots of data will be precisely estimated.
-For products with limited data the posterior distribution of time-to-failure will contain greater uncertainty, while being shrunk toward a "pooled" model.
-Finally, in order to compare the reliability across brands, the Bayesian approach is appealing as samples obtained from the posterior distribution via Markov Chain Monte Carlo allow for easy numerical integration for computing  various quantities of interest.
+As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters.
+This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information.
+We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained by  Markov Chain Monte Carlo.
 \\
@@
   \hline
-\pi & 0.03 & 0.05 & 0.10 \\ 
- \beta_1 & 0.47 & 1.13 & 1.72 \\ 
-  \beta_2 & 4.47 & 4.70 & 4.95 \\ 
-  \mu_1 & 7.46 & 8.06 & 8.76 \\ 
-  \mu_2 & 10.12 & 10.13 & 10.14 \\ 
+$\pi$ & 0.03 & 0.05 & 0.10 \\ 
+ $\beta_1$ & 0.47 & 1.13 & 1.72 \\ 
+  $\beta_2$ & 4.47 & 4.70 & 4.95 \\ 
+  $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
+  $\mu_2$ & 10.12 & 10.13 & 10.14 \\ 
    \hline

### 2017-07-13 paper/paper.tex
small edit

@@
 Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.
-The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.
-A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.
-Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.
+The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \cite{nelson}.
+A common assumption in series systems is the time to failure for each risk is statistically independent.
+Thus, the overall reliability of a unit is modeled using the product rule across all risks.
 Parameter estimation is straightforward if the cause of failure is known for each observation.
-With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the analyst.
-\\
+However, in many situations the cause of failure is unknown to the analyst.
+This is referred to in the literature as "masking".\\
 

### 2017-07-17 paper/paper.tex
Modify and expand the beginning of the Model section through the statement of the likelihood

@@
 
-$$ P(Y \le y) = 1 - \exp \left\{ -\exp \left( \frac{ \log y - \mu }{\sigma} \right) \right\}. $$
+\begin{equation} P(Y \le y) = 1 - \exp \left\{ -\exp \left( \frac{ \log y - \mu }{\sigma} \right) \right\}\\
+= \Phi_{SEV}\left(\frac{\log y - \mu}{\sigma}\right),
+\end{equation}
+where $\Phi_{SEV}$ is the cdf of the smallest extreme value distribution (SEV).
+Thus, the Weibull distribution is a log-location scale family, with $\log Y$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$.
+With this in mind, the corresponding density function is
+\begin{equation}
+p(y) = \frac{1}{\sigma y}\exp\left\{\frac{\log y - \mu}{\sigma} - \exp \left( \frac{\log y - \mu}{\sigma} \right)\right\}
+\end{equation}
 
-This make clear that $(\log Y - \mu)/\sigma$ follows the smallest extreme value distribution (SEV).
-The Weibull distribution is widely used in practice as it can model failure distributions with both increasing and decreasing hazard functions \cite{meeker}.
-The parameter $\beta$ captures this flexibility: if $\beta>1$ the Weibull distribution has an increasing hazard function, $\beta=1$ corresponds to constant hazard, and if $0<\beta<1$ the hazard is decreasing over time.
+An advantage conferred by writing the model in this way is that the log-quantile function is a simple linear function of the parameters, and this allows for more meaningful parameters through which to convey prior information.
+Using the inverse cdf of the SEV distribution,
+\begin{equation}
+\log t_p = \mu + \sigma \Phi^{-1}(p),
+\end{equation}
+where $t_p = \argmin\{y:\Phi(y)>p\}$.
+Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.
+% The Weibull distribution is widely used in practice as it can model failure distributions with both increasing and decreasing hazard functions \cite{meeker}.
+The parameter $\sigma$ controls this characteristic: if $\sigma < 1$ the Weibull distribution has an increasing hazard function, $\sigma=1$ corresponds to constant hazard, and if $\sigma > 1$ the hazard is decreasing.
 
@@
 Following \cite{chan}, let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$ respectively.
-The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $Y_1,\ldots Y_{n} \ind \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
-$$P(Y_i \le y) = 1 - (1-\pi\, F_{1}(y))(1 - F_{2}(y)),\, t>0,\, 0 < \pi < 1.$$
+The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $Y \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
+$$P(Y \le y) = F(y) = 1 - (1-\pi\, F_{1}(y))(1 - F_{2}(y)),\, t>0,\, 0 < \pi < 1.$$
 
-The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $D_i\ind \op Bern(\pi)$.
+The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $D_i\ind \op{Bernoulli}(\pi)$.
 $D_i$ is an indicator for whether or not an observational unit is defective.
@@
 \begin{align*}
-P(Y_i\le y | D_i=1) =& 1 -(1-F_1(y))(1-F_2(y))\\
-P(Y_i\le y | D_i=0) =& F_2(y).
+P(Y\le y | D_i=1) =& 1 -(1-F_1(y))(1-F_2(y))\\
+P(Y\le y | D_i=0) =& F_2(y).
 \end{align*}
 
-For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
+% For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode."
+The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
 Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.
+
+Taking the derivative of the (marginal) cdf, the density for the GLFP is
+\begin{equation}
+f(y|\pi, \mu_1,\sigma_1, \mu_2, \sigma_2) = \pi f_1(y|\mu_1,\sigma_1)\left(1-F_2(y|\mu_2,\sigma_2)\right) + f_2(y|\mu_2,\sigma_2)\left(1-\pi F_2(y|\mu_2,\sigma_2\right).
+\end{equation}
 
@@
 A common feature of lifetime data is right censoring.
-Due to time or budget constraints it is rare all units are observed until failure.
-Therefore, any units still in service when the data is analyzed are considered right censored.
-Products with long expected lifetimes, such as hard drives, are heavily right censored.
-Hard drives are built to last for decades under normal use conditions.
-Even when running continuously, hard drives can exhibit failure rates of less than two percent per year \cite{backblaze}.
-\\
+For various reasons failure times for observational units are known only to up to a lower bound.
+For example, observational units still surviving when the data is analyzed.
+Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.\\
 
-While right censoring imposes an upper bound on an observation, truncation means the existence of an observation is unknown if it falls within a certain interval \cite{meeker}.
-For example, a product may appear in a sample after operating for thousands of hours.
-Ignoring the left truncation and treating the data as identically distributed leads to biased estimates.
-Dropping left truncated observations substantially reduces the sample size.
-Therefore, both the right censoring and left truncation are incorporated into the likelihood function.
+When an observation is truncated, that implies that it would not have been observed had it not coincided with the observation window.
+This has the existence of an observation would have been unknown if it falls within a certain interval \cite{meeker}.
+While right censoring restricts information to a lower bound on survival, left truncation makes information conditional on survival up to the truncation point.
+This is a common feature of observational data, where the factors leading to inclusion in the data set are uncontrolled.
+Ignoring left truncation can lead to biased estimates.
+However, dropping left truncated observations should be avoided because it could substantially reduce the total available information.
+Therefore, we incorporate both right censoring and left truncation into the likelihood.
 
 \subsubsection{Likelihood}
-We now define the general form for the likelihood function that takes into account left truncation, as well as right censoring.
-Let $T_{i}$ denote the lifetime of product $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.
-Let $t_i^L$ define the left truncation time in hours.
-Specifically, for how many hours has a unit been running for before it is added to the sample for observation.
-For units that are right censored, $c_i$ denotes the number of hours a unit has survived without failing.
-We also specify the following indicator functions.
-Let $\nu_i=0$ if a product is left truncated and $\nu_i=1$ otherwise.
-Let $\delta_i=0$ if a product is right censored and $\delta_i=1$ otherwise.
-Given the two parametric forms for $F_1$ and $F_2$ the joint likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
-Assuming all products are independent, the complete likelihood for one product brand is
+We now give the general form for the likelihood function, taking into account left truncation and right censoring.
+Let $Y_{i}$ denote the end of the observed lifetime of unit $i$, in hours.
+Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.
+Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $Y_i$).
+The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
+Assuming the lifetimes of all units are independent, the likelihood for the data, $Y_1,\ldots,Y_N$ is given by
 \begin{equation*}
-L(\bm{\theta})= \prod_{i=1}^{n} f(T_i; \bm{\theta})^{\nu_i\delta_i} \left[\frac{f(T_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{(1-\nu_i) \delta_i}
-[1-F(c_i;\bm{\theta})]^{\nu_i (1-\delta_i)} \left[ \frac{1-F(c_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{(1-\nu_i)(1-\delta_i)}
+L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
 \end{equation*}

### 2017-07-17 paper/paper.tex
Edits to Example 1

--- 2017-07-13 paper/paper.tex
@@
 
+ADD PIECE ABOUT TRANSFORMATIONS
+
+For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.
+If $\text{log}(Y) \sim \text{SEV}(\mu, \sigma)$ then $Y \sim \text{Weibull}(\beta = 1/\sigma, \eta = \exp(\mu))$.
+This decision was chosen after encountering numerical issues when performing estimation using the standard Weibull parametrization.
+In addition, we transformed the parameters to the lower log quantile, $\log(t_{p_j})$ (where $p_{1} = 0.5$, and $p_2 = 0.2$), and log scale, $\log(\sigma_j)$.
+This parameterization was selected as the majority of failures occur at the lower tail of the distribution.
+\\
+
 \subsection{GLFP model}
@@
 \subsection{Censoring and Truncation}
-A common feature of lifetime data is right censoring.
+Right censoring is common in reliability data.
 Due to time or budget constraints it is rare all units are observed until failure.
-Therefore, any units still in service when the data is analyzed are considered right censored.
+If a unit has not yet failed when the data is analyzed it is considered right censored.
+In other words, right censoring puts a lower bound on the failure time.
 Products with long expected lifetimes, such as hard drives, are heavily right censored.
@@
 While right censoring imposes an upper bound on an observation, truncation means the existence of an observation is unknown if it falls within a certain interval \cite{meeker}.
-For example, a product may appear in a sample after operating for thousands of hours.
-Ignoring the left truncation and treating the data as identically distributed leads to biased estimates.
-Dropping left truncated observations substantially reduces the sample size.
-Therefore, both the right censoring and left truncation are incorporated into the likelihood function.
+In the case of left truncation, information about sample units are unknown before a certain time point.
+Therefore, these units follow a conditional lifetime distribution where the conditioning is based on having survived up until the time they enter the sample.
+Ignoring the left truncation and treating the data as identically distributed can lead to biased estimates whereas dropping left truncated observations substantially reduces the sample size.
+Therefore, units exhibiting right censoring and or left truncation should be properly incorporated into the likelihood function.
 
@@
 We now define the general form for the likelihood function that takes into account left truncation, as well as right censoring.
-Let $T_{i}$ denote the lifetime of product $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.
+Let $Y_{i}$ denote the lifetime of product $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.
 Let $t_i^L$ define the left truncation time in hours.
@@
 \begin{equation*}
-L(\bm{\theta})= \prod_{i=1}^{n} f(T_i; \bm{\theta})^{\nu_i\delta_i} \left[\frac{f(T_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{(1-\nu_i) \delta_i}
+L(\bm{\theta})= \prod_{i=1}^{n} f(Y_i; \bm{\theta})^{\nu_i\delta_i} \left[\frac{f(Y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{(1-\nu_i) \delta_i}
 [1-F(c_i;\bm{\theta})]^{\nu_i (1-\delta_i)} \left[ \frac{1-F(c_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{(1-\nu_i)(1-\delta_i)}
@@
 \subsection{Example 1}
-We fit the GLFP model on real hard drive failure data from the cloud-based storage company, Backblaze.
+We fit the GLFP model on real hard drive failure data from a cloud-based storage company.
 Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model \cite{chan}.
-Backblaze provides the number of hours the hard drive is in operation, and whether the drive failed or was right censored.
+The hard drive data contains the number of hours the hard drive was in operation, and whether the drive failed or was right censored.
 No information is available about the cause of failure.
 We first focus on the drive brand model with the most observed failures to demonstrate the GLFP model.
-\\
-
-For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.
-If $\text{log}(Y) \sim \text{SEV}(\mu, \sigma)$ then $Y \sim \text{Weibull}(\beta = 1/\sigma, \eta = \exp(\mu))$.
-This decision was chosen after encountering numerical issues when performing estimation using the standard Weibull parametrization.
-In addition, we transformed the parameters to the lower log quantile, $\log(t_{p_j})$ (where $p_{1} = 0.5$, and $p_2 = 0.2$), and log scale, $\log(\sigma_j)$.
-This parameterization was selected as the majority of failures occur at the lower tail of the distribution.
-\\
-
-Given the instability of maximum likelihood estimates for many of the hard drive brand models, we used a Bayesian approach and specified priors for following parameters where $\Phi^{-1}$ is the quantile function for the SEV distribution.
-
-$$t_{p_j} \equiv \mu_{j} + \sigma_{j}\,\Phi^{-1}(p_j) \quad j=1,2$$
+We fit the GLFP using a Bayesian framework and specified the following proper priors to ensure a proper posterior distribution.
 $$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
@@
 The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
+Given the interpretation of $\pi$ as the proportion of defective units, a range primarily covering the left side of 0.50 seems reasonable and weakly informative.
 The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.
-(Note that, due to symmetry, the Weibull shape parameter has the same prior distribution and that a value of 1, corresponding to a constant hazard, is the median of this prior distribution.) The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$.
+Again, this allows for a wide range of hazards, larger than we would expect to see in the data.
+The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$.
 For $t_{0.2,2}$ it is $(3.4,16.6)$.
 Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.
-They do, however, help identify the two failure modes.
+They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower average time to failure than the wearout mode.
+This is helpful for estimation since, as discussed in Chan & Meeker, ML had difficulty identifying the parameters of the GLFP model without making strong assumptions about the data.
 \\
-
-The model model was fit using the {\tt rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC) \cite{betancourt}.
-Multiple chains were run for 1500 iterations after 1500 burnin iterations.
-Convergence was assessed by examining posterior plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$.
 
@@
 
-To graphically examine the model fit, we plot the GLFP cdf using the median values of the posterior distribution along with 90\% credible interval bands.
-For reference, the solid line corresponds to a truncated adjusted non-parametric  estimator of $F(t)$.
-The posterior median follows the non-parametric estimate quite closely.
-Evidence eof at least two failure modes is seen in the probability plot having a kink occurring between 1 and 1.5 years.
-Moreover, the fitted GLFP curves bends with the non-parametric estimate, accurately following the curvature of the lifetime distribution.
+Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
+After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a visual check for distributional goodness of fit.
+We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \cite{kaplan}.
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see the Appendix for a more detailed description).
+
+In Figure 1 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.
+The solid stepwise curve corresponds to the observed failures.
+Censored drives are not plotted.
+Standard error bands are calculated using Greenwood's method \cite{green}.
+Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.
+The non-parametric and parametric model align up quite closely.
+Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.
+Therefore, a single Weibull model is clearly not flexible enough to model this lifetime distribution.
 
@@
 \centering
-  \includegraphics[width=.9\textwidth]{fig/drive14cred.pdf}
+  \includegraphics[width=.8\textwidth]{fig/drive14cred.pdf}
   \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.
@@
   \label{fig1}
-\end{figure}
-
-Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
-Identifying whether failure data are consistent with a specific family of distributions is difficult to do by examining a histogram or other density estimate.
-By properly transforming the axes of a plot of an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a graphical test for distributional goodness of fit.
-
-Applying this method to the hard drive data, we start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
-\cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
-$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
-$t_{d,\text{min}}^L$, the shortest reported running time of all units
-of drive-model $d$ for which records are available. To produce
-unconditional estimates, we adapt the adjustment method outlined by
-Turnbull, and given in more detail by Meeker and Escobar
-\cite{turnbull,meeker}.  For each drive-model we select
-$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
-By sampling from the full posterior distribution, since
-$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
-survived up to $t_{d,\text{min}}^L$) is a function of the model
-parameters, we can easily compute its posterior median,
-$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by
-
-$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_d(t)}_{KM},\; t>t_{d,\text{min}}^L.$$
-
-While this adjustment is negligible for the majority of drive-models, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
-2.3 years.
-
-In Figure 2 we plot the Kaplan-Meier adjusted cdf for drive-model 14 on Weibull paper.
-Drive-model 14 had the most observed failures of all the brands.
-Each point on the plot corresponds to an observed failure.
-Censored drives are not plotted.
-Standard error bands are calculated using Greenwood's method \cite{green}.
-As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.
-One mode is a result of manufacturing defects, which cause early failures, known as infant mortality.
-The second mode is non-defective hard drives that eventually fail due to wearout.
-Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring between 8000 and 20,000 hours.
-Therefore, fitting a single Weibull model is not flexible enough to model the failure distribution.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=.5\textwidth]{fig/mod14-KM-adj}
-\caption{Truncated Adjusted Kaplan-Meier plot showing nonparametric estimate of fraction failing for drive-model 14.  Dashed lines are Greenwood standard errors.  Plotted on Weibull paper.}
 \end{figure}
@@
 
+\section{Appendix}
+\label{sec:Appendix}
+Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
+Identifying whether failure data are consistent with a specific family of distributions is difficult to do by examining a histogram or other density estimate.
+By properly transforming the axes of a plot of an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a graphical test for distributional goodness of fit.
+
+Applying this method to the hard drive data, we start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
+\cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
+$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
+$t_{d,\text{min}}^L$, the shortest reported running time of all units
+of drive-model $d$ for which records are available. To produce
+unconditional estimates, we adapt the adjustment method outlined by
+Turnbull, and given in more detail by Meeker and Escobar
+\cite{turnbull,meeker}.  For each drive-model we select
+$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
+By sampling from the full posterior distribution, since
+$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
+survived up to $t_{d,\text{min}}^L$) is a function of the model
+parameters, we can easily compute its posterior median,
+$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by
+
+$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_d(t)}_{KM},\; t>t_{d,\text{min}}^L.$$
+
+While this adjustment is negligible for the majority of drive-models, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
+2.3 years.
+
 \pagebreak

### 2017-07-17 paper/paper.tex

@@
 [1-F(c_i;\bm{\theta})]^{\nu_i (1-\delta_i)} \left[ \frac{1-F(c_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{(1-\nu_i)(1-\delta_i)}
+
+A common feature of lifetime data is right censoring.
+For various reasons failure times for observational units are known only to up to a lower bound.
+For example, observational units still surviving when the data is analyzed.
+Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.\\
+
+When an observation is truncated, that implies that it would not have been observed had it not coincided with the observation window.
+This has the existence of an observation would have been unknown if it falls within a certain interval \cite{meeker}.
+While right censoring restricts information to a lower bound on survival, left truncation makes information conditional on survival up to the truncation point.
+This is a common feature of observational data, where the factors leading to inclusion in the data set are uncontrolled.
+Ignoring left truncation can lead to biased estimates.
+However, dropping left truncated observations should be avoided because it could substantially reduce the total available information.
+Therefore, we incorporate both right censoring and left truncation into the likelihood.
+
+\subsubsection{Likelihood}
+We now give the general form for the likelihood function, taking into account left truncation and right censoring.
+Let $Y_{i}$ denote the end of the observed lifetime of unit $i$, in hours.
+Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.
+Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $Y_i$).
+The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
+Assuming the lifetimes of all units are independent, the likelihood for the data, $Y_1,\ldots,Y_N$ is given by
+\begin{equation*}
+L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
 \end{equation*}

### 2017-07-17 paper/paper.tex
Merged my edits with Eric's from earlier in the day.

@@
 
-ADD PIECE ABOUT TRANSFORMATIONS
-
-For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.
-If $\text{log}(Y) \sim \text{SEV}(\mu, \sigma)$ then $Y \sim \text{Weibull}(\beta = 1/\sigma, \eta = \exp(\mu))$.
-This decision was chosen after encountering numerical issues when performing estimation using the standard Weibull parametrization.
-In addition, we transformed the parameters to the lower log quantile, $\log(t_{p_j})$ (where $p_{1} = 0.5$, and $p_2 = 0.2$), and log scale, $\log(\sigma_j)$.
-This parameterization was selected as the majority of failures occur at the lower tail of the distribution.
-\\
-
 \subsection{GLFP model}
@@
 \subsection{Censoring and Truncation}
-Right censoring is common in reliability data.
+A common feature of lifetime data is right censoring.
 Due to time or budget constraints it is rare all units are observed until failure.
@@
 In other words, right censoring puts a lower bound on the failure time.
-Products with long expected lifetimes, such as hard drives, are heavily right censored.
-Hard drives are built to last for decades under normal use conditions.
-Even when running continuously, hard drives can exhibit failure rates of less than two percent per year \cite{backblaze}.
-\\
-
-While right censoring imposes an upper bound on an observation, truncation means the existence of an observation is unknown if it falls within a certain interval \cite{meeker}.
-In the case of left truncation, information about sample units are unknown before a certain time point.
-Therefore, these units follow a conditional lifetime distribution where the conditioning is based on having survived up until the time they enter the sample.
-Ignoring the left truncation and treating the data as identically distributed can lead to biased estimates whereas dropping left truncated observations substantially reduces the sample size.
-Therefore, units exhibiting right censoring and or left truncation should be properly incorporated into the likelihood function.
-
-\subsubsection{Likelihood}
-We now define the general form for the likelihood function that takes into account left truncation, as well as right censoring.
-Let $Y_{i}$ denote the lifetime of product $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.
-Let $t_i^L$ define the left truncation time in hours.
-Specifically, for how many hours has a unit been running for before it is added to the sample for observation.
-For units that are right censored, $c_i$ denotes the number of hours a unit has survived without failing.
-We also specify the following indicator functions.
-Let $\nu_i=0$ if a product is left truncated and $\nu_i=1$ otherwise.
-Let $\delta_i=0$ if a product is right censored and $\delta_i=1$ otherwise.
-Given the two parametric forms for $F_1$ and $F_2$ the joint likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
-Assuming all products are independent, the complete likelihood for one product brand is
-\begin{equation*}
-L(\bm{\theta})= \prod_{i=1}^{n} f(Y_i; \bm{\theta})^{\nu_i\delta_i} \left[\frac{f(Y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{(1-\nu_i) \delta_i}
-[1-F(c_i;\bm{\theta})]^{\nu_i (1-\delta_i)} \left[ \frac{1-F(c_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{(1-\nu_i)(1-\delta_i)}
-
-A common feature of lifetime data is right censoring.
-For various reasons failure times for observational units are known only to up to a lower bound.
-For example, observational units still surviving when the data is analyzed.
-Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.\\
+Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.
 
 When an observation is truncated, that implies that it would not have been observed had it not coincided with the observation window.
-This has the existence of an observation would have been unknown if it falls within a certain interval \cite{meeker}.
 While right censoring restricts information to a lower bound on survival, left truncation makes information conditional on survival up to the truncation point.

### 2017-07-17 paper/paper.tex
Make truncation talk clearer and add subscripts when referring to SEV density, distribution functions

@@
 \end{equation}
-where $\Phi_{SEV}$ is the cdf of the smallest extreme value distribution (SEV).
+where $\Phi_{SEV}$ is the cdf of the standard smallest extreme value distribution (SEV).
 Thus, the Weibull distribution is a log-location scale family, with $\log Y$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$.
@@
 \begin{equation}
-p(y) = \frac{1}{\sigma y}\exp\left\{\frac{\log y - \mu}{\sigma} - \exp \left( \frac{\log y - \mu}{\sigma} \right)\right\}
+p(y) = \frac{1}{\sigma y}\exp\left\{\frac{\log y - \mu}{\sigma} - \exp \left( \frac{\log y - \mu}{\sigma} \right)\right\}\\
+=\frac{1}{\sigma y} \phi_{SEV}\left(\frac{\log y - \mu}{\sigma}\right)
 \end{equation}
 
-An advantage conferred by writing the model in this way is that the log-quantile function is a simple linear function of the parameters, and this allows for more meaningful parameters through which to convey prior information.
+An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which an analyst to apply prior information in a straightforward way.
 Using the inverse cdf of the SEV distribution,
 \begin{equation}
-\log t_p = \mu + \sigma \Phi^{-1}(p),
+\log t_p = \mu + \sigma \Phi_{SEV}^{-1}(p),
 \end{equation}
-where $t_p = \argmin\{y:\Phi(y)>p\}$.
+where $t_p = \argmin\{y:\Phi_{SEV}(y)>p\}$.
 Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.
@@
 
-The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $D_i\ind \op{Bernoulli}(\pi)$.
-$D_i$ is an indicator for whether or not an observational unit is defective.
-Expressed conditional on $D_i$,
+The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$.
+$\delta_i$ is an indicator for whether or not an observational unit is defective.
+Expressed conditional on $\delta_i$,
 \begin{align*}
-P(Y\le y | D_i=1) =& 1 -(1-F_1(y))(1-F_2(y))\\
-P(Y\le y | D_i=0) =& F_2(y).
+P(Y\le y | \delta_i=1) =& 1 -(1-F_1(y))(1-F_2(y))\\
+P(Y\le y | \delta_i=0) =& F_2(y).
 \end{align*}
@@
 \subsection{Censoring and Truncation}
-A common feature of lifetime data is right censoring.
+A common feature of lifetime data is right-censoring.
 Due to time or budget constraints it is rare all units are observed until failure.
-If a unit has not yet failed when the data is analyzed it is considered right censored.
-In other words, right censoring puts a lower bound on the failure time.
+If a unit has not yet failed when the data is analyzed it is considered right-censored.
+In other words, right-censoring puts a lower bound on the failure time.
 Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.
 
-When an observation is truncated, that implies that it would not have been observed had it not coincided with the observation window.
-While right censoring restricts information to a lower bound on survival, left truncation makes information conditional on survival up to the truncation point.
-This is a common feature of observational data, where the factors leading to inclusion in the data set are uncontrolled.
+When an observation is left-truncated, it would not have been observed if it had failed prior to a particular time, which we refer to as the left-truncation time.
+Left-truncation is a common feature of observational lifetime data, where the factors leading to inclusion in the data set are uncontrolled and/or the population of interest has a history prior to any data collection.
 Ignoring left truncation can lead to biased estimates.

### 2017-07-18 paper/paper.tex
Reorganized presentation around hierarchical model
- changed subscripts to more general "g" for group vs. "d" for drive-model

@@
 \end{equation}
-where $t_p = \argmin\{y:\Phi_{SEV}(y)>p\}$.
+where $t_p = \argmin\{y:\Phi_{SEV}(\log y)>p\}$.
 Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.
@@
 
-To borrow strength across different product populations, we can either share parameters across products, or model the product specific parameters hierarchically, allowing the data to inform the hyperparameters.
-For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
+To borrow strength across different populations, we can
+%either share parameters across populations, or 
+model the population specific parameters hierarchically, allowing the data to inform the hyperparameters.
+%For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
 
-$$\sigma_{d,1} \ind \op{Lognormal} \left( \eta_{\sigma,1}, \tau^2_{\sigma,1} \right) \mbox{ for } d=1,\ldots,D$$
-$$\sigma_{d,2} \ind \op{Truncated-Lognormal} \left( \eta_{\sigma,2}, \tau^2_{\sigma,2} , 0, 1 \right) \mbox{ for } d=1,\ldots,D$$
-$$t_{p_j,d,j} \equiv \mu_{d,j} + \sigma_{d,j}\,\Phi^{-1}(p_j)  \ind \op{Normal} \left(\eta_{t_{p_j},j}, \tau^2_{t_{p_j},j}\right) \mbox{ for } j=1,2\; d=1,\ldots,D$$
+Let
+\begin{equation}
+y_{g,i} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
+\end{equation}
+where $g=1,\ldots,G$ indexes populations. Next we model these parameters themselves as random variables. Let
+\begin{equation*}
+\sigma_{1g} \ind \op{LN} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
+\end{equation*}
+\begin{equation*}
+\sigma_{2g} \ind \op{LN} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
+\end{equation*}
+\begin{equation}
+\label{eq:hier-model}
+\log t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{N} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
+\end{equation}
+\begin{equation*}
+\log t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{N} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
+\end{equation*}
+\begin{equation*}
+\op{logit} \pi_g \ind \op{N}(\eta_\pi, \tau_\pi).
+\end{equation*}
 
-$$\op{logit} \pi_d \ind \op{N}(\eta_\pi, \tau_\pi) \mbox{ for } d=1,\ldots,D.$$
+As discussed in section \ref{sec:Weibull parameterization}, we reparameterize the Weibull in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution.
+We truncate the distribution of $\sigma_{2g}$ at $1$, restricting the wearout failure mode to have an increasing hazard function.
+We choose to model these parameters (or suitable transformations) with normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
 
-Here, $\Phi^{-1}$ is the quantile function of the standard log-Weibull distribution.
-We truncate $\sigma_{d,2}$ at $1$, appealing to the logic that, by nature, wearout produces an increasing hazard function.
-The decision to parameterize in terms of a quantile other than the log-location parameter, $\mu = t_{0.632}$, is that lifetime data often features heavy right-censoring where inferences about the location parameter are extrapolations beyond the range of the data.
-%For this data we selected $p_1=0.5,\mbox{ (the median), and } p_2 = 0.2$.
-
-We consider the models with the following set of restrictions (from most to least restrictive):
-
-\begin{enumerate}
-\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{1}=\sigma_1,\quad \mu_{d2} = \mu_2,\quad \sigma_{d2} = \sigma_2$
-\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1,\quad \sigma_{d2} = \sigma_2$
-\item $\pi_{d} = \pi,\quad \mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
-\item $\mu_{d1} = \mu_1,\quad \sigma_{d1}=\sigma_1$
-\end{enumerate}
-
-The set of model specifications were chosen based data, interpretation of the model, as well as estimation considerations.
-Product specific parameters for the the wearout failure mode $(\mu_{d2},\sigma_{d2})$, and the proportion defective $(\pi_d)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
-Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by product.
-
-For all of the models we consider, the parameters for the infant mortality failure mode are held in common across products.
-We found that there was often insufficient information to model these parameters hierarchically.
-Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_i$ and $\pi_j$ ($i \neq j$).
+In principle, we prefer this full model because we tend to believe that every population has a distribution with a distinct set of parameters.
+In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their estimation.
+For example, we may consider a model with a common $\sigma_{1}$, letting $\sigma_{1g}=\sigma_1$ for all $g=1,\ldots,G$.
+A practical strategy is to begin with a simple model and gradually add complexity, reevaluating model adequacy at each iteration.
+In section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.
 
@@
 
+\subsection{Candidate Models}
+We considered four models, all based on \ref{eq:hier-model}, which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):
+
+\begin{enumerate}
+\item $\pi_{g} = \pi,\quad \mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1,\quad \mu_{2g} = \mu_2,\quad \sigma_{2g} = \sigma_2$
+\item $\pi_{g} = \pi,\quad \mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1,\quad \sigma_{2g} = \sigma_2$
+\item $\pi_{g} = \pi,\quad \mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1$
+\item $\mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1$
+\end{enumerate}
+
+The set of model specifications were chosen based data, interpretation of the model, as well as estimation considerations.
+Product specific parameters for the the wearout failure mode $(\mu_{2g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
+Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by product.
+
+For all of the models we consider, the parameters for the infant mortality failure mode are held in common across products.
+We found that there was often insufficient information to model these parameters hierarchically.
+Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).
 \subsection{Prior Distributions}

### 2017-07-18 paper/paper.tex
Small edit to abstract.

--- 2017-07-17 paper/paper.tex
@@
 On the other hand, more suitable models are difficult to fit because of small sample sizes.
-If it were available, failure analysis could mitigate identifiability issues; however, consumers do not have this information, nor is it have practical value for them.
+If it were available, failure analysis could mitigate identifiability issues; however, consumers do not have this information, nor does it have practical value for them.
 

### 2017-07-20 plots/test_lifetime_plots.R
Add functions for visualizing lifetime data

@@ -1,0 +1,19 @@
+library(dplyr)
+library(ggplot2)
+
+#load data
+df <- readRDS("BB_data/clean_unit_summaries.rds")
+overvw <- readRDS("BB_data/overview.rds")
+overvw$stan_id <- NA
+overvw$stan_id[which(overvw$f>=3)] <- 1:44
+
+#model 14
+source("plotting_fns/lifetime_plot.R")
+
+filter(df, model==overvw$model[which(overvw$stan_id==14)]) %>%
+  lifetime_plot(n_to_show=800, lab="Drive-model 14", trans="log",
+                xlimits=c(50,40000), xlabels=c(.01,.05,.25,1,2,4))
+
+filter(df, model==overvw$model[which(overvw$stan_id==14)]) %>%
+  lifetime_plot2(n_to_show=800, lab="Drive-model 14", trans="log",
+                xlimits=c(50,40000), xlabels=c(.01,.05,.25,1,2,4))

### 2017-07-20 plotting_fns/test_KM_plot2.R

@@ -40,7 +40,7 @@ sampfull14 <- with(sampfull, data.frame(mu1 = mu1, sigma1 = sigma1,
                                         log_pi = log_pi[,14]))
 band14full <- bandFromPSamp(sampfull14, xlimits, length.out = 50, N = 1000, logscale = TRUE)
 p14base <- addBandToBaseplot(p14base, band14full, color="yellow", linetype=2, alpha=.2, label="Bayes hier. 90%")
-plotFinally(p14base, xbrks = c(.1, .5, 1, 2, 4)*365*24,
+plotFinally(p14base, xbrks = c(.1, .2, .5, 1, 2, 4)*365*24,
             ybrks = c(.01, .02, .05, .1, .2, .5), years=TRUE) +
   guides(fill = FALSE) +
   theme_bw()

### 2017-07-20 paper/paper.tex
Edits to first 5 sections.

--- 2017-07-18 paper/paper.tex
@@
 % Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
-While reliability is an important aspect of product quality, by its nature, the lifetime data available to consumers tends to be limited; high-reliability consumer products fail infrequently.
+While reliability is an important aspect of product quality, by its nature, the lifetime data available to consumers is limited since high-reliability consumer products fail infrequently.
 Due in part to multiple causes of failure, simple parametric models have limited application, being unable to capture relevant lifetime characteristics.
 On the other hand, more suitable models are difficult to fit because of small sample sizes.
-If it were available, failure analysis could mitigate identifiability issues; however, consumers do not have this information, nor is it have practical value for them.
+If failure analysis data were available it could mitigate identifiability issues; however, consumers do not have this information, nor does it have practical value for them.
 
@@
 Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-Moreover, using a Bayesian approach, comparison of different product brands is straightforward as estimation of arbitrary functionals are easily computable from the full posterior distribution.
+Moreover, using a Bayesian approach, comparison of different product brands is straightforward as estimation of arbitrary functionals are easily computable from full posterior distributions.
 \end{abstract}
@@
 Consumers have an interest in accurate assessment of product reliability.
-Toward this end, there is a need for models that can fit the data together with methods that account for uncertainty.
-While computationally convenient, simple parametric models can fail to capture key characteristics found in real data.
-
-% In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}.
-This 5-parameter model for lifetime data accommodates lifetime distributions displaying early failures due to "infant mortality" and wearout.
-By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.
-% 
-% We illustrate this method on hard drive reliability data from a cloud-based computing company.
-The data contains both left-truncation and right-censoring.
-Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous.
-% 
-% \subsection{Motivation}
-% 
-% Failure distributions of high-reliability consumer products can be difficult to assess.
-Even when field data are available, information may be limited for a variety of reasons.
-Certain brands may have few products in operation resulting in small sample sizes.
-Other products could have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.
-In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.
+Toward this end, there is a need for models that fit the data together with methods that account for uncertainty.
+While computationally convenient, simple lifetime models can fail to capture key characteristics found in real data.
 
 The reliability of many engineered products follow a similar pattern.
-Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process.
+Relatively high rates of failure occurring early ("infant mortality") due to manufacturing defects.
 After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.
@@
 Parametric models do not have these limitations.
-Truncated and/or censored data can be accomodated simply by adjusting the likelihood.
-Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength and incorporate prior information.
+Truncated and/or censored data can be accommodated simply by adjusting the likelihood.
+Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength across groups, and incorporate prior information.
 
-Often no failure analysis is available allowing objective discernment about when a failure is due to infant mortality.
+Often no information is available to identify if a failure is due to infant mortality.
 This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution.
@@
 
-The GLFP model has a meaningful parameterization while accomodating lifetime data with infant mortality and wearout but it can require a lot of data to fit because of this complexity.
-Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information.
+The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.
+Unfortunately, it can require a lot of data to fit due to the model's complexity.
+Bayesian methods offer a means of reducing the amount of data required to estimate the model by leveraging multiple sources of information.
 When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
@@
 This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information.
-We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained by  Markov Chain Monte Carlo.
-\\
-
-% the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.
-
-%\begin{shaded}
-  %THIS SHOULD BE SEPARATED OUT
-  %At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.
-Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are chosen.
-\\
-  %
-  %We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
-With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.
-Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.
-This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.\\
-  %
-  %
-  %Failure data for series systems may be collected at the system level
-  %by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
-  %model (GLFP), for
-  %failure times in a population where some units are susceptible to infant mortality. This model
-  %covers a subset of the class of distributions known as bathtub distributions
-  %\cite{sujata}. Bathtub distributions are characterized by a U-shaped
-  %hazard function, where early failures are due to infant mortality and
-  %late failure are due to wearout. While complex systems actually
-  %exhibit more than two modes of failure, suitable parametric models
-  %which can suitably approximate the overall lifetime distribution are
-  %useful. 
-  %
-  %In the case of complex, high-reliability systems where cause of failure is
-  %not available, a model which
-  %\begin{enumerate}[a.]
-  %\item has parameters that can be estimated from the observed data
-  %\item is flexible enough to fit the observed data
-  %\item in interpretable with respect to available theory of the underlying process
-  %\end{enumerate}
-  %is desired. We present a Bayesian, hierarchical modeling approach for
-  %grouped failure data. As a benefit of working with Monte Carlo
-  %samples, we can easily make inference on a wide range of quantities of
-  %interest while accounting for uncertainty. We demonstrate this approach on real data hard
-  %drive failure data from a
-  %cloud-based storage company.
-%\end{shaded}
-%
-%To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.
-In addition to describing the failure time distribution for products that are susceptible to infant mortality and wearout failure modes, it theorizes an unknown proportion of defective units which are susceptible to defects.
-While the GLFP model provides flexibility over a simple or piece-wise parametric model, it can be difficult to fit with limited data  As Chan and Meeker discuss in their original paper, few observed failures in addition to unknown causes of failure cause maximum likelihood estimates to be unstable or impossible to obtain.
-Therefore, when multiple populations are of interest, comparisons based on the standard GLFP model are restricted to those with sufficient data and observed failure from each failure mode.
-A naive approach to this limitation is to combine the data from all products and fit a single GLFP model.
-This model can be well-estimated, but the pooled GLFP model fits poorly.
-At the other extreme, fitting a separate GLFP model to each product is problematic for products where the data are weakly informative.
-Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are placed on the model parameters.
-\\
-%
-%We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.
-A hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for a whole class of products.
-Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: product brands with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other products to obtain stable estimates.
-This allows us to include all of the data and more importantly, make valid inferences for each individual product.
-%
-%
-%Failure data for series systems may be collected at the system level
-%by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
-%model (GLFP), for
-%failure times in a population where some units are susceptible to infant mortality. This model
-%covers a subset of the class of distributions known as bathtub distributions
-%\cite{sujata}. Bathtub distributions are characterized by a U-shaped
-%hazard function, where early failures are due to infant mortality and
-%late failure are due to wearout. While complex systems actually
-%exhibit more than two modes of failure, suitable parametric models
-%which can suitably approximate the overall lifetime distribution are
-%useful. 
-%
-%In the case of complex, high-reliability systems where cause of failure is
-%not available, a model which
-%\begin{enumerate}[a.]
-%\item has parameters that can be estimated from the observed data
-%\item is flexible enough to fit the observed data
-%\item in interpretable with respect to available theory of the underlying process
-%\end{enumerate}
-%is desired. We present a Bayesian, hierarchical modeling approach for
-%grouped failure data. As a benefit of working with Monte Carlo
-%samples, we can easily make inference on a wide range of quantities of
-%interest while accounting for uncertainty. We demonstrate this approach on real data hard
-%drive failure data from a
-%cloud-based storage company.
+We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.
 
@@
 
-An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which an analyst to apply prior information in a straightforward way.
+An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.
 Using the inverse cdf of the SEV distribution,
@@
 Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.
-% The Weibull distribution is widely used in practice as it can model failure distributions with both increasing and decreasing hazard functions \cite{meeker}.
-The parameter $\sigma$ controls this characteristic: if $\sigma < 1$ the Weibull distribution has an increasing hazard function, $\sigma=1$ corresponds to constant hazard, and if $\sigma > 1$ the hazard is decreasing.
 
@@
 
-% For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode."
 The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
@@
 In other words, right-censoring puts a lower bound on the failure time.
-Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.
+Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit right-censoring.
 
@@
 Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model \cite{chan}.
-The hard drive data contains the number of hours the hard drive was in operation, and whether the drive failed or was right censored.
+The data contains the number of hours the hard drive was in operation, and whether the drive failed or was right censored.
 No information is available about the cause of failure.
@@
 The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
-Given the interpretation of $\pi$ as the proportion of defective units, a range primarily covering the left side of 0.50 seems reasonable and weakly informative.
+Given the interpretation of $\pi$ as the proportion of defective units, which we would expect to be small, a range primarily covering the left side of 0.50 is reasonable and weakly informative.
 The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.
@@
 The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$.
-For $t_{0.2,2}$ it is $(3.4,16.6)$.
+For $t_{0.2,2}$ the range is $(3.4,16.6)$.
 Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.
-They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower average time to failure than the wearout mode.
-This is helpful for estimation since, as discussed in Chan & Meeker, ML had difficulty identifying the parameters of the GLFP model without making strong assumptions about the data.
+They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower mean time to failure than the wearout mode.
+This is helpful for estimation since, as discussed in Chan & Meeker, ML had difficulty identifying the two sets of GLFP parameters without making strong assumptions about the data.
 \\
@@
 Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, an increasing hazard.
-The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than for the wearout mode.
+The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.
 
@@
 Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
-After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a visual check for distributional goodness of fit.
+After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.
 We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \cite{kaplan}.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see the Appendix for a more detailed description).
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a more detailed description).
 
@@
 Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.
-Therefore, a single Weibull model is clearly not flexible enough to model this lifetime distribution.
+Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution.
 
@@
 
-To borrow strength across different populations, we can
-%either share parameters across populations, or 
-model the population specific parameters hierarchically, allowing the data to inform the hyperparameters.
+To borrow strength across different populations, we can model the population specific parameters hierarchically, allowing the data to inform the hyperparameters.
 %For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
@@
 In principle, we prefer this full model because we tend to believe that every population has a distribution with a distinct set of parameters.
-In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their estimation.
+In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their inclusion.
 For example, we may consider a model with a common $\sigma_{1}$, letting $\sigma_{1g}=\sigma_1$ for all $g=1,\ldots,G$.
@@
 Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).
+
 \subsection{Prior Distributions}
@@
 \label{sec:Appendix}
-Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
-Identifying whether failure data are consistent with a specific family of distributions is difficult to do by examining a histogram or other density estimate.
-By properly transforming the axes of a plot of an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a graphical test for distributional goodness of fit.
-
-Applying this method to the hard drive data, we start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
+We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
 \cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by

### 2017-07-21 paper/paper.tex
Play around with some plots, minor edits in intro

@@
 Consumers have an interest in accurate assessment of product reliability.
-Toward this end, there is a need for models that fit the data together with methods that account for uncertainty.
-While computationally convenient, simple lifetime models can fail to capture key characteristics found in real data.
+Toward this end, there is a need for models that reflect observable facts "on the ground", and methods which can make the most of available data and properly account for uncertainty.
+%While computationally convenient, simple lifetime models can fail at each of these.
 
@@
 Unfortunately, it can require a lot of data to fit due to the model's complexity.
-Bayesian methods offer a means of reducing the amount of data required to estimate the model by leveraging multiple sources of information.
+When multiple sources of information are available, partial pooling via hierarchical modeling is a way to reduce the amount of data required to produce stable estimates of the model parameters.
 When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.

### 2017-07-21 plots/indepedent_fits.R

@@ -1,6 +1,6 @@
 setwd("workflow2/")
 #3 8  14
-s <- readRDS("samples_nopool_data_dm14.rds")
+s <- readRDS("samples_nopool_data_dm8.rds")
 print(s)
 library(plyr)
 library(rstan)
@@ -46,3 +46,6 @@ plot_grid(plots[[1]]+bands[[1]]+ggtitle("6"),
           plots[[2]]+bands[[2]]+no_yaxis_nums+ggtitle("14"),
           plots[[3]]+bands[[3]]+no_yaxis_nums+ggtitle("23"), ncol=3)
 dev.off()
+
+plots[[1]] + geom_ribbon(data=bands[[1]]$band, inherit.aes = F,
+                         aes(x=x, y=y, ymin=lower, ymax=upper), alpha=.2)

### 2017-07-21 plots/test_lifetime_plots.R

@@ -1,3 +1,4 @@
+library(plyr)
 library(dplyr)
 library(ggplot2)
 
@@ -10,6 +11,8 @@ overvw$stan_id[which(overvw$f>=3)] <- 1:44
 #model 14
 source("plotting_fns/lifetime_plot.R")
 
+dfnew <- filter(df, model==overvw$model[which(overvw$stan_id==14)])
+
 filter(df, model==overvw$model[which(overvw$stan_id==14)]) %>%
   lifetime_plot(n_to_show=800, lab="Drive-model 14", trans="log",
                 xlimits=c(50,40000), xlabels=c(.01,.05,.25,1,2,4))
@@ -17,3 +20,16 @@ filter(df, model==overvw$model[which(overvw$stan_id==14)]) %>%
 filter(df, model==overvw$model[which(overvw$stan_id==14)]) %>%
   lifetime_plot2(n_to_show=800, lab="Drive-model 14", trans="log",
                 xlimits=c(50,40000), xlabels=c(.01,.05,.25,1,2,4))
+
+plf <- lifetime_plot3(dfnew, lab="Drive-model 14", trans="log",
+                 xlimits=c(50,40000), xlabels=c(.01,.05,.2,1,2,4))
+
+source("plotting_fns/KM_plot2.R")
+fit <- KM.survfit(filter(df, model==overvw$model[which(overvw$stan_id==14)]), greenwood = T)
+p <- baseKMplot(fit, logscale = T, xlimits = c(50, 40000), ylimits=c(.001, .75))
+pkm <- plotFinally(p, xbrks = c(.01, .05, .2, .5, 1, 2, 4)*365*24,
+            ybrks = c(.001, .01, .1, .25, .5, .75),
+            greenwood = T, years = T)
+
+library(cowplot)
+plot_grid(plf,pkm, ncol=1, align = "hv")

### 2017-07-21 plotting_fns/lifetime_plot.R

@@ -112,7 +112,7 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
   
   p <- ggplot(data, aes(x=ID, xend=ID, y=starttime, yend=endtime)) +
     geom_segment(aes(color=status), alpha=.5)+
-    coord_flip() +
+    coord_flip(expand=F) +
     theme_bw() +
     theme(axis.title.y= element_blank(),
           axis.ticks.y= element_blank(),
@@ -133,8 +133,8 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
                                   labels=xlabels,
                                   trans=trans)
   }
-  p <- p + scale_color_manual(values = c("censored"="black", "failed"="lightgray")) 
+  p <- p + scale_color_manual(values = c("censored"="black", "failed"="white")) 
   
-  if(is.null(n_to_show)) return(p+ggtitle(paste(mod)))
+  if(is.null(n_to_show)) return(p+ggtitle(paste(lab)))
   p + ggtitle(paste(lab, "(sample size = ", n_to_show, ", N = ", N, ")"))
 }

### 2017-07-28 paper/paper.tex
Add figure (nonparametric summaries of DM 14) to paper

@@
 \usepackage{url} % not crucial - just used below for the URL
+
+% no longer require relative path to figs
+\graphicspath{{fig/}}
 
@@
 \newcommand{\op}{\operatorname}
-
+\DeclareMathOperator*{\argmin}{arg\,min}
 \newcommand{\myequation}{\begin{equation}}
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics{np_ex1}
+\caption{Top: Lifetime plot of data from 75 randomly selected units of drive-model `14'. Bottom: Kaplan-Meier estimate for drive-model `14'.}
+\label{dm14}
+\end{figure}
 The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
@@
 They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower mean time to failure than the wearout mode.
-This is helpful for estimation since, as discussed in Chan & Meeker, ML had difficulty identifying the two sets of GLFP parameters without making strong assumptions about the data.
+This is helpful for estimation since, as discussed in Chan \& Meeker, ML had difficulty identifying the two sets of GLFP parameters without making strong assumptions about the data.
 \\

### 2017-07-28 plots/test_lifetime_plots.R
Tune the lifetime plot (using v.3)

@@ -21,15 +21,19 @@ filter(df, model==overvw$model[which(overvw$stan_id==14)]) %>%
   lifetime_plot2(n_to_show=800, lab="Drive-model 14", trans="log",
                 xlimits=c(50,40000), xlabels=c(.01,.05,.25,1,2,4))
 
-plf <- lifetime_plot3(dfnew, lab="Drive-model 14", trans="log",
-                 xlimits=c(50,40000), xlabels=c(.01,.05,.2,1,2,4))
+plf <- lifetime_plot3(dfnew, lab="Drive-model 14", trans="log",n_to_show = 75,
+                 xlimits=c(250,40000), xlabels=c(.02,.05,.2,.5,1,2,4))+
+  theme(legend.position="none")
 
 source("plotting_fns/KM_plot2.R")
 fit <- KM.survfit(filter(df, model==overvw$model[which(overvw$stan_id==14)]), greenwood = T)
-p <- baseKMplot(fit, logscale = T, xlimits = c(50, 40000), ylimits=c(.001, .75))
-pkm <- plotFinally(p, xbrks = c(.01, .05, .2, .5, 1, 2, 4)*365*24,
+p <- baseKMplot(fit, logscale = T, xlimits = c(250, 40000), ylimits=c(.001, .75))
+pkm <- plotFinally(p, xbrks = c(.02, .05, .2, .5, 1, 2, 4)*365*24,
             ybrks = c(.001, .01, .1, .25, .5, .75),
-            greenwood = T, years = T)
+            greenwood = T, years = T)+
+  theme(legend.position="none")+ggtitle("Nonparametric fit")
 
 library(cowplot)
+pdf("np_ex1.pdf", width=6, height=7)
 plot_grid(plf,pkm, ncol=1, align = "hv")
+dev.off()

### 2017-07-28 plotting_fns/lifetime_plot.R

@@ -106,17 +106,35 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
   }
   n <- nrow(data)
   data <- data %>% arrange(starttime) %>%
-    mutate(ID=1:n,
-           status=factor(ifelse(censored, "censored","failed")),
-           starttime=pmax(starttime, xlimits[1]))
+    mutate(ID=1:n)
+
+  data_c <- filter(data, censored==TRUE) %>%
+    mutate(status="censored",
+           starttime=endtime,
+           endtime=xlimits[2])
   
-  p <- ggplot(data, aes(x=ID, xend=ID, y=starttime, yend=endtime)) +
-    geom_segment(aes(color=status), alpha=.5)+
+  data_f <- filter(data, censored==FALSE) %>%
+    mutate(starttime=endtime,
+           endtime=xlimits[2],
+           status="failed")
+  
+  data_o <- mutate(data,
+                   starttime=pmax(starttime, xlimits[1]),
+                   status="observed")
+  
+  data_new <- rbind(data_c,data_f,data_o)
+  data_new$status <- factor(data_new$status)
+  
+  p <- ggplot(data_new, aes(x=ID, xend=ID, y=starttime, yend=endtime, color=status, linetype=status, alpha=status)) +
+    geom_segment()+
     coord_flip(expand=F) +
-    theme_bw() +
+    theme_dark() +
     theme(axis.title.y= element_blank(),
           axis.ticks.y= element_blank(),
-          axis.text.y = element_blank())
+          axis.text.y = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          panel.border = element_blank())
   
   transform <- switch(trans,
                       "log" = scales::log_trans,
@@ -133,7 +151,9 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
                                   labels=xlabels,
                                   trans=trans)
   }
-  p <- p + scale_color_manual(values = c("censored"="black", "failed"="white")) 
+  p <- p + scale_color_manual(values = c("censored"="white", "failed"="black", "observed"="white"))+
+    scale_linetype_manual(values = c("censored"="dashed", "failed"="solid", "observed"="solid"))+
+    scale_alpha_manual(values = c("censored"="1", "failed"=".5", "observed"="1"))
   
[…]

### 2017-08-02 paper/paper.tex
Add alternative plot for summarizing dms
Work on model assessment section

@@
 \subsection{Example 1}
+\label{subsec:ex1}
 We fit the GLFP model on real hard drive failure data from a cloud-based storage company.
@@
 Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution.
+The discrepancy in the right tail of the distribution appears to be real.
+We don't provide a solution to this problem here.
+Certainly, if there are clearly more that two well-separated failure modes, the GLFP will not fit the data well.
+However, we consider it satisfactory if we can provide a good approximation to the observed lifetime distribution over the bulk of the product life.
+In practical terms, the right tail is of less importance for electronic goods, due to obsolescence.
 
@@
 \centering
-  \includegraphics[width=.8\textwidth]{fig/drive14cred.pdf}
+  \includegraphics[width=.8\textwidth]{drive14cred.pdf}
   \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.
@@
 
-\section{Data analysis}
+\section{Real data analysis}
 \label{sec:Data analysis}
-\subsection{Data}
+\subsection{Backblaze hard drive data}
 \label{sec:Data}
@@
 \begin{figure}[H]
-  \includegraphics[width=.9\textwidth]{fig/data-sum.pdf}
+  \includegraphics[width=.9\textwidth]{data-sum.pdf}
+  \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
   \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data.
@@
 
-\subsection{Candidate Models}
+\subsection{Modeling}
+\subsubsection{Candidate Models}
 We considered four models, all based on \ref{eq:hier-model}, which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):
@@
 
-\subsection{Prior Distributions}
+\subsubsection{Prior Distributions}
 \label{sec:Prior Distributions}
@@
 
-\subsection{Model Comparisons}
+\subsection{Model Selection}
 \label{sec:Model Comparisons}
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{fig/heterogeneity-compare-preliminary}
+\includegraphics[width=\textwidth]{heterogeneity-compare-preliminary}
 \caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
@@
     \centering
-   \includegraphics[width=\textwidth]{fig/mod_compare_legend}
+   \includegraphics[width=\textwidth]{mod_compare_legend}
 		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  The solid step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
@@
 
-\subsection{Drive-model Comparisons}
-\label{sec:Drive-model Comparisons}
+\subsection{Results}
+\subsubsection{Model Assessment}
+To assess the goodness of fit, we use a graphical posterior predictive check.
+As stated by \citet{gelman1996postpred}, a graphical comparison of replicated data sets from the fitted model with the observed data can be very informative with regard to identification of problems with the model.
+
+To generate replicate data sets, we must first decide what to do with regard to censoring and truncation.
+The representation we choose to display the data, non-parametric probability plots, guides us in making this decision.
+The variability of this estimator arises primarily from the number of ``at risk" units at any given time, which depends in part on these processes.
+Since we are not modeling the mechanisms of truncation nor censoring, we want to try to duplicate these mechanisms in the replicated data.
+Our method for generating replicate a replicate data set $y_{rep,gi}^{(s)}$ is as follows:
+\begin{enumerate}
+\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)}$
+\item If $\tilde{y}_{rep,g,i} < t_{g,i}^L$, then remove this time as this record is not observed
+\item Define $m_{g,i}>0$ for each $g$ and $i$. If $c_{g,i}=0$, let $m_{g,i}=\max \{y_{G,i}: G=g\}$, else let $m_{g,i}=y_{g,i}$.\\
+If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}:=m_{g,i},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
+\end{enumerate}
+
+We can compare the realizations of $y_{rep,g}$ to $y_g$ by using them to reestimate the nonparametric estimator.
+The resulting plots are displayed in figure \ref{post-pred-KM}.
+For most drive-models, the observed estimates appear to behave similarly to those derived from the replicated data.
+As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there.
+Drive-moodel 4 also appears as an exception, which we discuss now.
+
+To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data using a version of the lifetime plot due to WHOM??
+along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate.
+The lifetime plot makes it clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures (4 or 5) that occur in the first quarter year.
+Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band.
+Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than the early failure mode (shared across drive models) would predict.
+
+Similar plots can help to illustrate the 
+and what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_pi, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$. These are shown in figure \ref{}.
+
+\begin{figure}
+\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all}
+\caption{Nonparametric estimates of proportion failing for each drive-model.
+Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
+\label{post-pred-KM}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all}
+\caption{}
+\label{mod4}
+\end{figure}
+
+\subsubsection{Drive-model Comparisons}
+\label{subsubsec:Drive-model Comparisons}
 We consider the problem of ranking drive-models. From a
@@
   \centering
-  \includegraphics[width=\textwidth]{fig/b10andpi}
+  \includegraphics[width=\textwidth]{b10andpi}
   \caption{Left: 50\% credible intervals for  $t_{0.10}$ (time in years till 10\% of the drives fail).
@@
   \centering
-  \includegraphics[width=.8\textwidth]{fig/dm-eval}
+  \includegraphics[width=.8\textwidth]{dm-eval}
   \caption{Left: Posterior predictive relative value-at-loss for new

### 2017-08-02 plotting_fns/KM_plot2.R

@@ -65,7 +65,8 @@ KM.survfit <- function(life.data, greenwood = FALSE, alpha = .05){
 }
 
 #Given a KM truncation adjustment estimate (Ft at min(starttime)), adjust myKM object
-tr.adj <- function(fit){
+tr.adj <- function(fit, tr_adj){
+  stopifnot(tr_adj>=0)
   if(!("myKM" %in% class(fit))) stop("fit must be output from KM.survfit")
   fit$Ft <- fit$Ft * (1 - tr_adj) + tr_adj
   fit
@@ -120,6 +121,11 @@ baseKMplot.multiple <- function(fit_list, xlimits=NULL, ylimits=NULL, color="bla
   if(!is.null(ylimits)){
     fit_list <- lapply(fit_list, function(li) trimKMy(li, ylimits))
   }
+  
+  #check to see if any KMs are now empty
+  empty_id <- which(sapply(fit_list, nrow)==0)
+  if(length(empty_id)>0) {fit_list <- fit_list[ -empty_id]}
+  
   fit_list <- lapply(1:length(fit_list), function(i) cbind(fit_list[[i]], rep=i))
   df.combined <- do.call(rbind, fit_list)
   class(df.combined) <- c("data.frame", "myKM")

### 2017-08-02 workflow/predKM1.R

@@ -71,34 +71,41 @@ get_plotlist <- function(dms){
   replicates <- make_replicate_data(orig = dat_dm, mcmc_samp = sampfull, dm=dm, samp_id = samp_id)
   
   
-  xlimits <- c(100,50000)
+  xlimits <- c(100,80000)
   ylimits <- c(.001, .9)
-  
+  xbrks <- c(.01, 0.1, 0.5, 1, 2, 5) * 24*365
+  ybrks <- c(.001,.01,.1,.25,.5,.75,.9)
   bp <- baseKMplot.multiple(replicates, xlimits=xlimits, ylimits=ylimits,
                             color="black",linetype = "dashed", logscale = TRUE,
                             label="replicates", alpha=.4)
   
   km_obs <- KM.survfit(dat_dm)
   bp <- addKmToBaseplot(bp, km_obs, color="red",linetype = "solid", label = "observed")
-  plist[[i]] <- plotFinally(bp, xbrks = c(100, 1000, 5000, 10000, 20000, 40000),
-              ybrks = c(.001,.01,.1,.25,.5,.75,.9), years=TRUE) +
+  plist[[i]] <- plotFinally(bp, xbrks=xbrks, ybrks=ybrks, years=TRUE) +
     theme_bw(base_size=14)+
-    theme(panel.grid=element_blank(),
-          legend.position = "none",
+    theme(legend.position = "none",
           axis.title.x=element_blank(),
-          axis.title.y=element_blank()) + ggtitle(as.character(dm))
+          axis.title.y=element_blank(),
+          axis.text.x=element_blank(),
+          axis.text.y=element_blank()) + ggtitle(as.character(dm))
   }
   plist
 }
 set.seed(10438)
-ran_set <- sort(sample(44, 6))
-plist <- get_plotlist(ran_set)
-
+# ran_set <- sort(sample(44, 6))
+all <- 1:44
+plist <- get_plotlist(all)
 library(cowplot)
-# length(plist)
-pdf("../paper/fig/pospredictKM.pdf")
-plot_grid(plist[[1]], plist[[2]],plist[[3]],plist[[4]],plist[[5]],plist[[6]], ncol=3)
+pdf("../paper/fig/post-pred-KM-all.pdf", height=14, width=9)
+plot_grid(plotlist = plist, ncol=5)
 dev.off()
+#debug
+# p <- get_plotlist(42)
[…]

### 2017-08-03 paper/paper.tex
Add to and edit paper
- flesh out section on depreciation aware metric
- add to model assessment section
- add pdfs

@@
 \label{sec:Model Comparisons}
-The right side of Figure \ref{fig2} shows that the set of estimated failure curves for all four models that we fit.
+The right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
 The upper left panel shows the estimated failure curve fit to all data ignoring drive-model as a factor (Model 1).
@@
 Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
-\label{fig2}
+\label{fig:fig2}
 \end{figure}
@@
 Using the set of parameters that maximize the posterior log probability, we examine the GLFP cdf for each of the 4 model specifications.
-The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure 4.
+The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure \ref{fig:mod_comp_leg}.
 The black step functions again corresponds to the adjusted non-parametric estimates.
@@
    \includegraphics[width=\textwidth]{mod_compare_legend}
-		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  The solid step function is the truncated adjusted Kaplan-Meier points.  \label{fig:first}} 
+		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  The solid step function is the truncated adjusted Kaplan-Meier points.  \label{fig:mod_comp_leg}} 
 \end{figure}
@@
 We can compare the realizations of $y_{rep,g}$ to $y_g$ by using them to reestimate the nonparametric estimator.
-The resulting plots are displayed in figure \ref{post-pred-KM}.
+The resulting plots are displayed in figure \ref{fig:post-pred-KM}.
 For most drive-models, the observed estimates appear to behave similarly to those derived from the replicated data.
@@
 
-To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data using a version of the lifetime plot due to WHOM??
-along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate.
-The lifetime plot makes it clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures (4 or 5) that occur in the first quarter year.
+To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}.
+From these plots, it is clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the first quarter year.
 Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band.
-Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than the early failure mode (shared across drive models) would predict.
+Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode.
 
-Similar plots can help to illustrate the 
-and what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_pi, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$. These are shown in figure \ref{}.
+We can use a similar plot to see how information is borrowed across drive-models.
+Drive-model 43 is an older model, so that the observed units had run for quite some time before data was made available.
+In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_pi, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
+Then, in the period where records exist for this drive model, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
 
@@
 Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
-\label{post-pred-KM}
+\label{fig:post-pred-KM}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all}
+\centering
+\includegraphics{dm4-exception}
 \caption{}
-\label{mod4}
+\label{fig:ex-mod-4}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{dm43-shrinkage}
+\caption{}
+\label{fig:ex-mod-43}
 \end{figure}
@@
 anticipation of advancements in technology, we can expect that the
-relative value of computer hardware will depreciate rather quickly. To account
-for this depreciation, rather than using TTF as the metric for drive-model comparison, we use the value at replacement. The IRS considers
+relative value of computer hardware will depreciate rather quickly.
+In simple terms, given two hard drives, I would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all.
+To account
+for depreciation, rather than using TTF as the metric for drive-model comparison, we use the value at replacement. The IRS considers
 computer hardware to be ``5-year'' equipment\cite{f4562}. Using``the declining
@@
 
+Posterior credible intervals for $L(t_{d,new})$ are shown in the left panel of figure \ref{fig5}.
+A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{d,new})|y)$ are shown in the bottom right panel.
+The panel above shows credible intervals for TTF.
+We can see that ranking by TTF would lead to a different result -- a new drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher that for the drive-models ranked highest by our depreciation-aware metric.
+The top four drive-models (4,5,2,3) seem very comparable, with a new drive-model 4 unit expected to depreciate the most.
+However, given the discrepancy of the model fit for drive-model 4, noted in the previous subsection, in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this drive-model.
+
 We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a central measure of tendency.
@@
 Drive-models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.
-Another interesting feature to compare is $\pi$, the proportion failing due to infant mortality.
+Another interesting feature to compare is $\pi$, the proportion defective.
 One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each drive-model.
@@
 
-Second, it is probably true that many modes of failure occur and that the heterogeneity that we observe among drive models is due to different combinations of these.
+Second, as discussed in subsection \ref{subsec:mod_assess}, if there are causes which lead to clusters of failures at spaced-out intervals, the GLFP is not flexible enough.
+What we assume is that the failures that occur from various causes can be approximately assigned to two phases of product life, so that the heterogeneity that we observe among drive models is due to different combinations of deficiencies.
 Our ignorance of the causes of failure leads us to approximate the main failure mode with a single drive-model specific Weibull distribution.
 Because of this, we suspect that our forecasts are too confident; especially for ``well-estimated" models, our uncertainty in the right tail of the distribution is probably too small.
-On the other hand, our model borrows strength across drive-models, including those with information about late failures.
+On the other hand, our model borrows strength across drive-models, including those with information about late failures -- information which is entirely missing in some cases.
 In addition, we suggest that using metrics for comparing drive-models that reduce the importance of the right tail of the distribution are appropriate for these data.

### 2017-08-03 plots/dm-summ_textscatter.R

@@ -1,0 +1,34 @@
+#dm summaries
+
+library(ggplot2)
+
+dat <- readRDS("BB_data/clean_unit_summaries.rds")
+ov <- readRDS("BB_data/overview.rds")
+
+id <- which(ov$f>=3)
+ov$stan_id <- NA
+ov$stan_id[id] <- 1:44
+
+dat <- dplyr::filter(dat, model %in% ov$model[id])
+
+# model now matches stan group ids
+dat$model <- ov$stan_id[match(dat$model, ov$model)]
+
+library(plyr)
+my_summary <- ddply(dat, .(model), function(x){
+  tt <- sum(x$endtime - x$starttime)
+  fs <- sum(x$censored == 0)
+  data.frame(tt=tt, fs=fs)
+})
+
+#in yrs
+xbrks=c(10, 100, 1000, 10000)
+
+
+setwd("paper/fig/")
+pdf("dm-summ-scatter.pdf", width=7, height=5)
+ggplot(my_summary, aes(x=tt, y=fs)) + geom_text(aes(label=model), size=5) +
+  scale_x_continuous(name="Total time observed", trans="log", breaks=xbrks*24*365, labels=xbrks)+
+  scale_y_continuous(name="Observed failures", trans="log", breaks = c(1, 10, 100, 1000))+
+  theme_bw(base_size=14)#+geom_smooth(method = "lm", se = F, linetype="dashed")
+dev.off()

### 2017-08-03 plotting_fns/KM_plot2.R

@@ -223,7 +223,7 @@ bandTrimy <- function(bandObj, ylimits){
 bandTrimx <- function(bandObj, xlimits){
   stopifnot("myBand" %in% class(bandObj))
   id <- which(bandObj$band$t > xlimits[2] | bandObj$band$t < xlimits[1])
-  bandObj$band <- bandObj$band[-id,]
+  if(length(id)>0) {bandObj$band <- bandObj$band[-id,]}
   bandObj
 }
 

### 2017-08-03 workflow/mod4exception.R

@@ -1,0 +1,92 @@
+setwd("workflow/")
+dat <- readRDS("../BB_data/clean_unit_summaries.rds")
+dat$model <- as.integer(dat$model)
+
+library(plyr)
+library(dplyr)
+library(rstan)
+library(ggplot2)
+overview <- ddply(dat, .(model), summarise,
+                  n=length(model),
+                  f=sum(failed>0),
+                  early_f = sum(failed>0 & endtime<365*24*1),
+                  late_f = sum(failed>0 & endtime>365*24*2))
+# id <- unique(dat$model)
+id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
+overview$stan_id <- NA
+overview[id,]$stan_id <- 1:length(id)
+
+# s <- readRDS("../workflow/samples_lor_only3fails.rds")
+
+sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
+tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")$median
+
+sampfull <- extract(sfull)
+source("../plotting_fns/KM_plot2.R")
+source("../predictive/sample_from_predictive.R")
+source("../plotting_fns/lifetime_plot.R")
+
+xlabels=c(.01,.05,.2,1,2,5)
+ylabels=c(.001,.01,.1,.5,.8)
+xlimits=c(100, 50000)
+ylimits=c(.001,.8)
+
+dat4 <- filter(dat, model==overview$model[which(overview$stan_id==4)])
+dat4 <- arrange(dat4, endtime)
+dat4_tr <- dat4[which(dat4$starttime>.2*24*365),]
+
+lt_plot <- lifetime_plot3(dat4, n_to_show=200, in_years=TRUE, lab="Drive-model 4", trans="log", xlimits = xlimits, xlabels=xlabels)
+
+km4 <- KM.survfit(dat4)
+km4 <- tr.adj(km4, tr_adj[4])
+km4_tr <- KM.survfit(dat4_tr)
+km4_tr <- tr.adj(km4_tr, tr_adj[4])
+
+bp <- baseKMplot(fit=km4, xlimits=xlimits, ylimits=ylimits, color="black",
+                 linetype = "solid", alpha = 1, logscale = TRUE, label="nonparametric")
+
+samp4 <- with(sampfull, list(mu1=mu1, sigma1=sigma1, log_pi=log_pi[,4], mu2=mu2[,4], sigma2=sigma2[,4]))
[…]

### 2017-08-03 workflow/shrinkage_plot.R

@@ -1,0 +1,74 @@
+setwd("workflow/")
+
+
+  dat <- readRDS("../BB_data/clean_unit_summaries.rds")
+  dat$model <- as.integer(dat$model)
+  
+  library(plyr)
+  library(dplyr)
+  library(rstan)
+  library(ggplot2)
+  overview <- ddply(dat, .(model), summarise,
+                    n=length(model),
+                    f=sum(failed>0),
+                    early_f = sum(failed>0 & endtime<365*24*1),
+                    late_f = sum(failed>0 & endtime>365*24*2))
+  # id <- unique(dat$model)
+  id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
+  overview$stan_id <- NA
+  overview[id,]$stan_id <- 1:length(id)
+  
+  # s <- readRDS("../workflow/samples_lor_only3fails.rds")
+  
+  sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
+  tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")$median
+  
+  sampfull <- extract(sfull)
+  source("../plotting_fns/KM_plot2.R")
+  source("../predictive/sample_from_predictive.R")
+  source("../plotting_fns/lifetime_plot.R")
+  
+  xlabels=c(1,2,5,10, 20)
+  ylabels=c(.001,.01,.1,.5,.9, .99, .999)
+  xlimits=c(5000, 170000)
+  ylimits=c(.001,.999)
+  
+  datdm <- filter(dat, model==overview$model[which(overview$stan_id==dm)])
+  lt_plot <- lifetime_plot3(datdm, xlabels = xlabels, in_years=TRUE,
+                            lab=paste(c("Drive-model ",dm),collapse=""), trans="log", xlimits = xlimits)
+  
+  kmdm <- KM.survfit(datdm, greenwood = FALSE)
+  kmdm <- tr.adj(kmdm, tr_adj[dm])
+  
+  bp <- baseKMplot(fit=kmdm, xlimits=xlimits, ylimits=ylimits, color="black",
+                   linetype = "solid", alpha = 1, logscale = TRUE, label="nonparametric", prob = .9)
+  
+  sampdm <- with(sampfull, list(mu1=mu1, sigma1=sigma1, log_pi=log_pi[,dm], mu2=mu2[,dm], sigma2=sigma2[,dm]))
[…]

### 2017-08-06 paper/paper.tex
Small edits to new section.

@@
 
-Often no information is available to identify if a failure is due to infant mortality.
-This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution.
+Often no information is available to identify why a unit failed.
+This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution.
 However, without any knowledge of cause of failure, traditional competing risk models are not identifiable.
@@
 Unfortunately, it can require a lot of data to fit due to the model's complexity.
-When multiple sources of information are available, partial pooling via hierarchical modeling is a way to reduce the amount of data required to produce stable estimates of the model parameters.
+When multiple sources of information are available, partial pooling via hierarchical modeling is a way to reduce the amount of data required to produce stable parameter estimates.
 When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics{np_ex1}
-\caption{Top: Lifetime plot of data from 75 randomly selected units of drive-model `14'. Bottom: Kaplan-Meier estimate for drive-model `14'.}
-\label{dm14}
-\end{figure}
+%\begin{figure}[H]
+%\centering
+%\includegraphics{np_ex1}
+%\caption{Top: Lifetime plot of data from 75 randomly selected units of drive-model `14'. Bottom: Kaplan-Meier estimate for drive-model `14'.}
+%\label{dm14}
+%\end{figure}
+
 The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
@@
 The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
-Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive-models with at least 3 failures.
+Figure \ref{fig1} shows a scatterplot of the total observed running time in hours versus the total number of failures for drive-models with at least 3 failures.
 For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analyses.
@@
 \begin{figure}[H]
-  \includegraphics[width=.9\textwidth]{data-sum.pdf}
+  %\includegraphics[width=.9\textwidth]{data-sum.pdf}
   \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
-  \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data.
-Right: Total accumulated time in operation (log scale) for each of the drive-models in the Backblaze data.
-The plots are sorted based on the total number of failures on the left.}
+  \caption{Scatterplot of total observed running time (hours) versus total number of observed failures.  Each number labels a unique drive brand model.}
   \label{fig1}
@@
 \subsubsection{Model Assessment}
-To assess the goodness of fit, we use a graphical posterior predictive check.
+To assess the goodness of fit for our final model (Model 4), we use a graphical posterior predictive check.
 As stated by \citet{gelman1996postpred}, a graphical comparison of replicated data sets from the fitted model with the observed data can be very informative with regard to identification of problems with the model.
@@
 Since we are not modeling the mechanisms of truncation nor censoring, we want to try to duplicate these mechanisms in the replicated data.
-Our method for generating replicate a replicate data set $y_{rep,gi}^{(s)}$ is as follows:
+Our method for generating a replicate data set  $y_{rep,g,i}^{(s)}}$ for $s = 1 \dots 19$ is as follows:
 \begin{enumerate}
-\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)}$
+\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.
+Where each set of parameters is randomly drawn from the full posterior distribution.
 \item If $\tilde{y}_{rep,g,i} < t_{g,i}^L$, then remove this time as this record is not observed
 \item Define $m_{g,i}>0$ for each $g$ and $i$. If $c_{g,i}=0$, let $m_{g,i}=\max \{y_{G,i}: G=g\}$, else let $m_{g,i}=y_{g,i}$.\\
-If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}:=m_{g,i},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
+If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=m_{g,i},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
 \end{enumerate}
 
-We can compare the realizations of $y_{rep,g}$ to $y_g$ by using them to reestimate the nonparametric estimator.
+We can compare the 19 simulated data sets $y_{rep,g, i}}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator.
 The resulting plots are displayed in figure \ref{fig:post-pred-KM}.
@@
 As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there.
-Drive-moodel 4 also appears as an exception, which we discuss now.
+Drive-model 4 also appears as an exception, which we discuss now.
 
-To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}.
+\begin{figure}[H]
+\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all}
+\caption{Nonparametric estimates of proportion failing for each drive-model.
+Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
+\label{fig:post-pred-KM}
+\end{figure}
+
+To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}).
 From these plots, it is clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the first quarter year.
@@
 Drive-model 43 is an older model, so that the observed units had run for quite some time before data was made available.
-In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_pi, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
+In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
 Then, in the period where records exist for this drive model, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
-
-\begin{figure}
-\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all}
-\caption{Nonparametric estimates of proportion failing for each drive-model.
-Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
-\label{fig:post-pred-KM}
-\end{figure}
 

### 2017-08-07 paper/paper.tex
Edits to abstract, intro and model assessment
- Fix model 4 plot

@@
 % Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
-While reliability is an important aspect of product quality, by its nature, the lifetime data available to consumers is limited since high-reliability consumer products fail infrequently.
-Due in part to multiple causes of failure, simple parametric models have limited application, being unable to capture relevant lifetime characteristics.
-On the other hand, more suitable models are difficult to fit because of small sample sizes.
-If failure analysis data were available it could mitigate identifiability issues; however, consumers do not have this information, nor does it have practical value for them.
+When the information carried by lifetime data are limited by sample size, censoring and truncation, lifetime analysis can suffer from imprecision.
+These limitations are not atypical; for example, lifetime data for high-reliability products will tend to be right-censored.
+
+With field data on consumer products, model-based appproaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
+The causes of these, while not interesting to the consumer per se, can lead to various observed lifetime distributions.
+Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate.
 
@@
 This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout.
-We fit the GLFP model using a hierarchical Bayesian approach.
+We fit the GLFP model using a hierarchical modeling approach.
 Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-Moreover, using a Bayesian approach, comparison of different product brands is straightforward as estimation of arbitrary functionals are easily computable from full posterior distributions.
+Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using samples from the posterior distribution.
 \end{abstract}
@@
 Consumers have an interest in accurate assessment of product reliability.
-Toward this end, there is a need for models that reflect observable facts "on the ground", and methods which can make the most of available data and properly account for uncertainty.
+Toward this end, there is a need for models that can accomodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
 %While computationally convenient, simple lifetime models can fail at each of these.
@@
 
-Nonparametric methods, which provide such flexibility are well-known, Kaplan-Meier for example.
-However, truncation, censoring and finite samples limit the scope of inference for these methods.
-Parametric models do not have these limitations.
-Truncated and/or censored data can be accommodated simply by adjusting the likelihood.
-Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength across groups, and incorporate prior information.
+Nonparametric methods require relatively few assumptions, but they are not suitable for forecasting.
+Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be practical.
+Appropriate parametric models can make better use of available data.
+For inference, likelihood-based methods are well-equipped to deal with truncated and/or censored data.
+Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across groups, and incorporate prior information.
 
@@
 Unfortunately, it can require a lot of data to fit due to the model's complexity.
-When multiple sources of information are available, partial pooling via hierarchical modeling is a way to reduce the amount of data required to produce stable parameter estimates.
+When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates.
 When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
@@
 
-\subsection{Overview}
 The structure of the paper is as follows.
@@
   rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
-\cite{betancourt}. Multiple chain were run for 1500 iterations after 1500 warmup iterations. Convergence was assessed by examining posterior
-plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$. Four chains were run for Models 1,2 and 3 and 16 chains were run for Model 4.
+\cite{betancourt}.
+HMC jointly updates all model parameters by simulating paths along the posterior density.
+This is done to reduce autocorrelation and efficiently explore the posterior.
+Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1,2 and 3 and 16 chains were run for Model 4.
+The Gelman-Rubin potential scale reduction factors, ``R-hat", which test for lack of convergence by comparing the between versus within chain variability were used.
+Upon convergence, R-hat converges to 1.
+For the four model fits, the  maximum values over all the model parameters were 1.005, 1.028, 1.009 and 1.004, respectively.
+The minimum estimated effective sample size over all the model parameters for the four model fits were 567, 205, 672 and 3330, respectively.
+No divergent transitions occurred after warmup.
+
+% HMC generates proposals by simulating paths along the posterior density.
+This is done to reduce autocorrelation between samples, but doing so also provides a useful diagnostic.
+\textit{Divergent transitions} will tend to occur when the posterior has regions of high curvature, which can present problems for MCMC by making it difficult to fully explore the posterior distribution.
+This feedback can be useful
 
@@
 Since we are not modeling the mechanisms of truncation nor censoring, we want to try to duplicate these mechanisms in the replicated data.
-Our method for generating a replicate data set  $y_{rep,g,i}^{(s)}}$ for $s = 1 \dots 19$ is as follows:
+Our method for generating a replicate data set  $y_{rep,g,i}^{(s)}$ for $s = 1 \dots 19$ is as follows:
 \begin{enumerate}
@@
 
-We can compare the 19 simulated data sets $y_{rep,g, i}}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator.
+We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator.
 The resulting plots are displayed in figure \ref{fig:post-pred-KM}.
@@
 Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode.
+In any case, this example demonstrates that, conditional on the model, the weight of evidence in all the data overpowers the signal from the small sample for this drive-model, which suggests high rates infant mortality in the first quarter year.
 
-We can use a similar plot to see how information is borrowed across drive-models.
-Drive-model 43 is an older model, so that the observed units had run for quite some time before data was made available.
+For a clear example of this, consider drive-model 43.
+The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available.
 In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
-Then, in the period where records exist for this drive model, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
+Then, in the period where records exist for this drive model, at around $1\frac{1}{2}$ years, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
 
@@
 \includegraphics{dm4-exception}
-\caption{}
+\caption{Top: Graphical representation of lifetime for random sample of 200 drives of drive-model 4 from the Backblaze data set.
+Solid white lines indicate the observed window of life.
+Right-censored observations changed to a dashed line and failures mark the transition to a solid black line.
+Bottom: Solid step function is truncation-adjusted nonparametric estimate of the proportion failing.
+The dotted line is an estimate obtained after truncating the data to those starting after .25 years, with the corresponding truncation adjustment.
+The dashed line with gray band is pointwise posterior median, and 90\% credible interval, of proportion failing.}
+
 \label{fig:ex-mod-4}

### 2017-08-07 workflow/mod4exception.R

@@ -20,6 +20,7 @@ overview[id,]$stan_id <- 1:length(id)
 
 sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
 tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")$median
+tr_adj2 <- readRDS("../BB_data/tr_adj_fullmodel2.rds")
 
 sampfull <- extract(sfull)
 source("../plotting_fns/KM_plot2.R")
@@ -33,14 +34,16 @@ ylimits=c(.001,.8)
 
 dat4 <- filter(dat, model==overview$model[which(overview$stan_id==4)])
 dat4 <- arrange(dat4, endtime)
-dat4_tr <- dat4[which(dat4$starttime>.2*24*365),]
+dat4_tr <- dat4[which(dat4$starttime>.25*24*365),]
+
+set.seed(80717)
 
 lt_plot <- lifetime_plot3(dat4, n_to_show=200, in_years=TRUE, lab="Drive-model 4", trans="log", xlimits = xlimits, xlabels=xlabels)
 
 km4 <- KM.survfit(dat4)
 km4 <- tr.adj(km4, tr_adj[4])
 km4_tr <- KM.survfit(dat4_tr)
-km4_tr <- tr.adj(km4_tr, tr_adj[4])
+km4_tr <- tr.adj(km4_tr, tr_adj2$median[tr_adj2$modelid==4])
 
 bp <- baseKMplot(fit=km4, xlimits=xlimits, ylimits=ylimits, color="black",
                  linetype = "solid", alpha = 1, logscale = TRUE, label="nonparametric")

### 2017-08-07 workflow/shrinkage_plot.R

@@ -1,5 +1,5 @@
 setwd("workflow/")
-
+dm=1
 
   dat <- readRDS("../BB_data/clean_unit_summaries.rds")
   dat$model <- as.integer(dat$model)

### 2017-08-07 workflow/tr_adj.R

@@ -34,3 +34,17 @@ tr_adj_df <- ddply(dat2, .(model), function(x){
 
 saveRDS(tr_adj_df, "../BB_data/tr_adj_tp2s2pi.rds")
 
+
+dat3 <- filter(dat2, starttime >= .25*365*24)
+tr_adj_df_quarteryr <- ddply(dat3, .(model), function(x){
+  start = min(x$starttime)
+  modelid = overview$stan_id[which(overview$model==x$model[1])]
+  samp_tr = get_tr_adj(start, exp(as.numeric(samp$log_pi[,modelid])),
+                       as.numeric(samp$mu1),
+                       as.numeric(samp$sigma1),
+                       as.numeric(samp$mu2[,modelid]), 
+                       as.numeric(samp$sigma2[modelid]))
+  adj = quantile(samp_tr, c(.25,.5,.75))
+  data.frame(modelid=modelid, start=start, lower=adj[1], median=adj[2], upper=adj[3])
+})
+saveRDS(tr_adj_df_quarteryr, "../BB_data/tr_adj_fullmodel2.rds")

### 2017-08-09 paper/paper.tex
Fix plot for Example 1.  Started work on new Conclusion.

@@
 
-With field data on consumer products, model-based appproaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
+With field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
 The causes of these, while not interesting to the consumer per se, can lead to various observed lifetime distributions.
@@
 
-%\begin{figure}[H]
-%\centering
-%\includegraphics{np_ex1}
-%\caption{Top: Lifetime plot of data from 75 randomly selected units of drive-model `14'. Bottom: Kaplan-Meier estimate for drive-model `14'.}
-%\label{dm14}
-%\end{figure}
-
 The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
@@
 \centering
-  \includegraphics[width=.8\textwidth]{drive14cred.pdf}
+  \includegraphics[width=.8\textwidth]{mod14_example1.pdf}
   \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.
 Dashed line corresponds to the median of the posterior draws.
-Solid blue line is a non-parametric Kaplan-Meier estimator.
+Solid line is a non-parametric Kaplan-Meier estimator.
 Credible 90\% bands are in red.
-Greenwood non-parametric error bands are in blue.}
+Greenwood non-parametric error bands are in gray.}
   \label{fig1}
@@
 
-\section{Conclusions}
-\label{sec:Conclusions}
+\section{Discussion}
+\label{sec:Discussion}
+This paper has shown how to model and make inference for the kind of product reliability data commonly found in the field.
+Products with long lifetimes often have few failures and when failures do occur the cause is frequently unknown to the analyst.
+The hierarchical GLFP model provides a unified framework to accurately model data with multiple failure modes, and provide a measure of uncertainty.
+Moreover, when there are multiple brands within the same product class, the hierarchical approach allows the model to borrow strength from brands with lots of failures in order to fit and make inference for groups with little information.
+This is extremely advantageous since many products have time (e.g., left truncation and right censoring) and resource constraints (unidentified failure modes) that limit the amount of information in the data.
+This makes standard parametric lifetime models inaccurate and difficult to estimate using classical estimation techniques such as ML.
+
 We should point out some assumptions made in this analysis.

### 2017-08-10 paper/paper.tex
Change bib style
Updated 2nd half of references.
Edits to Conclusion.  Also changed some D's to G's for group.

@@
   \title{\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes}
-  \author{Eric Mittman\thanks{
-    The authors gratefully acknowledge Bill Meeker for his comments and suggestions}\hspace{.2cm}\\
+  \author{Eric Mittman \\
     Department of Statistics, Iowa State University\\
@@
 
-\item We allow $t_{pd}$ to vary by drive-model. To help with model identifiably, we tighten the priors on the defective mode:
+\item We allow $t_{pg}$ to vary by drive-model. To help with model identifiably, we tighten the priors on the defective mode:
 $$ \mbox{logit}^{-1}\pi \sim \op{Normal}(-3,1),\quad \sigma_1 \sim \op{Lognormal}(0, 1), \quad t_{p1} \sim \op{Normal}(7,2),$$
@@
 
-\item $\sigma_{d2}$ is allowed to vary. Priors for constrained parameters are the same as for model 2.
+\item $\sigma_{g2}$ is allowed to vary. Priors for constrained parameters are the same as for model 2.
 
-\item $\pi_d$ varies. Priors for constrained parameters are the same as for model 2.
+\item $\pi_g$ varies. Priors for constrained parameters are the same as for model 2.
 
@@
 
-Second, as discussed in subsection \ref{subsec:mod_assess}, if there are causes which lead to clusters of failures at spaced-out intervals, the GLFP is not flexible enough.
-What we assume is that the failures that occur from various causes can be approximately assigned to two phases of product life, so that the heterogeneity that we observe among drive models is due to different combinations of deficiencies.
-Our ignorance of the causes of failure leads us to approximate the main failure mode with a single drive-model specific Weibull distribution.
-Because of this, we suspect that our forecasts are too confident; especially for ``well-estimated" models, our uncertainty in the right tail of the distribution is probably too small.
+Second, as discussed in subsection \ref{subsec:mod_assess}, if there are causes which lead to clusters of failures at spaced-out intervals, the proposed GLFP is not flexible enough.
+We assume the failures that occur from various causes can be approximately assigned to two phases of product life.
+Heterogeneity observed across drive models is due to different combinations of the two lifetime distributions.
+In practice there might be groups of products or individual brands that exhibit more than two failure modes.
+Extending the GLFP model to incorporate more than two failure distributions or a mixture of different lifetime models is possible.
+However, additional information about the failure modes or more informative priors may be necessary for model identification and estimation.
+
+Another limitation is our ignorance of the causes of failure.
+We approximated the main failure mode with a single drive-model specific Weibull distribution.
+Because of this, we suspect our forecasts are too confident; especially for ``well-estimated" models, our uncertainty in the right tail of the distribution is probably too small.
 On the other hand, our model borrows strength across drive-models, including those with information about late failures -- information which is entirely missing in some cases.
@@
 With respect to decisions based on such metrics, we believe the assumptions in the previous paragraph to be the most critical.
-Those issues are typical to observational data such as these.
 
-Consumers of high-tech manufactured goods usually lack the expertise to perform a failure analysis to determine the cause of failure.
-The ability to fit marginal models, such as GLFP, can allow investigators to answer many pertinent questions while still appropriately accounting for uncertainty and avoiding substantial model bias.
+Consumers of manufactured goods usually lack the data and expertise to perform a failure analysis to determine the cause of failure.
+Moreover, their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.
+The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model bathtub hazard lifetime distributions and make comparisons across different product groups.
+Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics meaningful for the context of the data.
 
-Advances in technology have made computationally intensive Bayesian methods for fitting hierarchical models feasible in practice.
-Hierarchical modeling can be an effective way to regularize a large number of models fits by borrowing information.
-Because all the data contribute some information about the global parameters, it is often sufficient to use weakly informative priors on the global parameters, making the resulting shrinkage data dependent.
-Using samples from the full joint posterior avoids the need for potential problematic asymptotic approximations.
+Advances in technology have made computationally intensive Bayesian methods for fitting hierarchical models available to practitioners.
+When there are many product groups and model parameters efficient MCMC algorithms, such as HMC, can speed up computation due to its joint updating scheme that efficiently explores the parameter space.
+In addition, black box Bayesian software, like R Stan, makes programing HMC straightforward once the full model and likelihood are specified.
+We hopes this makes the hierarchical GLFP model a useful tool for engineers and reliability analysts.
 
-Analysis using HMC is accessible to practitioners by Stan. HMC has been shown to be very efficient in effective samples per iteration because its joint updating scheme.
+\section{Acknowledgments}    
+The authors gratefully acknowledge Bill Meeker for his comments and suggestions.
+
+%Hierarchical modeling can be an effective way to regularize a large number of models fits by borrowing information.
+%Because all the data contribute some information about the global parameters, it is often sufficient to use weakly %informative priors on the global parameters, making the resulting shrinkage data dependent.
+Analysis using HMC is %accessible to practitioners by Stan.
+HMC has been shown to be very efficient in effective samples per iteration because %its joint updating scheme.
 
@@
 \bibliographystyle{plainnat}
+%\bibliographystyle{apalike}
 \bibliography{sample}

### 2017-08-10 paper/paper.tex
Rearrange abstract

--- 2017-08-09 paper/paper.tex
@@
 % Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
-When the information carried by lifetime data are limited by sample size, censoring and truncation, lifetime analysis can suffer from imprecision.
-These limitations are not atypical; for example, lifetime data for high-reliability products will tend to be right-censored.
-
-With field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
+When conducting lifetime analysis with field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
 The causes of these, while not interesting to the consumer per se, can lead to various observed lifetime distributions.
 Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate.
+Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision.
+These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{apalike}
 \bibliography{sample}

### 2017-08-11 paper/paper.tex
Work on references
Clean up appendix

@@
 \cite{betancourt}.
-HMC jointly updates all model parameters by simulating paths along the posterior density.
+HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density.
 This is done to reduce autocorrelation and efficiently explore the posterior.
@@
 HMC has been shown to be very efficient in effective samples per iteration because %its joint updating scheme.
+\clearpage
 
-\section{Appendix}
-\label{sec:Appendix}
+\appendix
+\section{Truncation adjusted nonparametric estimate of lifetime}
+\label{sec:trunc-adj}
 We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
@@
 
-\begin{center}
-{\large\bf SUPPLEMENTARY MATERIAL}
-\end{center}
+% \begin{center}
+% {\large\bf SUPPLEMENTARY MATERIAL}
+% \end{center}
+% 
+% \begin{description}
+% 
+% \item Put R Stan code here
+% 
+% \end{description}
 
-\begin{description}
-
-\item Put R Stan code here
-
-\end{description}
-
-\begin{appendix}
-\scriptsize
+\section{Stan code for model 4}
+\label{sec:stan-code}
+{\scriptsize
 \begin{verbatim}
@@
 \end{verbatim}
-\end{appendix}
+}
 
@@
 
-\bibliographystyle{plainnat}
-%\bibliographystyle{apalike}
+% \bibliographystyle{plainnat}
+\bibliographystyle{apalike}
 \bibliography{sample}

### 2017-08-11 paper/paper.tex
Reworking the explanation of model assessment section

@@
 \subsubsection{Model Assessment}
-To assess the goodness of fit for our final model (Model 4), we use a graphical posterior predictive check.
-As stated by \citet{gelman1996postpred}, a graphical comparison of replicated data sets from the fitted model with the observed data can be very informative with regard to identification of problems with the model.
+To assess the goodness of fit for our final model (Model 4), we use a graphical posterior predictive check, wherein replicate data sets are generated from the posterior predictive distribution and compared with the observed data.
+This is a general approach recommended by \citet{gelman1996postpred} for models where no classical goodness-of-fit test is available.
+Because of the rich information gleaned from plotting the data, this method can be very effective at identifying limitations of the model.
 
-To generate replicate data sets, we must first decide what to do with regard to censoring and truncation.
-The representation we choose to display the data, non-parametric probability plots, guides us in making this decision.
-The variability of this estimator arises primarily from the number of ``at risk" units at any given time, which depends in part on these processes.
-Since we are not modeling the mechanisms of truncation nor censoring, we want to try to duplicate these mechanisms in the replicated data.
-Our method for generating a replicate data set  $y_{rep,g,i}^{(s)}$ for $s = 1 \dots 19$ is as follows:
+Due to censoring, it is not entirely clear how to generate the replicate data sets because we have not modeled the censoring mechanism.
+Since we plan to compare nonparametric lifetime estimates, we chose to replicate as best as possible the pattern of censoring that was observed -- to generate the data conditional on the observed pattern of censoring.
+The problem in doing so is that the censoring mechanism itself is not fully observed.
+Our solution was to assume, for units whose failures were observed (and whose censoring time was not), that had they been right-censored, the censoring time would have been the latest censoring or failure time in the data for that drive model.
+We chose to generate 19 replicates for each drive model.
+This gives a 1/20 chance under the hypothesis of no difference that the most extreme data set would be the observed one and is not too many so as to produce an overly cluttered plot.
+Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each drive-model $g=1,\ldots,44$, is as follows:
 \begin{enumerate}
 \item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.
-Where each set of parameters is randomly drawn from the full posterior distribution.
-\item If $\tilde{y}_{rep,g,i} < t_{g,i}^L$, then remove this time as this record is not observed
-\item Define $m_{g,i}>0$ for each $g$ and $i$. If $c_{g,i}=0$, let $m_{g,i}=\max \{y_{G,i}: G=g\}$, else let $m_{g,i}=y_{g,i}$.\\
-If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=m_{g,i},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
+Where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right) $, is a random draw from the full posterior distribution.
+\item Repeat step 1 until $\tilde{y}_{rep,g,i} > t_{g,i}^L$.
+\item If $c_{g,i}=0$, let $c_{rep,g,i}^(s)=\max \{y_{g,i}: i \ge 1\}$, otherwise let $c_{rep,g,i}^{(s)}=y_{g,i}$.\\
+If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=c_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
 \end{enumerate}
@@
 The resulting plots are displayed in figure \ref{fig:post-pred-KM}.
-For most drive-models, the observed estimates appear to behave similarly to those derived from the replicated data.
+For most drive-models, the estimates from the observed data appear to behave similarly to those from the replicated data.
 As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there.
-Drive-model 4 also appears as an exception, which we discuss now.
+Drive-models 4 and 34 appear to be exceptions as well.
+We will focus our attention on drive-model 4, but our diagnosis of 34 is similar, so the following comments extend to that drive-model as well.
 
@@
 Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band.
-Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode.
+Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode as estimated by the global parameters, $\mu_1$ and $\sigma_1$.
 In any case, this example demonstrates that, conditional on the model, the weight of evidence in all the data overpowers the signal from the small sample for this drive-model, which suggests high rates infant mortality in the first quarter year.
 
-For a clear example of this, consider drive-model 43.
+Generally we consider this behavior a feature of hierarchical modeling.
+For another example of this, consider drive-model 43.
 The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available.

### 2017-08-11 paper/paper.tex
A couple small edits in Conclusion.

@@
 \label{sec:Discussion}
-This paper has shown how to model and make inference for the kind of product reliability data commonly found in the field.
+This paper has shown how to model and make inference for consumer product reliability data commonly found in the field.
 Products with long lifetimes often have few failures and when failures do occur the cause is frequently unknown to the analyst.
 The hierarchical GLFP model provides a unified framework to accurately model data with multiple failure modes, and provide a measure of uncertainty.
-Moreover, when there are multiple brands within the same product class, the hierarchical approach allows the model to borrow strength from brands with lots of failures in order to fit and make inference for groups with little information.
-This is extremely advantageous since many products have time (e.g., left truncation and right censoring) and resource constraints (unidentified failure modes) that limit the amount of information in the data.
+Moreover, when there are multiple brands within the same product class, the hierarchical approach allows the model to borrow strength from brands with lots of failures to make inference for groups with little information.
+This is extremely advantageous since many products have time (left truncation and right censoring) and resource constraints (unidentified failure modes) that limit the information in the data.
 This makes standard parametric lifetime models inaccurate and difficult to estimate using classical estimation techniques such as ML.
@@
 This assumption is due to our ignorance about the causes of failure and of potentially important covariates.
-If we had information about distinguishing characteristics of the two identified failure modes, then we should use this information.
+If we had information about distinguishing characteristics of the two identified failure modes we should use this information.
 Also, there could be many environmental or human-related factors that impact lifetime, such as operating conditions associated with location in the data center, which are likely confounded with drive-model.
 
-Second, as discussed in subsection \ref{subsec:mod_assess}, if there are causes which lead to clusters of failures at spaced-out intervals, the proposed GLFP is not flexible enough.
+Second, as discussed in subsection \ref{subsec: Model Assessment}, if there are causes which lead to clusters of failures at spaced-out intervals, the proposed GLFP is too rigid.
 We assume the failures that occur from various causes can be approximately assigned to two phases of product life.
-Heterogeneity observed across drive models is due to different combinations of the two lifetime distributions.
+Heterogeneity observed across drive models is due to different combinations of these two lifetime distributions.
 In practice there might be groups of products or individual brands that exhibit more than two failure modes.
@@
 Advances in technology have made computationally intensive Bayesian methods for fitting hierarchical models available to practitioners.
-When there are many product groups and model parameters efficient MCMC algorithms, such as HMC, can speed up computation due to its joint updating scheme that efficiently explores the parameter space.
+When there are many product groups (and model parameters) efficient MCMC algorithms, such as HMC, can speed up computation due to its joint updating scheme that efficiently explores the parameter space.
 In addition, black box Bayesian software, like R Stan, makes programing HMC straightforward once the full model and likelihood are specified.

### 2017-08-11 paper/paper.tex
Crop conclusion.

@@
 
-Advances in technology have made computationally intensive Bayesian methods for fitting hierarchical models available to practitioners.
-When there are many product groups (and model parameters) efficient MCMC algorithms, such as HMC, can speed up computation due to its joint updating scheme that efficiently explores the parameter space.
-In addition, black box Bayesian software, like R Stan, makes programing HMC straightforward once the full model and likelihood are specified.
-We hopes this makes the hierarchical GLFP model a useful tool for engineers and reliability analysts.
-
 \section{Acknowledgments}    

### 2017-08-14 paper/paper.tex
Rework discussion section, refine plots

@@
   \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.
-Dashed line corresponds to the median of the posterior draws.
-Solid line is a non-parametric Kaplan-Meier estimator.
-Credible 90\% bands are in red.
-Greenwood non-parametric error bands are in gray.}
+Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.
+Solid line is a non-parametric estimate with pointwise 90\% uncertainty band in gray.}
   \label{fig1}
@@
 \item Repeat step 1 until $\tilde{y}_{rep,g,i} > t_{g,i}^L$.
-\item If $c_{g,i}=0$, let $c_{rep,g,i}^(s)=\max \{y_{g,i}: i \ge 1\}$, otherwise let $c_{rep,g,i}^{(s)}=y_{g,i}$.\\
+\item If $c_{g,i}=0$, let $c_{rep,g,i}^{(s)}=\max \{y_{g,i}: i \ge 1\}$, otherwise let $c_{rep,g,i}^{(s)}=y_{g,i}$.\\
 If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=c_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all}
+\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all2}
 \caption{Nonparametric estimates of proportion failing for each drive-model.
@@
 \label{sec:Discussion}
-This paper has shown how to model and make inference for consumer product reliability data commonly found in the field.
+This paper offers a new approach to modeling and making inference for the lifetime of consumer products using incomplete field reliability data.
 Products with long lifetimes often have few failures and when failures do occur the cause is frequently unknown to the analyst.
-The hierarchical GLFP model provides a unified framework to accurately model data with multiple failure modes, and provide a measure of uncertainty.
-Moreover, when there are multiple brands within the same product class, the hierarchical approach allows the model to borrow strength from brands with lots of failures to make inference for groups with little information.
-This is extremely advantageous since many products have time (left truncation and right censoring) and resource constraints (unidentified failure modes) that limit the information in the data.
-This makes standard parametric lifetime models inaccurate and difficult to estimate using classical estimation techniques such as ML.
+The GLFP model provides a framework to accurately model real data with evidence of multiple failure modes.
+Moreover, when there are multiple product populations within the same product class, our hierarchical modeling approach borrows strength across brands with many observed failures to help make inference for those populations with little information.
+This can be important when trying to leverage multiple sources of data with various sample sizes and levels of censoring and truncation.
+We found that the use of weakly informative priors on the hyperparameters can increase the overall efficiency of the MCMC.
+%This makes standard parametric lifetime models inaccurate and difficult to estimate using classical estimation techniques such as ML.
 
-We should point out some assumptions made in this analysis.
-First, we assume exchangability of units within drive-models.
-This assumption is due to our ignorance about the causes of failure and of potentially important covariates.
-If we had information about distinguishing characteristics of the two identified failure modes we should use this information.
-Also, there could be many environmental or human-related factors that impact lifetime, such as operating conditions associated with location in the data center, which are likely confounded with drive-model.
+We should point out some important caveats of our analyses, given certain model assumptions.
 
-Second, as discussed in subsection \ref{subsec: Model Assessment}, if there are causes which lead to clusters of failures at spaced-out intervals, the proposed GLFP is too rigid.
-We assume the failures that occur from various causes can be approximately assigned to two phases of product life.
-Heterogeneity observed across drive models is due to different combinations of these two lifetime distributions.
-In practice there might be groups of products or individual brands that exhibit more than two failure modes.
-Extending the GLFP model to incorporate more than two failure distributions or a mixture of different lifetime models is possible.
-However, additional information about the failure modes or more informative priors may be necessary for model identification and estimation.
+First, we assume the failures occurring from various causes can be approximately assigned to two phases of product life.
+In practice, there might be additional phases, clearly supported by the data, making the GLFP model too rigid.
+This rigidity is illustrated by the right tail of the estimated distribution in figure \ref{fig1}.
+Extrapolations of the fitted model appear to be very conservative and overly confident.
 
-Another limitation is our ignorance of the causes of failure.
-We approximated the main failure mode with a single drive-model specific Weibull distribution.
-Because of this, we suspect our forecasts are too confident; especially for ``well-estimated" models, our uncertainty in the right tail of the distribution is probably too small.
-On the other hand, our model borrows strength across drive-models, including those with information about late failures -- information which is entirely missing in some cases.
-In addition, we suggest that using metrics for comparing drive-models that reduce the importance of the right tail of the distribution are appropriate for these data.
-With respect to decisions based on such metrics, we believe the assumptions in the previous paragraph to be the most critical.
+Second, we assume exchangability of units within a population.
+This assumption is due in part to ignorance about which units are defective but also of potentially important covariates.
+We do not account for batch effects or varying conditions in the facility which have impacted the observed failure rates.
+These factors are confounded with drive-model in our analysis.
+This second issue is a more fundamental issue with observational data like the hard drive data we have analyzed here.
 
@@
 Moreover, their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.
-The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model bathtub hazard lifetime distributions and make comparisons across different product groups.
-Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics meaningful for the context of the data.
+The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product groups.
+Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics meaningful in the context of the data.
 
-\section{Acknowledgments}    
-The authors gratefully acknowledge Bill Meeker for his comments and suggestions.
+\section*{Acknowledgment}    
+The authors express their gratitude to Dr. \hspace{-.5em} Meeker for his comments and suggestions.
 

### 2017-08-14 plots/dm-summ_textscatter.R

@@ -27,8 +27,8 @@ xbrks=c(10, 100, 1000, 10000)
 
 setwd("paper/fig/")
 pdf("dm-summ-scatter.pdf", width=7, height=5)
-ggplot(my_summary, aes(x=tt, y=fs)) + geom_text(aes(label=model), size=5) +
-  scale_x_continuous(name="Total time observed", trans="log", breaks=xbrks*24*365, labels=xbrks)+
+ggplot(my_summary, aes(x=tt, y=fs)) + geom_text(aes(label=model), size=4) +
+  scale_x_continuous(name="Total time observed (years)", trans="log", breaks=xbrks*24*365, labels=xbrks)+
   scale_y_continuous(name="Observed failures", trans="log", breaks = c(1, 10, 100, 1000))+
   theme_bw(base_size=14)#+geom_smooth(method = "lm", se = F, linetype="dashed")
 dev.off()

### 2017-08-14 plotting_fns/lifetime_plot.R

@@ -96,12 +96,12 @@ lifetime_plot2 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
 }
 
 lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans="identity",
-                           xlabels=c(.01,.05,.25,1,2,5), xlimits=c(100, 50000)){
+                           xlabels=c(.01,.05,.25,1,2,5), xlimits=c(100, 50000), font_size=10){
   if(!in_years){
     xlabels <- signif(xlabels * 24*365, 0)
   } 
   if(!is.null(n_to_show)){
-    N = nrow(data)
+    N = nrow(data) #referenced in plot main title
     data <- sample_n(data, size=n_to_show)
   }
   n <- nrow(data)
@@ -128,20 +128,21 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
   p <- ggplot(data_new, aes(x=ID, xend=ID, y=starttime, yend=endtime, color=status, linetype=status, alpha=status)) +
     geom_segment()+
     coord_flip(expand=F) +
-    theme_dark() +
+    theme_dark(base_size=font_size) +
     theme(axis.title.y= element_blank(),
           axis.ticks.y= element_blank(),
           axis.text.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
-          panel.border = element_blank())
+          panel.border = element_blank()) +
+    scale_x_continuous(limits=c(0, n+1))
   
   transform <- switch(trans,
                       "log" = scales::log_trans,
                       "identity" = scales::identity_trans)
   
   if(in_years){
-    p <-   p + scale_y_continuous(name="time(years)",limits = xlimits,
+    p <-   p + scale_y_continuous(name="time(years)", limits=xlimits,
                                   breaks=xlabels*365*24,
                                   labels=xlabels,
                                   trans=trans)

### 2017-08-14 workflow/mod4exception.R

@@ -11,12 +11,10 @@ overview <- ddply(dat, .(model), summarise,
                   f=sum(failed>0),
                   early_f = sum(failed>0 & endtime<365*24*1),
                   late_f = sum(failed>0 & endtime>365*24*2))
-# id <- unique(dat$model)
+
 id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
 overview$stan_id <- NA
 overview[id,]$stan_id <- 1:length(id)
-
-# s <- readRDS("../workflow/samples_lor_only3fails.rds")
 
 sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
 tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")$median
@@ -31,6 +29,7 @@ xlabels=c(.01,.05,.2,1,2,5)
 ylabels=c(.001,.01,.1,.5,.8)
 xlimits=c(100, 50000)
 ylimits=c(.001,.8)
+font_size=12
 
 dat4 <- filter(dat, model==overview$model[which(overview$stan_id==4)])
 dat4 <- arrange(dat4, endtime)
@@ -38,7 +37,11 @@ dat4_tr <- dat4[which(dat4$starttime>.25*24*365),]
 
 set.seed(80717)
 
-lt_plot <- lifetime_plot3(dat4, n_to_show=200, in_years=TRUE, lab="Drive-model 4", trans="log", xlimits = xlimits, xlabels=xlabels)
+lt_plot <- lifetime_plot3(dat4, n_to_show=200, in_years=TRUE, lab="Drive-model 4",
+                          trans="log", xlimits = xlimits, xlabels=xlabels, font_size = font_size) +
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank())
+  
 
 km4 <- KM.survfit(dat4)
 km4 <- tr.adj(km4, tr_adj[4])
@@ -56,14 +59,12 @@ bpp <- addBandToBaseplot(baseplot=bp, bandObj=band4, color="black",
 bppp <- addKmToBaseplot(baseplot = bpp, fitObj = km4_tr, color="black",
                         linetype="dotted", label="NP (truncated)")
 
-bpppp <- addKmToBaseplot(baseplot = bppp, fitObj = km4p, color="black",
-                         linetype="dotdash", label="nonparametric(excl. 4)")
-
[…]

### 2017-08-23 paper/paper.tex
Small edits.
Edits to abstract and intro

@@
 \usepackage{enumerate}
-\usepackage[numbers]{natbib}
+\usepackage{natbib}
 \usepackage{url} % not crucial - just used below for the URL
@@
 % Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
-When conducting lifetime analysis with field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
-The causes of these, while not interesting to the consumer per se, can lead to various observed lifetime distributions.
+When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
+The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions.
 Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate.
@@
 
-In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}.
-This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout.
+In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model.
+This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to ``infant mortality'' and failure due to wearout.
 We fit the GLFP model using a hierarchical modeling approach.
 Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using samples from the posterior distribution.
+Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using draws from the joint posterior distribution of the model parameters.
 \end{abstract}
@@
 The reliability of many engineered products follow a similar pattern.
-Relatively high rates of failure occurring early ("infant mortality") due to manufacturing defects.
-After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.
+Relatively high rates of failure occurring early (``infant mortality'') due to manufacturing defects.
+After this ``burn-in'' period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.
 Finally, after prolonged use, rates of failure increase due to wearout.
@@
 
-Often no information is available to identify why a unit failed.
+Often no information is available to identify why a unit failed and even when the failed units are available, physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive.
 This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution.
@@
 When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
-As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters.
+As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters.
 This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information.
@@
 The structure of the paper is as follows.
-Section~\ref{sec:Background} describes some of the potential applications of our method and summarizes previous related work found in the literature.
-Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates an application in an example.
-Section~\ref{sec:Hierarchical GLFP model} introduces a hierarchical model for extending the GLFP model to multiple populations.
+In Section~\ref{Motivating example}, we describe, as a motivating example, the Backblaze Hard Drive data set.
+In Section~\ref{sec:Background}, we introduce our proposed method and potential applications and examine previous related work in the literature.
+Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data.
+In section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations.
 In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set.
-Subsection~\ref{sec:Data} describes and summarizes the Backblaze Hard Drive data set.
-Subsection~\ref{sec:Model Comparisons} shows the use of approximate leave-one-out cross-validation as a way to do model selection between different hierarchical model specifications.
-We also address computational issues and modeling decisions that lead to a final model.
-In subsection~\ref{sec:Drive-model Comparisons} we present an analysis comparing drive-model lifetime distributions.
+First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation.
+Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data.
+Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
 Finally, in section~\ref{sec:Conclusions}, we discuss the strengths and limitations of our method and consider extensions and alternatives.
@@
 % \bibliographystyle{plainnat}
-\bibliographystyle{apalike}
+\bibliographystyle{plain}
 \bibliography{sample}

### 2017-08-23 paper/paper.tex
Replace y with t, make corrections based on feedback to Background, GLFP model

@@
 \label{sec:Weibull parameterization}
-For a random variable $Y$ following a Weibull distribution, the standard parameterization of the cumulative distribution function, in terms of scale parameter $\alpha$ and shape parameter $\beta$, is
+For a random variable $T$ following a Weibull distribution, the standard parameterization of the cumulative distribution function, in terms of scale parameter $\alpha$ and shape parameter $\beta$, is
 
-$$ P(Y \le y) = 1 - \exp \left\{ \left( \frac{y}{\alpha} \right)^\beta \right\}. $$
+$$ P(T \le t) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$
 
-An alternative parameterization, in terms of log-location parameter $\mu = \log \alpha$ and log-scale parameter $\sigma = 1/\beta$ is
+An alternative parameterization, in terms of log-location parameter $\mu = \log (\alpha)$ and log-scale parameter $\sigma = 1/\beta$ is
 
-\begin{equation} P(Y \le y) = 1 - \exp \left\{ -\exp \left( \frac{ \log y - \mu }{\sigma} \right) \right\}\\
-= \Phi_{SEV}\left(\frac{\log y - \mu}{\sigma}\right),
+\begin{equation} P(T \le t) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
+= \Phi_{SEV}\left[\frac{\log (t) - \mu}{\sigma}\right],
 \end{equation}
 where $\Phi_{SEV}$ is the cdf of the standard smallest extreme value distribution (SEV).
-Thus, the Weibull distribution is a log-location scale family, with $\log Y$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$.
+Thus, the Weibull distribution is a log-location scale family, with $\log(T)$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$.
 With this in mind, the corresponding density function is
 \begin{equation}
-p(y) = \frac{1}{\sigma y}\exp\left\{\frac{\log y - \mu}{\sigma} - \exp \left( \frac{\log y - \mu}{\sigma} \right)\right\}\\
-=\frac{1}{\sigma y} \phi_{SEV}\left(\frac{\log y - \mu}{\sigma}\right)
+p(t) = \frac{1}{\sigma t}\exp\left\{\frac{\log(t) - \mu}{\sigma} - \exp \left[ \frac{\log(t) - \mu}{\sigma} \right]\right\}\\
+=\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
 \end{equation}
@@
 \begin{equation}
-\log t_p = \mu + \sigma \Phi_{SEV}^{-1}(p),
+\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p).
 \end{equation}
-where $t_p = \argmin\{y:\Phi_{SEV}(\log y)>p\}$.
-Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.
+Because $\Phi(0)\approx 0.632$, we have $\log{\eta} = \mu  =t_{0.63}$.
+When there is more prior knowledge about a lower quantile, as is often the case with high-reliability products, using this parameterization allows the analyst to put prior distributions directly on those quantities.
+In addition, in the presence of censoring, $p$ can be chosen to make $\log(t_p)$ and $\sigma$ approximately independent.
+Since the choice of independent priors is common in practice, this observation lends further credence to the use of these reparameterizations.
 
@@
 \label{subsec:GLFP model}
-Following \cite{chan}, let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$ respectively.
-The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $Y \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
-$$P(Y \le y) = F(y) = 1 - (1-\pi\, F_{1}(y))(1 - F_{2}(y)),\, t>0,\, 0 < \pi < 1.$$
+Let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$, respectively.
+The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
+$$P(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
 
@@
 \begin{align*}
-P(Y\le y | \delta_i=1) =& 1 -(1-F_1(y))(1-F_2(y))\\
-P(Y\le y | \delta_i=0) =& F_2(y).
+P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
+P(T\le t | \delta_i=0) =& F_2(t).
 \end{align*}
 
-The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
+The parameter $\pi$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
 Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.
@@
 \begin{equation}
-f(y|\pi, \mu_1,\sigma_1, \mu_2, \sigma_2) = \pi f_1(y|\mu_1,\sigma_1)\left(1-F_2(y|\mu_2,\sigma_2)\right) + f_2(y|\mu_2,\sigma_2)\left(1-\pi F_2(y|\mu_2,\sigma_2\right).
+f(t|\pi, \mu_1,\sigma_1, \mu_2, \sigma_2) = \pi f_1(t|\mu_1,\sigma_1)\left(1-F_2(t|\mu_2,\sigma_2)\right) + f_2(t|\mu_2,\sigma_2)\left(1-\pi F_2(t|\mu_2,\sigma_2\right).
 \end{equation}
@@
 We now give the general form for the likelihood function, taking into account left truncation and right censoring.
-Let $Y_{i}$ denote the end of the observed lifetime of unit $i$, in hours.
+Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.
 Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.
@@
 We fit the GLFP model on real hard drive failure data from a cloud-based storage company.
-Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model \cite{chan}.
-The data contains the number of hours the hard drive was in operation, and whether the drive failed or was right censored.
+Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model.
+The data contain the number of hours the hard drive was in operation, and whether the drive failed or was right censored.
 No information is available about the cause of failure.
-We first focus on the drive brand model with the most observed failures to demonstrate the GLFP model.
-We fit the GLFP using a Bayesian framework and specified the following proper priors to ensure a proper posterior distribution.
+We first focus on the drive-model with the most observed failures to demonstrate the GLFP model.
+We the GLFP using a Bayesian computational approach, after specifying prior distributions on the model parameters.
+We selected proper prior distributions to improve identification of the model parameters, which ensures a proper posterior distribution.
+We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}).
 $$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$

### 2017-08-23 paper/paper.tex
Modify notation for prior specification

@@
 No information is available about the cause of failure.
-We first focus on the drive-model with the most observed failures to demonstrate the GLFP model.
-We the GLFP using a Bayesian computational approach, after specifying prior distributions on the model parameters.
-We selected proper prior distributions to improve identification of the model parameters, which ensures a proper posterior distribution.
-We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}).
-$$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
-$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad \sigma_2 \sim \op{Lognormal}(0, 2.5) $$
+To illustate the GLFP model, we will present an analysis of the drive-model with the most observed failures.
+We use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution.
+We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}$) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}$).
+Although non-standard, for purposes of exposition, we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.
+
+%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
+$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,7.6\times 10^6 \rangle,\quad
+t_{0.2,2} \sim \op{Lognormal}\langle 8.6, 5.6\times 10^7 \rangle$$
+$$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
+\sigma_1 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle, \quad
+\sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
 
@@
 Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 models.
-
+\begin{figure}
 \begin{table}[H]
@@
 \end{table}
-
+\end{figure}
 \subsection{Results}

### 2017-08-23 paper/paper.tex
Fix notation of priors, add to computation section.

@@
 
-\begin{enumerate}
-\item This model constrains all drive-models to the same failure distribution. For this ``reduced" model we assume the same priors as in Example 1.
+\paragraph{Model 1.} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same priors as in Example 1.
 
-\item We allow $t_{pg}$ to vary by drive-model. To help with model identifiably, we tighten the priors on the defective mode:
-$$ \mbox{logit}^{-1}\pi \sim \op{Normal}(-3,1),\quad \sigma_1 \sim \op{Lognormal}(0, 1), \quad t_{p1} \sim \op{Normal}(7,2),$$
-implying now a prior $90\%$ probability that $\sigma_1$ is between $0.19$ and $5.18$ and $t_{0.5,1}$ between $3.7$ and $10.3$, which, translating from the log-scale, is equivalent to an interval from 1.5 days to 3.4 years.
+\paragraph{Model 2.} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
+$$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
+$$\sigma_1 \sim \op{Log-normal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
+$$t_{p1} \sim \op{Log-normal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
 
-\item $\sigma_{g2}$ is allowed to vary. Priors for constrained parameters are the same as for model 2.
+\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for model 2.
 
-\item $\pi_g$ varies. Priors for constrained parameters are the same as for model 2.
+\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for model 2.
 
-\end{enumerate}
-
-For the scale hyper-parameters we follow the recommendations of \citet{gelman2014bayesian} and use half-Cauchy priors.
-As for the location hyper-parameters, we select weakly informative priors consistent with our prior information on hard-drives.
-The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification?
+Where applicable prior distributions on hyperparameters are as listed below.
+We follow the recommendation of \citet{gelman2014bayesian} to use half-Cauchy priors on hierarchical scale parameters.
+As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
+% The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification?
 Should maybe be lower.) For $\eta_{\sigma, 2}$, $95\%$ of prior mass is on values greater than $0.037$.
@@
   \eta_{\pi} & \sim \op{Normal}(-3, 1)\\
-  \tau_{\pi} & \sim \op{Cauchy}^+(0, 1)\\
+  \tau_{\pi} & \sim \op{Half-Cauchy}(0, 1)\\
   \eta_{\sigma ,2} & \sim \op{Normal}(0, 2)\\
-  \tau_{\sigma ,2} & \sim \op{Cauchy}^+(0, 1)\\
+  \tau_{\sigma ,2} & \sim \op{Half-Cauchy}(0, 1)\\
   \eta_{t_{.2_2},2} & \sim \op{Normal}(9, 2)\\
-  \tau_{t_{.2_2},2} & \sim \op{Cauchy}^+(0, 1)
+  \tau_{t_{.2_2},2} & \sim \op{Half-Cauchy}(0, 1).
  \end{align*}
@@
 Upon convergence, R-hat converges to 1.
-For the four model fits, the  maximum values over all the model parameters were 1.005, 1.028, 1.009 and 1.004, respectively.
-The minimum estimated effective sample size over all the model parameters for the four model fits were 567, 205, 672 and 3330, respectively.
+For the four model fits, the  maximum values over all 140 model parameters were 1.005, 1.028, 1.009 and 1.004, respectively.
+The minimum estimated effective sample size over the model parameters were 567, 205, 672 and 3330 for Models 1 through 4, respectively.
 No divergent transitions occurred after warmup.

### 2017-08-23 util/getImpliedInterval.R

@@ -1,0 +1,5 @@
+getImpliedInterval <- function(distr, a=0, b=1, prob=.95, alpha=1-prob){
+  callingFun <- paste("q",distr, sep="")
+  qtls <- c(alpha/2, 1-alpha/2)
+  eval(call(callingFun, qtls, a, b))
+}

### 2017-08-24 paper/paper.tex
Small edits to computation paragraph. Remove reference to effective sample sizes.

@@
 Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1,2 and 3 and 16 chains were run for Model 4.
-The Gelman-Rubin potential scale reduction factors, ``R-hat", which test for lack of convergence by comparing the between versus within chain variability were used.
-Upon convergence, R-hat converges to 1.
-For the four model fits, the  maximum values over all 140 model parameters were 1.005, 1.028, 1.009 and 1.004, respectively.
-The minimum estimated effective sample size over the model parameters were 567, 205, 672 and 3330 for Models 1 through 4, respectively.
-No divergent transitions occurred after warmup.
+The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains.
+Upon convergence, $\hat{R}$, converges to 1.
+The number of model parameters for the four model fits are .
+The  maximum values of $\hat{R}$ across all model parameters for Models 1 throught 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004.
+Other diagnostics provided by the software (tree-depth, and divergent transitions) were clean.
+Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.
 

### 2017-08-24 plots/km_overplot.R
Change scales on plots Add sample of posterior predictive KM plots, as well as larger versions for all drive-models for supplementary material.

@@ -33,101 +33,160 @@ df1 <- dat %>% filter(model %in% id) %>%
   select(-start_date, -end_date) %>%
   mutate(model = overview$stan_id[match(model, overview$model)])
 
-source("../plotting_fns/KM_plot.R")
+source("../plotting_fns/KM_plot2.R")
 
 xlimits <- c(5, 50000)
-ylimits <- c(.00001, .9)
+ylimits <- c(0, .9)
+x.breaks <- 0:5*1e4
+y.breaks <- 0:4*.25
 
-x.breaks <- xbrks(xlimits[1], xlimits[2], N=10, prec=-2)
-y.breaks <- ybrks(ylimits[1], ylimits[2], model="weibull")
+all_kms <- ldply(unique(df1$model), function(m){
+  fit <- KM.survfit(filter(df1, model==m))
+  fit <- tr.adj(fit, tr_adj = tr_adj[m])
+  fit <- trimKMx(fit,xlimits)
+  fit <- trimKMy(fit,ylimits)
+  data.frame(model=m,fit)})
 
-x.scale <- scale_x_continuous(name="Hours", breaks=x.breaks, limits=xlimits)
-y.scale <- scale_y_continuous(name="Fraction failing", breaks=y.breaks, limits=ylimits)
-
-no_legend <- theme(legend.position="none")
-x_log <- scale_x_continuous(trans="log")
-
-p1 <- KM_plot_multi(df1, tr_adj = tr_adj, linear_axes = F,
-                    xlimits = xlimits, ylimits=ylimits, grayscale = TRUE, size=1)
+p <- ggplot(all_kms, aes(x=t, y=Ft, color=factor(model), group=factor(model))) + geom_step(size=1) +
+  theme_bw(base_size = 14) + scale_colour_grey(start = 0, end = .9) +
+  scale_x_continuous(name="Thousands of hours", breaks = x.breaks, labels=x.breaks/1000,
+                     limits = xlimits) +
+  scale_y_continuous(name="Fraction failing", breaks=y.breaks, limits=ylimits)+
+  theme(legend.position = "none") + 
+  ggtitle("Adjusted Kaplan-Meier")
 
 null_s <- readRDS("../workflow_null/samples_null_model_3_29.rds")
-null_samp <- extract(null_s)
[…]

### 2017-08-24 plotting_fns/KM_plot2.R

@@ -33,7 +33,6 @@ my_pweibull <- function(x, location, scale, ...){
   pweibull(x, 1/scale, exp(location), ...)
 }
 
-
 ## Compute Kaplan-Meier estimate for possibly left-trunctated,
 # right-censored univariate lifetime data (s3 class "myKM")
 KM.survfit <- function(life.data, greenwood = FALSE, alpha = .05){
@@ -294,13 +293,13 @@ plotFinally <- function(plotList, xbrks, ybrks, years=FALSE, greenwood=FALSE, al
     p <- p + scale_x_continuous(name="Time (Years)", trans="log",
                                             breaks = xbrks, labels=signif(xbrks/365/24, 0), limits = plotList$xlimits)
   } else if(plotList$logscale & !years) {
-    p <- p + scale_x_continuous(name="Time (Hours)", trans="log",
+    p <- p + scale_x_continuous(name="Thousands of hours", trans="log", labels = xbrks/1000,
                                             breaks = xbrks, limits = plotList$xlimits)
   } else if(years){
     p <- p + scale_x_continuous(name="Time (Years)", breaks = xbrks,
                                             labels=signif(xbrks/365/24, 0), limits = plotList$xlimits)
   } else{
-    p <- p + scale_x_continuous(name="Time (Hours)", breaks = xbrks
+    p <- p + scale_x_continuous(name="Thousands of hours", breaks = xbrks, labels=xbrks/1000
                                             , limits = plotList$xlimits)
   }
   

### 2017-08-24 plotting_fns/lifetime_plot.R

@@ -113,22 +113,22 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
            starttime=endtime,
            endtime=xlimits[2])
   
-  data_f <- filter(data, censored==FALSE) %>%
-    mutate(starttime=endtime,
-           endtime=xlimits[2],
-           status="failed")
+  # data_f <- filter(data, censored==FALSE) %>%
+  #   mutate(starttime=endtime,
+  #          endtime=xlimits[2],
+  #          status="failed")
   
   data_o <- mutate(data,
                    starttime=pmax(starttime, xlimits[1]),
                    status="observed")
   
-  data_new <- rbind(data_c,data_f,data_o)
+  data_new <- rbind(data_c,data_o)
   data_new$status <- factor(data_new$status)
   
   p <- ggplot(data_new, aes(x=ID, xend=ID, y=starttime, yend=endtime, color=status, linetype=status, alpha=status)) +
     geom_segment()+
     coord_flip(expand=F) +
-    theme_dark(base_size=font_size) +
+    theme_bw(base_size=font_size) +
     theme(axis.title.y= element_blank(),
           axis.ticks.y= element_blank(),
           axis.text.y = element_blank(),
@@ -147,14 +147,14 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
                                   labels=xlabels,
                                   trans=trans)
   } else{
-    p <-   p + scale_y_continuous(name="time(years)",limits = xlimits,
-                                  breaks=xlabels,
+    p <-   p + scale_y_continuous(name="Thousands of hours",limits = xlimits,
+                                  breaks=xlabels*1000,
                                   labels=xlabels,
                                   trans=trans)
   }
-  p <- p + scale_color_manual(values = c("censored"="white", "failed"="black", "observed"="white"))+
-    scale_linetype_manual(values = c("censored"="dashed", "failed"="solid", "observed"="solid"))+
-    scale_alpha_manual(values = c("censored"="1", "failed"=".5", "observed"="1"))
+  p <- p + scale_color_manual(values = c("censored"="black", "observed"="black"))+
+    scale_linetype_manual(values = c("censored"="dashed", "observed"="solid"))+
+    scale_alpha_manual(values = c("censored"="1", "observed"="1"))
   
[…]

### 2017-08-24 workflow/mod4exception.R

@@ -25,20 +25,21 @@ source("../plotting_fns/KM_plot2.R")
 source("../predictive/sample_from_predictive.R")
 source("../plotting_fns/lifetime_plot.R")
 
-xlabels=c(.01,.05,.2,1,2,5)
-ylabels=c(.001,.01,.1,.5,.8)
+xlabels=c(.1,.5,2.5,10,50)
+ylabels=c(.001,.005, .01,.05,.1)
 xlimits=c(100, 50000)
-ylimits=c(.001,.8)
+ylimits=c(.001,.2)
 font_size=12
 
 dat4 <- filter(dat, model==overview$model[which(overview$stan_id==4)])
 dat4 <- arrange(dat4, endtime)
-dat4_tr <- dat4[which(dat4$starttime>.25*24*365),]
+dat4_tr <- dat4[which(dat4$starttime>2500),]
 
 set.seed(80717)
 
-lt_plot <- lifetime_plot3(dat4, n_to_show=200, in_years=TRUE, lab="Drive-model 4",
-                          trans="log", xlimits = xlimits, xlabels=xlabels, font_size = font_size) +
+lt_plot <- lifetime_plot3(data = dat4, n_to_show=60, in_years=FALSE, lab="Drive-model 4",
+                          trans="log", xlimits = xlimits, xlabels=xlabels,
+                          font_size = font_size) +
   theme(axis.text.x = element_blank(),
         axis.title.x = element_blank())
   
@@ -46,10 +47,19 @@ lt_plot <- lifetime_plot3(dat4, n_to_show=200, in_years=TRUE, lab="Drive-model 4
 km4 <- KM.survfit(dat4)
 km4 <- tr.adj(km4, tr_adj[4])
 km4_tr <- KM.survfit(dat4_tr)
-km4_tr <- tr.adj(km4_tr, tr_adj2$median[tr_adj2$modelid==4])
+#calculate adjustment: estimate of proportion failing at first starttime
+first_start <- min(dat4_tr$starttime)
+draws <- with(sampfull, sapply(1:24000, function(iter){
+  1 - (1 - exp(log_pi[iter,4]) * 
+         my_pweibull(first_start, mu1[iter], sigma1[iter])) * 
+    (1 - my_pweibull(first_start, mu2[iter,4], sigma2[iter,4]))}))
+
+summary(draws)
+
+km4_tr <- tr.adj(km4_tr, median(draws))
 
 bp <- baseKMplot(fit=km4, xlimits=xlimits, ylimits=ylimits, color="black",
[…]

### 2017-08-24 workflow/predKM1.R

@@ -25,6 +25,7 @@ sampfull <- extract(sfull)
 source("../plotting_fns/KM_plot2.R")
 source("../predictive/sample_from_predictive.R")
 
+set.seed(82417)
 
 make_replicate_data <- function(orig, mcmc_samp, dm, samp_id){
   n <- nrow(orig)
@@ -58,47 +59,51 @@ make_replicate_data <- function(orig, mcmc_samp, dm, samp_id){
   replicates
 }
 
-get_plotlist <- function(dms, firstcol, lastrow){
+reps <- 19
+
+y_reps <- lapply(1:44, function(dm){
+  dat_dm <- filter(dat, model==overview$model[which(overview$stan_id==dm)])
+  samp_id <- sample(length(sampfull$mu1), reps)
+  replicates <- make_replicate_data(orig = dat_dm, mcmc_samp = sampfull, dm=dm, samp_id = samp_id)
+  replicates
+})
+
+saveRDS(y_reps,"../plots/y_rep.rds")
+
+get_plotlist <- function(y_reps, dat){
   plist <- list()
   xaxis <- NULL
   yaxis <- NULL
   legend <- NULL
-  n <- length(dms)
+  n <- length(y_reps)
   for(i in 1:n){
-    dm <- dms[i]
-    dat_dm <- filter(dat, model==overview$model[which(overview$stan_id==dm)])
-  
-  
-  
-    reps <- 19
-    samp_id <- sample(length(sampfull$mu1), reps)
-    replicates <- make_replicate_data(orig = dat_dm, mcmc_samp = sampfull, dm=dm, samp_id = samp_id)
-  
-  
-    xlimits <- c(100,80000)
+    xlimits <- c(100,100000)
     ylimits <- c(.001, .9)
-    xbrks <- c(.01, 0.1, 0.5, 1, 2, 5) * 24*365
+    xbrks <- 0:10 * 10000
     ybrks <- c(.001,.01,.1,.25,.5,.75,.9)
-    bp <- baseKMplot.multiple(replicates, xlimits=xlimits, ylimits=ylimits,
-                              color="black",linetype = "dashed", logscale = TRUE,
-                              label="replicates", alpha=.4)
-  
+    bp <- baseKMplot.multiple(y_reps[[i]], xlimits=xlimits, ylimits=ylimits,
+                              color="black",linetype = "dashed", logscale = FALSE,
+                              label=expression(KM(y[rep])), alpha=.4)
[…]

### 2017-08-25 paper/paper.tex
Change scales on plots Add sample of posterior predictive KM plots, as well as larger versions for all drive-models for supplementary material.
Editing section 5

@@
 \label{sec:Model Comparisons}
-The right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
-The upper left panel shows the estimated failure curve fit to all data ignoring drive-model as a factor (Model 1).
-The upper left plot, which represents Model 2, demonstrates that there is appreciable differentiation between drive-models; note that, because the scale parameter, $\tau_{t_{p2}}$ is learned hierarchically from the data, we can ascertain that there is evidence in the data that this variability is real.
+As we have discussed, while we prefer to allow all of the GLFP model parameters to be drive-model specific, there are practical issues with fitting the full model.
+For example, due to heavy left-truncation, there may be insufficient information for particular drive models to estimate all of the parameters.
+Furthermore, we prefer not to introduce unnecessary complexity into our analyses.
+Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of time to failure.
+The large left panel shows the Kaplan-Meier nonparametric estimate of the same.
+the Kaplan-Meier estimates right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
+The smaller plots correspond to Models 1 through 4 reading from left to right, top to bottom.
+Note that while all the paths for the posterior estimates necessarily pass through the origin, the nonparametric curves do not, but only because, for the Kaplan-Meier estimator, left-truncation disallows any kind of interpolation.
 
-The lower left panel, representing Model 3, suggests differing shape parameters in the wearout failure mode which means that the ranking of drive-models with respect to fraction failing differs over time.
-
-Finally, the lower right panel, corresponding to Model 4, shows that, by allowing $\pi_d$ to vary by drive-model, there is greater variability in the early rate of failure and, for some drive-models, the forecasted fraction failing is adjusted downward.
+%Should we be brief discussing the plot?
+These plots seem to suggest that Model 1 is far too naive, Model 2 is still too restrictive and Models 3 and 4 are fairly similar.
 
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{heterogeneity-compare-preliminary}
+\includegraphics[width=\textwidth]{heterogeneity-compare}
 \caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
@@
 Conversely, Model 4 follows the observed failures quite closely--accurately capturing failures at the lower and upper ends of the distribution.
-The difference in accuracy between Models 3 and 4 are and Models 1 and 2 are visually quite apparent.
+The difference in accuracy between Models 3 and 4 and Models 1 and 2 are visually quite apparent.
 Differentiating between Model 3 and 4 is less clear; for instance, the fits from Model 3 and 4 line almost exactly on top of each other for drive-model 40.
@@
 
+The question of which Model is best requires the choice of some criteria.
+From the point of view of an operator, the goal of an analysis like this is likely to rate manufacturers and/or to inform future purchaes based on predicted product performance.
+In this case, it seems that the best model would provide the most accurate prediction of lifetime for future units.
+One such metric is log posterior predictive density (lpd), $\log[p(t_new|t)]$.
+Thus we might choose a model as the one most likely to produce a high lpd, one maximizing $\op{E}_h \log[p(t_new|t)]$, where the expectation is with respect to the true distribution of $t$.
+Although $h$ is unknown, it can be estimated using leave-one-out cross-validation:
+$$\frac{1}{G}\frac{1}{n_g}\sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{g,i}^L,c_{g,i},y_{-(g,i)})].$$ Using recently developed software provided in the \texttt{loo} package in \texttt{R},
 To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
@@
 \subsubsection{Model Assessment}
-To assess the goodness of fit for our final model (Model 4), we use a graphical posterior predictive check, wherein replicate data sets are generated from the posterior predictive distribution and compared with the observed data.
+Having chosen Model 4 as the best of the candidate models we considered, we would like to assess whether our model provides an adequate fit to the observed data.
+One way to do this is to draw replicate data sets from the posterior predictive distribution.
 This is a general approach recommended by \citet{gelman1996postpred} for models where no classical goodness-of-fit test is available.
+In the hierarchical setting, there can be different formulations of the predictive distribution.
+In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
+We choose to focus on the populations for which we have data collected.
+There are many possible approaches to the posterior predictive check.
+For this problem, we chose to consider the posterior predictive distribution of the Kaplan-Meier estimate for each drive-model $g$.
+
+% Graphical comparisons can be highly informative in revealing limitations of the model.
+as in our cas a graphical posterior predictive check, wherein replicate data sets are generated from the posterior predictive distribution and compared with the observed data.
 Because of the rich information gleaned from plotting the data, this method can be very effective at identifying limitations of the model.
 
-Due to censoring, it is not entirely clear how to generate the replicate data sets because we have not modeled the censoring mechanism.
-Since we plan to compare nonparametric lifetime estimates, we chose to replicate as best as possible the pattern of censoring that was observed -- to generate the data conditional on the observed pattern of censoring.
-The problem in doing so is that the censoring mechanism itself is not fully observed.
-Our solution was to assume, for units whose failures were observed (and whose censoring time was not), that had they been right-censored, the censoring time would have been the latest censoring or failure time in the data for that drive model.
+Because the variability in the Kaplan-Meier is primarily a function of the number of at-risk units, we want to replicate the pattern of censoring that was observed.
+That is, for unit $i$ of drive-model $g$,
+$$p(t_{rep,g}) = \int \prod_{i=1}^{n_g}f(t_{new}|t_{g,i}^L,c_{g,i},\theta)\,p(\theta|t)d\theta.$$
+Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing $\theta^{(s)}$ at random, then conditionally draw $t_{rep,g}$ from $f(t_{new}|t_{g,i}^L,c_{g,i},\theta^{(s)}.$
+
+An obstacle remains, in that the censoring mechanism itself is not fully observed.
+Our admittedly ad hoc approach is to assume that failed units would have been the censored had they survived until the latest censoring or failure time for that drive model.
 We chose to generate 19 replicates for each drive model.
-This gives a 1/20 chance under the hypothesis of no difference that the most extreme data set would be the observed one and is not too many so as to produce an overly cluttered plot.
+This gives a 1/20 chance under the hypothesis of no difference that the Kaplan-Meier estimate based on the observed data would be ``most extreme" relative all the replicated  from the those one and is not too many so as to produce an overly cluttered plot.
 Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each drive-model $g=1,\ldots,44$, is as follows:
 \begin{enumerate}
-\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.
 Where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right) $, is a random draw from the full posterior distribution.
-\item Repeat step 1 until $\tilde{y}_{rep,g,i} > t_{g,i}^L$.
-\item If $c_{g,i}=0$, let $c_{rep,g,i}^{(s)}=\max \{y_{g,i}: i \ge 1\}$, otherwise let $c_{rep,g,i}^{(s)}=y_{g,i}$.\\
-If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=c_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
+\item Repeat step 1 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
+\item If $c_{g,i}=0$, let $c_{rep,g,i}^{(s)}=\max \{t_{g,i}: i \ge 1\}$, otherwise let $c_{rep,g,i}^{(s)}=t_{g,i}$.\\
+If $\tilde{t}_{rep,g,i}^{(s)}>m_{g,i}$, set $t_{rep,g,i}^{(s)}=c_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=1$, otherwise set $t_{rep,g,i}^{(s)}=\tilde{t}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
 \end{enumerate}
 
-We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator.
-The resulting plots are displayed in figure \ref{fig:post-pred-KM}.
-For most drive-models, the estimates from the observed data appear to behave similarly to those from the replicated data.
-As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there.
-Drive-models 4 and 34 appear to be exceptions as well.
-We will focus our attention on drive-model 4, but our diagnosis of 34 is similar, so the following comments extend to that drive-model as well.
-
 \begin{figure}[H]
-\includegraphics[width=\textwidth,height=.9\textheight]{post-pred-KM-all2}
+\includegraphics[width=\textwidth]{ppcheck-sample}
 \caption{Nonparametric estimates of proportion failing for each drive-model.
@@
 
-To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}).
-From these plots, it is clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the first quarter year.
-Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band.
-Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode as estimated by the global parameters, $\mu_1$ and $\sigma_1$.
-In any case, this example demonstrates that, conditional on the model, the weight of evidence in all the data overpowers the signal from the small sample for this drive-model, which suggests high rates infant mortality in the first quarter year.
+We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the Kaplan-Meier nonparametric estimator.
+The resulting plots are displayed in figure \ref{fig:post-pred-KM}.
+For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.
+Only a representative sample of plots are shown here for brevity.
+There are a few drive-models, however where this is not the case.
+As we addressed previously, in Subsection \ref{subsec:ex1}, drive-model 14 has a discrepancy between the GFLP and the Kaplan-Meier fits in the right tail.
+Due to the large number of failures for this drive-model, the posterior for the hierarchical model is very similar to the stand-alone model, so all the same comments made earlier still apply.
+In addition, drive-models 4, 11 and 34 appear to be exceptions as well.
+We will focus our attention on drive-model 4, but our diagnosis of 34 is similar, so the following comments extend to that drive-model as well.
 
-Generally we consider this behavior a feature of hierarchical modeling.
-For another example of this, consider drive-model 43.
+To attempt to diagnose the cause of the discrepancies between the observed data for drive-model 4 and its posterior predictive distribution, we plot an event plot for a random sample of units for this drive-model, along with the Kaplan-Meier estimate, and the posterior estimate of time to failure (see figure \ref{fig:ex-mod-4}).
+From these plots, it appears that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the several thousand hours, at a time where very few units were observed to be at risk.
+In further support of this hypothesis, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band.
+While it might be tempting to treat those early failures as anomalous, they also point to a weakness in our model.
+By virtue of having all drive-models share in common the Weibull parameters for the infant mortality failure mode, an early rash of failures indicating a subpopulation with, say, an increasing hazard function,  would be problematic.
+While the proportion at risk of early failure can vary via $\pi_g$, the plausible values of the common parameters, $\mu_1$ and $\sigma_1$, do not ``line up" in a way so to influence the posterior as they might have had they been allowed to vary.
+In other words, the problem seems to go away much quicker than would be expected under Model 4.
+
+As another example, to further appreciate how information is being borrowed across groups, consider drive-model 43.
 The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available.
 In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
-Then, in the period where records exist for this drive model, at around $1\frac{1}{2}$ years, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
+Then, in the period where records exist for this drive model, at around 10,000 hours, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
 
@@
 \pagebreak
-
+\section{Posterior predictive graphical checks}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{ppcheck-1.pdf}
+\end{figure}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{ppcheck-2}
+\end{figure}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{ppcheck-3.pdf}
+\end{figure}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{ppcheck-4.pdf}
+\end{figure}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{ppcheck-5.pdf}
+\end{figure}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{ppcheck-6.pdf}
+\end{figure}
+\clearpage
 % \bibliographystyle{plainnat}
+% %\bibliographystyle{apalike}
 \bibliographystyle{plain}

### 2017-08-25 workflow/predKM1.R

@@ -137,9 +137,9 @@ for(j in 1:6){
 set.seed(12122017)
 smplID <- sort(sample(44, 4))
 plot_grid(plot_grid(plotlist=plist$plots[c(smplID)], ncol=2, align="hv"), plist$legend,
-          nrow=1, rel_widths = c(4,.5))
+          nrow=1, rel_widths = c(1,.2))
 
-ggsave("../paper/fig/ppcheck-sample.pdf", width=5, height=6)
+ggsave("../paper/fig/ppcheck-sample.pdf", width=11, height=8)
 #debug
 # p <- get_plotlist(42)
 # 

### 2017-08-25 paper/paper.tex
Revisions to Model Selection

@@
 For example, due to heavy left-truncation, there may be insufficient information for particular drive models to estimate all of the parameters.
-Furthermore, we prefer not to introduce unnecessary complexity into our analyses.
+
 Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of time to failure.
@@
 The smaller plots correspond to Models 1 through 4 reading from left to right, top to bottom.
-Note that while all the paths for the posterior estimates necessarily pass through the origin, the nonparametric curves do not, but only because, for the Kaplan-Meier estimator, left-truncation disallows any kind of interpolation.
+Note that while all the paths for the posterior estimates necessarily pass through the origin, the nonparametric curves do not, but only because, for the Kaplan-Meier estimator, inference is entirely conditional on survival up to left-most left-trunctation time; to make the plot comparable, we use an adjustment based on a parametric model (Model 4).
 
 %Should we be brief discussing the plot?
-These plots seem to suggest that Model 1 is far too naive, Model 2 is still too restrictive and Models 3 and 4 are fairly similar.
+These plots suggest that Model 1 is insufficient to explain the observed data.
+Model 2 certainly captures more of the heterogeneity observed in the Kaplan-Meier estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in the rate of failure, which may be too restrictive.
+Model 3, which removes this restriction on $\sigma_{2,g}$ suggests that there is evidence to support variation in this parameter among drive-models.
+While Model 3 and 4 do display differences, their impact is relatively subtle.
 
@@
 
-We can also compare the model fits for each drive-model individually.
-Using the set of parameters that maximize the posterior log probability, we examine the GLFP cdf for each of the 4 model specifications.
+We can also compare the four Model fits for each drive-model individually.
+For these comparisons, we use the set of parameters which maximize the marginal posterior log probability.
+Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.
 The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure \ref{fig:mod_comp_leg}.
-The black step functions again corresponds to the adjusted non-parametric estimates.
+The black step functions again correspond to the adjusted Kaplan-Meier estimates.
 As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.
-All of the GLFP models appear to overestimate the proportion of failed hard drives for drive-model 2, but Model 4 agrees best with the nonparametric estimate.
-For drive-model 14, Models 1-3 underestimate the proportion of drives failing, and for drive-model 40, they slightly overestimate the data.
-Conversely, Model 4 follows the observed failures quite closely--accurately capturing failures at the lower and upper ends of the distribution.
-The difference in accuracy between Models 3 and 4 and Models 1 and 2 are visually quite apparent.
-Differentiating between Model 3 and 4 is less clear; for instance, the fits from Model 3 and 4 line almost exactly on top of each other for drive-model 40.
+All of the GLFP models appear to produce lower estimates, relative to Kaplan-Meier, of the proportion of failed hard drives for drive-model 2, with Model 4 being in closest agreement.
+For drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for drive-model 40, they produce higher estimates.
+On the other hand, Model 4 follows the nonparametric estimate quite closely across stages of lifetime.
+The differences between Models 3 and 4 and Models 1 and 2 are visually quite apparent.
+Differentiating between Model 3 and 4 is less clear; for instance, the estimates from Model 3 and 4 are nearly identical for drive-model 40.
 \begin{figure}[H]
@@
 
-The question of which Model is best requires the choice of some criteria.
-From the point of view of an operator, the goal of an analysis like this is likely to rate manufacturers and/or to inform future purchaes based on predicted product performance.
-In this case, it seems that the best model would provide the most accurate prediction of lifetime for future units.
-One such metric is log posterior predictive density (lpd), $\log[p(t_new|t)]$.
-Thus we might choose a model as the one most likely to produce a high lpd, one maximizing $\op{E}_h \log[p(t_new|t)]$, where the expectation is with respect to the true distribution of $t$.
-Although $h$ is unknown, it can be estimated using leave-one-out cross-validation:
-$$\frac{1}{G}\frac{1}{n_g}\sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{g,i}^L,c_{g,i},y_{-(g,i)})].$$ Using recently developed software provided in the \texttt{loo} package in \texttt{R},
+From the point of view of an operator, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
+In this case, we should choose the Model that will provide the most accurate prediction for future units of previously observed drive-models.
+One appropriate criterion for model selection, then, is log predictive density (lpd),
+
+\begin{equation}
+\mbox{lpd} = \log[p(t_{g, new}|t, t^L, c)] = \log \int p(t_{g, new}|\theta) p(\theta|t,t^L,c) d\theta
+\end{equation}
+
+Using this criterion, we should choose the Model with the highest expected lpd:
+\begin{equation}
+\op{E}_h \mbox{lpd} = \int p(t_{g,new}|t,t^L,c) h(t_{g,new}) d t_{g,new}.
+\end{equation}
+Where $h$ is the true distribution of $t_{new}$. Since $h$ is unknown, leave-one-out cross-validation can be used to estimate it:
+\begin{equation}
+\widehat{\op{E}_h \mbox{lpd}} = \frac{1}{G}\frac{1}{n_g}\sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{g,i}^L,c_{g,i},-t_{-(g,i)})].
+\end{equation}
+Using recently developed software provided in the \texttt{loo} package in \texttt{R},
 To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.

### 2017-08-25 paper/paper.tex
Edits to Bkround.

--- 2017-08-25 paper/paper.tex
@@
 However, in many situations the cause of failure is unknown to the analyst.
-This is referred to in the literature as "masking".\\
+This is referred to in the literature as ``masking''.\\
 
@@
 This helped identify parameters when using maximum likelihood (ML) estimation.
-Extending Chan and Meeker's analysis, Basu et.al performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.
+Extending Chan and Meeker's analysis, Basu et al.
+performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure \cite{basu}.
 Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.
@@
 
-\subsection{Example 1}
+\section{Motivating Example}
+\subsection{Backblaze hard drive data}
+\label{sec:Data}
+Backblaze is a company that offers cloud backup storage to protect against data loss.
+Since 2013 it has been collecting data on hard drives operating at its facility.
+The purpose is to provide consumers and businesses with reliability information on different hard drive-models.
+The hard drives continuously spin in controlled storage pods.
+Drives are run until failure.
+When a hard drive fails it is permanently removed and replaced.
+In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.
+Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
+In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.
+No other analysis or modeling of the failure data is provided.
+Backblaze does, however, encourage others to further analyze its data.
+
+As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.
+Some drive-models have been running since 2013 or before, while others were added at a later date.
+Data have been reported on 75,297 different hard drives that are or where in operation.
+The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
+Figure \ref{fig1} shows a scatterplot of the total observed running time in hours versus the total number of failures for drive-models with at least 3 failures.
+For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analyses.
+
+\begin{figure}[H]
+  %\includegraphics[width=.9\textwidth]{data-sum.pdf}
+  \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
+  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.
+Each number labels a unique drive brand model.
+Axes are on the log scale.}
+  \label{fig1}
+\end{figure}
+
+\subsection{Single GLFP Model}
 \label{subsec:ex1}
@@
 We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \cite{kaplan}.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a more detailed description).
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a detailed description).
 
@@
 
-\section{Real data analysis}
+\section{Analysis of the Backblaze hard drive data}
 \label{sec:Data analysis}
-\subsection{Backblaze hard drive data}
-\label{sec:Data}
-Backblaze is a company that offers cloud backup storage to protect against data loss.
-Since 2013 it has been collecting data on hard drives operating at its facility.
-The purpose is to provide consumers and businesses with reliability information on different hard drive-models.
-The hard drives continuously spin in controlled storage pods.
-Drives are run until failure.
-When a hard drive fails it is permanently removed and replaced.
-In addition, the number of storage pods is increasing as Backblaze adds to its storage capacity.
-Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
-In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.
-No other analysis or modeling of the failure data is provided.
-Backblaze does, however, encourage others to further analyze its data.
-
-As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.
-Some drive-models have been running since 2013 or before, while others were added at a later date.
-Data have been reported on 75,297 different hard drives that are or where in operation.
-The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
-Figure \ref{fig1} shows a scatterplot of the total observed running time in hours versus the total number of failures for drive-models with at least 3 failures.
-For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analyses.
-
-\begin{figure}[H]
-  %\includegraphics[width=.9\textwidth]{data-sum.pdf}
-  \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
-  \caption{Scatterplot of total observed running time (hours) versus total number of observed failures.  Each number labels a unique drive brand model.}
-  \label{fig1}
-\end{figure}
-
 \subsection{Modeling}
-\subsubsection{Candidate Models}
 We considered four models, all based on \ref{eq:hier-model}, which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):
@@
 of drive-model $d$ for which records are available. To produce
-unconditional estimates, we adapt the adjustment method outlined by
-Turnbull, and given in more detail by Meeker and Escobar
-\cite{turnbull,meeker}.  For each drive-model we select
+unconditional estimates, we adapt the adjustment method outlined by by Meeker and Escobar  in Chapter 11 of Reliability Analysis \cite{meeker}.  For each drive-model we select
 $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.

### 2017-08-25 paper/paper.tex
Edits and a couple fixes to plotting functions.

@@
 As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters.
-This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information.
+This enables estimation of the lifetime distribution for all populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of population specific information.
 We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.
@@
 The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
-Figure \ref{fig1} shows a scatterplot of the total observed running time in hours versus the total number of failures for drive-models with at least 3 failures.
-For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analyses.
+Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.
+For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.
 
@@
 \label{subsec:ex1}
-We fit the GLFP model on real hard drive failure data from a cloud-based storage company.
-Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model.
-The data contain the number of hours the hard drive was in operation, and whether the drive failed or was right censored.
-No information is available about the cause of failure.
-To illustate the GLFP model, we will present an analysis of the drive-model with the most observed failures.
-We use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution.
+To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures (1,713), drive-model 14.
+The non-parametric Kaplan-Meier estimate plotted below does not appear linear, suggesting different rates of failure in the data.
+In addition, computer hard drives are known to have a mixture of early and late failures \cite{chan}.
+Thus, as a first step, the GLFP model seems like an appropriate parametric model.
+
+\begin{figure}[H]
+\centering
+  \includegraphics[width=.8\textwidth]{KM14.pdf}
+  \caption{Kaplan Meier estimate for drive model 14.  Bands correspond to Greenwood standard errors.}
+  \label{fig1}
+\end{figure}
+
+To estimate the GLFP we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution.
 We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}$) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}$).
@@
 $$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle, \quad
-\sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
+\sigma_1, \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle 
+%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
 
@@
 Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.
-They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower mean time to failure than the wearout mode.
-This is helpful for estimation since, as discussed in Chan \& Meeker, ML had difficulty identifying the two sets of GLFP parameters without making strong assumptions about the data.
 \\
@@
 The parameter estimates for the two Weibull distributions also match intuition.
-The wearout failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.
-Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, an increasing hazard.
+The wearout failure mode puts posterior probability on a values of $\sigma_1$ greater than 1, which corresponds to a decreasing hazard.
+Conversely, the credible interval for $\sigma_2$, the wearout mode, is strictly below 1, an increasing hazard.
 The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.
@@
 $\pi$ & 0.03 & 0.05 & 0.10 \\ 
- $\beta_1$ & 0.47 & 1.13 & 1.72 \\ 
-  $\beta_2$ & 4.47 & 4.70 & 4.95 \\ 
+ $\sigma_1$ & 1.25 & 1.96 & 15.54 \\ 
+  $\sigma_2$ & 0.48 & 0.65 & 0.78 \\ 
   $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
@@
 
-In Figure 1 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.
+In Figure 3 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.
 The solid stepwise curve corresponds to the observed failures.
 Censored drives are not plotted.
-Standard error bands are calculated using Greenwood's method \cite{green}.
+Standard error bands are again calculated using Greenwood's method \cite{green}.
 Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.

### 2017-08-25 plots/dm-summ_textscatter.R

@@ -21,14 +21,17 @@ my_summary <- ddply(dat, .(model), function(x){
   data.frame(tt=tt, fs=fs)
 })
 
+my_summary$tt100k <- my_summary$tt/100000
+
 #in yrs
-xbrks=c(10, 100, 1000, 10000)
+#xbrks = c(10, 100, 1000, 10000)
+xbrks2 = c(1, 10, 100, 1000)
 
 
 setwd("paper/fig/")
 pdf("dm-summ-scatter.pdf", width=7, height=5)
-ggplot(my_summary, aes(x=tt, y=fs)) + geom_text(aes(label=model), size=4) +
-  scale_x_continuous(name="Total time observed (years)", trans="log", breaks=xbrks*24*365, labels=xbrks)+
-  scale_y_continuous(name="Observed failures", trans="log", breaks = c(1, 10, 100, 1000))+
-  theme_bw(base_size=14)#+geom_smooth(method = "lm", se = F, linetype="dashed")
+ggplot(my_summary, aes(x=tt100k, y=fs)) + geom_text(aes(label=model), size=4) + 
+  scale_x_continuous(name="Total time observed (hundred thousands of hours)", trans="log", breaks = xbrks2, labels = xbrks2) +
+  scale_y_continuous(name="Observed failures", trans="log", breaks = c(1, 10, 100, 1000)) +
+  theme_bw(base_size=14) #+geom_smooth(method = "lm", se = F, linetype="dashed")
 dev.off()

### 2017-08-25 plotting_fns/KM_plot2.R

@@ -141,7 +141,7 @@ baseKMplot.multiple <- function(fit_list, xlimits=NULL, ylimits=NULL, color="bla
 }
 
 #Make base KMplot
-baseKMplot <- function(fit, xlimits=NULL, ylimits=NULL, color="black", linetype=1, logscale=NULL, prob=TRUE, label="nonparametric", alpha=1){
+baseKMplot <- function(fit, xlimits=NULL, ylimits=NULL, color="black", linetype=1, logscale = FALSE, prob=TRUE, label="nonparametric", alpha=1){
   require(ggplot2)
   if(!("myKM" %in% class(fit))) stop("fit must be output from KM.survfit")
   if(!is.null(xlimits)){

### 2017-08-25 plotting_fns/test_KM_plot2.R

@@ -3,9 +3,9 @@ library(rstan)
 library(plyr)
 library(dplyr)
 
-source("plotting_fns/KM_plot2.R")
+source("/plotting_fns/KM_plot2.R")
 
-s <- readRDS("workflow2/samples_nopool_data_dm8.rds") #number 8 after applying filter (f>=5, ef>=1, lf>=1)
+s <- readRDS("../../workflow2/samples_nopool_data_dm14.rds") #number 8 after applying filter (f>=5, ef>=1, lf>=1)
 samp <- extract(s)
 
 dat14 <- readRDS("BB_data/clean_unit_summaries.rds") %>%
@@ -25,12 +25,11 @@ band14 <- bandFromPSamp(samp, xlimits, length.out = 50, N = 1000, logscale = TRU
 p14base <- addBandToBaseplot(p14base, band14, color="red", linetype=3, alpha=.2, label="Bayes 90%")
 
 plotFinally(p14base, xbrks = c(.1, .5, 1, 2, 4)*365*24,
-            ybrks = c(.01, .02, .05, .1, .2, .5), years=TRUE) +
+            ybrks = c(.01, .02, .05, .1, .2, .5), years=FALSE) +
   theme_bw()
 
-plotFinally(p14base, xbrks = c(.1, .5, 1, 2, 4)*365*24,
-            ybrks = c(.01, .02, .05, .1, .2, .5), years=TRUE, greenwood = TRUE) +
-  theme_bw()
+KMDrive14 <- plotFinally(p14base, xbrks = c(1000, 5000, 10000, 15000, 20000, 25000),
+            ybrks = c(.01, .02, .05, .1, .2, .5), years = FALSE, greenwood = TRUE) + theme_bw()
 
 ##################################################
 sfull <- readRDS("workflow/samples_lor_only3fails.rds")

### 2017-08-25 paper/paper.tex
More work on model selection section

--- 2017-08-25 paper/paper.tex
@@
 In this case, we should choose the Model that will provide the most accurate prediction for future units of previously observed drive-models.
-One appropriate criterion for model selection, then, is log predictive density (lpd),
+One appropriate criterion for model selection, then, is log pointwise predictive density (lpd),
 
 \begin{equation}
-\mbox{lpd} = \log[p(t_{g, new}|t, t^L, c)] = \log \int p(t_{g, new}|\theta) p(\theta|t,t^L,c) d\theta
+\mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] = \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
 \end{equation}
 
-Using this criterion, we should choose the Model with the highest expected lpd:
+Using this criterion, we should choose the Model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$. Suppressing the conditioning on $t^L$ and $c$,
 \begin{equation}
-\op{E}_h \mbox{lpd} = \int p(t_{g,new}|t,t^L,c) h(t_{g,new}) d t_{g,new}.
+elpd = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
+\label{elpd}
 \end{equation}
-Where $h$ is the true distribution of $t_{new}$. Since $h$ is unknown, leave-one-out cross-validation can be used to estimate it:
+where $h$ is a density for the true distribution of $t_{new,g,i}$. Although $h$ is unknown, leave-one-out cross-validation provides a robust estimator:
 \begin{equation}
-\widehat{\op{E}_h \mbox{lpd}} = \frac{1}{G}\frac{1}{n_g}\sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{g,i}^L,c_{g,i},-t_{-(g,i)})].
+\widehat{elpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
+\label{loo-est}
 \end{equation}
-Using recently developed software provided in the \texttt{loo} package in \texttt{R},
-To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
+A recently published \texttt{R} package, \texttt{loo}\cite{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
+It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\frac{\sum_{i=1}^S w_{i,s} p(t_{g,i})}{\sum_{s=1}^S w_{i,s}}$.
+In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.
+For details, refer to \citet{vehtari}.
+%When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
+% To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
 As outlined in Vehtari et.
@@
 For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.
-The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
-We computed the elpd for all 4 models using the R package loo \cite{loo}.
-When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically.
+% The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).  
+Results showing estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:1}.
  We calculated the difference in expected predictive accuracy for Model 4 vs 3, Model 3 vs 2, and Model 2 vs 1, as well as the standard error of the difference (Table 1).  
-As the model complexity increased, the expected predictive accuracy improved.
-Of all the models, Model 4 had the best predictive accuracy and was significantly better than the other 3 models.
+For the Models we considered, each successive increase in model complexity increases elpd.  Of all the models, Model 4 has significantly better elpd compared to the the other 3 Models.
 \begin{figure}
@@
   \hline
- & ELPD \ & Difference in ELPD \ & SE of the Difference \\ 
+ & $\widehat{elpd}$ \ & Difference in $\widehat{elpd}$ \ & SE of the Difference \\ 
   \hline
@@
 \end{tabular}
-\caption{Expected Log Pointwise Predictive Density (ELPD) for each model specification.
+\caption{Expected Log Pointwise Predictive Density (elpd) for each model specification.
 Each model is compared to the more parsimonious model below.
@@
 
-As another example, to further appreciate how information is being borrowed across groups, consider drive-model 43.
+As another example, to further appreciate how information is being borrowed across groups, consider drive-model 43 (Figure \ref{fig:ex-mod-43}).
 The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available.

### 2017-08-25 paper/paper.tex
Fix some captions

@@
 $$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1, \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle 
+\sigma_1, \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 \includegraphics{dm4-exception}
-\caption{Top: Graphical representation of lifetime for random sample of 200 drives of drive-model 4 from the Backblaze data set.
-Solid white lines indicate the observed window of life.
-Right-censored observations changed to a dashed line and failures mark the transition to a solid black line.
-Bottom: Solid step function is truncation-adjusted nonparametric estimate of the proportion failing.
-The dotted line is an estimate obtained after truncating the data to those starting after .25 years, with the corresponding truncation adjustment.
-The dashed line with gray band is pointwise posterior median, and 90\% credible interval, of proportion failing.}
+\caption{Top: Event plot for uniform random sample of 60 units of drive-model 4.
+Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_4,i, \infty)|c_{4,i}=1$, and right endpoints without joined dashed lines are observed failure times, i.e.
+$t_{4,i}|c_{4,i}=0$.
+
+Bottom: The adjusted Kaplan-Meier estimate (solid step function) substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval) due to several early failures while few were at risk.
+Using the same Kaplan-Meier estimator, but only the data truncated after 2500 hours (dotted line) shows close agreement with the posterior median.}
 
@@
   \includegraphics[width=.8\textwidth]{dm-eval}
-  \caption{Left: Posterior predictive relative value-at-loss for new
-    unit, by drive-model. Right: TTF (top) and VAF based on 40\% annual depreciation (bottom), for best 11 drive-models.}
+  \caption{Left: Summary of posterior predictive distributions for $e^{-.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each drive-model.
+  Right: A comparison of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
   \label{fig5}
@@
 % \bibliographystyle{plainnat}
-% %\bibliographystyle{apalike}
-\bibliographystyle{plain}
+\bibliographystyle{apalike}
+% \bibliographystyle{plain}
 \bibliography{sample}

### 2017-08-26 paper/paper.tex
Edits to paper for next Thursday meeting.

@@
 Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.
-Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $Y_i$).
+Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$).
 The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
-Assuming the lifetimes of all units are independent, the likelihood for the data, $Y_1,\ldots,Y_N$ is given by
+Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
 \begin{equation*}
-L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
+L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
 \end{equation*}
@@
 
-\subsection{Single GLFP Model}
+\subsection{Single Drive-Model Analysis}
 \label{subsec:ex1}
-To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures (1,713), drive-model 14.
-The non-parametric Kaplan-Meier estimate plotted below does not appear linear, suggesting different rates of failure in the data.
+To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, drive-model 14.
+The Kaplan-Meier estimate plotted below suggests at least two failure modes.
+The curve has a slow ramp up in failures till about 10,000 hours, increases from about 10,000 to 20,000 hours, and then starts to gradually flatten out.
 In addition, computer hard drives are known to have a mixture of early and late failures \cite{chan}.
-Thus, as a first step, the GLFP model seems like an appropriate parametric model.
+Therefore, there is both an empirical and theoretical justification to apply the GLFP model.
 
@@
 \centering
-  \includegraphics[width=.8\textwidth]{KM14.pdf}
+  \includegraphics[width=.8\textwidth]{mod14km.pdf}
   \caption{Kaplan Meier estimate for drive model 14.  Bands correspond to Greenwood standard errors.}
@@
 
-The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.
-Given the interpretation of $\pi$ as the proportion of defective units, which we would expect to be small, a range primarily covering the left side of 0.50 is reasonable and weakly informative.
-The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.
-Again, this allows for a wide range of hazards, larger than we would expect to see in the data.
-The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$.
-For $t_{0.2,2}$ the range is $(3.4,16.6)$.
-Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.
-\\
+These priors put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.
+Thus, we consider the priors to be relatively uninformative.
 
@@
 
-To borrow strength across different populations, we can model the population specific parameters hierarchically, allowing the data to inform the hyperparameters.
-%For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:
-
+In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population specific parameters hierarchically, borrowing strength across the sub-populations.
 Let
 \begin{equation}
-y_{g,i} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
+y_{ig} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
 \end{equation}
-where $g=1,\ldots,G$ indexes populations. Next we model these parameters themselves as random variables. Let
+where $g=1,\ldots,G$ indexes the sub-populations.
+The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g}, \pi_{g})$.
+Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
 \begin{equation*}
-\sigma_{1g} \ind \op{LN} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
+L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
+\end{equation*}
+
+Next, we model these parameters themselves as random variables, varying across the different sub-populations.  Let
+\begin{equation*}
+\sigma_{1g} \ind \op{Lognormal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
 \end{equation*}
 \begin{equation*}
-\sigma_{2g} \ind \op{LN} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
+\sigma_{2g} \ind \op{Lognormal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
 \end{equation*}
@@
 \label{eq:hier-model}
-\log t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{N} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
+t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{Lognormal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
 \end{equation}
 \begin{equation*}
-\log t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{N} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
+t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{Lognormal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
 \end{equation*}

### 2017-08-27 paper/paper.tex
Working on bib.
Modify caption for drive-model 4 figure
Fixing references to ASA style.

@@
 Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.
-The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \cite{nelson}.
+The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep{nelson}.
 A common assumption in series systems is the time to failure for each risk is statistically independent.
@@
 When modeling computer system failures Reiser et al.
-assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.
+assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \citep{reiser}.
 Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.
@@
 Extending Chan and Meeker's analysis, Basu et al.
-performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure \cite{basu}.
-Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.
+performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure \citep{basu}.
+Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \citep{berger}.
 More recently, Ranjan et al.
-considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.
+considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \citep{ranjan}.
 Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.
@@
 In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.
-Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}.
+Every quarter Backblaze makes its data publicly available through their website \citep{backblaze}.
 In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.
@@
 The curve has a slow ramp up in failures till about 10,000 hours, increases from about 10,000 to 20,000 hours, and then starts to gradually flatten out.
-In addition, computer hard drives are known to have a mixture of early and late failures \cite{chan}.
+In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.
 Therefore, there is both an empirical and theoretical justification to apply the GLFP model.
@@
 After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.
-We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \cite{kaplan}.
+We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \citep{kaplan}.
 With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a detailed description).
@@
 Censored drives are not plotted.
-Standard error bands are again calculated using Greenwood's method \cite{green}.
+Standard error bands are again calculated using Greenwood's method \citep{green}.
 Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.
@@
 Each model was fit using the {\tt
-  rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
-\cite{betancourt}.
+  rstan} \citep{rstan} package in {\tt R} \citep{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
+\citep{betancourt}.
 HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density.
@@
 \end{equation}
-A recently published \texttt{R} package, \texttt{loo}\cite{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
+A recently published \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
 It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\frac{\sum_{i=1}^S w_{i,s} p(t_{g,i})}{\sum_{s=1}^S w_{i,s}}$.
 In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.
-For details, refer to \citet{vehtari}.
+For details, refer to \cite{vehtari}.
 %When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
@@
 One way to do this is to draw replicate data sets from the posterior predictive distribution.
-This is a general approach recommended by \citet{gelman1996postpred} for models where no classical goodness-of-fit test is available.
+This is a general approach recommended by \cite{gelman1996postpred} for models where no classical goodness-of-fit test is available.
 In the hierarchical setting, there can be different formulations of the predictive distribution.
@@
 for depreciation, rather than using TTF as the metric for drive-model comparison, we use the value at replacement. The IRS considers
-computer hardware to be ``5-year'' equipment\cite{f4562}. Using``the declining
+computer hardware to be ``5-year'' equipment \citep{f4562}. Using``the declining
 balance method'' the rate of depreciation is 40\% per year.
@@
 We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
-\cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
+\citep{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
 $\widehat{F_d(t)}_{KM}$, is conditional on survival up to
@@
 of drive-model $d$ for which records are available. To produce
-unconditional estimates, we adapt the adjustment method outlined by by Meeker and Escobar  in Chapter 11 of Reliability Analysis \cite{meeker}.  For each drive-model we select
+unconditional estimates, we adapt the adjustment method outlined by by Meeker and Escobar  in Chapter 11 of Reliability Analysis \citep{meeker}.  For each drive-model we select
 $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
@@
 \clearpage
-% \bibliographystyle{plainnat}
-\bibliographystyle{apalike}
+%\bibliographystyle{plainnat}
+\bibliographystyle{asa}
 % \bibliographystyle{plain}

### 2017-08-27 plots/single-drive-4-Models.R
Scripts for plots

@@ -1,0 +1,127 @@
+library(plyr)
+library(dplyr)
+library(rstan)
+library(ggplot2)
+
+setwd("plots/")
+dat <- readRDS("../BB_data/clean_unit_summaries.rds")
+dat$model <- as.integer(dat$model)
+
+overview <- readRDS("../BB_data/overview.rds")
+id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
+overview$stan_id <- NA
+overview[id,]$stan_id <- 1:length(id)
+
+source("../plotting_fns/KM_plot2.R")
+
+ssigtp <- readRDS("../workflow_sig2_and_tp2_vary/vary_s2_and_tp2_4_17.rds")
+stp2 <- readRDS("../workflow_tp2_vary/samples_tp2_vary.rds")
+snull <- readRDS("../workflow_null/samples_null_model_3_29.rds")
+sfull <- readRDS("../workflow/samples_lor_only3fails.rds")
+tr_adj <- readRDS("../BB_data/tr_adj_tp2s2pi.rds")$median
+
+samps2tp <- extract(ssigtp)
+samptp2 <- extract(stp2)
+sampnull <- extract(snull)
+sampfull <- extract(sfull)
+
+xlimits <- c(200, 30000)
+ylimits <- c(.001, .8)
+xbreaks <- 1:6*5000
+ybreaks <- 1:8*0.1
+
+#local function to produce ggplot with 4 curves for each drive-model
+single_drive.all.4.Models <- function(dm, prob.log.scales){
+  local_dat <- filter(dat, model==overview$model[which(overview$stan_id==dm)])
+  
+  km <- KM.survfit(local_dat, greenwood = FALSE)
+  
+  if(!prob.log.scales){
+    xlimits <- range(km$t) + c(-1000,1000)
+    xlimits[1] <- max(xlimits[1], 0)
+    ylimits <- range(km$Ft) + c(-.5, .2)
+    ylimits[1] <- max(ylimits[1], 0)
+    ylimits[2] <- min(ylimits[2], 1)
+    xbreaks <- 0:ceiling(xlimits[2]/1000)*1000
+    ybreaks <- 0:ceiling(ylimits[2]*20) /20
+  } else{
+    xlimits <- exp(range(log(km$t))+c(-.1,.1))
+    ylimits <- psev(qsev(range(km$Ft)) + c(-1, 1))
[…]

### 2017-08-27 plots/single-drive-plots.R

@@ -1,0 +1,61 @@
+library(plyr)
+library(dplyr)
+library(rstan)
+library(ggplot2)
+
+setwd("plots/")
+dat <- readRDS("../BB_data/clean_unit_summaries.rds")
+dat$model <- as.integer(dat$model)
+
+overview <- readRDS("../BB_data/overview.rds")
+id <- with(overview, which(overview$early >= 0 & overview$late_f >= 0 & f >=3))
+overview$stan_id <- NA
+overview[id,]$stan_id <- 1:length(id)
+
+source("../plotting_fns/KM_plot2.R")
+
+#Figure 14 (index 8 in independent fits)
+indFit <- extract(readRDS("../workflow2/samples_nopool_data_dm8.rds"))
+samp14 <- with(indFit, list(mu1=mu1, sigma1=sigma1, mu2=mu2, sigma2=sigma2, log_pi=log_pi))
+
+#Linear scales
+xlimits <- c(200, 30000)
+ylimits <- c(.001, .8)
+xbreaks <- 1:6*5000
+ybreaks <- 1:8*0.1
+
+km14 <- KM.survfit(filter(dat, model==overview$model[which(overview$stan_id==14)]),
+                   greenwood = TRUE, alpha=.1)
+base_p14 <- baseKMplot(km14, xlimits=xlimits, ylimits=ylimits, color="blue", linetype="solid",
+                       logscale = FALSE, prob=FALSE, label = "Kaplan-Meier")
+p14 <- plotFinally(base_p14, xbrks = xbreaks, ybrks = ybreaks, years=FALSE,
+                   greenwood = TRUE, alpha = .1)
+p14 + theme_bw(base_size=14) + theme(legend.position="none")
+ggsave("../paper/fig/km14-natural.pdf", width = 6, height = 6)
+
+#Add independent fit
+band14_nat <- bandFromPSamp(samp14, range=xlimits, length.out = 40, N=1000, logscale = FALSE)
+comb_p14 <- addBandToBaseplot(baseplot = base_p14, bandObj = band14_nat,
+                              color="red", linetype="dashed", alpha=.1, label="GLFP posterior")
+p14plus <- plotFinally(comb_p14, xbrks = xbreaks, ybrks = ybreaks, years=FALSE,
+                       greenwood = TRUE, alpha = .1)
+p14plus + theme_bw(base_size=14)
[…]

### 2017-08-28 paper/paper.tex
Replace figures, add captions

@@
 \centering
-  \includegraphics[width=.8\textwidth]{mod14km.pdf}
-  \caption{Kaplan Meier estimate for drive model 14.  Bands correspond to Greenwood standard errors.}
+  \includegraphics[width=.8\textwidth]{km14-prob}
+  \caption{Kaplan Meier estimate for drive model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
   \label{fig1}
@@
 \centering
-  \includegraphics[width=.8\textwidth]{mod14_example1.pdf}
+  \includegraphics[width=.8\textwidth]{km14-prob-plus}
   \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.
 Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.
-Solid line is a non-parametric estimate with pointwise 90\% uncertainty band in gray.}
+Solid line is a non-parametric estimate with pointwise 90\% uncertainty band in blue.}
   \label{fig1}
@@
 Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.
+
+\begin{figure}[H]
+{\centering
+\includegraphics[width=\textwidth]{single-drive-4-Models-ex}}
+\caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
+The solid step function is the truncated adjusted Kaplan-Meier points.}
+\label{fig:mod_comp_leg}
+\end{figure}
+
 The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure \ref{fig:mod_comp_leg}.
@@
 Differentiating between Model 3 and 4 is less clear; for instance, the estimates from Model 3 and 4 are nearly identical for drive-model 40.
-\begin{figure}[H]
-    \centering
-   \includegraphics[width=\textwidth]{mod_compare_legend}
-		\caption{Estimated GLFP for 4 drive-models plotted on Weibull paper.  The solid step function is the truncated adjusted Kaplan-Meier points.  \label{fig:mod_comp_leg}} 
-\end{figure}
 
@@
 \includegraphics{dm4-exception}
-\caption{Top: Event plot for uniform random sample of 60 units of drive-model 4.
-Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_4,i, \infty)|c_{4,i}=1$, and right endpoints without joined dashed lines are observed failure times, i.e.
-$t_{4,i}|c_{4,i}=0$.
-
-Bottom: The adjusted Kaplan-Meier estimate (solid step function) substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval) due to several early failures while few were at risk.
+\caption{\small Top: Event plot for uniform random sample of 60 units of drive-model 4.
+Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_4,i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
+$t_{4,i}\mbox{, where }c_{4,i}=0$.\\ Bottom: The adjusted Kaplan-Meier estimate (solid step function) substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval) due to several early failures while few were at risk.
 Using the same Kaplan-Meier estimator, but only the data truncated after 2500 hours (dotted line) shows close agreement with the posterior median.}
@@
 \begin{figure}
-\centering
-\includegraphics{dm43-shrinkage}
-\caption{}
+{\centering
+\includegraphics{dm43-shrinkage}}
+\caption{\small Top: Event plot for all units of drive-model 4.
+Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_{43},i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
+$t_{43,i}\mbox{, where }c_{43,i}=0$.\\ Bottom: The solid step function correponds to the adjusted Kaplan-Meier estimate.
+Due to the absence of collected information about early life for this particular model,due to left-truncation, the posterior median (dashed line) follows the ``global model" (dotted line), i.e.
+the posterior median of proportion failing for a new drive-model, approximately until the beginning of the earliest observational period.}
 \label{fig:ex-mod-43}

### 2017-08-28 paper/paper.tex
Fix broken refs, address various compilation problems (compiles)

@@
 \usepackage{natbib}
+\bibliographystyle{asa}
 \usepackage{url} % not crucial - just used below for the URL
@@
 The structure of the paper is as follows.
-In Section~\ref{Motivating example}, we describe, as a motivating example, the Backblaze Hard Drive data set.
 In Section~\ref{sec:Background}, we introduce our proposed method and potential applications and examine previous related work in the literature.
+In Section~\ref{sec:Data}, we describe the Backblaze Hard Drive data set, which is used throughout as a motivating example.
 Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data.
@@
 Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
-Finally, in section~\ref{sec:Conclusions}, we discuss the strengths and limitations of our method and consider extensions and alternatives.
+Finally, in section~\ref{sec:Discussion}, we discuss the strengths and limitations of our method and consider extensions and alternatives.
 
@@
 \section{Motivating Example}
+\label{sec:Data}
 \subsection{Backblaze hard drive data}
-\label{sec:Data}
 Backblaze is a company that offers cloud backup storage to protect against data loss.
@@
 Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_4,i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
-$t_{4,i}\mbox{, where }c_{4,i}=0$.\\ Bottom: The adjusted Kaplan-Meier estimate (solid step function) substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval) due to several early failures while few were at risk.
+$t_{4,i}\mbox{, where }c_{4,i}=0$.\hspace{\textwidth}Bottom: The adjusted Kaplan-Meier estimate (solid step function) substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval) due to several early failures while few were at risk.
 Using the same Kaplan-Meier estimator, but only the data truncated after 2500 hours (dotted line) shows close agreement with the posterior median.}
@@
 \begin{figure}
-{\centering
-\includegraphics{dm43-shrinkage}}
+\centering
+\includegraphics{dm43-shrinkage}
 \caption{\small Top: Event plot for all units of drive-model 4.
 Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_{43},i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
-$t_{43,i}\mbox{, where }c_{43,i}=0$.\\ Bottom: The solid step function correponds to the adjusted Kaplan-Meier estimate.
+$t_{43,i}\mbox{, where }c_{43,i}=0$.\hspace{\textwidth}Bottom: The solid step function correponds to the adjusted Kaplan-Meier estimate.
 Due to the absence of collected information about early life for this particular model,due to left-truncation, the posterior median (dashed line) follows the ``global model" (dotted line), i.e.
@@
 %\bibliographystyle{plainnat}
-\bibliographystyle{asa}
 % \bibliographystyle{plain}
-\bibliography{sample}
+\bibliography{./sample}
 

### 2017-08-28 plots/single-drive-4-Models.R
Add plots

@@ -35,7 +35,7 @@ single_drive.all.4.Models <- function(dm, prob.log.scales){
   local_dat <- filter(dat, model==overview$model[which(overview$stan_id==dm)])
   
   km <- KM.survfit(local_dat, greenwood = FALSE)
-  
+  km <- tr.adj(km, tr_adj = tr_adj[dm])  
   if(!prob.log.scales){
     xlimits <- range(km$t) + c(-1000,1000)
     xlimits[1] <- max(xlimits[1], 0)
@@ -48,8 +48,7 @@ single_drive.all.4.Models <- function(dm, prob.log.scales){
     xlimits <- exp(range(log(km$t))+c(-.1,.1))
     ylimits <- psev(qsev(range(km$Ft)) + c(-1, 1))
     ylimits[1] <- min(ylimits,.001)
-    xbreaks <- signif(1000*10^seq.int(from=floor(log10(xlimits[1]/1000)),
-                                to=ceiling(log10(xlimits[2])/1000), by=.5), 1)
+    xbreaks <- c(100, 500, 1000, 2000, 5000,10000, 20000, 40000, 80000, 160000)
     ybreaks <- signif(psev(seq.int(from=floor(qsev(ylimits[1])),
                                    ceiling(qsev(ylimits[2])), by=.5)),1)
   }
@@ -117,11 +116,11 @@ save(plots.prob, file="plots-prob.RData")
 library(cowplot)
 lgnd <- get_legend(plots.prob[[1]])
 lgnd_off <- theme(legend.position="none")
-plot_grid(plot_grid(plots.prob[[2]] + ggtitle("2") + lgnd_off,
+to.save=plot_grid(plot_grid(plots.prob[[2]] + ggtitle("2") + lgnd_off,
                     plots.prob[[9]] + ggtitle("9") + lgnd_off,
                     plots.prob[[14]] + ggtitle("14") + lgnd_off,
                     plots.prob[[40]] + ggtitle("40") + lgnd_off,ncol=2),
-          lgnd, rel_widths = c(10,1), nrow=1)
-
-plots.prob <- lapply(1:44, function(dm) single_drive.all.4.Models(dm, FALSE))
-save(plots.nat, "ggplots-nat.RData")
+          lgnd, rel_widths = c(10,2), nrow=1)
+ggsave("../paper/fig/single-drive-4-Models-ex.pdf", width=12, height=10)
[…]

### 2017-08-28 paper/paper.tex
Edit to end of single-drive example
- Modify section titles

@@
 
-In Figure 3 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.
-The solid stepwise curve corresponds to the observed failures.
-Censored drives are not plotted.
-Standard error bands are again calculated using Greenwood's method \citep{green}.
-Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.
-The non-parametric and parametric model align up quite closely.
-Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.
-Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution.
-The discrepancy in the right tail of the distribution appears to be real.
-We don't provide a solution to this problem here.
-Certainly, if there are clearly more that two well-separated failure modes, the GLFP will not fit the data well.
-However, we consider it satisfactory if we can provide a good approximation to the observed lifetime distribution over the bulk of the product life.
-In practical terms, the right tail is of less importance for electronic goods, due to obsolescence.
+In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted Kaplan-Meier estimate with axes on the Weibull probability scale.
+Pointwise uncertainty bands associated with each estimator are included.
+While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours.
+The discrepancy in the right tail of the distribution suggests that the GLFP may overestimate the proportion failing in the right tail of the distribution.
+While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.
 
@@
 Solid line is a non-parametric estimate with pointwise 90\% uncertainty band in blue.}
-  \label{fig1}
+  \label{ex1-overlay}
 \end{figure}

### 2017-08-28 paper/paper.tex
Modify section names
A new title

@@
 {
-  \title{\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes}
+  \title{\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
   \author{Eric Mittman \\
@@
 
-\subsection{Single Drive-Model Analysis}
+\subsection{Analysis of a Single Drive-Model}
 \label{subsec:ex1}
@@
 
-For models 1-4, different sets of restrictions required different prior specifications, which were assigned as follows:
+For Models 1-4, different sets of restrictions required different prior specifications, which were assigned as follows:
 
@@
 
-\section{Stan code for model 4}
+\section{Stan code for Model 4}
 \label{sec:stan-code}

### 2017-09-24 paper/paper.tex
Paper and References Edits per Meeker Meeting on 9.20

@@
 In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model.
-This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to ``infant mortality'' and failure due to wearout.
+This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failures due to ``infant mortality'' and failures due to wearout.
 We fit the GLFP model using a hierarchical modeling approach.
@@
 The structure of the paper is as follows.
-In Section~\ref{sec:Background}, we introduce our proposed method and potential applications and examine previous related work in the literature.
+Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work in the literature.
+Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data.
 In Section~\ref{sec:Data}, we describe the Backblaze Hard Drive data set, which is used throughout as a motivating example.
-Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data.
 In section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations.
-In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set.
+Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard disk drive data.
 First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation.
@@
 For example, a computer system can break if the mother board, disc drive or power supply stop working.
-Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.
-The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep{nelson}.
+Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout of certain components.
+The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.
 A common assumption in series systems is the time to failure for each risk is statistically independent.
@@
 Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.
-When modeling computer system failures Reiser et al.
-assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \citep{reiser}.
-Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.
+When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.
+\cite{chan} labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.
 This helped identify parameters when using maximum likelihood (ML) estimation.
-Extending Chan and Meeker's analysis, Basu et al.
-performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure \citep{basu}.
-Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \citep{berger}.
-More recently, Ranjan et al.
-considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \citep{ranjan}.
+Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure.
+\cite{berger} introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions.
+More recently, \cite{ranjan} considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions.
 Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.
@@
 
-$$ P(T \le t) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$
+$$ \Pr(T \le t) = F(t) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$
 
@@
 
-\begin{equation} P(T \le t) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
+\begin{equation} \Pr(T \le t) = F(t) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
 = \Phi_{SEV}\left[\frac{\log (t) - \mu}{\sigma}\right],
@@
 \begin{equation}
-p(t) = \frac{1}{\sigma t}\exp\left\{\frac{\log(t) - \mu}{\sigma} - \exp \left[ \frac{\log(t) - \mu}{\sigma} \right]\right\}\\
-=\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
+f(t) =\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
 \end{equation}
 
-An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.
+An advantage conferred by writing the model in this way is that the log of the $p$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.
 Using the inverse cdf of the SEV distribution,
@@
 \end{equation}
-Because $\Phi(0)\approx 0.632$, we have $\log{\eta} = \mu  =t_{0.63}$.
+Because $\Phi(0)\approx 0.632$, we have $\log{\eta} = \mu  \cong t_{0.63}$.
 When there is more prior knowledge about a lower quantile, as is often the case with high-reliability products, using this parameterization allows the analyst to put prior distributions directly on those quantities.
 In addition, in the presence of censoring, $p$ can be chosen to make $\log(t_p)$ and $\sigma$ approximately independent.
-Since the choice of independent priors is common in practice, this observation lends further credence to the use of these reparameterizations.
+Since the choice of independent marginal priors is common in practice, this observation lends further credence to the use of these reparameterizations.
 
@@
 The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
-$$P(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
+$$\Pr(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
 
@@
 A common feature of lifetime data is right-censoring.
-Due to time or budget constraints it is rare all units are observed until failure.
-If a unit has not yet failed when the data is analyzed it is considered right-censored.
+In the analysis of reliability field data, it is rare all units are observed until failure.
+If a unit has not yet failed when the data are analyzed it is considered right-censored.
 In other words, right-censoring puts a lower bound on the failure time.
-Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit right-censoring.
 
@@
 
-As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.
-Some drive-models have been running since 2013 or before, while others were added at a later date.
+As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different Drive-models.
+Some Drive-models have been running since 2013 or before, while others were added at a later date.
 Data have been reported on 75,297 different hard drives that are or where in operation.
-The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
-Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.
+The number of drives varies by Drive-model; some Drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
+Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for Drive-models with at least 3 failures.
 For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.
@@
 \label{subsec:ex1}
-To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, drive-model 14.
+To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.
 The Kaplan-Meier estimate plotted below suggests at least two failure modes.
-The curve has a slow ramp up in failures till about 10,000 hours, increases from about 10,000 to 20,000 hours, and then starts to gradually flatten out.
+The curve has a slow ramp up in failures till about 10,000 hours (early failures), increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).
 In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.
@@
   \includegraphics[width=.8\textwidth]{km14-prob}
-  \caption{Kaplan Meier estimate for drive model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
+  \caption{Kaplan Meier estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
   \label{fig1}
@@
 To estimate the GLFP we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution.
-We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}$) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}$).
-Although non-standard, for purposes of exposition, we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.
+We reparameterize the component Weibull distributions using the $0.50$ quantile for the infant mortality failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$).
+\\
+
+When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the parameters of the distribution.
+Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.
 
@@
 
-These priors put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.
+These prior distributions put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.
 Thus, we consider the priors to be relatively uninformative.
 
-Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for drive model 14.
+Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.
 The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.
 As expected, this fraction is small with a median value of 0.05.
-The parameter estimates for the two Weibull distributions also match intuition.
-The wearout failure mode puts posterior probability on a values of $\sigma_1$ greater than 1, which corresponds to a decreasing hazard.
-Conversely, the credible interval for $\sigma_2$, the wearout mode, is strictly below 1, an increasing hazard.
+The parameter $\pi$ estimates for the two Weibull distributions also match intuition.
+The infant mortality failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.
+Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.
 The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.
@@
 $\pi$ & 0.03 & 0.05 & 0.10 \\ 
- $\sigma_1$ & 1.25 & 1.96 & 15.54 \\ 
-  $\sigma_2$ & 0.48 & 0.65 & 0.78 \\ 
+ $\beta_1$ & 0.47 & 1.13 & 1.72 \\ 
+  $\beta_2$ & 4.47 & 4.70 & 4.95 \\ 
   $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
@@
 \end{tabular}
-\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive Model 14.}
+\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.}
 \label{table:1}
@@
 After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.
-We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \citep{kaplan}.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a detailed description).
+We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf.
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \cite[Chapter 11]{meeker} (see Appendix for a detailed description).
 
@@
 While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours.
-The discrepancy in the right tail of the distribution suggests that the GLFP may overestimate the proportion failing in the right tail of the distribution.
+The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.
 While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.
@@
   \includegraphics[width=.8\textwidth]{km14-prob-plus}
-  \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.
+  \caption{Estimated GLFP for Drive-model 14 plotted on Weibull paper.
 Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.
@@
 \end{equation*}
+where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L}$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.
+All times are again in hours.
 
@@
 
-As discussed in section \ref{sec:Weibull parameterization}, we reparameterize the Weibull in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution.
+As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution.
 We truncate the distribution of $\sigma_{2g}$ at $1$, restricting the wearout failure mode to have an increasing hazard function.
@@
 A practical strategy is to begin with a simple model and gradually add complexity, reevaluating model adequacy at each iteration.
-In section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.
+In Section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.
 
@@
 \subsection{Modeling}
-We considered four models, all based on \ref{eq:hier-model}, which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):
+We considered four models, all based on (\ref{eq:hier-model}), which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):
 
@@
 Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by product.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by Drive-model.
 
-For all of the models we consider, the parameters for the infant mortality failure mode are held in common across products.
+For all of the models we consider, the parameters for the infant mortality failure mode are held in common across Drive-models.
 We found that there was often insufficient information to model these parameters hierarchically.
@@
 \label{sec:Prior Distributions}
-To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior.
+To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model.
+We select proper prior distributions to ensure a proper posterior distributions.
 
-For Models 1-4, different sets of restrictions required different prior specifications, which were assigned as follows:
+For Models 1-4, different sets of restrictions required different prior distribution specifications, which were assigned as follows:
 
-\paragraph{Model 1.} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same priors as in Example 1.
+\paragraph{Model 1.} Constrain all Drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.
 
-\paragraph{Model 2.} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
+\paragraph{Model 2.} $t_{p,g}$ varies by Drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
 $$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
@@
 
-\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for model 2.
+\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
 
-\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for model 2.
+\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
 
 Where applicable prior distributions on hyperparameters are as listed below.
-We follow the recommendation of \citet{gelman2014bayesian} to use half-Cauchy priors on hierarchical scale parameters.
+We follow the recommendation of \citet{gelman2014bayesian} to use half-Cauchy prior distributions on hierarchical scale parameters.
 As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
@@
 This is done to reduce autocorrelation and efficiently explore the posterior.
-Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1,2 and 3 and 16 chains were run for Model 4.
+Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 and 16 chains were run for Model 4.
 The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains.
 Upon convergence, $\hat{R}$, converges to 1.
-The number of model parameters for the four model fits are .
 The  maximum values of $\hat{R}$ across all model parameters for Models 1 throught 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004.
@@
 \label{sec:Model Comparisons}
-As we have discussed, while we prefer to allow all of the GLFP model parameters to be drive-model specific, there are practical issues with fitting the full model.
-For example, due to heavy left-truncation, there may be insufficient information for particular drive models to estimate all of the parameters.
+As we have discussed, while we prefer to allow all of the GLFP model parameters to be Drive-model specific, there are practical issues with fitting the full model.
+For example, due to heavy left-truncation, there may be insufficient information for particular Drive-models to estimate all of the parameters.
 
@@
 The large left panel shows the Kaplan-Meier nonparametric estimate of the same.
-the Kaplan-Meier estimates right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
+The Kaplan-Meier estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
 The smaller plots correspond to Models 1 through 4 reading from left to right, top to bottom.
@@
 \includegraphics[width=\textwidth]{heterogeneity-compare}
-\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
+\caption{Left: Kaplan-Meier estimates of the time to failure for each of the Drive-models in the Backblaze data.
 Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
@@
 
-We can also compare the four Model fits for each drive-model individually.
+We can also compare the four Model fits for each Drive-model individually.
 For these comparisons, we use the set of parameters which maximize the marginal posterior log probability.
@@
 \includegraphics[width=\textwidth]{single-drive-4-Models-ex}}
-\caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
+\caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of Drive-models, with axes scales as for Weibull probability plotting.
 The solid step function is the truncated adjusted Kaplan-Meier points.}
@@
 
-The GLFP fit for drive-models 2, 9, 14, and 40 are presented in Figure \ref{fig:mod_comp_leg}.
+The GLFP fit for a representative selection of Drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for all plots).
 The black step functions again correspond to the adjusted Kaplan-Meier estimates.
 As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.
-All of the GLFP models appear to produce lower estimates, relative to Kaplan-Meier, of the proportion of failed hard drives for drive-model 2, with Model 4 being in closest agreement.
-For drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for drive-model 40, they produce higher estimates.
+All of the GLFP models appear to produce lower estimates, relative to Kaplan-Meier, of the proportion of failed hard drives for Drive-model 2, with Model 4 being in closest agreement.
+For Drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for Drive-model 40, they produce higher estimates.
 On the other hand, Model 4 follows the nonparametric estimate quite closely across stages of lifetime.
 The differences between Models 3 and 4 and Models 1 and 2 are visually quite apparent.
-Differentiating between Model 3 and 4 is less clear; for instance, the estimates from Model 3 and 4 are nearly identical for drive-model 40.
+Differentiating between Model 3 and 4 is less clear; for instance, the estimates from Model 3 and 4 are nearly identical for Drive-model 40.
 
-From the point of view of an operator, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
-In this case, we should choose the Model that will provide the most accurate prediction for future units of previously observed drive-models.
+From the point of view of a consumer of hard disk drives, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
+In this case, we should choose the model that will provide the most accurate prediction for future units of previously observed Drive-models.
 One appropriate criterion for model selection, then, is log pointwise predictive density (lpd),
@@
 \begin{equation}
-elpd = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
+\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
 \label{elpd}
@@
 \begin{equation}
-\widehat{elpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
+\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
 \label{loo-est}
 \end{equation}
-A recently published \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
+The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
 It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\frac{\sum_{i=1}^S w_{i,s} p(t_{g,i})}{\sum_{s=1}^S w_{i,s}}$.
@@
 Results showing estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:1}.
- We calculated the difference in expected predictive accuracy for Model 4 vs 3, Model 3 vs 2, and Model 2 vs 1, as well as the standard error of the difference (Table 1).  
-For the Models we considered, each successive increase in model complexity increases elpd.  Of all the models, Model 4 has significantly better elpd compared to the the other 3 Models.
+ We calculated the difference in expected predictive accuracy for Model 4 versus 3, Model 3 versus 2, and Model 2 versus 1, as well as the standard error of the difference (Table 1).  
+For the models we considered, each successive increase in model complexity increases elpd.  Of all the models, Model 4 has significantly better elpd compared to the the other 3 models.
 \begin{figure}
@@
   \hline
- & $\widehat{elpd}$ \ & Difference in $\widehat{elpd}$ \ & SE of the Difference \\ 
+ & $\widehat{\text{elpd}}$ \ & Difference in $\widehat{\text{elpd}}$ \ & SE of the Difference \\ 
   \hline
@@
 \subsubsection{Model Assessment}
-Having chosen Model 4 as the best of the candidate models we considered, we would like to assess whether our model provides an adequate fit to the observed data.
+Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate fit to the observed data.
 One way to do this is to draw replicate data sets from the posterior predictive distribution.
@@
 In the hierarchical setting, there can be different formulations of the predictive distribution.
-In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
-We choose to focus on the populations for which we have data collected.
+In the present context, one could predict the lifetime of a new unit of a previously observed Drive-model or a new Drive-model.
+We choose to focus on the former.
 There are many possible approaches to the posterior predictive check.
-For this problem, we chose to consider the posterior predictive distribution of the Kaplan-Meier estimate for each drive-model $g$.
+For this problem, we chose to consider the posterior predictive distribution of the Kaplan-Meier estimate for each Drive-model $g$.
 
@@
 $$p(t_{rep,g}) = \int \prod_{i=1}^{n_g}f(t_{new}|t_{g,i}^L,c_{g,i},\theta)\,p(\theta|t)d\theta.$$
-Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing $\theta^{(s)}$ at random, then conditionally draw $t_{rep,g}$ from $f(t_{new}|t_{g,i}^L,c_{g,i},\theta^{(s)}.$
+Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing $\theta^{(s)}$ at random, then conditionally draw $t_{rep,g}$ from $f(t_{new}|t_{g,i}^L,c_{g,i},\theta^{(s)}).$
 
 An obstacle remains, in that the censoring mechanism itself is not fully observed.
-Our admittedly ad hoc approach is to assume that failed units would have been the censored had they survived until the latest censoring or failure time for that drive model.
+Our admittedly simple approach is to assume that failed units would have been the censored had they survived until the latest censoring or failure time for that drive model.
 We chose to generate 19 replicates for each drive model.
 This gives a 1/20 chance under the hypothesis of no difference that the Kaplan-Meier estimate based on the observed data would be ``most extreme" relative all the replicated  from the those one and is not too many so as to produce an overly cluttered plot.
-Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each drive-model $g=1,\ldots,44$, is as follows:
+Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each Drive-model $g=1,\ldots,44$, is as follows:
 \begin{enumerate}
-\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.
-Where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right) $, is a random draw from the full posterior distribution.
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right) $, is a random draw from the full posterior distribution.
 \item Repeat step 1 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
@@
 \includegraphics[width=\textwidth]{ppcheck-sample}
-\caption{Nonparametric estimates of proportion failing for each drive-model.
+\caption{Nonparametric estimates of proportion failing for a representative subset of Drive-models.
 Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
@@
 We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the Kaplan-Meier nonparametric estimator.
-The resulting plots are displayed in figure \ref{fig:post-pred-KM}.
-For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.
-Only a representative sample of plots are shown here for brevity.
-There are a few drive-models, however where this is not the case.
-As we addressed previously, in Subsection \ref{subsec:ex1}, drive-model 14 has a discrepancy between the GFLP and the Kaplan-Meier fits in the right tail.
-Due to the large number of failures for this drive-model, the posterior for the hierarchical model is very similar to the stand-alone model, so all the same comments made earlier still apply.
-In addition, drive-models 4, 11 and 34 appear to be exceptions as well.
-We will focus our attention on drive-model 4, but our diagnosis of 34 is similar, so the following comments extend to that drive-model as well.
+The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
+Only a representative sample of plots are shown here for brevity (see the Supplementary Material for the complete set of Drive-models).
+For most Drive-models, the estimates based on the observed data look similar to those based on the replicated data.
+There are a few Drive-models, however where this is not the case.
+As we addressed previously, in Subsection \ref{subsec:ex1}, Drive-model 14 has a discrepancy between the GFLP and the Kaplan-Meier fits in the right tail.
+Due to the large number of failures for this Drive-model, the posterior for the hierarchical model is very similar to the stand-alone model, so all the same comments made earlier still apply.
+In addition, Drive-models 4, 11 and 34 appear to be exceptions as well.
+We will focus our attention on Drive-model 4, but our diagnosis of 34 is similar, so the following comments extend to that Drive-model as well.
 
-To attempt to diagnose the cause of the discrepancies between the observed data for drive-model 4 and its posterior predictive distribution, we plot an event plot for a random sample of units for this drive-model, along with the Kaplan-Meier estimate, and the posterior estimate of time to failure (see figure \ref{fig:ex-mod-4}).
+To attempt to diagnose the cause of the discrepancies between the observed data for Drive-model 4 and its posterior predictive distribution, we plot an event plot for a random sample of units for this Drive-model, along with the Kaplan-Meier estimate, and the posterior estimate of time to failure (see Figure \ref{fig:ex-mod-4}).
 From these plots, it appears that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the several thousand hours, at a time where very few units were observed to be at risk.
@@
 While it might be tempting to treat those early failures as anomalous, they also point to a weakness in our model.
-By virtue of having all drive-models share in common the Weibull parameters for the infant mortality failure mode, an early rash of failures indicating a subpopulation with, say, an increasing hazard function,  would be problematic.
+By virtue of having all Drive-models share in common the Weibull parameters for the infant mortality failure mode, an early rash of failures indicating a subpopulation with, say, an increasing hazard function,  would be problematic.
 While the proportion at risk of early failure can vary via $\pi_g$, the plausible values of the common parameters, $\mu_1$ and $\sigma_1$, do not ``line up" in a way so to influence the posterior as they might have had they been allowed to vary.
@@
 
-As another example, to further appreciate how information is being borrowed across groups, consider drive-model 43 (Figure \ref{fig:ex-mod-43}).
-The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available.
-In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
-Then, in the period where records exist for this drive model, at around 10,000 hours, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.
+As another example, to further appreciate how information is being borrowed across groups, consider Drive-model 43 (Figure \ref{fig:ex-mod-43}).
+The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available.
+In Figure \ref{fig:ex-mod-43} we see that the lack of data for this Drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
+Then, in the period where records exist for this Drive-model, at around 10,000 hours, the posterior deviates from the global model, achieving a compromise between it and the nonparametric estimate.
 
@@
 \includegraphics{dm4-exception}
-\caption{\small Top: Event plot for uniform random sample of 60 units of drive-model 4.
+\caption{\small Top: Event plot for uniform random sample of 60 units of Drive-model 4.
 Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_4,i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
@@
 \includegraphics{dm43-shrinkage}
-\caption{\small Top: Event plot for all units of drive-model 4.
+\caption{\small Top: Event plot for all units of Drive-model 43.
 Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_{43},i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
 $t_{43,i}\mbox{, where }c_{43,i}=0$.\hspace{\textwidth}Bottom: The solid step function correponds to the adjusted Kaplan-Meier estimate.
-Due to the absence of collected information about early life for this particular model,due to left-truncation, the posterior median (dashed line) follows the ``global model" (dotted line), i.e.
-the posterior median of proportion failing for a new drive-model, approximately until the beginning of the earliest observational period.}
+Due to the absence of collected information about early life for this particular model, due to left-truncation, the posterior median (dashed line) follows the ``global model" (dotted line), i.e.
+the posterior median of proportion failing for a new Drive-model, approximately until the beginning of the earliest observational period.}
 \label{fig:ex-mod-43}
@@
 \label{subsubsec:Drive-model Comparisons}
-We consider the problem of ranking drive-models. From a
+We consider the problem of ranking Drive-models. From a
 business perspective, it is clear
-that we should prefer the drive-models which will provide more years
+that we should prefer the Drive-models which will provide more years
 of service. For ease of exposition, we will assume that the purchase
-price of a hard-drive is the same across drive-models.
+price of a hard-drive is the same across Drive-models.
 
@@
 \begin{equation*}
-  p(t_{d,new}|t) = \int p(t_{d,new}|\theta_d) p(\theta_d|t) \, d\theta_d
+  p(t_{d,new}|t) = \int p(t_{d,new}|\theta_d) p(\theta_d|t_d) \, d\theta_d
 \end{equation*}
@@
 $s=1,...,S$, using the saved posterior draws for the model
-parameters. The expected time-to-failure (TTF) for drive-model $d$ can be estimated by
+parameters. The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
 $$\frac{1}{S} \sum_{s=1}^{S} t_{d,new}^{(s)}.$$
 
-While TTF is clearly important, due to the
+While MTTF is clearly important, due to the
 anticipation of advancements in technology, we can expect that the
 relative value of computer hardware will depreciate rather quickly.
-In simple terms, given two hard drives, I would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all.
+In simple terms, given two hard drives, one would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all.
 To account
-for depreciation, rather than using TTF as the metric for drive-model comparison, we use the value at replacement. The IRS considers
+for depreciation, rather than using MTTF as the metric for Drive-model comparison, we use the value at replacement. The US Internal Revenue Service considers
 computer hardware to be ``5-year'' equipment \citep{f4562}. Using``the declining
@@
 
-Let $L(t) = e^{-.4t}$ represent the value at the time of failure
-relative to a new unit. We can rank the drive-models by 
+Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
+relative to a new unit. We can rank the Drive-models by 
 $$\op{E}(L(t_{d,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{d,new}^{(s)}).$$
 
-Posterior credible intervals for $L(t_{d,new})$ are shown in the left panel of figure \ref{fig5}.
-A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{d,new})|y)$ are shown in the bottom right panel.
-The panel above shows credible intervals for TTF.
-We can see that ranking by TTF would lead to a different result -- a new drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher that for the drive-models ranked highest by our depreciation-aware metric.
-The top four drive-models (4,5,2,3) seem very comparable, with a new drive-model 4 unit expected to depreciate the most.
-However, given the discrepancy of the model fit for drive-model 4, noted in the previous subsection, in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this drive-model.
+Posterior credible intervals for $L(t_{d,new})$ are shown in the left panel of Figure \ref{fig5}.
+A zoomed version of the top 8 Drive-models, as ranked by $\op{E}(L(t_{d,new})|y)$ are shown in the bottom right panel.
+The panel above shows credible intervals for MTTF.
+We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher than for the Drive-models ranked highest by our depreciation-aware metric.
+The top four Drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
+However, given the discrepancy of the model fit for Drive-model 4, noted in the previous subsection, in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.
 
-We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a central measure of tendency.
-On the left panel of Figure 5 we plot $t_{0.10}$, the .10 quantile (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
-The best six drive-models ranked by TTF are also the drive-models with the longest time to $t_{0.10}$.
-Drive-models 5, 2, 4, and 3 seem to really stick out as the most reliable as there is a larger gap separating them from the rest of the group.
+We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a measure of central tendency.
+The left panel of Figure 5 shows $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all Drive-models.
+The best six Drive-models ranked by MTTF are also the Drive-models with the largest $t_{0.10}$.
+Drive-models 5, 2, 4, and 3 seem to really stand out as the most reliable as there is a larger gap separating them from the rest of the group.
 Another interesting feature to compare is $\pi$, the proportion defective.
-One the right panel of Figure 5 are the posterior credible intervals of $\pi$ for each drive-model.
+The right panel of Figure 5 shows the posterior credible intervals of $\pi$ for each Drive-model.
 The plot is again sorted according to $t_{0.10}$ so it can be compared to the left hand side.
-While for many of the drive-models the ordinal ranking is the same as on the left, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
+While for many of the Drive-models the ordinal ranking is the same as on the left, there are some Drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
 Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.
@@
   \includegraphics[width=.8\textwidth]{dm-eval}
-  \caption{Left: Summary of posterior predictive distributions for $e^{-.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each drive-model.
-  Right: A comparison of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
+  \caption{Left: Summary of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
+  Right: A comparison of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 Drive-models, shows that rankings produced by the two measures are not equivalent.}
   \label{fig5}
@@
 We do not account for batch effects or varying conditions in the facility which have impacted the observed failure rates.
-These factors are confounded with drive-model in our analysis.
+These factors are confounded with Drive-model in our analysis.
 This second issue is a more fundamental issue with observational data like the hard drive data we have analyzed here.
@@
 Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics meaningful in the context of the data.
+
+Besides comparing different Drive-models for purchasing decisions, another important application would be prediction of future failures (say, week-by-week) to assure that there is always a sufficient number of replacement drives available.
+Methods similar to those in \cite{hmm} can be adapted to Bayesian prediction, as described in Section 15.5 of \cite{intervals}.
 
@@
 \label{sec:trunc-adj}
-We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
-\citep{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
+We first start with a non-parametric estimate of the empirical cdf for each Drive-model using the Kaplan-Meier estimator.
+With left truncation, however, the standard Kaplan-Meier estimator for Drive-model $d$, denoted by
 $\widehat{F_d(t)}_{KM}$, is conditional on survival up to
@@
 of drive-model $d$ for which records are available. To produce
-unconditional estimates, we adapt the adjustment method outlined by by Meeker and Escobar  in Chapter 11 of Reliability Analysis \citep{meeker}.  For each drive-model we select
+unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each Drive-model we select
 $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
@@
 parameters, we can easily compute its posterior median,
-$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by
+$\widehat{A}_{\text{med}} = \widehat{\Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by
 
@@
 
-\pagebreak
+\section{Supplementary Material}
+\label{sec:supple}
+
+\begin{description}
+
+\item[Stan Code] R Stan code for Model 4 (stan file)
+
+\item[Graphs of Model Comparison Fits for all Drive-models] Posterior median of proportion failing as a function of time under Models 1-4 for all Drive-models. (pdf file)
+
+\item[Posterior Predictive Graphical Checks for all Drive-Models] Nonparametric estimates of proportion failing for original and replicated data sets for all Drive-models. (pdf file)
+
+\end{description}
 
@@
 % \end{description}
-
-\section{Stan code for Model 4}
-\label{sec:stan-code}
-{\scriptsize
-\begin{verbatim}
-functions {
-  real sev_logpdf(real y, real mu, real sigma){
-    real z;
-    z = (y - mu) / sigma;
-    return -log(sigma) + z - exp(z);
-  }
-
-  real sev_logccdf(real y, real mu, real sigma){
-    return -exp((y - mu) / sigma);
-  }
-
-  real sev_logcdf(real y, real mu, real sigma){
-    return log1m_exp(-exp((y - mu) / sigma));
-  }
-}
-
-data {
-  int M;
-  int N_obs;
-  int N_cens;
-  real endtime_obs[N_obs];
-  real endtime_cens[N_cens];
-  real starttime_obs[N_obs];
-  real starttime_cens[N_cens];
-  int<lower=1, upper=M> dm_obs[N_obs];
-  int<lower=1, upper=M> dm_cens[N_cens];
-  vector<lower=0, upper=1>[2] p; # quantiles to model
-}
-transformed data{
-  vector[2] z_corr;
-  for(i in 1:2)
-    z_corr[i] = log(-1.0 * log1m(p[i])); # used to get location(mu) from quantile(t_p)
-}
-parameters{
-  real log_tp1;
-  real<lower=0> sigma1;
-  real eta_tp2;
-  real eta_s2;
-  real eta_pi;
-  real<lower=0> tau_tp2;
-  real<lower=0> tau_s2;
-  real<lower=0> tau_pi;
-  vector[M] log_tp2_raw;
-  real<lower=0, upper=1> sigma2[M];
-  vector[M] logit_pi_raw;
-}
-
-transformed parameters{
-  real mu1;
-  vector[M] mu2;
-  vector[M] log_pi;
-  mu1 = log_tp1 - sigma1 * z_corr[1];
-  for(m in 1:M){
-    mu2[m] = (eta_tp2 + tau_tp2 * log_tp2_raw[m]) - (sigma2[m] * z_corr[2]);
-  }
-  for(m in 1:M)
-    log_pi[m] = log_inv_logit(eta_pi + tau_pi * logit_pi_raw[m]);
-}
-
-model{
-  real tmp[2];
-  int m;
-  real logpi;
-  real mu_1;
-  real mu_2;
-  real sig_1;
-  real sig_2;
-
-  for(i in 1:N_obs){
-    m = dm_obs[i];
-    logpi = log_pi[m];
-    mu_2 = mu2[m];
-    sig_2 = sigma2[m];
-    // numerator:   log( p * f1 * (1 - F2) + f2 * (1 - p * F1) )
-    //            = log( exp(log(p) + log(f1) + log(1 - F2)) + exp(log(f2) + log(1 - exp(log(p) + log(F1)))) )
-    tmp[1] = log_sum_exp(logpi + sev_logpdf(endtime_obs[i], mu1, sigma1) +
-               sev_logccdf(endtime_obs[i], mu_2, sig_2),
-               sev_logpdf( endtime_obs[i], mu_2, sig_2) + 
-               log1m_exp(logpi + sev_logcdf(endtime_obs[i], mu1, sigma1)
-               )
-             );
-    // denominator:  log((1 - p * F1) * (1 - F2))
-    //            =  log(1 - p * F1) + log(1 - F2)
-    tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_obs[i], mu1, sigma1)) + 
-             sev_logccdf(starttime_obs[i], mu_2, sig_2);
-
-    target += tmp[1] - tmp[2];
-  }
-
-  for(i in 1:N_cens){
-    m = dm_cens[i];
-    logpi = log_pi[m];
-    mu_2 = mu2[m];
-    sig_2 = sigma2[m];
-
-    // numerator:   log((1 - p * F1) * (1 - F2))
-    //            =  log(1 - p * F1) + log(1 - F2)
-    tmp[1] = log1m_exp(logpi + sev_logcdf(endtime_cens[i], mu1, sigma1)) + 
-             sev_logccdf(endtime_cens[i], mu_2, sig_2);
-    // denominator:  log((1 - p * F1) * (1 - F2))
-    //            =  log(1 - p * F1) + log(1 - F2)
-    tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_cens[i], mu1, sigma1)) + 
-             sev_logccdf(starttime_cens[i], mu_2, sig_2);
-
-    target += tmp[1] - tmp[2];
-  }
-
-  log_tp1        ~ normal(7, 2);
-  log_tp2_raw    ~ normal(0, 1);
-  sigma1         ~ lognormal(0, 1);
-  sigma2         ~ lognormal(eta_s2, tau_s2);
-  logit_pi_raw   ~ normal(0, 1);
-  eta_tp2        ~ normal(9, 2);
-  eta_s2         ~ normal(0, 2);
-  eta_pi         ~ normal(-3, 1);
-  tau_tp2        ~ cauchy(0,1);
-  tau_s2         ~ cauchy(0,1);
-  tau_pi         ~ cauchy(0,1);
-}
-
-\end{verbatim}
-}
-
-\pagebreak
-\section{Posterior predictive graphical checks}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-1.pdf}
-\end{figure}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-2}
-\end{figure}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-3.pdf}
-\end{figure}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-4.pdf}
-\end{figure}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-5.pdf}
-\end{figure}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-6.pdf}
-\end{figure}
-\clearpage
+%
+%\section{Stan code for Model 4}
+%\label{sec:stan-code}
+%{\scriptsize
+%\begin{verbatim}
+%functions {
+  %real sev_logpdf(real y, real mu, real sigma){
+    %real z;
+    %z = (y - mu) / sigma;
+    %return -log(sigma) + z - exp(z);
+  %}
+  %
+  %real sev_logccdf(real y, real mu, real sigma){
+    %return -exp((y - mu) / sigma);
+  %}
+  %
+  %real sev_logcdf(real y, real mu, real sigma){
+    %return log1m_exp(-exp((y - mu) / sigma));
+  %}
+%}
+%
+%data {
+  %int M;
+  %int N_obs;
+  %int N_cens;
+  %real endtime_obs[N_obs];
+  %real endtime_cens[N_cens];
+  %real starttime_obs[N_obs];
+  %real starttime_cens[N_cens];
+  %int<lower=1, upper=M> dm_obs[N_obs];
+  %int<lower=1, upper=M> dm_cens[N_cens];
+  %vector<lower=0, upper=1>[2] p; # quantiles to model
+%}
+%transformed data{
+  %vector[2] z_corr;
+  %for(i in 1:2)
+    %z_corr[i] = log(-1.0 * log1m(p[i])); # used to get location(mu) from quantile(t_p)
+%}
+%parameters{
+  %real log_tp1;
+  %real<lower=0> sigma1;
+  %real eta_tp2;
+  %real eta_s2;
+  %real eta_pi;
+  %real<lower=0> tau_tp2;
+  %real<lower=0> tau_s2;
+  %real<lower=0> tau_pi;
+  %vector[M] log_tp2_raw;
+  %real<lower=0, upper=1> sigma2[M];
+  %vector[M] logit_pi_raw;
+%}
+%
+%transformed parameters{
+  %real mu1;
+  %vector[M] mu2;
+  %vector[M] log_pi;
+  %mu1 = log_tp1 - sigma1 * z_corr[1];
+  %for(m in 1:M){
+    %mu2[m] = (eta_tp2 + tau_tp2 * log_tp2_raw[m]) - (sigma2[m] * z_corr[2]);
+  %}
+  %for(m in 1:M)
+    %log_pi[m] = log_inv_logit(eta_pi + tau_pi * logit_pi_raw[m]);
+%}
+%
+%model{
+  %real tmp[2];
+  %int m;
+  %real logpi;
+  %real mu_1;
+  %real mu_2;
+  %real sig_1;
+  %real sig_2;
+  %
+  %for(i in 1:N_obs){
+    %m = dm_obs[i];
+    %logpi = log_pi[m];
+    %mu_2 = mu2[m];
+    %sig_2 = sigma2[m];
+    %// numerator:   log( p * f1 * (1 - F2) + f2 * (1 - p * F1) )
+    %//            = log( exp(log(p) + log(f1) + log(1 - F2)) + exp(log(f2) + log(1 - exp(log(p) + log(F1)))) )
+    %tmp[1] = log_sum_exp(logpi + sev_logpdf(endtime_obs[i], mu1, sigma1) +
+               %sev_logccdf(endtime_obs[i], mu_2, sig_2),
+               %sev_logpdf( endtime_obs[i], mu_2, sig_2) + 
+               %log1m_exp(logpi + sev_logcdf(endtime_obs[i], mu1, sigma1)
+               %)
+             %);
+    %// denominator:  log((1 - p * F1) * (1 - F2))
+    %//            =  log(1 - p * F1) + log(1 - F2)
+    %tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_obs[i], mu1, sigma1)) + 
+             %sev_logccdf(starttime_obs[i], mu_2, sig_2);
+             %
+    %target += tmp[1] - tmp[2];
+  %}
+  %
+  %for(i in 1:N_cens){
+    %m = dm_cens[i];
+    %logpi = log_pi[m];
+    %mu_2 = mu2[m];
+    %sig_2 = sigma2[m];
+  %
+    %// numerator:   log((1 - p * F1) * (1 - F2))
+    %//            =  log(1 - p * F1) + log(1 - F2)
+    %tmp[1] = log1m_exp(logpi + sev_logcdf(endtime_cens[i], mu1, sigma1)) + 
+             %sev_logccdf(endtime_cens[i], mu_2, sig_2);
+    %// denominator:  log((1 - p * F1) * (1 - F2))
+    %//            =  log(1 - p * F1) + log(1 - F2)
+    %tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_cens[i], mu1, sigma1)) + 
+             %sev_logccdf(starttime_cens[i], mu_2, sig_2);
+             %
+    %target += tmp[1] - tmp[2];
+  %}
+  %
+  %log_tp1        ~ normal(7, 2);
+  %log_tp2_raw    ~ normal(0, 1);
+  %sigma1         ~ lognormal(0, 1);
+  %sigma2         ~ lognormal(eta_s2, tau_s2);
+  %logit_pi_raw   ~ normal(0, 1);
+  %eta_tp2        ~ normal(9, 2);
+  %eta_s2         ~ normal(0, 2);
+  %eta_pi         ~ normal(-3, 1);
+  %tau_tp2        ~ cauchy(0,1);
+  %tau_s2         ~ cauchy(0,1);
+  %tau_pi         ~ cauchy(0,1);
+%}
+%
+%\end{verbatim}
+%}
+%
+%\pagebreak
+%\section{Posterior predictive graphical checks}
+%\begin{figure}[H]
+%\includegraphics[width=\textwidth]{ppcheck-1.pdf}
+%\end{figure}
+%\begin{figure}[H]
+%\includegraphics[width=\textwidth]{ppcheck-2}
+%\end{figure}
+%\begin{figure}[H]
+%\includegraphics[width=\textwidth]{ppcheck-3.pdf}
+%\end{figure}
+%\begin{figure}[H]
+%\includegraphics[width=\textwidth]{ppcheck-4.pdf}
+%\end{figure}
+%\begin{figure}[H]
+%\includegraphics[width=\textwidth]{ppcheck-5.pdf}
+%\end{figure}
+%\begin{figure}[H]
+%\includegraphics[width=\textwidth]{ppcheck-6.pdf}
+%\end{figure}
+%\clearpage
 %\bibliographystyle{plainnat}

### 2017-09-25 paper/paper.tex
A few more edits and Figure 7 update.

@@
 \usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf,float, bm}
+\usepackage{graphicx,psfrag,epsf,float, bm, subcaption}
 \graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
@@
 However, in many situations the cause of failure is unknown to the analyst.
-This is referred to in the literature as ``masking''.\\
+This is referred to in the literature as masking.\\
 
@@
 While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours.
-The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.
+%The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.
 While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.
@@
 Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by Drive-model.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by Drive-model.
 
@@
 We can also compare the four Model fits for each Drive-model individually.
-For these comparisons, we use the set of parameters which maximize the marginal posterior log probability.
-Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.
+For these comparisons, we use the posterior median of the MCMC draws.
+%Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.
 
@@
 \centering
-  \includegraphics[width=\textwidth]{b10andpi}
-  \caption{Left: 50\% credible intervals for  $t_{0.10}$ (time in years till 10\% of the drives fail).
+   \begin{subfigure}[b]{0.80\textwidth}
+   \includegraphics[width=1\linewidth]{b10new.pdf}
+   \caption{50\% CI for $t_{0.10}$}
+   \label{fig:Ng1} 
+\end{subfigure}
+
+\begin{subfigure}[b]{0.80\textwidth}
+   \includegraphics[width=1\linewidth]{pinew.pdf}
+   \caption{50\% CI for $\pi$}
+   \label{fig:Ng2}
+\end{subfigure}
+
+\caption{(a) 50\% CI (in years) for $t_{0.10}$ (b) 50\% CI for $\pi$ plotted on the logit scale.  Both plots are sorted based on the median value of $t_{0.10}$.}
+\end{figure}
+
+%\begin{figure}[H]
+  %\centering
+  %\includegraphics[width=\textwidth]{b10andpi}
+  %\caption{Left: 50\% credible intervals for  $t_{0.10}$ (time in years till 10\% of the drives fail).
 Right: 50\% credible intervals for $\pi$ plotted on the logit scale.
 Both plots are sorted based on the median value of $t_{0.10}$.}
-  \label{fig4}
-\end{figure}
+  %\label{fig4}
+%\end{figure}
 
@@
 
-\item[Stan Code] R Stan code for Model 4 (stan file)
+\item[Stan Code] Stan code for Model 4 (pdf file)
 
@@
 
-\item[Posterior Predictive Graphical Checks for all Drive-Models] Nonparametric estimates of proportion failing for original and replicated data sets for all Drive-models. (pdf file)
+\item[Posterior Predictive Graphical Checks for all Drive-models] Nonparametric estimates of proportion failing for original and replicated data sets for all Drive-models. (pdf file)
 

### 2017-09-25 plots/compare_fits.R
small edits to b10 and pi plots.

@@ -158,10 +158,10 @@ b10$model <- factor(b10$model,levels=b10$model[order(b10$med)])  #Sort By Lower 
 #Make Catepillar Plot for B10; Perhaps Sort by Sample Size?
 b10.plot <- ggplot(b10, aes(x=model, y=med, ymin=lb, ymax=ub)) +
   geom_pointrange() +  
-  coord_cartesian(ylim = c(0, 8)) + 
-  xlab('Drive Model') + theme_bw() + 
+  coord_cartesian(xlim = c(0, 8)) + 
+  xlab('Drive Model') + theme_bw(base_size = 14) + 
   ylab("Years")   +
-  scale_x_discrete(breaks=seq(1,44,1))
+  scale_x_discrete(breaks=seq(1,44,1)) + coord_flip(ylim = c(0,8))
 b10.plot
 
 
@@ -197,10 +197,10 @@ pi.dat$model <- factor(pi.dat$model,levels=pi.dat$model[order(pi.dat$bmed)])  #S
 #Make Catepillar Plot for B10; Perhaps Sort by Sample Size?
 pi.plot <- ggplot(pi.dat, aes(x=as.factor(model), y=med, ymin=lower, ymax=upper)) +
   geom_pointrange() +  
-  xlab('Drive Model') + theme_bw() + 
+  xlab('Drive Model') + theme_bw(base_size = 14) + 
   ylab(expression(pi))   +
-  scale_x_discrete(breaks=seq(1,44,1))+
-  scale_y_continuous(trans="logit", breaks=c(.01, .02, .05, .1, .2, .4, .6, .8, .9))
+  scale_x_discrete(breaks=seq(1,44,1)) +
+  scale_y_continuous(trans="logit", breaks=c(.01, .02, .05, .1, .2, .4, .6, .8, .9)) + coord_flip()
 pi.plot
 
 #Make Cow Plot

### 2017-09-27 paper/paper.tex
Add separate supplementary section with plots
- change figure 6 to corresponding Drive-models of figure 5

- fix labeling clashes

@@
 Axes are on the log scale.}
-  \label{fig1}
+  \label{drive-scatter}
 \end{figure}
@@
 \end{equation*}
-where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L}$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.
+where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.
 All times are again in hours.
@@
 % The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).  
-Results showing estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:1}.
+Results showing estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:2}.
  We calculated the difference in expected predictive accuracy for Model 4 versus 3, Model 3 versus 2, and Model 2 versus 1, as well as the standard error of the difference (Table 1).  
@@
 The estimated difference of expected leave-one-out prediction errors between the two models, as well as the standard error of the difference, is also presented.}
-\label{table:1}
+\label{table:2}
 \end{table}
@@
 
-% \begin{center}
-% {\large\bf SUPPLEMENTARY MATERIAL}
-% \end{center}
-% 
-% \begin{description}
-% 
-% \item Put R Stan code here
-% 
-% \end{description}
-%
-%\section{Stan code for Model 4}
-%\label{sec:stan-code}
-%{\scriptsize
-%\begin{verbatim}
-%functions {
-  %real sev_logpdf(real y, real mu, real sigma){
-    %real z;
-    %z = (y - mu) / sigma;
-    %return -log(sigma) + z - exp(z);
-  %}
-  %
-  %real sev_logccdf(real y, real mu, real sigma){
-    %return -exp((y - mu) / sigma);
-  %}
-  %
-  %real sev_logcdf(real y, real mu, real sigma){
-    %return log1m_exp(-exp((y - mu) / sigma));
-  %}
-%}
-%
-%data {
-  %int M;
-  %int N_obs;
-  %int N_cens;
-  %real endtime_obs[N_obs];
-  %real endtime_cens[N_cens];
-  %real starttime_obs[N_obs];
-  %real starttime_cens[N_cens];
-  %int<lower=1, upper=M> dm_obs[N_obs];
-  %int<lower=1, upper=M> dm_cens[N_cens];
-  %vector<lower=0, upper=1>[2] p; # quantiles to model
-%}
-%transformed data{
-  %vector[2] z_corr;
-  %for(i in 1:2)
-    %z_corr[i] = log(-1.0 * log1m(p[i])); # used to get location(mu) from quantile(t_p)
-%}
-%parameters{
-  %real log_tp1;
-  %real<lower=0> sigma1;
-  %real eta_tp2;
-  %real eta_s2;
-  %real eta_pi;
-  %real<lower=0> tau_tp2;
-  %real<lower=0> tau_s2;
-  %real<lower=0> tau_pi;
-  %vector[M] log_tp2_raw;
-  %real<lower=0, upper=1> sigma2[M];
-  %vector[M] logit_pi_raw;
-%}
-%
-%transformed parameters{
-  %real mu1;
-  %vector[M] mu2;
-  %vector[M] log_pi;
-  %mu1 = log_tp1 - sigma1 * z_corr[1];
-  %for(m in 1:M){
-    %mu2[m] = (eta_tp2 + tau_tp2 * log_tp2_raw[m]) - (sigma2[m] * z_corr[2]);
-  %}
-  %for(m in 1:M)
-    %log_pi[m] = log_inv_logit(eta_pi + tau_pi * logit_pi_raw[m]);
-%}
-%
-%model{
-  %real tmp[2];
-  %int m;
-  %real logpi;
-  %real mu_1;
-  %real mu_2;
-  %real sig_1;
-  %real sig_2;
-  %
-  %for(i in 1:N_obs){
-    %m = dm_obs[i];
-    %logpi = log_pi[m];
-    %mu_2 = mu2[m];
-    %sig_2 = sigma2[m];
-    %// numerator:   log( p * f1 * (1 - F2) + f2 * (1 - p * F1) )
-    %//            = log( exp(log(p) + log(f1) + log(1 - F2)) + exp(log(f2) + log(1 - exp(log(p) + log(F1)))) )
-    %tmp[1] = log_sum_exp(logpi + sev_logpdf(endtime_obs[i], mu1, sigma1) +
-               %sev_logccdf(endtime_obs[i], mu_2, sig_2),
-               %sev_logpdf( endtime_obs[i], mu_2, sig_2) + 
-               %log1m_exp(logpi + sev_logcdf(endtime_obs[i], mu1, sigma1)
-               %)
-             %);
-    %// denominator:  log((1 - p * F1) * (1 - F2))
-    %//            =  log(1 - p * F1) + log(1 - F2)
-    %tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_obs[i], mu1, sigma1)) + 
-             %sev_logccdf(starttime_obs[i], mu_2, sig_2);
-             %
-    %target += tmp[1] - tmp[2];
-  %}
-  %
-  %for(i in 1:N_cens){
-    %m = dm_cens[i];
-    %logpi = log_pi[m];
-    %mu_2 = mu2[m];
-    %sig_2 = sigma2[m];
-  %
-    %// numerator:   log((1 - p * F1) * (1 - F2))
-    %//            =  log(1 - p * F1) + log(1 - F2)
-    %tmp[1] = log1m_exp(logpi + sev_logcdf(endtime_cens[i], mu1, sigma1)) + 
-             %sev_logccdf(endtime_cens[i], mu_2, sig_2);
-    %// denominator:  log((1 - p * F1) * (1 - F2))
-    %//            =  log(1 - p * F1) + log(1 - F2)
-    %tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_cens[i], mu1, sigma1)) + 
-             %sev_logccdf(starttime_cens[i], mu_2, sig_2);
-             %
-    %target += tmp[1] - tmp[2];
-  %}
-  %
-  %log_tp1        ~ normal(7, 2);
-  %log_tp2_raw    ~ normal(0, 1);
-  %sigma1         ~ lognormal(0, 1);
-  %sigma2         ~ lognormal(eta_s2, tau_s2);
-  %logit_pi_raw   ~ normal(0, 1);
-  %eta_tp2        ~ normal(9, 2);
-  %eta_s2         ~ normal(0, 2);
-  %eta_pi         ~ normal(-3, 1);
-  %tau_tp2        ~ cauchy(0,1);
-  %tau_s2         ~ cauchy(0,1);
-  %tau_pi         ~ cauchy(0,1);
-%}
-%
-%\end{verbatim}
-%}
-%
-%\pagebreak
-%\section{Posterior predictive graphical checks}
-%\begin{figure}[H]
-%\includegraphics[width=\textwidth]{ppcheck-1.pdf}
-%\end{figure}
-%\begin{figure}[H]
-%\includegraphics[width=\textwidth]{ppcheck-2}
-%\end{figure}
-%\begin{figure}[H]
-%\includegraphics[width=\textwidth]{ppcheck-3.pdf}
-%\end{figure}
-%\begin{figure}[H]
-%\includegraphics[width=\textwidth]{ppcheck-4.pdf}
-%\end{figure}
-%\begin{figure}[H]
-%\includegraphics[width=\textwidth]{ppcheck-5.pdf}
-%\end{figure}
-%\begin{figure}[H]
-%\includegraphics[width=\textwidth]{ppcheck-6.pdf}
-%\end{figure}
-%\clearpage
-%\bibliographystyle{plainnat}
-% \bibliographystyle{plain}
 \bibliography{./sample}

### 2017-09-27 paper/supplement.tex
Add separate supplementary section with plots
- change figure 6 to corresponding Drive-models of figure 5

- fix labeling clashes

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf,float, bm, subcaption}
\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
\usepackage{enumerate}
\usepackage{natbib}
\bibliographystyle{asa}
\usepackage{url} % not crucial - just used below for the URL

% no longer require relative path to figs
\graphicspath{{fig/}}

\usepackage[x11names]{xcolor}
\usepackage{framed} % provides framed env., should be removed in final draft
\definecolor{shadecolor}{rgb}{1,.5,.5}

% math macros

\newcommand{\ind}{\stackrel{ind.}{\sim}}
\newcommand{\op}{\operatorname}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\myequation}{\begin{equation}}
\newcommand{\myendequation}{\end{equation}}
\let\[\myequation
       \let\]\myendequation

\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

\begin{description}

\item Put R Stan code here

\end{description}

\section{Stan code for Model 4}
\label{sec:stan-code}
{\scriptsize
\begin{verbatim}
functions {
real sev_logpdf(real y, real mu, real sigma){
real z;
z = (y - mu) / sigma;
return -log(sigma) + z - exp(z);
}

real sev_logccdf(real y, real mu, real sigma){
return -exp((y - mu) / sigma);
}

real sev_logcdf(real y, real mu, real sigma){
return log1m_exp(-exp((y - mu) / sigma));
}
}

data {
int M;
int N_obs;
int N_cens;
real endtime_obs[N_obs];
real endtime_cens[N_cens];
real starttime_obs[N_obs];
real starttime_cens[N_cens];
int<lower=1, upper=M> dm_obs[N_obs];
int<lower=1, upper=M> dm_cens[N_cens];
vector<lower=0, upper=1>[2] p; # quantiles to model
}
transformed data{
vector[2] z_corr;
for(i in 1:2)
z_corr[i] = log(-1.0 * log1m(p[i])); # used to get location(mu) from quantile(t_p)
}
parameters{
real log_tp1;
real<lower=0> sigma1;
real eta_tp2;
real eta_s2;
real eta_pi;
real<lower=0> tau_tp2;
real<lower=0> tau_s2;
real<lower=0> tau_pi;
vector[M] log_tp2_raw;
real<lower=0, upper=1> sigma2[M];
vector[M] logit_pi_raw;
}

transformed parameters{
real mu1;
vector[M] mu2;
vector[M] log_pi;
mu1 = log_tp1 - sigma1 * z_corr[1];
for(m in 1:M){
mu2[m] = (eta_tp2 + tau_tp2 * log_tp2_raw[m]) - (sigma2[m] * z_corr[2]);
}
for(m in 1:M)
log_pi[m] = log_inv_logit(eta_pi + tau_pi * logit_pi_raw[m]);
}

model{
real tmp[2];
int m;
real logpi;
real mu_1;
real mu_2;
real sig_1;
real sig_2;

for(i in 1:N_obs){
m = dm_obs[i];
logpi = log_pi[m];
mu_2 = mu2[m];
sig_2 = sigma2[m];
// numerator:   log( p * f1 * (1 - F2) + f2 * (1 - p * F1) )
//            = log( exp(log(p) + log(f1) + log(1 - F2)) + exp(log(f2) + log(1 - exp(log(p) + log(F1)))) )
tmp[1] = log_sum_exp(logpi + sev_logpdf(endtime_obs[i], mu1, sigma1) +
sev_logccdf(endtime_obs[i], mu_2, sig_2),
sev_logpdf( endtime_obs[i], mu_2, sig_2) +
log1m_exp(logpi + sev_logcdf(endtime_obs[i], mu1, sigma1)
)
);
// denominator:  log((1 - p * F1) * (1 - F2))
//            =  log(1 - p * F1) + log(1 - F2)
tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_obs[i], mu1, sigma1)) +
sev_logccdf(starttime_obs[i], mu_2, sig_2);

target += tmp[1] - tmp[2];
}

for(i in 1:N_cens){
m = dm_cens[i];
logpi = log_pi[m];
mu_2 = mu2[m];
sig_2 = sigma2[m];

// numerator:   log((1 - p * F1) * (1 - F2))
//            =  log(1 - p * F1) + log(1 - F2)
tmp[1] = log1m_exp(logpi + sev_logcdf(endtime_cens[i], mu1, sigma1)) +
sev_logccdf(endtime_cens[i], mu_2, sig_2);
// denominator:  log((1 - p * F1) * (1 - F2))
//            =  log(1 - p * F1) + log(1 - F2)
tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_cens[i], mu1, sigma1)) +
sev_logccdf(starttime_cens[i], mu_2, sig_2);

target += tmp[1] - tmp[2];
}

log_tp1        ~ normal(7, 2);
log_tp2_raw    ~ normal(0, 1);
sigma1         ~ lognormal(0, 1);
sigma2         ~ lognormal(eta_s2, tau_s2);
logit_pi_raw   ~ normal(0, 1);
eta_tp2        ~ normal(9, 2);
eta_s2         ~ normal(0, 2);
eta_pi         ~ normal(-3, 1);
tau_tp2        ~ cauchy(0,1);
tau_s2         ~ cauchy(0,1);
tau_pi         ~ cauchy(0,1);
}

\end{verbatim}
}

\pagebreak
\section{Posterior predictive graphical checks}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-1.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-2}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-3.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-4.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-5.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-6.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-7.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-8.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-9.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-10.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-11.pdf}
\end{figure}
\clearpage

\end{document}

### 2017-09-27 plotting_fns/KM_plot2.R

@@ -1,3 +1,7 @@
+
+
+
+
 ###################################################################################
 #### Suite of functions for plotting K-M estimate of cdf with truncation correction
 #### overlaying point estimate of cdf with optional pointwise credible/confidence band

### 2017-09-27 workflow/predKM1.R

@@ -70,14 +70,14 @@ y_reps <- lapply(1:44, function(dm){
 
 saveRDS(y_reps,"../plots/y_rep.rds")
 
-get_plotlist <- function(y_reps, dat){
+get_plotlist <- function(y_reps, dat, xlimits=c(100,100000)){
   plist <- list()
   xaxis <- NULL
   yaxis <- NULL
   legend <- NULL
   n <- length(y_reps)
   for(i in 1:n){
-    xlimits <- c(100,100000)
+    xlimits <- xlimits
     ylimits <- c(.001, .9)
     xbrks <- 0:10 * 10000
     ybrks <- c(.001,.01,.1,.25,.5,.75,.9)
@@ -89,9 +89,10 @@ get_plotlist <- function(y_reps, dat){
     bp <- addKmToBaseplot(bp, km_obs, color="red",linetype = "solid",
                           label = expression(KM(y[obs])))
     pnew <- plotFinally(bp, xbrks=xbrks, ybrks=ybrks, years=FALSE) +
-      theme_bw(base_size=14) + theme(axis.title.y=element_blank(),
-                                     plot.margin = unit(c(0,0,0,0), "cm")) + 
-      ggtitle(as.character(i))
+      theme_bw(base_size=14) + #theme(#axis.title.y=element_blank(),
+                                     #plot.margin = unit(c(0,0,0,0), "cm")
+        #) + 
+      ggtitle(as.character(i)) + ylab("Proportion failing")
   
     if(i==1) legend <- cowplot::get_legend(pnew)
     
@@ -112,9 +113,9 @@ set.seed(10438)
 # ran_set <- sort(sample(44, 6))
 all <- 1:44
 
-plist <- get_plotlist(y_reps, dat)
+plist <- get_plotlist(y_reps, dat, xlimits=c(100,100000))
 
-plist <- get_plotlist(all, firstcol = 1+(0:8)*5, lastrow = 41:44)
+# plist <- get_plotlist(all, firstcol = 1+(0:8)*5, lastrow = 41:44)
 library(cowplot)
 pdf("../paper/fig/post-pred-KM-all2.pdf", height=14, width=9)
 plot_grid(plotlist = plist$plots, ncol=5, align="hv")
@@ -123,20 +124,30 @@ dev.off()
 legnd <- get_legend(plist$plots[[1]] + theme(legend.position="right"))
 
[…]

### 2017-09-28 plotting_fns/lifetime_plot.R
Update figures 6 and 7

@@ -159,3 +159,58 @@ lifetime_plot3 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans=
   if(is.null(n_to_show)) return(p+ggtitle(paste(lab)))
   p + ggtitle(paste(lab, "(sample size = ", n_to_show, ", N = ", N, ")"))
 }
+lifetime_plot4 <- function(data, n_to_show=NULL, in_years=TRUE, lab = "", trans="identity",
+                           xlabels=c(.01,.05,.25,1,2,5), xlimits=c(100, 50000), font_size=10){
+  if(in_years){
+    xlabels <- signif(xlabels * 24*365, 0)
+  } 
+  if(!is.null(n_to_show)){
+    N = nrow(data) #referenced in plot main title
+    data <- sample_n(data, size=n_to_show)
+  }
+  n <- nrow(data)
+  data <- data %>% arrange(starttime) %>%
+    mutate(ID=1:n,
+           status = ifelse(censored, "censored", "failure"))
+
+  p <- ggplot(data, aes(x=ID, xend=ID, y=xlimits[1], yend=endtime)) +
+    geom_segment()+
+    geom_point(aes(y=endtime, pch=status), size=2.5)+
+    geom_point(aes(y=starttime, pch="trunc"), size=2)+
+    coord_flip(expand=F) +
+    theme_bw(base_size=font_size) +
+    theme(axis.title.y= element_blank(),
+          axis.ticks.y= element_blank(),
+          axis.text.y = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          panel.border = element_blank()) +
+    scale_x_continuous(limits=c(0, n+1))+
+    scale_shape_manual(name="",
+                       values = c("censored"='\u27A4',
+                                  "failure"='\u2B58',
+                                  "trunc"='\u007C'),
+                       labels = c("censored",
+                                  "failure",
+                                  "truncation time"))
+  
+  transform <- switch(trans,
+                      "log" = scales::log_trans,
+                      "identity" = scales::identity_trans)
+  
+  if(in_years){
+    p <-   p + scale_y_continuous(name="time(years)", limits=xlimits,
+                                  breaks=xlabels*365*24,
+                                  labels=xlabels,
+                                  trans=trans)
+  } else{
+    p <-   p + scale_y_continuous(name="Thousands of hours",limits = xlimits,
[…]

### 2017-09-28 workflow/mod4exception.R

@@ -33,13 +33,13 @@ font_size=12
 
 dat4 <- filter(dat, model==overview$model[which(overview$stan_id==4)])
 dat4 <- arrange(dat4, endtime)
-dat4_tr <- dat4[which(dat4$starttime>2500),]
+dat4_tr <- dat4[which(dat4$starttime>800),]
 
 set.seed(80717)
 
-lt_plot <- lifetime_plot3(data = dat4, n_to_show=60, in_years=FALSE, lab="Drive-model 4",
+lt_plot <- lifetime_plot4(data = dat4, n_to_show=60, in_years=FALSE, lab="Drive-model 4",
                           trans="log", xlimits = xlimits, xlabels=xlabels,
-                          font_size = font_size) +
+                          font_size = font_size) #+
   theme(axis.text.x = element_blank(),
         axis.title.x = element_blank())
   
@@ -59,19 +59,18 @@ summary(draws)
 km4_tr <- tr.adj(km4_tr, median(draws))
 
 bp <- baseKMplot(fit=km4, xlimits=xlimits, ylimits=ylimits, color="black",
-                 linetype = "solid", alpha = 1, logscale = TRUE, label="nonparametric\nestimate")
+                 linetype = 4242, alpha = 1, logscale = TRUE, label="nonparametric\nestimate (all drives)")
 
 samp4 <- with(sampfull, list(mu1=mu1, sigma1=sigma1, log_pi=log_pi[,4], mu2=mu2[,4], sigma2=sigma2[,4]))
 band4 <- bandFromPSamp(samples=samp4, range=xlimits, length.out = 50, N=300, logscale = TRUE)
 bpp <- addBandToBaseplot(baseplot=bp, bandObj=band4, color="black",
-                         linetype="dashed", label="posterior median\n(90% interval)", alpha = .3)
+                         linetype=11, label="\nposterior median\n(90% interval)", alpha = .3)
 
-bppp <- addKmToBaseplot(baseplot = bpp, fitObj = km4_tr, color="black",
-                        linetype="dotted", label="nonparam. est\n of truncated data")
+bppp <- addKmToBaseplot(baseplot = bpp, fitObj = km4_tr, color="black",size=1,
+                        linetype="solid", label="\nnonparametric \nestimate excluding\noldest 75 drives")
 
 combined <- plotFinally(plotList=bppp, xbrks=xlabels*1000, ybrks=ylabels, years=FALSE) +
[…]

### 2017-09-29 paper/paper.tex
Update figures 6 and 7
Edits to abstract, discussion, captions for figs. 6 and 7
-add co-author
-add citation

@@
   \title{\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
-  \author{Eric Mittman \\
-    Department of Statistics, Iowa State University\\
+  \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
     and \\
-    Colin Lewis-Beck \\
-    Department of Statistics, Iowa State University}
+    Colin Lewis-Beck\footnotemark[1]\\
+    and \\
+    William Q. Meeker\footnotemark[1]}
   \maketitle
@@
 Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using draws from the joint posterior distribution of the model parameters.
+Potential applications include assessment and comparison of reliability to inform purchasing decisions.
 \end{abstract}
@@
 Toward this end, there is a need for models that can accomodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
+Doing these things well enables good decision-making in matters related to product life, including purchase decisions and contingency planning.
 %While computationally convenient, simple lifetime models can fail at each of these.
@@
 Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
-Finally, in section~\ref{sec:Discussion}, we discuss the strengths and limitations of our method and consider extensions and alternatives.
+Finally, in section~\ref{sec:Discussion}, we suggest potential applications of our model and consider limitation for inference implied by the model assumptions.
 
@@
 \subsubsection{Model Assessment}
+\label{model-assessment}
 Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate fit to the observed data.
@@
 There are a few Drive-models, however where this is not the case.
-As we addressed previously, in Subsection \ref{subsec:ex1}, Drive-model 14 has a discrepancy between the GFLP and the Kaplan-Meier fits in the right tail.
-Due to the large number of failures for this Drive-model, the posterior for the hierarchical model is very similar to the stand-alone model, so all the same comments made earlier still apply.
-In addition, Drive-models 4, 11 and 34 appear to be exceptions as well.
-We will focus our attention on Drive-model 4, but our diagnosis of 34 is similar, so the following comments extend to that Drive-model as well.
+Drive-model 14 shows a discrepancy between the GFLP fit and the Kaplan-Meier fit in the right tail.
+In particular, the GLFP predictions appear conservative, overpredicting the proportion failing after about 2.5 years.
+Due to the large number of failures for this Drive-model, the posterior for the hierarchical model is very similar to the stand-alone model.
+Other Drive-models showing a lack of fit are 4 and 34.
+We first address Drive-model 4; our diagnosis of 34 is analogous, so the following comments extend to that Drive-model as well.
 
-To attempt to diagnose the cause of the discrepancies between the observed data for Drive-model 4 and its posterior predictive distribution, we plot an event plot for a random sample of units for this Drive-model, along with the Kaplan-Meier estimate, and the posterior estimate of time to failure (see Figure \ref{fig:ex-mod-4}).
-From these plots, it appears that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the several thousand hours, at a time where very few units were observed to be at risk.
-In further support of this hypothesis, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band.
-While it might be tempting to treat those early failures as anomalous, they also point to a weakness in our model.
-By virtue of having all Drive-models share in common the Weibull parameters for the infant mortality failure mode, an early rash of failures indicating a subpopulation with, say, an increasing hazard function,  would be problematic.
-While the proportion at risk of early failure can vary via $\pi_g$, the plausible values of the common parameters, $\mu_1$ and $\sigma_1$, do not ``line up" in a way so to influence the posterior as they might have had they been allowed to vary.
-In other words, the problem seems to go away much quicker than would be expected under Model 4.
-
-As another example, to further appreciate how information is being borrowed across groups, consider Drive-model 43 (Figure \ref{fig:ex-mod-43}).
-The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available.
-In Figure \ref{fig:ex-mod-43} we see that the lack of data for this Drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$.
-Then, in the period where records exist for this Drive-model, at around 10,000 hours, the posterior deviates from the global model, achieving a compromise between it and the nonparametric estimate.
+To attempt to diagnose the cause of the discrepancies between the observed data for Drive-model 4 and its posterior predictive distribution, we consider the event plot for a random sample of units for this Drive-model while looking at estimates of lifetime (Figure \ref{fig:ex-mod-4}).
+From these plots, we can see that the discrepancy between the posterior and the nonparametric estimate is driven by failure among a relatively small subset of units, namely those with left-truncation times smaller than 2500 hours.
+Further investigation reveals that, if we exclude the newest 75 drives, those with truncation times less than 800 hours, the nonparametric estimate falls within the posterior 90\% pointwise credible band.
+This situation suggests that, despite belonging to the same Drive-model, there may be something different about the newer drives leading to worse reliability.
 
@@
 \centering
-\includegraphics{dm4-exception}
-\caption{\small Top: Event plot for uniform random sample of 60 units of Drive-model 4.
-Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_4,i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
-$t_{4,i}\mbox{, where }c_{4,i}=0$.\hspace{\textwidth}Bottom: The adjusted Kaplan-Meier estimate (solid step function) substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval) due to several early failures while few were at risk.
-Using the same Kaplan-Meier estimator, but only the data truncated after 2500 hours (dotted line) shows close agreement with the posterior median.}
+\includegraphics[height=.8\textheight]{dm4-exception.png}
+\caption{\footnotesize Top: Event plot for 60 randomly sampled drives of Drive-model 4.
+Note that all drives in the sample are left truncated well past 1000 hours.\hspace{\textwidth}Bottom: The adjusted Kaplan-Meier estimate (dashed step function) is substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval).
+However, this discrepancy is due to several early failures among the small set of drives with the earliest left-truncation times (the newest set).
+The same adjusted Kaplan-Meier estimator after the exclusion of the newest 75 drives (solid step function) shows close agreement with the posterior median.}
 
@@
 
+A similar pair of plots for Drive-model 43 illustrate (Figure \ref{fig:ex-mod-43}).
+The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.
+In Figure \ref{fig:ex-mod-43} we see that the lack of data for this Drive-model during early life results in a diffuse posterior centered around the what we call the ``global mean." The global mean is corresponds to $\op{GFLP}$ parameters taking the value of the hierarchical means, i.e., $\pi=\eta_{\pi}, \, \mu_2=\eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2)$ and $\sigma_2= \eta_{\sigma_2})$.
+This continues until about 10,000 hours at which time the posterior diverges from the global model, arriving at a compromise between it and the nonparametric estimate.
+
 \begin{figure}
 \centering
-\includegraphics{dm43-shrinkage}
-\caption{\small Top: Event plot for all units of Drive-model 43.
-Left endpoints are left-truncation times, dashed lines extending to the right indicate censoring intervals, $[t_{43},i, \infty)$, and right endpoints without joined dashed lines are observed failure times, i.e.
-$t_{43,i}\mbox{, where }c_{43,i}=0$.\hspace{\textwidth}Bottom: The solid step function correponds to the adjusted Kaplan-Meier estimate.
-Due to the absence of collected information about early life for this particular model, due to left-truncation, the posterior median (dashed line) follows the ``global model" (dotted line), i.e.
-the posterior median of proportion failing for a new Drive-model, approximately until the beginning of the earliest observational period.}
+\includegraphics[height=.8\textheight]{dm43-shrinkage.png}
+\caption{\footnotesize Top: Event plot for all units of Drive-model 43.
+Data for all units are left-truncated, being observed conditional on survival beyond 12,000 hours.\hspace{\textwidth}Bottom: The dashed step function correponds to the adjusted Kaplan-Meier estimate.
+Due to the heavy left-truncation, the posterior median (dashed line) follows the ``global model" (dotted line), defined by the GLFP parameters corresponding to the means of the hierarchical distribution, until approximately the first left-truncation time.}
 \label{fig:ex-mod-43}
@@
 
-We should point out some important caveats of our analyses, given certain model assumptions.
+We now address some important model assumptions and their potential for impacting statistical inference.
 
@@
 This rigidity is illustrated by the right tail of the estimated distribution in figure \ref{fig1}.
-Extrapolations of the fitted model appear to be very conservative and overly confident.
 
@@
 These factors are confounded with Drive-model in our analysis.
-This second issue is a more fundamental issue with observational data like the hard drive data we have analyzed here.
+An example of a possible batch effect is noted in Section \ref{model-assessment} and illustrated by Figure \ref{fig:ex-mod-4}.
 
@@
 The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product groups.
-Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics meaningful in the context of the data.
+Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the data.
 
-Besides comparing different Drive-models for purchasing decisions, another important application would be prediction of future failures (say, week-by-week) to assure that there is always a sufficient number of replacement drives available.
-Methods similar to those in \cite{hmm} can be adapted to Bayesian prediction, as described in Section 15.5 of \cite{intervals}.
+While not presented here, another potential application is forecasting the number of future failures over a fixed period of time for a current population of drives.
+\citet{hmm} proposed a method, based on a Poisson binomial distribution.
+It would be straightforward to adapt their approach to work with our method.
+For details and examples see \citet[Sect.~6]{hmm} and \citet{xu2015assessing}.
 
-\section*{Acknowledgment}    
-The authors express their gratitude to Dr. \hspace{-.5em} Meeker for his comments and suggestions.
+% \section*{Acknowledgment}    
+% The authors express their gratitude to Dr. \hspace{-.5em} Meeker for his comments and suggestions.
 
@@
 
-\item[Stan Code] Stan code for Model 4 (pdf file)
+\item[Stan Code] Stan code for Model 4 (model4.pdf)
 
-\item[Graphs of Model Comparison Fits for all Drive-models] Posterior median of proportion failing as a function of time under Models 1-4 for all Drive-models. (pdf file)
-
-\item[Posterior Predictive Graphical Checks for all Drive-models] Nonparametric estimates of proportion failing for original and replicated data sets for all Drive-models. (pdf file)
+\item[Complete set of model comparison and model assessment plots] Posterior median of proportion failing as a function of time under Models 1-4 and posterior predictive graphical checks for all Drive-models.
+(allplots.pdf)
+\item[Backblaze hard-drive data] Lifetime data for all 44 Drive-models used in our analysis. (backblaze\_hd\_data.csv)
 

### 2017-09-29 plots/km_overplot.R

@@ -96,7 +96,7 @@ p_mod3 <- ggplot(all_mod3, aes(x=t,y=est, group=model)) +
   geom_line(size=1, alpha=.5) +
   scale_x_continuous(name="", breaks = x.breaks, labels=x.breaks/1000)+
   scale_y_continuous(name="", breaks=y.breaks, limits=ylimits)+
-  theme_bw(base_size = 14) + ggtitle("Model 2")+
+  theme_bw(base_size = 14) + ggtitle("Model 3")+
   theme(axis.title = element_blank())
 
 mod4_s <- readRDS("../workflow/samples_lor_only3fails.rds")

### 2017-09-29 workflow/mod4exception.R

@@ -67,7 +67,7 @@ bpp <- addBandToBaseplot(baseplot=bp, bandObj=band4, color="black",
                          linetype=11, label="\nposterior median\n(90% interval)", alpha = .3)
 
 bppp <- addKmToBaseplot(baseplot = bpp, fitObj = km4_tr, color="black",size=1,
-                        linetype="solid", label="\nnonparametric \nestimate excluding\noldest 75 drives")
+                        linetype="solid", label="\nnonparametric \nestimate excluding\nnewest 75 drives")
 
 combined <- plotFinally(plotList=bppp, xbrks=xlabels*1000, ybrks=ylabels, years=FALSE) +
   guides(fill=FALSE) + ggtitle("Drive-model 4, lifetime estimates") +theme_bw(base_size=12)

### 2017-10-04 paper/paper.tex
Edits to bibtek and paper per last round of Meeker comments.

@@
 
-$$ \Pr(T \le t) = F(t) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$
+$$ \Pr(T \le t) = F(t|\alpha,\beta) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$
 
@@
 
-\begin{equation} \Pr(T \le t) = F(t) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
+\begin{equation} \Pr(T \le t) = F(t|\mu, \sigma) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
 = \Phi_{SEV}\left[\frac{\log (t) - \mu}{\sigma}\right],
@@
 \begin{equation}
-f(t) =\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
+f(t|\mu, \sigma) =\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
 \end{equation}
@@
 An advantage conferred by writing the model in this way is that the log of the $p$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.
-Using the inverse cdf of the SEV distribution,
+Using the inverse cdf of the SEV distribution, $\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p)$, this provides the following reparameterized version of the Weibull distribution
 \begin{equation}
-\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p).
+\Pr(T \le t) = F(t|\text{log}(t_p), \sigma) = \Phi_{SEV}\left[\frac{\log(\frac{t}{t_p}) + \sigma \Phi_{SEV}^{-1}(p)}{\sigma}\right]
 \end{equation}
-Because $\Phi(0)\approx 0.632$, we have $\log{\eta} = \mu  \cong t_{0.63}$.
 When there is more prior knowledge about a lower quantile, as is often the case with high-reliability products, using this parameterization allows the analyst to put prior distributions directly on those quantities.
@@
 
-As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different Drive-models.
+As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.
 Some Drive-models have been running since 2013 or before, while others were added at a later date.
 Data have been reported on 75,297 different hard drives that are or where in operation.
-The number of drives varies by Drive-model; some Drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
-Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for Drive-models with at least 3 failures.
+The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
+Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.
 For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.
@@
 In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.
-Therefore, there is both an empirical and theoretical justification to apply the GLFP model.
+Therefore, there is both an empirical and a theoretical justification to apply the GLFP model.
 
@@
 
-To estimate the GLFP we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution.
+To estimate the parameters of the GLFP model we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution.
 We reparameterize the component Weibull distributions using the $0.50$ quantile for the infant mortality failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$).
@@
 Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.
-
+Using this convention the prior distributions used in our analyses are
 %$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
@@
 $$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1, \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle $$
+\sigma_1, \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
 
-These prior distributions put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.
+\noindent These prior distributions put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.
 Thus, we consider the priors to be relatively uninformative.
@@
 Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.
-The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.
+The two quantiles are also well identified with the infant mortality failure mode having an much earlier time to reach the 0.50 quantile compared to the time to reach the 0.20 quantile for the wearout mode.
 
@@
 $\pi$ & 0.03 & 0.05 & 0.10 \\ 
- $\beta_1$ & 0.47 & 1.13 & 1.72 \\ 
-  $\beta_2$ & 4.47 & 4.70 & 4.95 \\ 
-  $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
-  $\mu_2$ & 10.12 & 10.13 & 10.14 \\ 
+ $\beta_1 = 1/\sigma_1 $ & 0.47 & 1.13 & 1.72 \\ 
+  $\beta_2 = 1/\sigma_2$ & 4.47 & 4.70 & 4.95 \\ 
+  $t_{0.5,1}$ & 1,029.42 & 2,279.09 & 3,989.76 \\ 
+  $t_{0.2,2}$ & 18,010.30 & 18,273.22 & 18,553.22 \\ 
    \hline
@@
 We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \cite[Chapter 11]{meeker} (see Appendix for a detailed description).
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to\cite[Chapter 11]{meeker} (see Appendix for a detailed description).
 
@@
 
-For all of the models we consider, the parameters for the infant mortality failure mode are held in common across Drive-models.
+For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.
 We found that there was often insufficient information to model these parameters hierarchically.
@@
 
-\paragraph{Model 1.} Constrain all Drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.
+\paragraph{Model 1.} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.
 
@@
 As we have discussed, while we prefer to allow all of the GLFP model parameters to be Drive-model specific, there are practical issues with fitting the full model.
-For example, due to heavy left-truncation, there may be insufficient information for particular Drive-models to estimate all of the parameters.
+For example, due to heavy left-truncation, there may be insufficient information for particular drive-models to estimate all of the parameters.
 
@@
 \includegraphics[width=\textwidth]{heterogeneity-compare}
-\caption{Left: Kaplan-Meier estimates of the time to failure for each of the Drive-models in the Backblaze data.
+\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
 Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
@@
 \includegraphics[width=\textwidth]{single-drive-4-Models-ex}}
-\caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of Drive-models, with axes scales as for Weibull probability plotting.
+\caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
 The solid step function is the truncated adjusted Kaplan-Meier points.}
@@
 
-The GLFP fit for a representative selection of Drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for all plots).
+The GLFP fit for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for all plots).
 The black step functions again correspond to the adjusted Kaplan-Meier estimates.
@@
 From the point of view of a consumer of hard disk drives, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
-In this case, we should choose the model that will provide the most accurate prediction for future units of previously observed Drive-models.
+In this case, we should choose the model that will provide the most accurate prediction for future units of previously observed drive-models.
 One appropriate criterion for model selection, then, is log pointwise predictive density (lpd),
@@
 \includegraphics[width=\textwidth]{ppcheck-sample}
-\caption{Nonparametric estimates of proportion failing for a representative subset of Drive-models.
+\caption{Nonparametric estimates of proportion failing for a representative subset of drive-models.
 Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
@@
 The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
-Only a representative sample of plots are shown here for brevity (see the Supplementary Material for the complete set of Drive-models).
-For most Drive-models, the estimates based on the observed data look similar to those based on the replicated data.
-There are a few Drive-models, however where this is not the case.
+Only a representative sample of plots are shown here for brevity (see the Supplementary Material for the complete set of drive-models).
+For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.
+There are a few drive-models, however where this is not the case.
 Drive-model 14 shows a discrepancy between the GFLP fit and the Kaplan-Meier fit in the right tail.
@@
 Due to the large number of failures for this Drive-model, the posterior for the hierarchical model is very similar to the stand-alone model.
-Other Drive-models showing a lack of fit are 4 and 34.
+Other drive-models showing a lack of fit are 4 and 34.
 We first address Drive-model 4; our diagnosis of 34 is analogous, so the following comments extend to that Drive-model as well.
@@
 \label{subsubsec:Drive-model Comparisons}
-We consider the problem of ranking Drive-models. From a
+We consider the problem of ranking drive-models. From a
 business perspective, it is clear
-that we should prefer the Drive-models which will provide more years
+that we should prefer the drive-models which will provide more years
 of service. For ease of exposition, we will assume that the purchase
-price of a hard-drive is the same across Drive-models.
+price of a hard-drive is the same across drive-models.
 
@@
 Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
-relative to a new unit. We can rank the Drive-models by 
+relative to a new unit. We can rank the drive-models by 
 $$\op{E}(L(t_{d,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{d,new}^{(s)}).$$
@@
 Posterior credible intervals for $L(t_{d,new})$ are shown in the left panel of Figure \ref{fig5}.
-A zoomed version of the top 8 Drive-models, as ranked by $\op{E}(L(t_{d,new})|y)$ are shown in the bottom right panel.
+A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{d,new})|y)$ are shown in the bottom right panel.
 The panel above shows credible intervals for MTTF.
-We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher than for the Drive-models ranked highest by our depreciation-aware metric.
-The top four Drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
+We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric.
+The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
 However, given the discrepancy of the model fit for Drive-model 4, noted in the previous subsection, in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.
@@
 We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a measure of central tendency.
-The left panel of Figure 5 shows $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all Drive-models.
-The best six Drive-models ranked by MTTF are also the Drive-models with the largest $t_{0.10}$.
+The left panel of Figure 5 shows $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
+The best six drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
 Drive-models 5, 2, 4, and 3 seem to really stand out as the most reliable as there is a larger gap separating them from the rest of the group.
@@
 The plot is again sorted according to $t_{0.10}$ so it can be compared to the left hand side.
-While for many of the Drive-models the ordinal ranking is the same as on the left, there are some Drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
+While for many of the drive-models the ordinal ranking is the same as on the left, there are some Drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
 Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.
@@
   \caption{Left: Summary of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
-  Right: A comparison of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 Drive-models, shows that rankings produced by the two measures are not equivalent.}
+  Right: A comparison of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
   \label{fig5}
@@
 
-\item[Complete set of model comparison and model assessment plots] Posterior median of proportion failing as a function of time under Models 1-4 and posterior predictive graphical checks for all Drive-models.
+\item[Complete set of model comparison and model assessment plots] Posterior median of proportion failing as a function of time under Models 1-4 and posterior predictive graphical checks for all drive-models.
 (allplots.pdf)
-\item[Backblaze hard-drive data] Lifetime data for all 44 Drive-models used in our analysis. (backblaze\_hd\_data.csv)
+\item[Backblaze hard-drive data] Lifetime data for all 44 drive-models used in our analysis. (backblaze\_hd\_data.csv)
 

### 2017-10-04 paper/paper.tex
More edits to bibtek and paper.  focused on results section.

@@
 Finally, after prolonged use, rates of failure increase due to wearout.
-Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability.
+Ignoring this pattern in failure rates can lead to spurious inferences about a product's reliability and suboptimal decisions.
 This highlights the importance of choosing a sufficiently flexible model.
 
-Nonparametric methods require relatively few assumptions, but they are not suitable for forecasting.
+This paper presents a general framework for statistically modeling reliability field data from a heterogeneous population of components or subsystems operating within a larger population or fleet of systems.
+Such a model would be useful for applications such as making purchase decisions from among different suppliers of the components or subsystems.
+Another application is making predictions for the number of needed replacement components or subsystems for spare part provisioning.
+
+Nonparametric methods require relatively few assumptions, but they are not suitable for prediction.
 Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be practical.
@@
 
-Often no information is available to identify why a unit failed and even when the failed units are available, physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive.
+Often no information is available to identify why a unit failed, even when the failed units are available.
+Physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive.
 This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution.
@@
 In Section~\ref{sec:Data}, we describe the Backblaze Hard Drive data set, which is used throughout as a motivating example.
-In section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations.
+In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations.
 Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard disk drive data.
@@
 Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
-Finally, in section~\ref{sec:Discussion}, we suggest potential applications of our model and consider limitation for inference implied by the model assumptions.
+Finally, in Section~\ref{sec:Discussion}, we suggest potential applications of our model and consider limitation for inference implied by the model assumptions.
 
@@
 
-Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.
+Previous papers have employed various assumptions and methods to model masked lifetime failure data.
 When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.
@@
 Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure.
-\cite{berger} introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions.
+\cite{berger} introduced the Poly-Weibull distribution where the failure time is the minimum of a several Weibull distributions, corresponding to different failure modes.
 More recently, \cite{ranjan} considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions.
@@
 
-In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population specific parameters hierarchically, borrowing strength across the sub-populations.
+In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population-specific parameters hierarchically, borrowing strength across the sub-populations.
 Let
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution.
-We truncate the distribution of $\sigma_{2g}$ at $1$, restricting the wearout failure mode to have an increasing hazard function.
-We choose to model these parameters (or suitable transformations) with normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
+We truncate the distribution of $\sigma_{2g}$ at $1$, restricting the wearout failure mode to have an increasing hazard function (e.g., $0<\sigma_{2g}<1$).
+We choose to model these parameters with lognormal and normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
 
@@
 
-The set of model specifications were chosen based data, interpretation of the model, as well as estimation considerations.
+The set of model specifications was chosen based data, interpretation of the model, as well as estimation considerations.
 Product specific parameters for the the wearout failure mode $(\mu_{2g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
 Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by Drive-model.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by drive-model.
 
@@
 
-\paragraph{Model 2.} $t_{p,g}$ varies by Drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
+\paragraph{Model 2.} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
 $$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
@@
 \label{sec:Model Comparisons}
-As we have discussed, while we prefer to allow all of the GLFP model parameters to be Drive-model specific, there are practical issues with fitting the full model.
+As we have discussed, while we prefer to allow all of the GLFP model parameters to be drive-model specific, there are practical issues with fitting the full model.
 For example, due to heavy left-truncation, there may be insufficient information for particular drive-models to estimate all of the parameters.
@@
 The Kaplan-Meier estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
-The smaller plots correspond to Models 1 through 4 reading from left to right, top to bottom.
+The smaller plots correspond to Models 1 through 4.
 Note that while all the paths for the posterior estimates necessarily pass through the origin, the nonparametric curves do not, but only because, for the Kaplan-Meier estimator, inference is entirely conditional on survival up to left-most left-trunctation time; to make the plot comparable, we use an adjustment based on a parametric model (Model 4).
@@
 Model 3, which removes this restriction on $\sigma_{2,g}$ suggests that there is evidence to support variation in this parameter among drive-models.
-While Model 3 and 4 do display differences, their impact is relatively subtle.
+While Model 3 and 4 do display differences, they are relatively subtle.
 
@@
 
-We can also compare the four Model fits for each Drive-model individually.
+We can also compare the four Model fits for each drive-model individually.
 For these comparisons, we use the posterior median of the MCMC draws.
@@
 
-The GLFP fit for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for all plots).
+The GLFP fit for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for similar plots for the other drive-models).
 The black step functions again correspond to the adjusted Kaplan-Meier estimates.
-As model complexity increases, the GLFP curves become more flexible and are better able to fit the observed failure data.
+As model complexity increases, the GLFP curves become more flexible and are better able to describe the observed failure data.
 All of the GLFP models appear to produce lower estimates, relative to Kaplan-Meier, of the proportion of failed hard drives for Drive-model 2, with Model 4 being in closest agreement.
@@
 In this case, we should choose the model that will provide the most accurate prediction for future units of previously observed drive-models.
-One appropriate criterion for model selection, then, is log pointwise predictive density (lpd),
+One appropriate criterion for model selection, then, is the log pointwise predictive density (lpd),
 
@@
 
-Using this criterion, we should choose the Model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$. Suppressing the conditioning on $t^L$ and $c$,
+\noident Using this criterion, we should choose the Model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
+Suppressing the conditioning on $t^L$ and $c$,
 \begin{equation}
@@
  We calculated the difference in expected predictive accuracy for Model 4 versus 3, Model 3 versus 2, and Model 2 versus 1, as well as the standard error of the difference (Table 1).
-For the models we considered, each successive increase in model complexity increases elpd.  Of all the models, Model 4 has significantly better elpd compared to the the other 3 models.
+For the models we considered, each successive increase in model complexity increases elpd.
+Of all the models, Model 4 has significantly better elpd compared to the the other 3 models.
+
 \begin{figure}
@@
 \end{figure}
+
 \subsection{Results}
@@
 \label{model-assessment}
-Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate fit to the observed data.
+Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate description of the observed data.
 One way to do this is to draw replicate data sets from the posterior predictive distribution.
@@
 In the hierarchical setting, there can be different formulations of the predictive distribution.
-In the present context, one could predict the lifetime of a new unit of a previously observed Drive-model or a new Drive-model.
+In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
 We choose to focus on the former.
 There are many possible approaches to the posterior predictive check.
-For this problem, we chose to consider the posterior predictive distribution of the Kaplan-Meier estimate for each Drive-model $g$.
+For this problem, we chose to consider the posterior predictive distribution of the Kaplan-Meier estimate for each drive-model $g$.
 
@@
 Our admittedly simple approach is to assume that failed units would have been the censored had they survived until the latest censoring or failure time for that drive model.
-We chose to generate 19 replicates for each drive model.
+We generated 19 replicates for each drive model.
 This gives a 1/20 chance under the hypothesis of no difference that the Kaplan-Meier estimate based on the observed data would be ``most extreme" relative all the replicated  from the those one and is not too many so as to produce an overly cluttered plot.
-Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each Drive-model $g=1,\ldots,44$, is as follows:
+Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each drive-model $g=1,\ldots,44$, is as follows:
 \begin{enumerate}
-\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right) $, is a random draw from the full posterior distribution.
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right)$, is a random draw from the full posterior distribution.
 \item Repeat step 1 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
@@
 \includegraphics[width=\textwidth]{ppcheck-sample}
-\caption{Nonparametric estimates of proportion failing for a representative subset of drive-models.
+\caption{Nonparametric estimates of the proportion failing for a representative subset of drive-models.
 Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
@@
 The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
-Only a representative sample of plots are shown here for brevity (see the Supplementary Material for the complete set of drive-models).
+Only a representative sample of plots is shown here for brevity (see the Supplementary Material for the complete set of drive-models).
 For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.
@@
 In particular, the GLFP predictions appear conservative, overpredicting the proportion failing after about 2.5 years.
-Due to the large number of failures for this Drive-model, the posterior for the hierarchical model is very similar to the stand-alone model.
+Due to the large number of failures for this Drive-model, the posterior for the hierarchical model is similar to the stand-alone model.
 Other drive-models showing a lack of fit are 4 and 34.
@@
 We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a measure of central tendency.
-The left panel of Figure 5 shows $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
+The top panel of Figure 9 shows $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
+In this application 50\% CIs are more meaningful as there is lots of uncertainty in some of the parameters: especially for those with few drives in operation.
 The best six drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
-Drive-models 5, 2, 4, and 3 seem to really stand out as the most reliable as there is a larger gap separating them from the rest of the group.
 Another interesting feature to compare is $\pi$, the proportion defective.
-The right panel of Figure 5 shows the posterior credible intervals of $\pi$ for each Drive-model.
-The plot is again sorted according to $t_{0.10}$ so it can be compared to the left hand side.
-While for many of the drive-models the ordinal ranking is the same as on the left, there are some Drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
+The bottom panel of Figure 10 shows the posterior credible intervals of $\pi$ for each drive-model.
+The plot is again sorted according to $t_{0.10}$ so it can be compared to the plot above.
+While for many of the drive-models the ordinal ranking is the same as on the top, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
 Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.
@@
 \label{sec:trunc-adj}
-We first start with a non-parametric estimate of the empirical cdf for each Drive-model using the Kaplan-Meier estimator.
-With left truncation, however, the standard Kaplan-Meier estimator for Drive-model $d$, denoted by
+We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator.
+With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
 $\widehat{F_d(t)}_{KM}$, is conditional on survival up to
@@
 of drive-model $d$ for which records are available. To produce
-unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each Drive-model we select
+unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each drive-model we select
 $t_{d,\text{min}}^L$, the smallest left truncated time in the sample.

### 2017-10-05 paper/paper.tex
Edits to paper (Eric, Thursday afternoon)

@@
 This paper presents a general framework for statistically modeling reliability field data from a heterogeneous population of components or subsystems operating within a larger population or fleet of systems.
-Such a model would be useful for applications such as making purchase decisions from among different suppliers of the components or subsystems.
-Another application is making predictions for the number of needed replacement components or subsystems for spare part provisioning.
+Such a model would be useful for applications such as
+
+\begin{itemize}
+\item Making purchase decisions from among different suppliers of the components or subsystems
+\item Making predictions for the number of needed replacement components or subsystems for spare part provisioning
+\end{itemize}
 
@@
 Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
-Finally, in Section~\ref{sec:Discussion}, we suggest potential applications of our model and consider limitation for inference implied by the model assumptions.
+In Section~\ref{sec:Discussion}, we suggest potential applications of our model and consider limitation for inference implied by the model assumptions.
 
@@
 Thus, the Weibull distribution is a log-location scale family, with $\log(T)$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$.
-With this in mind, the corresponding density function is
-\begin{equation}
-f(t|\mu, \sigma) =\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
-\end{equation}
+% With this in mind, the corresponding density function is
+% \begin{equation}
+% f(t|\mu, \sigma) =\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
+% \end{equation}
 
-An advantage conferred by writing the model in this way is that the log of the $p$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.
-Using the inverse cdf of the SEV distribution, $\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p)$, this provides the following reparameterized version of the Weibull distribution
+An advantage conferred by writing the model in this way is that the log of the $p$ quantile, $t_p$, is a linear function of the parameters, namely, $\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p)$.
+This allows an analyst to apply prior information in a straightforward way.
+Reparameterizing in terms of $(\log(t_p),\sigma)$, we can write the Weibull cdf as
 \begin{equation}
@@
 As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.
-Some Drive-models have been running since 2013 or before, while others were added at a later date.
+Some drive-models have been running since 2013 or before, while others were added at a later date.
 Data have been reported on 75,297 different hard drives that are or where in operation.
-The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.
+The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.
 Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.
@@
   \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
-  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.
-Each number labels a unique drive brand model.
-Axes are on the log scale.}
+  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.  Each number labels a unique drive-model.   Axes are on the log scale.}
   \label{drive-scatter}
@@
 
-\subsection{Analysis of a Single Drive-Model}
+\subsection{Analysis of a single drive-model}
 \label{subsec:ex1}
@@
 $$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1, \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
+\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to\cite[Chapter 11]{meeker} (see Appendix for a detailed description).
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to\citet[Chapter 11, ]{meeker} (see Appendix for a detailed description).
 
@@
 \begin{equation*}
-\sigma_{1g} \ind \op{Lognormal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
+\sigma_{1g} \ind \op{Log-normal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
 \end{equation*}
 \begin{equation*}
-\sigma_{2g} \ind \op{Lognormal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
+\sigma_{2g} \ind \op{Log-normal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
 \end{equation*}
@@
 \label{eq:hier-model}
-t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{Lognormal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
+t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
 \end{equation}
 \begin{equation*}
-t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{Lognormal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
+t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
 \end{equation*}
 \begin{equation*}
-\op{logit} \pi_g \ind \op{N}(\eta_\pi, \tau_\pi).
+\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi).
 \end{equation*}
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution.
-We truncate the distribution of $\sigma_{2g}$ at $1$, restricting the wearout failure mode to have an increasing hazard function (e.g., $0<\sigma_{2g}<1$).
-We choose to model these parameters with lognormal and normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
+We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by``$T(0,1)$" above) restricting the wearout failure mode to have an increasing hazard function (e.g., $\beta_{2g} = 1/\sigma_{2g}>1$).
+We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
 
@@
 This is done to reduce autocorrelation and efficiently explore the posterior.
-Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 and 16 chains were run for Model 4.
+Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 for a total of 6,000 post burn-in iterations and 16 chains were run for Model 4 for a total of 24,000 post burn-in iterations.
 The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains.
 Upon convergence, $\hat{R}$, converges to 1.
-The  maximum values of $\hat{R}$ across all model parameters for Models 1 throught 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004.
-Other diagnostics provided by the software (tree-depth, and divergent transitions) were clean.
+The  respective maximum values of $\hat{R}$ across all model parameters for Models 1 through 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004.
+Other diagnostics provided by the software (tree-depth, and divergent transitions) did not indicate problems with sampling.
 Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.
@@
 
-\subsection{Model Selection}
+\subsection{Model selection}
 \label{sec:Model Comparisons}
@@
 
-We can also compare the four Model fits for each drive-model individually.
-For these comparisons, we use the posterior median of the MCMC draws.
+We can also compare posterior estimates of time to failure under Models 1-4 for each drive-model individually.
+For these comparisons, we calculate, pointwise, the posterior median of proportion failing over a grid.
+That is, for every model, and for a fixed set, $\tilde{t}=\{\tilde{t}_1,\ldots,\tilde{t}_M\}$, we compute
+\begin{equation}
+\op{median}\left\{\op{GLFP}\left(\tilde{t}_m|\pi_{g}^{(s)},\mu_{1g}^{(s)},\sigma_{1g}^{(s)},\mu_{2g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
+\end{equation}
+where $\theta^{(s)}$ is the $s^{th}$ posterior sample from $p(\theta|y)$.
 %Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.
@@
 As model complexity increases, the GLFP curves become more flexible and are better able to describe the observed failure data.
-All of the GLFP models appear to produce lower estimates, relative to Kaplan-Meier, of the proportion of failed hard drives for Drive-model 2, with Model 4 being in closest agreement.
+All of the GLFP models appear to produce higher estimates of the proportion failing, relative to Kaplan-Meier, for Drive-model 2.
+This behavior is not unexpected, however, since hierarchical models improve model stability through shrinkage; in this context, shrinkage toward an average cumulative time to failure.
+Model 4 is in closest agreement.
+
+Drive-model 9 shows a high proportion of  of early failures relative to Model 1.
+Model 2 addresses this by estimating an earlier time to wearout, but still results in a poor fit to the data.
+Models 3 and 4 conform more closely to the Kaplan-Meier.
+The disagreement between Models 3 and 4 for Drive-model 9, is interesting: Model 4 allows the proportion of defective units to vary by Drive-model, but Model 3 appears to fit just as well in this case, by increasing the scale parameter for the wearout failure mode.
+
 For Drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for Drive-model 40, they produce higher estimates.
@@
 \end{equation}
-
-\noident Using this criterion, we should choose the Model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
+\noindent Using this criterion, we should choose the Model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
 Suppressing the conditioning on $t^L$ and $c$,
@@
 In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
-We choose to focus on the former.
+We choose to focus on the first of these.
 There are many possible approaches to the posterior predictive check.
@@
 
-To attempt to diagnose the cause of the discrepancies between the observed data for Drive-model 4 and its posterior predictive distribution, we consider the event plot for a random sample of units for this Drive-model while looking at estimates of lifetime (Figure \ref{fig:ex-mod-4}).
+To diagnose the cause of the discrepancies between the observed data for Drive-model 4 and its posterior predictive distribution, we consider the event plot for a random sample of units for this Drive-model while looking at estimates of lifetime (Figure \ref{fig:ex-mod-4}).
 From these plots, we can see that the discrepancy between the posterior and the nonparametric estimate is driven by failure among a relatively small subset of units, namely those with left-truncation times smaller than 2500 hours.
@@
 
-A similar pair of plots for Drive-model 43 illustrate (Figure \ref{fig:ex-mod-43}).
+Figure 8 shows a similar pair of plots for Drive-model 43 illustrate (Figure \ref{fig:ex-mod-43}).
 The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.

### 2017-10-05 paper/paper.tex
End of day commit to paper

@@
 Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
-After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.
-We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf.
+After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.
+We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.
 With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to\citet[Chapter 11, ]{meeker} (see Appendix for a detailed description).
@@
 \label{eq:hier-model}
-t_{p_{1}g} \equiv \mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
+t_{p_{1}g} \equiv \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
 \end{equation}
 \begin{equation*}
-t_{p_{2}g} \equiv \mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
+t_{p_{2}g} \equiv \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
 \end{equation*}
@@
 Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of time to failure.
-The large left panel shows the Kaplan-Meier nonparametric estimate of the same.
+The large left panel shows the Kaplan-Meier estimate of the same.
 The Kaplan-Meier estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
 The smaller plots correspond to Models 1 through 4.
-Note that while all the paths for the posterior estimates necessarily pass through the origin, the nonparametric curves do not, but only because, for the Kaplan-Meier estimator, inference is entirely conditional on survival up to left-most left-trunctation time; to make the plot comparable, we use an adjustment based on a parametric model (Model 4).
+Note that while all the paths for the posterior estimates necessarily pass through the origin, the Kaplan-Meier estimates do not, because, for the Kaplan-Meier estimator, inference is entirely conditional on survival up to left-most left-trunctation time.
+To make the plot comparable, we use an adjustment based on a parametric model (Model 4).
 
@@
 \begin{equation}
+\label{pointwise-medians}
 \op{median}\left\{\op{GLFP}\left(\tilde{t}_m|\pi_{g}^{(s)},\mu_{1g}^{(s)},\sigma_{1g}^{(s)},\mu_{2g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
@@
 For Drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for Drive-model 40, they produce higher estimates.
-On the other hand, Model 4 follows the nonparametric estimate quite closely across stages of lifetime.
+On the other hand, Model 4 follows the Kaplan-Meier quite closely across stages of lifetime.
 The differences between Models 3 and 4 and Models 1 and 2 are visually quite apparent.
@@
 \begin{enumerate}
-\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, $\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right)$, is a random draw from the full posterior distribution.
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, \\$\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right)$, is a random draw from the full posterior distribution.
 \item Repeat step 1 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
@@
 \includegraphics[width=\textwidth]{ppcheck-sample}
-\caption{Nonparametric estimates of the proportion failing for a representative subset of drive-models.
+\caption{Adjusted Kaplan-Meier estimates of the proportion failing for a representative subset of drive-models.
 Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
@@
 
-We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the Kaplan-Meier nonparametric estimator.
+We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the Kaplan-Meier estimator.
 The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
@@
 To diagnose the cause of the discrepancies between the observed data for Drive-model 4 and its posterior predictive distribution, we consider the event plot for a random sample of units for this Drive-model while looking at estimates of lifetime (Figure \ref{fig:ex-mod-4}).
-From these plots, we can see that the discrepancy between the posterior and the nonparametric estimate is driven by failure among a relatively small subset of units, namely those with left-truncation times smaller than 2500 hours.
-Further investigation reveals that, if we exclude the newest 75 drives, those with truncation times less than 800 hours, the nonparametric estimate falls within the posterior 90\% pointwise credible band.
+From these plots, we can see that the discrepancy between the posterior and Kaplan-Meier estimates is driven by failures among a relatively small subset of units, namely those with left-truncation times smaller than 2500 hours.
+Further investigation reveals that, if we exclude the newest 75 drives, those with truncation times less than 800 hours, the Kaplan-Meier estimate falls within the posterior 90\% pointwise credible band.
 This situation suggests that, despite belonging to the same Drive-model, there may be something different about the newer drives leading to worse reliability.
@@
 The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.
-In Figure \ref{fig:ex-mod-43} we see that the lack of data for this Drive-model during early life results in a diffuse posterior centered around the what we call the ``global mean." The global mean is corresponds to $\op{GFLP}$ parameters taking the value of the hierarchical means, i.e., $\pi=\eta_{\pi}, \, \mu_2=\eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2)$ and $\sigma_2= \eta_{\sigma_2})$.
-This continues until about 10,000 hours at which time the posterior diverges from the global model, arriving at a compromise between it and the nonparametric estimate.
+In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for Drive-model 43 to what we call the `global average.' We get this by setting the $\op{GLFP}$ parameters equal to the median of their hierarchical distribution, i.e., $\pi=\op{logit}^{-1}(\eta_{\pi}),\, \mu_1=\mu_1,\, \sigma_1=\sigma_1,\, \mu_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2)$ and $\sigma_2= m_{\sigma_2}$, where $m_{\sigma_2}=\Phi_{Log-Normal}^{-1}[0.5 \cdot \Phi_{Log-Normal}(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}]$ is the population median of the $\sigma_{2g}$.
+The curve was estimated using pointwise medians as in Equation \ref{pointwise-medians}.
+We can see that the lack of data for this Drive-model during early life results in a diffuse posterior centered around the global average until about 10,000 hours when they diverge.
+Although the posterior is close to the adjusted Kaplan-Meier, it shows shrinkage toward the global average.
 
@@
 \appendix
-\section{Truncation adjusted nonparametric estimate of lifetime}
+\section{Truncation adjusted Kaplan-Meier estimate of lifetime}
 \label{sec:trunc-adj}

### 2017-10-05 plots/single-drive-plots.R

@@ -46,7 +46,7 @@ ggsave("../paper/fig/km14-natural-plus.pdf", width = 7.5, height=6)
 ybreaks<- c(.002,.01,.05,.1,.2,.4,.55)
 xbreaks<- c(500, 1000, 2000, 4000, 8000, 15000, 30000)
 base_p14_p <- baseKMplot(km14, xlimits=xlimits, ylimits=ylimits, color="blue", linetype="solid",
-                         logscale = TRUE, prob=TRUE)
+                         logscale = TRUE, prob=TRUE, label = "adjusted\nKaplan-Meier")
 p14_p <- plotFinally(base_p14_p, xbrks = xbreaks, ybrks = ybreaks, years=FALSE,
                    greenwood = TRUE, alpha = .1)
 p14_p + theme_bw(base_size=14) + theme(legend.position="none")
@@ -54,8 +54,9 @@ ggsave("../paper/fig/km14-prob.pdf", width = 6, height = 6)
 
 band14_nat <- bandFromPSamp(samp14, range=xlimits, length.out = 40, N=1000, logscale = TRUE)
 comb_p14_p <- addBandToBaseplot(baseplot = base_p14_p, bandObj = band14_nat,
-                          color="red", linetype="dashed", alpha=.1, label="GLFP posterior")
+                          color="red", linetype="dashed", alpha=.1, label="\nGLFP posterior")
 p14_pplus <- plotFinally(comb_p14_p, xbrks = xbreaks, ybrks = ybreaks, years=FALSE,
                          greenwood = TRUE, alpha = .1)
-p14_pplus + theme_bw(base_size=14)
+p14_pplus + theme_bw(base_size=14)+
+            theme(legend.text = element_text(size=12))
 ggsave("../paper/fig/km14-prob-plus.pdf", width=7.5, height=6)

### 2017-10-05 workflow/mod4exception.R

@@ -29,7 +29,7 @@ xlabels=c(.1,.5,2.5,10,50)
 ylabels=c(.001,.005, .01,.05,.1)
 xlimits=c(100, 50000)
 ylimits=c(.001,.2)
-font_size=12
+font_size=14
 
 dat4 <- filter(dat, model==overview$model[which(overview$stan_id==4)])
 dat4 <- arrange(dat4, endtime)
@@ -39,9 +39,10 @@ set.seed(80717)
 
 lt_plot <- lifetime_plot4(data = dat4, n_to_show=60, in_years=FALSE, lab="Drive-model 4",
                           trans="log", xlimits = xlimits, xlabels=xlabels,
-                          font_size = font_size) #+
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank())
+                          font_size = font_size)
+  # theme(axis.text.x = element_blank(),
+        # axis.title.x = element_blank()
+        # )
   
 
 km4 <- KM.survfit(dat4)
@@ -59,7 +60,7 @@ summary(draws)
 km4_tr <- tr.adj(km4_tr, median(draws))
 
 bp <- baseKMplot(fit=km4, xlimits=xlimits, ylimits=ylimits, color="black",
-                 linetype = 4242, alpha = 1, logscale = TRUE, label="nonparametric\nestimate (all drives)")
+                 linetype = 4242, alpha = 1, logscale = TRUE, label="adjusted KM\nestimate (all drives)")
 
 samp4 <- with(sampfull, list(mu1=mu1, sigma1=sigma1, log_pi=log_pi[,4], mu2=mu2[,4], sigma2=sigma2[,4]))
 band4 <- bandFromPSamp(samples=samp4, range=xlimits, length.out = 50, N=300, logscale = TRUE)
@@ -67,16 +68,16 @@ bpp <- addBandToBaseplot(baseplot=bp, bandObj=band4, color="black",
                          linetype=11, label="\nposterior median\n(90% interval)", alpha = .3)
 
 bppp <- addKmToBaseplot(baseplot = bpp, fitObj = km4_tr, color="black",size=1,
-                        linetype="solid", label="\nnonparametric \nestimate excluding\nnewest 75 drives")
+                        linetype="solid", label="\nadjusted KM\nestimate excluding\nnewest 75 drives")
 
 combined <- plotFinally(plotList=bppp, xbrks=xlabels*1000, ybrks=ylabels, years=FALSE) +
[…]

### 2017-10-06 plots/single-drive-plots.R
Single drive plot edits.

@@ -18,6 +18,13 @@ source("../plotting_fns/KM_plot2.R")
 indFit <- extract(readRDS("../workflow2/samples_nopool_data_dm8.rds"))
 samp14 <- with(indFit, list(mu1=mu1, sigma1=sigma1, mu2=mu2, sigma2=sigma2, log_pi=log_pi))
 
+#Get Summary Quantiles for Paper
+quantile(1/samp14$sigma1, probs=c(.025, .5, .975))
+q1 <- exp(samp14$mu1 + samp14$sigma1*log(-log(1-.5)))
+quantile(q1, probs=c(.025, .5, .975))
+q2 <- exp(samp14$mu2 + samp14$sigma2*log(-log(1-.2)))
+quantile(q2, probs=c(.025, .5, .975))
+
 #Linear scales
 xlimits <- c(200, 30000)
 ylimits <- c(.001, .8)

### 2017-10-06 paper/paper.tex

@@
 \centering
-  \includegraphics[width=.8\textwidth]{km14-prob}
-  \caption{Kaplan Meier estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
+  \includegraphics[width=.6\textwidth]{km14-prob}
+  \caption{Kaplan-Meier estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
   \label{fig1}
@@
 Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.
-Solid line is a non-parametric estimate with pointwise 90\% uncertainty band in blue.}
+Solid line is adjusted Kaplan-Meier estimate with pointwise 90\% credible band in blue.}
   \label{ex1-overlay}
@@
 \caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
-The solid step function is the truncated adjusted Kaplan-Meier points.}
+The solid step function is the adjusted Kaplan-Meier.}
 \label{fig:mod_comp_leg}

### 2017-10-06 paper/paper.tex

@@
 Data for all units are left-truncated, being observed conditional on survival beyond 12,000 hours.\hspace{\textwidth}Bottom: The dashed step function correponds to the adjusted Kaplan-Meier estimate.
-Due to the heavy left-truncation, the posterior median (dashed line) follows the ``global model" (dotted line), defined by the GLFP parameters corresponding to the means of the hierarchical distribution, until approximately the first left-truncation time.}
+Due to the heavy left-truncation, the posterior median for Drive-model 43 (dashed line) coincides with the posterior median of the ``global model" (dotted line) until the first left-truncation time.}
 \label{fig:ex-mod-43}

### 2017-10-06 paper/paper.tex
Fix notation in replicate data method

@@
 \item Repeat step 1 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
-\item If $c_{g,i}=0$, let $c_{rep,g,i}^{(s)}=\max \{t_{g,i}: i \ge 1\}$, otherwise let $c_{rep,g,i}^{(s)}=t_{g,i}$.\\
-If $\tilde{t}_{rep,g,i}^{(s)}>m_{g,i}$, set $t_{rep,g,i}^{(s)}=c_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=1$, otherwise set $t_{rep,g,i}^{(s)}=\tilde{t}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
+\item Set $d_{g,i}^{(s)}= \begin{cases} \max \{t_{g,i}: i \ge 1\}, & \mbox{ if }c_{g,i}=0\\
+t_{g,i}, & \mbox{ if }c_{g, i}=1. \end{cases}$\\
+Lastly, set $(t_{rep,g,i}^{(s)},c_{rep,g,i}^{(s)}) = \begin{cases}
+  (d_{g,i}^{(s)},1), & \mbox{ if } \tilde{t}_{rep,g,i}^{(s)}>d_{g,i}^{(s)}\\
+  (\tilde{t}_{rep,g,i}^{(s)},0), & \mbox{ if }\tilde{t}_{rep,g,i}^{(s)} \le d_{g,i}^{(s)}.
+\end{cases}$
 \end{enumerate}

### 2017-10-06 paper/paper.tex
Replace later instances of ``Kaplan-Meier" with `K-M'

@@
 To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.
-The Kaplan-Meier estimate plotted below suggests at least two failure modes.
+The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.
 The curve has a slow ramp up in failures till about 10,000 hours (early failures), increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).
@@
   \includegraphics[width=.6\textwidth]{km14-prob}
-  \caption{Kaplan-Meier estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
+  \caption{K-M estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
   \label{fig1}
@@
 After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.
-We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to\citet[Chapter 11, ]{meeker} (see Appendix for a detailed description).
+We applied this method using the K-M estimate of the cdf.
+With left truncation, however, the standard K-M estimator is biased, so we used an adjusted version due to\citet[Chapter 11, ]{meeker} (see Appendix for a detailed description).
 
-In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted Kaplan-Meier estimate with axes on the Weibull probability scale.
-Pointwise uncertainty bands associated with each estimator are included.
+In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted K-M estimate with axes on the Weibull probability scale.
+90\% pointwise uncertainty bands associated with each estimator are included.
 While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours.
@@
 Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.
-Solid line is adjusted Kaplan-Meier estimate with pointwise 90\% credible band in blue.}
+Solid line is adjusted K-M estimate with pointwise 90\% credible band in blue.}
   \label{ex1-overlay}
@@
 Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of time to failure.
-The large left panel shows the Kaplan-Meier estimate of the same.
-The Kaplan-Meier estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
+The large left panel shows the adjusted K-M estimate of the same.
+The adjusted K-M estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
 The smaller plots correspond to Models 1 through 4.
-Note that while all the paths for the posterior estimates necessarily pass through the origin, the Kaplan-Meier estimates do not, because, for the Kaplan-Meier estimator, inference is entirely conditional on survival up to left-most left-trunctation time.
+Note that while all the paths for the posterior estimates necessarily pass through the origin, the K-M estimates do not, because, for the K-M estimator, inference is entirely conditional on survival up to left-most left-trunctation time.
 To make the plot comparable, we use an adjustment based on a parametric model (Model 4).
@@
 These plots suggest that Model 1 is insufficient to explain the observed data.
-Model 2 certainly captures more of the heterogeneity observed in the Kaplan-Meier estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in the rate of failure, which may be too restrictive.
+Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in the rate of failure, which may be too restrictive.
 Model 3, which removes this restriction on $\sigma_{2,g}$ suggests that there is evidence to support variation in this parameter among drive-models.
@@
 \includegraphics[width=\textwidth]{heterogeneity-compare}
-\caption{Left: Kaplan-Meier estimates of the time to failure for each of the drive-models in the Backblaze data.
+\caption{Left: K-M estimates of the time to failure for each of the drive-models in the Backblaze data.
 Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
@@
 \caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
-The solid step function is the adjusted Kaplan-Meier.}
+The solid step function is the adjusted K-M estimate.}
 \label{fig:mod_comp_leg}
@@
 The GLFP fit for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for similar plots for the other drive-models).
-The black step functions again correspond to the adjusted Kaplan-Meier estimates.
+The black step functions again correspond to the adjusted K-M estimates.
 As model complexity increases, the GLFP curves become more flexible and are better able to describe the observed failure data.
-All of the GLFP models appear to produce higher estimates of the proportion failing, relative to Kaplan-Meier, for Drive-model 2.
+All of the GLFP models appear to produce higher estimates of the proportion failing, relative to K-M, for Drive-model 2.
 This behavior is not unexpected, however, since hierarchical models improve model stability through shrinkage; in this context, shrinkage toward an average cumulative time to failure.
@@
 Model 2 addresses this by estimating an earlier time to wearout, but still results in a poor fit to the data.
-Models 3 and 4 conform more closely to the Kaplan-Meier.
+Models 3 and 4 conform more closely to the K-M estimate.
 The disagreement between Models 3 and 4 for Drive-model 9, is interesting: Model 4 allows the proportion of defective units to vary by Drive-model, but Model 3 appears to fit just as well in this case, by increasing the scale parameter for the wearout failure mode.
@@
 For Drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for Drive-model 40, they produce higher estimates.
-On the other hand, Model 4 follows the Kaplan-Meier quite closely across stages of lifetime.
+On the other hand, Model 4 follows the K-M quite closely across stages of lifetime.
 The differences between Models 3 and 4 and Models 1 and 2 are visually quite apparent.
@@
 There are many possible approaches to the posterior predictive check.
-For this problem, we chose to consider the posterior predictive distribution of the Kaplan-Meier estimate for each drive-model $g$.
+For this problem, we chose to consider the posterior predictive distribution of the K-M estimate for each drive-model $g$.
 
@@
 
-Because the variability in the Kaplan-Meier is primarily a function of the number of at-risk units, we want to replicate the pattern of censoring that was observed.
+Because the variability in the K-M estimator is primarily a function of the number of at-risk units, we want to replicate the pattern of censoring that was observed.
 That is, for unit $i$ of drive-model $g$,
@@
 We generated 19 replicates for each drive model.
-This gives a 1/20 chance under the hypothesis of no difference that the Kaplan-Meier estimate based on the observed data would be ``most extreme" relative all the replicated  from the those one and is not too many so as to produce an overly cluttered plot.
+This gives a 1/20 chance under the hypothesis of no difference that the K-M estimate based on the observed data would be ``most extreme" relative to those from the replicated data, but not too many so as to produce an overly cluttered plot.
 Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each drive-model $g=1,\ldots,44$, is as follows:
@@
 \includegraphics[width=\textwidth]{ppcheck-sample}
-\caption{Adjusted Kaplan-Meier estimates of the proportion failing for a representative subset of drive-models.
+\caption{Adjusted K-M estimates of the proportion failing for a representative subset of drive-models.
 Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
@@
 
-We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the Kaplan-Meier estimator.
+We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the K-M estimator.
 The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
@@
 There are a few drive-models, however where this is not the case.
-Drive-model 14 shows a discrepancy between the GFLP fit and the Kaplan-Meier fit in the right tail.
+Drive-model 14 shows a discrepancy between the GFLP fit and the K-M fit in the right tail.
 In particular, the GLFP predictions appear conservative, overpredicting the proportion failing after about 2.5 years.
@@
 To diagnose the cause of the discrepancies between the observed data for Drive-model 4 and its posterior predictive distribution, we consider the event plot for a random sample of units for this Drive-model while looking at estimates of lifetime (Figure \ref{fig:ex-mod-4}).
-From these plots, we can see that the discrepancy between the posterior and Kaplan-Meier estimates is driven by failures among a relatively small subset of units, namely those with left-truncation times smaller than 2500 hours.
-Further investigation reveals that, if we exclude the newest 75 drives, those with truncation times less than 800 hours, the Kaplan-Meier estimate falls within the posterior 90\% pointwise credible band.
+From these plots, we can see that the discrepancy between the posterior and K-M estimates is driven by failures among a relatively small subset of units, namely those with left-truncation times smaller than 2500 hours.
+Further investigation reveals that, if we exclude the newest 75 drives, those with truncation times less than 800 hours, the K-M estimate falls within the posterior 90\% pointwise credible band.
 This situation suggests that, despite belonging to the same Drive-model, there may be something different about the newer drives leading to worse reliability.
@@
 \caption{\footnotesize Top: Event plot for 60 randomly sampled drives of Drive-model 4.
-Note that all drives in the sample are left truncated well past 1000 hours.\hspace{\textwidth}Bottom: The adjusted Kaplan-Meier estimate (dashed step function) is substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval).
+Note that all drives in the sample are left truncated well past 1000 hours.\hspace{\textwidth}Bottom: The adjusted K-M estimate (dashed step function) is substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval).
 However, this discrepancy is due to several early failures among the small set of drives with the earliest left-truncation times (the newest set).
-The same adjusted Kaplan-Meier estimator after the exclusion of the newest 75 drives (solid step function) shows close agreement with the posterior median.}
+The same adjusted K-M estimator after the exclusion of the newest 75 drives (solid step function) shows close agreement with the posterior median.}
 
@@
 We can see that the lack of data for this Drive-model during early life results in a diffuse posterior centered around the global average until about 10,000 hours when they diverge.
-Although the posterior is close to the adjusted Kaplan-Meier, it shows shrinkage toward the global average.
+Although the posterior is close to the adjusted K-M, it shows shrinkage toward the global average.
 
@@
 \caption{\footnotesize Top: Event plot for all units of Drive-model 43.
-Data for all units are left-truncated, being observed conditional on survival beyond 12,000 hours.\hspace{\textwidth}Bottom: The dashed step function correponds to the adjusted Kaplan-Meier estimate.
+Data for all units are left-truncated, being observed conditional on survival beyond 12,000 hours.\hspace{\textwidth}Bottom: The dashed step function correponds to the adjusted K-M estimate.
 Due to the heavy left-truncation, the posterior median for Drive-model 43 (dashed line) coincides with the posterior median of the ``global model" (dotted line) until the first left-truncation time.}
@@
 \label{sec:trunc-adj}
-We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator.
+We first start with a nonparametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator.
 With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by

### 2017-10-06 paper/paper.tex
edits in morning with eric.

--- 2017-10-06 paper/paper.tex
@@
 \usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf,float, bm, subcaption}
+\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url}
 \graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
@@
 
-\subsection{Censoring and Truncation}
+\subsection{Censoring and truncation}
 A common feature of lifetime data is right-censoring.
@@
 We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to\citet[Chapter 11, ]{meeker} (see Appendix for a detailed description).
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (see Appendix for a detailed description).
 
@@
 \subsection{Modeling}
+\subsubsection{Hierarchical models}
 We considered four models, all based on (\ref{eq:hier-model}), which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):
@@
 
-\subsubsection{Prior Distributions}
-\label{sec:Prior Distributions}
+\subsubsection{Prior distributions}
+\label{sec:Prior distributions}
 To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model.
@@
 \subsection{Results}
-\subsubsection{Model Assessment}
+\subsubsection{Model assessment}
 \label{model-assessment}
@@
 
-\subsubsection{Drive-model Comparisons}
-\label{subsubsec:Drive-model Comparisons}
+\subsubsection{Drive-model comparisons}
+\label{subsubsec:Drive-model comparisons}
 We consider the problem of ranking drive-models. From a
@@
 In this application 50\% CIs are more meaningful as there is lots of uncertainty in some of the parameters: especially for those with few drives in operation.
-The best six drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
+The top seven drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
 Another interesting feature to compare is $\pi$, the proportion defective.
-The bottom panel of Figure 10 shows the posterior credible intervals of $\pi$ for each drive-model.
+The bottom panel of Figure 9 shows the posterior credible intervals of $\pi$ for each drive-model.
 The plot is again sorted according to $t_{0.10}$ so it can be compared to the plot above.
@@
   \includegraphics[width=.8\textwidth]{dm-eval}
-  \caption{Left: Summary of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
-  Right: A comparison of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
+  \caption{Left: 90\% CI of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each drive-model.
+  Right: A comparison (90\% CI) of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
   \label{fig5}
@@
 In practice, there might be additional phases, clearly supported by the data, making the GLFP model too rigid.
-This rigidity is illustrated by the right tail of the estimated distribution in figure \ref{fig1}.
+This rigidity is illustrated by the right tail of the estimated distribution in Figure \ref{fig1}.
 
@@
 While not presented here, another potential application is forecasting the number of future failures over a fixed period of time for a current population of drives.
-\citet{hmm} proposed a method, based on a Poisson binomial distribution.
+\citet{hmm} proposed a method, based on the Poisson-binomial distribution.
 It would be straightforward to adapt their approach to work with our method.
@@
 
-\item[Stan Code] Stan code for Model 4 (model4.pdf)
+\item[Stan Code and Plots] Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6 (supplementary.pdf)
 
-\item[Complete set of model comparison and model assessment plots] Posterior median of proportion failing as a function of time under Models 1-4 and posterior predictive graphical checks for all drive-models.
-(allplots.pdf)
-\item[Backblaze hard-drive data] Lifetime data for all 44 drive-models used in our analysis. (backblaze\_hd\_data.csv)
+\item[Backblaze hard-drive data] Lifetime data set for all 44 drive-models used in our analysis. (backblaze\_hd\_data.csv)
 

### 2017-10-06 paper/paper.tex

--- 2017-10-06 paper/paper.tex
@@
 After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.
-We applied this method using the K-M estimate of the cdf.
-With left truncation, however, the standard K-M estimator is biased, so we used an adjusted version due to\citet[Chapter 11, ]{meeker} (see Appendix for a detailed description).
+We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (see Appendix for a detailed description).
 

### 2017-10-12 paper/aosgflp.tex
New Tek formatting (aosgflp.tex. Use this from now on.

--- 2017-10-06 paper/paper.tex
@@
-\documentclass[12pt]{article}
-\usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url}
+% Template for the submission to:
+%   The Annals of Probability           [aop]
+%   The Annals of Applied Probability   [aap]
+%   The Annals of Statistics            [aos] 
+%   The Annals of Applied Statistics    [aoas]
+%   Stochastic Systems                  [ssy]
+%
+%Author: In this template, the places where you need to add information
+%        (or delete line) are indicated by {???}.  Mostly the information
+%        required is obvious, but some explanations are given in lines starting
+%Author:
+%All other lines should be ignored.  After editing, there should be
+%no instances of ??? after this line.
+
+% use option [preprint] to remove info line at bottom
+% journal options: aop,aap,aos,aoas,ssy
+% natbib option: authoryear
+\documentclass[aoas]{imsart}
+\RequirePackage{natbib}
+\usepackage{amsthm,amsmath,natbib,enumerate}
+\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url, imsart}
 \graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
-\usepackage{enumerate}
-\usepackage{natbib}
-\bibliographystyle{asa}
-\usepackage{url} % not crucial - just used below for the URL
+\bibliographystyle{imsart-nameyear}
+%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+
+% provide arXiv number if available:
+%\arxiv{arXiv:0000.0000}
 
@@
 \let\]\myendequation
-
-\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
-
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
+% put your definitions there:
+\startlocaldefs
+\endlocaldefs
 
@@
 
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
+\begin{frontmatter}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% "Title of the paper"
+\title{A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
 
-\if0\blind
-{
-  \title{\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
-  \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
-    and \\
-    Colin Lewis-Beck\footnotemark[1]\\
-    and \\
-    William Q. Meeker\footnotemark[1]}
-  \maketitle
-} \fi
+\begin{aug}
+  \author{\fnms{Eric}  \snm{Mittman}\corref{}\ead[label=e1]{first@somewhere.com}},
+  \author{\fnms{Colin} \snm{Lewis-Beck}\ead[label=e2]{second@somewhere.com}}
+  \and
+  \author{\fnms{William Q.}  \snm{Meeker}%
+  \ead[label=e3]{third@somewhere.com}%
+  \ead[label=u1,url]{http://www.foo.com}}
 
-\if1\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Title}
-\end{center}
-  \medskip
-} \fi
+  %\thankstext{t2}{Footnote to the first author with the `thankstext' command.}
 
-\bigskip
+  %\runauthor{F. Author et al.}
+
+  \affiliation{Iowa State University}
+
+  \address{Address of the First and Second authors,\\ 
+          \printead{e1,e2}}
+
+  \address{Address of the Third author,\\
+          \printead{e3,u1}}
+
+\end{aug}
+
+% indicate corresponding author with \corref{}
+% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
+% \thankstext{t1}{Thanks to somebody} 
+% \address{line 1\\ line 2\\ printead{e1}}
+%\author{\fnms{Eric Mittman and Colin Lewis-Beck and William Q. Meeker}}
+%\affiliation{Iowa State University}
+%\address{\printead{e1}}
+%\and
+%\author{\fnms{???} \snm{???}\ead[label=e2]{???}}
+%\address{\printead{e2}}
+%\affiliation{???}
+
+%\runauthor{???}
+
 \begin{abstract}
-% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
 When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
@@
 Potential applications include assessment and comparison of reliability to inform purchasing decisions.
+
 \end{abstract}
 
-\noindent%
-{\it Keywords:} GLFP, Bathtub hazard, Censored data, Hierarchical models, Bayesian estimation
-\vfill
+%\begin{keyword}[class=MSC]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
 
-\newpage
-\tableofcontents
-\newpage
-\spacingset{1.45} % DON'T change the spacing!
+\begin{keyword}
+\kwd{GLFP, Bathtub hazard, Censored data, Hierarchical models, Bayesian estimation}
+%\kwd{}
+\end{keyword}
+
+\end{frontmatter}
+
 \section{Introduction}
@@
 Nonparametric methods require relatively few assumptions, but they are not suitable for prediction.
-Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be practical.
+Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be useful.
 Appropriate parametric models can make better use of available data.
@@
 \cite{berger} introduced the Poly-Weibull distribution where the failure time is the minimum of a several Weibull distributions, corresponding to different failure modes.
-More recently, \cite{ranjan} considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions.
+\cite{ranjan} considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions.
 Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.
@@
 \label{sec:Weibull parameterization}
-For a random variable $T$ following a Weibull distribution, the standard parameterization of the cumulative distribution function, in terms of scale parameter $\alpha$ and shape parameter $\beta$, is
+The Weibull cumulative
+ distribution function (cdf) is
+\begin{align}
+\label{equation:weibull.cdf}
+\Pr(T \leq t|\alpha,\beta ) &= F(t|\alpha,\beta)=1-
+\exp \left [-\left (\frac{t}{\alpha} \right )^{\beta}
+\right ], \quad t > 0,
+\end{align}
+where $\beta>0$ is the Weibull shape parameter and $\alpha>0$ is a
+scale parameter. Because $\log(T)$ has a smallest extreme value
+distribution (a member of the location-scale family of
+distributions), the Weibull cdf can also be written as
+\begin{align*}
+\Pr(T \leq t| \mu,\sigma ) &= F(t| \mu,\sigma)= \Phi_{SEV}\left[\frac{\log(t)-\mu}{\sigma}\right], \quad t > 0,
+\end{align*}
+where $\Phi_{SEV}(z)=1-\exp[-\exp(z)]$ is the standard smallest extreme value
+distribution cdf and
+$\mu=\log(\alpha)$ and $\sigma=1/\beta$ are, respectively, location
+and scale parameters for the distribution of $\log(T)$. Thus the Weibull
+distribution is a member of the log-location-scale family of
+distributions.
 
-$$ \Pr(T \le t) = F(t|\alpha,\beta) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$
-
-An alternative parameterization, in terms of log-location parameter $\mu = \log (\alpha)$ and log-scale parameter $\sigma = 1/\beta$ is
-
-\begin{equation} \Pr(T \le t) = F(t|\mu, \sigma) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
-= \Phi_{SEV}\left[\frac{\log (t) - \mu}{\sigma}\right],
-\end{equation}
-where $\Phi_{SEV}$ is the cdf of the standard smallest extreme value distribution (SEV).
-Thus, the Weibull distribution is a log-location scale family, with $\log(T)$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$.
-% With this in mind, the corresponding density function is
-% \begin{equation}
-% f(t|\mu, \sigma) =\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
-% \end{equation}
-
-An advantage conferred by writing the model in this way is that the log of the $p$ quantile, $t_p$, is a linear function of the parameters, namely, $\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p)$.
-This allows an analyst to apply prior information in a straightforward way.
-Reparameterizing in terms of $(\log(t_p),\sigma)$, we can write the Weibull cdf as
-\begin{equation}
-\Pr(T \le t) = F(t|\text{log}(t_p), \sigma) = \Phi_{SEV}\left[\frac{\log(\frac{t}{t_p}) + \sigma \Phi_{SEV}^{-1}(p)}{\sigma}\right]
-\end{equation}
-When there is more prior knowledge about a lower quantile, as is often the case with high-reliability products, using this parameterization allows the analyst to put prior distributions directly on those quantities.
-In addition, in the presence of censoring, $p$ can be chosen to make $\log(t_p)$ and $\sigma$ approximately independent.
-Since the choice of independent marginal priors is common in practice, this observation lends further credence to the use of these reparameterizations.
+For purposes of specifying a prior distribution for the Weibull
+distribution parameters, we will use an alternative parameterization
+where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
+[-\log(1-p)\right ]^{1/\beta}$. Replacing $\alpha$ in (1) with the
+$\alpha=t_{p}/[-\log(1-p)]^{1/\beta}$ gives
+\begin{align*}
+\nonumber
+\Pr(T \leq t|t_{p},\beta ) = F(t|t_{p},\beta)&=1-
+\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{1/\beta}} \right )^{\beta}
+\right ]\\
+\label{equation:reparameterized.weibull.cdf}
+&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{\beta}\right], \quad t > 0.
+\end{align*}
+There are two important reasons for using this parameterization.
+\begin{itemize}
+\item
+Especially with a high-reliability product, it will be easier to
+elicit prior information about a quantile in the lower tail of the
+distribution than it will be to elicit prior information about
+$\alpha$. There is, however, generally available
+information the shape parameter $\beta$ for a given failure
+mechanism (e.g., if the failure is due to a wearout mechanism, the
+it is known that $\beta>1$). 
+\item
+Because of heavy censoring in reliability field data, the parameter
+estimates of the $\alpha$ and $\beta$ parameters will generally be
+highly correlated and thus specification of independent marginal
+prior distributions would be be inappropriate. On the other hand,
+estimates of $t_{p}$ and $\beta$, for some appropriately chosen
+value of $p$, will be approximately independent, allowing the easier
+elicitation and specification of independent marginal prior
+distributions. For example, if a typical field data set has 10\%
+failing, then choosing $t_{0.05}$ would 
+work well.
+\end{itemize}
 
@@
 $$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
+\beta_1, \beta_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.7 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 \centering
-\begin{tabular}{rrrr}
+\begin{tabular}{lrrr}
   \hline
@@
   \hline
-$\pi$ & 0.03 & 0.05 & 0.10 \\ 
+$\pi$ & 0.033 & 0.054 & 0.099 \\ 
  $\beta_1 = 1/\sigma_1 $ & 0.47 & 1.13 & 1.72 \\ 
   $\beta_2 = 1/\sigma_2$ & 4.47 & 4.70 & 4.95 \\ 
-  $t_{0.5,1}$ & 1,029.42 & 2,279.09 & 3,989.76 \\ 
-  $t_{0.2,2}$ & 18,010.30 & 18,273.22 & 18,553.22 \\ 
+  $t_{0.5,1}$ & $1.03\times 10^3$  & $2.28 \times 10^3$ & $3.99 \times 10^3$\\ 
+  $t_{0.2,2}$ & $1.80 \times 10^4$ & $1.82 \times 10^4$ & $1.86 \times 10^4$ \\ 
    \hline
@@
 In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted K-M estimate with axes on the Weibull probability scale.
-90\% pointwise uncertainty bands associated with each estimator are included.
+The plot also contains 90\% pointwise uncertainty bands associated with each estimate.
 While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours.
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution.
-We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by``$T(0,1)$" above) restricting the wearout failure mode to have an increasing hazard function (e.g., $\beta_{2g} = 1/\sigma_{2g}>1$).
+We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (e.g., $\beta_{2g} = 1/\sigma_{2g}>1$).
 We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
@@
 We can also compare posterior estimates of time to failure under Models 1-4 for each drive-model individually.
-For these comparisons, we calculate, pointwise, the posterior median of proportion failing over a grid.
+For these comparisons, we calculate, pointwise, the posterior median of the proportion failing over a grid.
 That is, for every model, and for a fixed set, $\tilde{t}=\{\tilde{t}_1,\ldots,\tilde{t}_M\}$, we compute
@@
 On the other hand, Model 4 follows the K-M quite closely across stages of lifetime.
-The differences between Models 3 and 4 and Models 1 and 2 are visually quite apparent.
+The differences between Models 3 and 4 and Models 1 and 2 are visually apparent.
 Differentiating between Model 3 and 4 is less clear; for instance, the estimates from Model 3 and 4 are nearly identical for Drive-model 40.
@@
 \end{equation}
-\noindent Using this criterion, we should choose the Model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
-Suppressing the conditioning on $t^L$ and $c$,
+\noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
+Notationally suppressing the conditioning on $t^L$ and $c$,
 \begin{equation}
@@
 The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
-It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\frac{\sum_{i=1}^S w_{i,s} p(t_{g,i})}{\sum_{s=1}^S w_{i,s}}$.
+It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\sum_{i=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$.
 In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.
@@
 The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
-Only a representative sample of plots is shown here for brevity (see the Supplementary Material for the complete set of drive-models).
+Only a representative sample of plots is shown here for brevity (see the Supplementary Material for similar plots for all of the drive-models).
 For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.
@@
 Other drive-models showing a lack of fit are 4 and 34.
-We first address Drive-model 4; our diagnosis of 34 is analogous, so the following comments extend to that Drive-model as well.
+We first address Drive-model 4; our diagnosis of 34 is analogous, so the following comments extend to that drive-model as well.
 
@@
 This situation suggests that, despite belonging to the same Drive-model, there may be something different about the newer drives leading to worse reliability.
+
+A possible explanation for the discrepancy shown in Figure 7 is that
+the units of Drive-model 4 were introduced into the field over a
+period of xx months. Manufactures of hard drives (and most other
+products) often make changes to the product design over time to
+either improve reliability or to reduce cost. Such changes will
+sometimes have an important effect on the products failure-time
+distribution, invalidating one of our important assumptions that
+failure-time distribution for a drive-model does not depend on the
+date of manufacture.
 
@@
 \label{subsubsec:Drive-model comparisons}
-We consider the problem of ranking drive-models. From a
+We consider the problem of ranking drive-models with respect to economical value. From a
 business perspective, it is clear
@@
 
-There are two sources of variation at play; the
+There are two sources of variation at play: the
 posterior uncertainty in the parameters, which largely depends on the
@@
 
-\section{Supplementary Material}
+\section{Online Supplementary Material}
 \label{sec:supple}
+\begin{supplement}[id=suppA]
+  \sname{Supplement A}
+  \stitle{Stan Code and Plots}
+  \slink[doi]{COMPLETED BY THE TYPESETTER}
+  \sdatatype{.pdf}
+  \sdescription{Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6.}
+\end{supplement}
 
-\begin{description}
-
-\item[Stan Code and Plots] Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6 (supplementary.pdf)
-
-\item[Backblaze hard-drive data] Lifetime data set for all 44 drive-models used in our analysis. (backblaze\_hd\_data.csv)
-
-\end{description}
+\begin{supplement}[id=suppB]
+  \sname{Supplement B}
+  \stitle{Backblaze Data}
+  \slink[doi]{COMPLETED BY THE TYPESETTER}
+  \sdatatype{.csv}
+  \sdescription{Lifetime data set for all 44 drive-models used in our analysis.}
+\end{supplement}
 
@@
 
+% AOS,AOAS: If there are supplements please fill:
+%\begin{supplement}[id=suppA]
+%  \sname{Supplement A}
+%  \stitle{Title}
+%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
+%  \sdatatype{.pdf}" 
+%  \sdescription{Some text}
+%\end{supplement}
+
 \end{document}

### 2017-10-12 paper/ims-template.tex
New Tek formatting (aosgflp.tex. Use this from now on.

--- 2017-10-12 paper/aosgflp.tex
@@
 
-\begin{aug}
-  \author{\fnms{Eric}  \snm{Mittman}\corref{}\ead[label=e1]{first@somewhere.com}},
-  \author{\fnms{Colin} \snm{Lewis-Beck}\ead[label=e2]{second@somewhere.com}}
-  \and
-  \author{\fnms{William Q.}  \snm{Meeker}%
-  \ead[label=e3]{third@somewhere.com}%
-  \ead[label=u1,url]{http://www.foo.com}}
-
-  %\thankstext{t2}{Footnote to the first author with the `thankstext' command.}
-
-  %\runauthor{F. Author et al.}
-
-  \affiliation{Iowa State University}
-
-  \address{Address of the First and Second authors,\\ 
-          \printead{e1,e2}}
-
-  \address{Address of the Third author,\\
-          \printead{e3,u1}}
-
-\end{aug}
-
 % indicate corresponding author with \corref{}
@@
 % \address{line 1\\ line 2\\ printead{e1}}
-%\author{\fnms{Eric Mittman and Colin Lewis-Beck and William Q. Meeker}}
-%\affiliation{Iowa State University}
+\author{\fnms{Eric Mittman and Colin Lewis-Beck and William Q. Meeker}}
+ \affiliation{Iowa State University}
 %\address{\printead{e1}}

### 2017-10-12 paper/aosgflp.tex
Move definition of `global average' to the Appendix and improve the clarity of the def.

@@
 The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.
-In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for Drive-model 43 to what we call the `global average.' We get this by setting the $\op{GLFP}$ parameters equal to the median of their hierarchical distribution, i.e., $\pi=\op{logit}^{-1}(\eta_{\pi}),\, \mu_1=\mu_1,\, \sigma_1=\sigma_1,\, \mu_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2)$ and $\sigma_2= m_{\sigma_2}$, where $m_{\sigma_2}=\Phi_{Log-Normal}^{-1}[0.5 \cdot \Phi_{Log-Normal}(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}]$ is the population median of the $\sigma_{2g}$.
-The curve was estimated using pointwise medians as in Equation \ref{pointwise-medians}.
+In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for Drive-model 43 to an average GLFP cdf for the entire population, which we call the `global average'; a description of the global average is given in Appendix \ref{global-avg}.
 We can see that the lack of data for this Drive-model during early life results in a diffuse posterior centered around the global average until about 10,000 hours when they diverge.
-Although the posterior is close to the adjusted K-M, it shows shrinkage toward the global average.
+Although the posterior is close to the adjusted K-M, it is shrunk toward the global average.
 
@@
 
+\section{Definition of global average for Model 4}
+\label{global-avg}
+In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a `global average' which represents an average GLFP cdf for the entire population, $\op{GLFP}\left(\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
+For the parameters that vary across drive models, we select values corresponding to the medians of the hierarchical distributions (Eq.
+\ref{eq:hier-model}) conditional on the hyperparameters.
+In particular, setting
+$$\bar{\pi}=\op{logit}^{-1}(\eta_{\pi}),\;\bar{\mu}_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2) \mbox{ and } \bar{\sigma}_2= m_{\sigma_2}.$$
+Let $F(\cdot|a,b),\;F^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a log-normal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
+$$m_{\sigma_2}=F^{-1}[0.5 \cdot F(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}],$$
+which is the median of a log-normal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.
+
+We estimate the global average pointwise, using draws from the joint posterior distribution, $\left(\eta_\pi^{(s)}, \mu_1^{(s)}, \sigma_1^{(s)}, \eta_{t_2}^{(s)},\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$, $s=1,2,\ldots,S$.
+The computation is thus similar to that shown in Equation \ref{pointwise-medians}.
+ % We get this by setting the $\op{GLFP}$ parameters equal to the median of their hierarchical distribution, i.e., $\pi=\op{logit}^{-1}(\eta_{\pi}^{(s)}),\, \mu_1=\mu_1^{(s)},\, \sigma_1=\sigma_1^{(s)},\, \mu_2=\eta_{t_2}^{(s)} - m_{\sigma_2}^{(s)} \Phi_{SEV}^{-1}(.2)$ and $\sigma_2= m_{\sigma_2}^{(s)}$, where $m_{\sigma_2}^{(s)}=F^{-1}[0.5 \cdot F(1;\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)});\eta_{\sigma_2}^{(s)}, \tau_{\sigma_2}^{(s)}]$ is the population median of the $\sigma_{2g}$, conditioning on $\left(\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$.
+The curve was estimated using pointwise medians as in Equation \ref{pointwise-medians}.
 \section{Online Supplementary Material}

### 2017-10-13 paper/supplement.tex
Edits to supplement.  Think about adding a folder to the Git Repo with Supplemental Material.

@@
 
-\item Put R Stan code here
+\item Stan code for Model 4, estimated using the Backblaze data set (also available as a supplementary .csv file).
 
@@
 \section{Posterior predictive graphical checks}
+\item Full set of Figures 5 and 6 for all 44 drive-models.
+Numbering in the top left indicates drive-model.
+Left: Adjusted K-M estimates of the proportion failing.
+The original data (bold line) and 19 ``replicated" data sets from the posterior predictive distribution (dashed lines) are shown.
+Right: Posterior median of proportion failing as a function of time for Models 1-4.
+Plotted on Weibull probability scales.
+The solid step function is the adjusted K-M estimate.
 \begin{figure}[H]

### 2017-10-18 paper/aosgflp.tex
Edits to supplemental section.
First pass at Nick and WQM edits.  Fixed bib too, and added cities.

@@
 
-  \address{Address of the First and Second authors,\\ 
-          \printead{e1,e2}}
+  %\address{Address of the First and Second authors,\\ 
+   %       \printead{e1,e2}}
 
-  \address{Address of the Third author,\\
-          \printead{e3,u1}}
+  %\address{Address of the Third author,\\
+   %       \printead{e3,u1}}
 
@@
 Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate.
+Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
 Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision.
@@
 Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work in the literature.
-Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data.
-In Section~\ref{sec:Data}, we describe the Backblaze Hard Drive data set, which is used throughout as a motivating example.
+Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of hard drive failure data from the company, Backblaze.
+In Section~\ref{sec:Data}, we describe the complete Backblaze Hard Drive data set, which is used throughout as a motivating example.
 In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations.
@@
 A common assumption in series systems is the time to failure for each risk is statistically independent.
-Thus, the overall reliability of a unit is modeled using the product rule across all risks.
+Thus, the overall reliability of a unit is the product of all the risks.
 Parameter estimation is straightforward if the cause of failure is known for each observation.
@@
 
-\section{Generalized Limited Failure Population model}
+\section{Generalized Limited Failure Population Model}
 \label{sec:GLFP model}
@@
 
-For purposes of specifying a prior distribution for the Weibull
-distribution parameters, we will use an alternative parameterization
+We will use an alternative parameterization
 where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
-[-\log(1-p)\right ]^{1/\beta}$. Replacing $\alpha$ in (1) with the
-$\alpha=t_{p}/[-\log(1-p)]^{1/\beta}$ gives
+[-\log(1-p)\right ]^{\sigma}$. Replacing $\alpha$ in (1) with the
+$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
 \begin{align*}
 \nonumber
-\Pr(T \leq t|t_{p},\beta ) = F(t|t_{p},\beta)&=1-
-\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{1/\beta}} \right )^{\beta}
+\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
+\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
 \right ]\\
 \label{equation:reparameterized.weibull.cdf}
-&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{\beta}\right], \quad t > 0.
+&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
 \end{align*}
@@
 Especially with a high-reliability product, it will be easier to
-elicit prior information about a quantile in the lower tail of the
+elicit prior information about a quantile  ($t_p$) in the lower tail of the
 distribution than it will be to elicit prior information about
 $\alpha$. There is, however, generally available
-information the shape parameter $\beta$ for a given failure
+information the shape parameter $\sigma$ for a given failure
 mechanism (e.g., if the failure is due to a wearout mechanism, the
-it is known that $\beta>1$). 
+it is known that $\sigma<1$). 
 \item
 Because of heavy censoring in reliability field data, the parameter
-estimates of the $\alpha$ and $\beta$ parameters will generally be
+estimates of the $\mu$ and $\sigma$ parameters will generally be
 highly correlated and thus specification of independent marginal
 prior distributions would be be inappropriate. On the other hand,
-estimates of $t_{p}$ and $\beta$, for some appropriately chosen
+estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
 value of $p$, will be approximately independent, allowing the easier
@@
 \label{subsec:GLFP model}
-Let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$, respectively.
-The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
+Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p1},\sigma_1)$ and $(t_{p2}, \sigma_2)$, respectively.
+The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$. Then
 $$\Pr(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
@@
 \begin{equation}
-f(t|\pi, \mu_1,\sigma_1, \mu_2, \sigma_2) = \pi f_1(t|\mu_1,\sigma_1)\left(1-F_2(t|\mu_2,\sigma_2)\right) + f_2(t|\mu_2,\sigma_2)\left(1-\pi F_2(t|\mu_2,\sigma_2\right).
+f(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
 \end{equation}
@@
 Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$).
-The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.
+The likelihood for the GLFP is a function of $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.
 Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
@@
 $$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\beta_1, \beta_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.7 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
+\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.
-The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.
+The parameter $\pi$ is an estimate of the proportion of drives susceptible to to infant mortality.
 As expected, this fraction is small with a median value of 0.05.
@@
 \begin{equation}
-y_{ig} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
+y_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
 \end{equation}
 where $g=1,\ldots,G$ indexes the sub-populations.
-The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g}, \pi_{g})$.
+The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.
 Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
@@
 \begin{enumerate}
-\item $\pi_{g} = \pi,\quad \mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1,\quad \mu_{2g} = \mu_2,\quad \sigma_{2g} = \sigma_2$
-\item $\pi_{g} = \pi,\quad \mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1,\quad \sigma_{2g} = \sigma_2$
-\item $\pi_{g} = \pi,\quad \mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1$
-\item $\mu_{1g} = \mu_1,\quad \sigma_{1g}=\sigma_1$
+\item $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad t_{p_{2}g} = t_{p2},\quad \sigma_{2g} = \sigma_2$
+\item $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad \sigma_{2g} = \sigma_2$
+\item $\pi_{g} = \pi,\quad  t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
+\item $t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
 \end{enumerate}
@@
 The set of model specifications was chosen based data, interpretation of the model, as well as estimation considerations.
-Product specific parameters for the the wearout failure mode $(\mu_{2g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
+Product specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
 Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
@@
 \label{pointwise-medians}
-\op{median}\left\{\op{GLFP}\left(\tilde{t}_m|\pi_{g}^{(s)},\mu_{1g}^{(s)},\sigma_{1g}^{(s)},\mu_{2g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
+\op{median}\left\{\op{GLFP}\left(\tilde{t}_m|\pi_{g}^{(s)},t_{p_{1}g}^{(s)},\sigma_{1g}^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
 \end{equation}
@@
 \begin{enumerate}
-\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, \\$\left( \pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)} \right)$, is a random draw from the full posterior distribution.
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, \\$\left( \pi_{g}^{(s)},t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)}, \sigma_{2g}^{(s)} \right)$, is a random draw from the full posterior distribution.
 \item Repeat step 1 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
@@
 the units of Drive-model 4 were introduced into the field over a
-period of xx months. Manufactures of hard drives (and most other
+period of 5 months. Manufactures of hard drives (and most other
 products) often make changes to the product design over time to
@@
 
-Figure 8 shows a similar pair of plots for Drive-model 43 illustrate (Figure \ref{fig:ex-mod-43}).
+Figure 8 shows a similar pair of plots for Drive-model 43 illustrate.
 The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.
@@
 The curve was estimated using pointwise medians as in Equation \ref{pointwise-medians}.
-\section{Online Supplementary Material}
-\label{sec:supple}
-\begin{supplement}[id=suppA]
+
+\begin{supplement}
 \sname{Supplement A}
-  \stitle{Stan Code and Plots}
-  \slink[doi]{COMPLETED BY THE TYPESETTER}
-  \sdatatype{.pdf}
+\stitle{Model code and plots}
+\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
 \sdescription{Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6.}
 \end{supplement}
-
-\begin{supplement}[id=suppB]
+\begin{supplement}
 \sname{Supplement B}
-  \stitle{Backblaze Data}
-  \slink[doi]{COMPLETED BY THE TYPESETTER}
-  \sdatatype{.csv}
+\stitle{Backblaze data set}
+\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
 \sdescription{Lifetime data set for all 44 drive-models used in our analysis.}
 \end{supplement}
+
+%\begin{supplement}
+  %\sname{Supplement A}
+  %\stitle{Stan Code and Plots}
+  %%\slink[doi]{COMPLETED BY THE TYPESETTER}
+  %\sdatatype{.pdf}
+  %\sdescription{Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6.}
+%\end{supplement}
+%
+%\begin{supplement}[id=suppB]
+  %\sname{Supplement B}
+  %\stitle{Backblaze Data}
+  %\slink[doi]{COMPLETED BY THE TYPESETTER}
+  %\sdatatype{.csv}
+  %\sdescription{Lifetime data set for all 44 drive-models used in our analysis.}
+%\end{supplement}
 

### 2017-10-18 paper/aosgflp.tex
Fix G to H for GLFP CDF.  Also make changes to d to g for groups.

@@
 The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$. Then
-$$\Pr(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
+$$\Pr(T \le t) = H(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
 
@@
 \begin{equation}
-f(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
+h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
 \end{equation}
@@
 \begin{equation*}
-L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(t_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
+L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
 \end{equation*}
@@
 \begin{equation*}
-L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
+L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
 \end{equation*}
@@
 
-\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
+\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
 
-\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
+\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
 
@@
 \begin{equation*}
-  p(t_{d,new}|t) = \int p(t_{d,new}|\theta_d) p(\theta_d|t_d) \, d\theta_d
+  p(t_{g,new}|t) = \int p(t_{g,new}|\theta_g) p(\theta_g|t_g) \, d\theta_g
 \end{equation*}
 
-We can sample from this distribution by drawing $t_{d,new}^{(s)}$ from
-$\operatorname{GLFP}(\pi_d^{(s)},
-\mu_1^{(s)},\sigma_1^{(s)},\mu_{2,d}^{(s)},\sigma_{2,d}^{(s)})$, for
+We can sample from this distribution by drawing $t_{g,new}^{(s)}$ from
+$\operatorname{GLFP}(\pi_g^{(s)},
+t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)})$, for
 $s=1,...,S$, using the saved posterior draws for the model
 parameters. The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
-$$\frac{1}{S} \sum_{s=1}^{S} t_{d,new}^{(s)}.$$
+$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}.$$
 
@@
 relative to a new unit. We can rank the drive-models by 
-$$\op{E}(L(t_{d,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{d,new}^{(s)}).$$
+$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)}).$$
 
-Posterior credible intervals for $L(t_{d,new})$ are shown in the left panel of Figure \ref{fig5}.
-A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{d,new})|y)$ are shown in the bottom right panel.
+Posterior credible intervals for $L(t_{g,new})$ are shown in the left panel of Figure \ref{fig5}.
+A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel.
 The panel above shows credible intervals for MTTF.
@@
 We first start with a nonparametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator.
-With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
-$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
-$t_{d,\text{min}}^L$, the shortest reported running time of all units
-of drive-model $d$ for which records are available. To produce
+With left truncation, however, the standard Kaplan-Meier estimator for drive-model $g$, denoted by
+$\widehat{F_g(t)}_{KM}$, is conditional on survival up to
+$t_{g,\text{min}}^L$, the shortest reported running time of all units
+of drive-model $g$ for which records are available. To produce
 unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each drive-model we select
-$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
+$t_{g,\text{min}}^L$, the smallest left truncated time in the sample.
 By sampling from the full posterior distribution, since
-$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
-survived up to $t_{d,\text{min}}^L$) is a function of the model
+$\Pr(T>t_{g,\text{min}}^L|\theta_g)$ (the probability a hard drive has
+survived up to $t_{g,\text{min}}^L$) is a function of the model
 parameters, we can easily compute its posterior median,
-$\widehat{A}_{\text{med}} = \widehat{\Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by
+$\widehat{A}_{\text{med}} = \widehat{\Pr}(T>t_{g,\text{min}}^L|\theta_g)$. We compute the adjusted estimate by
 
-$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_d(t)}_{KM},\; t>t_{d,\text{min}}^L.$$
+$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_g(t)}_{KM},\; t>t_{g,\text{min}}^L.$$
 

### 2017-10-19 paper/aosgflp.tex
Modify appendix

--- 2017-10-12 paper/aosgflp.tex
@@
 
-% \section*{Acknowledgment}    
-% The authors express their gratitude to Dr. \hspace{-.5em} Meeker for his comments and suggestions.
+\section*{Acknowledgment}    
+The authors express their gratitude to Nick Clark for his comments and suggestions.
 
@@
 \label{global-avg}
-In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a `global average' which represents an average GLFP cdf for the entire population, $\op{GLFP}\left(\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
-For the parameters that vary across drive models, we select values corresponding to the medians of the hierarchical distributions (Eq.
-\ref{eq:hier-model}) conditional on the hyperparameters.
+In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a `global average' which represents an average GLFP cdf for the entire population, $\op{H}\left(\cdot|\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
+Since $\mu_1$ and $\sigma_1$ are shared across drive-models, these can already be interpreted as `global'.
+For the parameters that vary across drive models, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.
 In particular, setting
@@
 
-We estimate the global average pointwise, using draws from the joint posterior distribution, $\left(\eta_\pi^{(s)}, \mu_1^{(s)}, \sigma_1^{(s)}, \eta_{t_2}^{(s)},\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$, $s=1,2,\ldots,S$.
-The computation is thus similar to that shown in Equation \ref{pointwise-medians}.
+We estimate the global average pointwise, using draws from the joint posterior distribution, $H\left(\tilde{t}|\eta_\pi^{(s)}, \mu_1^{(s)}, \sigma_1^{(s)}, \eta_{t_2}^{(s)},\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$, $s=1,2,\ldots,S$.
+The computation is thus similar to that shown in (\ref{pointwise-medians}).
  % We get this by setting the $\op{GLFP}$ parameters equal to the median of their hierarchical distribution, i.e., $\pi=\op{logit}^{-1}(\eta_{\pi}^{(s)}),\, \mu_1=\mu_1^{(s)},\, \sigma_1=\sigma_1^{(s)},\, \mu_2=\eta_{t_2}^{(s)} - m_{\sigma_2}^{(s)} \Phi_{SEV}^{-1}(.2)$ and $\sigma_2= m_{\sigma_2}^{(s)}$, where $m_{\sigma_2}^{(s)}=F^{-1}[0.5 \cdot F(1;\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)});\eta_{\sigma_2}^{(s)}, \tau_{\sigma_2}^{(s)}]$ is the population median of the $\sigma_{2g}$, conditioning on $\left(\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$.

### 2017-10-19 paper/aosgflp.tex
Small edits, months correction

@@
 information the shape parameter $\sigma$ for a given failure
-mechanism (e.g., if the failure is due to a wearout mechanism, the
+mechanism (e.g., if the failure is due to a wearout mechanism, then
 it is known that $\sigma<1$). 
@@
 %$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
-$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,7.6\times 10^6 \rangle,\quad
-t_{0.2,2} \sim \op{Lognormal}\langle 8.6, 5.6\times 10^7 \rangle$$
-$$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
+$$t_{0.5,1} \sim \op{Log-normal} \langle 1.7,7.6\times 10^6 \rangle,\quad
+t_{0.2,2} \sim \op{Log-normal}\langle 8.6, 5.6\times 10^7 \rangle$$
+$$\pi \sim \op{Logit-normal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
+\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Log-normal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 \label{pointwise-medians}
-\op{median}\left\{\op{GLFP}\left(\tilde{t}_m|\pi_{g}^{(s)},t_{p_{1}g}^{(s)},\sigma_{1g}^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
+\op{median}\left\{\op{H}\left(\tilde{t}_m|\pi_{g}^{(s)},t_{p_{1}g}^{(s)},\sigma_{1g}^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
 \end{equation}
@@
 the units of Drive-model 4 were introduced into the field over a
-period of 5 months. Manufactures of hard drives (and most other
+period of 22 months. Manufactures of hard drives (and most other
 products) often make changes to the product design over time to
@@
 For the parameters that vary across drive models, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.
-In particular, setting
+In particular, we set
 $$\bar{\pi}=\op{logit}^{-1}(\eta_{\pi}),\;\bar{\mu}_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2) \mbox{ and } \bar{\sigma}_2= m_{\sigma_2}.$$
-Let $F(\cdot|a,b),\;F^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a log-normal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
-$$m_{\sigma_2}=F^{-1}[0.5 \cdot F(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}],$$
+Let $J(\cdot|a,b),\;J^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a log-normal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
+$$m_{\sigma_2}=J^{-1}[0.5 \cdot J(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}],$$
 which is the median of a log-normal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.

### 2017-10-30 paper/aosgflp.tex
Incorporate edits to paper

@@
 
-The reliability of many engineered products follow a similar pattern.
-Relatively high rates of failure occurring early (``infant mortality'') due to manufacturing defects.
-After this ``burn-in'' period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.
+The reliability of many engineered products follows a similar pattern.
+Relatively high rates of failure occurring early (``infant mortality'') is due to manufacturing defects.
+After this ``burn-in'' period, failure rates stabilize after the majority of defective units have failed and are no longer in the population.
 Finally, after prolonged use, rates of failure increase due to wearout.
@@
 
-Nonparametric methods require relatively few assumptions, but they are not suitable for prediction.
+Nonparametric methods require relatively few assumptions, but may not be suitable for prediction when extrapolation is required.
 Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be useful.
 Appropriate parametric models can make better use of available data.
-For inference, likelihood-based methods are well-equipped to deal with truncated and/or censored data.
+For inference, likelihood-based methods are able to deal with truncated and/or censored data.
 Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across groups, and incorporate prior information.
@@
 It avoids non-identifiability by introducing a parameter representing the population proportion defective.
-Where this parameter is zero, the GLFP model reduces to a unimodal distribution.
+When this parameter is zero, the GLFP model reduces to a unimodal distribution.
 \\
 
-The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.
+The GLFP model has a meaningful parameterization and accommodates lifetime data with both infant mortality and wearout.
 Unfortunately, it can require a lot of data to fit due to the model's complexity.
@@
 The structure of the paper is as follows.
-Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work in the literature.
+Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work.
 Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of hard drive failure data from the company, Backblaze.
@@
 Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
-In Section~\ref{sec:Discussion}, we suggest potential applications of our model and consider limitation for inference implied by the model assumptions.
+Section~\ref{sec:Discussion} describes potential applications of our model and consider limitation for inference implied by the model assumptions.
 
@@
 
-The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$.
-$\delta_i$ is an indicator for whether or not an observational unit is defective.
-Expressed conditional on $\delta_i$,
+The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$.
+$\delta_i$ is an indicator for whether or not a unit $i$ is defective.
+Expressed conditional on~$\delta_i$,
 \begin{align*}
@@
 Taking the derivative of the (marginal) cdf, the density for the GLFP is
-\begin{equation}
-h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
-\end{equation}
+\begin{align*}
+h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
+\end{align*}
 
@@
 Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.
-Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.
+Let $t_i^L$ be the left truncation time, the age of unit $i$ when data reporting commenced.
 Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$).
-The likelihood for the GLFP is a function of $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.
+The likelihood for the GLFP is a function of the parameter $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.
 Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
@@
 Drives are run until failure.
-When a hard drive fails it is permanently removed and replaced.
+When a hard drive fails it is removed and replaced.
 In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.
-Every quarter Backblaze makes its data publicly available through their website \citep{backblaze}.
+Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, Accessed April 1, 2016)\nocite{backblaze}.
 In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.
 No other analysis or modeling of the failure data is provided.
-Backblaze does, however, encourage others to further analyze its data.
+Backblaze does, however, encourage others to further analyze their data.
 
@@
 
-\noindent These prior distributions put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.
-Thus, we consider the priors to be relatively uninformative.
+\noindent These prior distributions put probability mass on a wide range of values for all model parameters --- much larger than we would expect for a typical Weibull distribution.
+Thus, we consider these prior distributions to be relatively uninformative.
 
-Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.
-The parameter $\pi$ is an estimate of the proportion of drives susceptible to to infant mortality.
-As expected, this fraction is small with a median value of 0.05.
-The parameter $\pi$ estimates for the two Weibull distributions also match intuition.
+Table \ref{table:1} gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.
+The parameter $\pi$ is an estimate of the proportion of drives susceptible to infant mortality.
+As expected, this proportion is small with a median value of 0.05.
+The shape parameter estimates for the two Weibull distributions also match intuition.
 The infant mortality failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.
@@
 
-Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.
+Probability plotting is a simple method to assess and compare the adequacy of members of the log-location-scale families of distributions.
 After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.
@@
 The plot also contains 90\% pointwise uncertainty bands associated with each estimate.
-While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours.
+While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to adequately describe the rapid increase in the empirical cdf between 8000 and 20,000 hours.
 %The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.
@@
   \includegraphics[width=.8\textwidth]{km14-prob-plus}
-  \caption{Estimated GLFP for Drive-model 14 plotted on Weibull paper.
-Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.
-Solid line is adjusted K-M estimate with pointwise 90\% credible band in blue.}
+  \caption{The estimated GLFP for Drive-model 14 plotted on Weibull paper.
+The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are shown in red.
+The solid line is an adjusted K-M estimate with pointwise 90\% credible band in blue.}
   \label{ex1-overlay}
@@
 where $g=1,\ldots,G$ indexes the sub-populations.
-The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.
+The likelihood for the hierarchical GLFP is a function of the parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.
 Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
@@
 \end{equation*}
-where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.
+where $t_{ig}$ is the observed failure or survival time of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.
 All times are again in hours.
 
-Next, we model these parameters themselves as random variables, varying across the different sub-populations.  Let
+In the hierarchical model, the parameters are modeled as random variables, varying across the different sub-populations.  Let
 \begin{equation*}
@@
 
-As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution.
+As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of a quantile and the shape parameter because the data we encounter are more informative about the lower tail of their respective lifetime distribution (note that if $T$ has a Weibull distribution, $\sigma$ is a shape parameter for the distribution of $T$ and a scale parameter for the distribution of $\log(T)$).
 We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (e.g., $\beta_{2g} = 1/\sigma_{2g}>1$).
-We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
+%We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
 
@@
 \subsubsection{Hierarchical models}
-We considered four models, all based on (\ref{eq:hier-model}), which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):
+We considered four models, all based on (\ref{eq:hier-model}), differing by which model parameters are held constant across drive-models. These are (from most to least restrictive):
 
@@
 The set of model specifications was chosen based data, interpretation of the model, as well as estimation considerations.
-Product specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.
-Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.
+Drive-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across drive-models in the right tails of the failure distribution.
+Going from a common model for all product brands and gradually increasing the complexity of the model, we conclude that Model 4 provides the best description of the data.
 Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by drive-model.
@@
 Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).
+Also the failure-time distribution of the defective subpopulation is not of high interest (as opposed to the proportion defective.)
 
@@
 
-For Models 1-4, different sets of restrictions required different prior distribution specifications, which were assigned as follows:
+For Models 1--4, different sets of restrictions required different prior distribution specifications, which were assigned as follows:
 
-\paragraph{Model 1.} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.
+\paragraph{Model 1} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.
 
-\paragraph{Model 2.} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
+\paragraph{Model 2} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
 $$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
@@
 
-\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
+\paragraph{Model 3} $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Prior distributions for the constrained parameters are the same as for Model 2.
 
-\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
+\paragraph{Model 4} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
 
@@
 This is done to reduce autocorrelation and efficiently explore the posterior.
-Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 for a total of 6,000 post burn-in iterations and 16 chains were run for Model 4 for a total of 24,000 post burn-in iterations.
+Multiple chains were run, each with 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 for a total of 6,000 post burn-in iterations and 16 chains were run for Model 4 for a total of 24,000 post burn-in iterations.
 The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains.
@@
 
-Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of time to failure.
+Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of the cdf for each drive-model.
+To make the plot comparable, we use an adjustment based on a parametric model (Model 4).
 The large left panel shows the adjusted K-M estimate of the same.
@@
 Note that while all the paths for the posterior estimates necessarily pass through the origin, the K-M estimates do not, because, for the K-M estimator, inference is entirely conditional on survival up to left-most left-trunctation time.
-To make the plot comparable, we use an adjustment based on a parametric model (Model 4).
 
@@
 These plots suggest that Model 1 is insufficient to explain the observed data.
-Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in the rate of failure, which may be too restrictive.
+Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in cdf, which may be too restrictive.
 Model 3, which removes this restriction on $\sigma_{2,g}$ suggests that there is evidence to support variation in this parameter among drive-models.
@@
 \includegraphics[width=\textwidth]{heterogeneity-compare}
-\caption{Left: K-M estimates of the time to failure for each of the drive-models in the Backblaze data.
+\caption{Left: Adjusted K-M estimates of the time to failure for each of the drive-models in the Backblaze data.
 Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
@@
 We can also compare posterior estimates of time to failure under Models 1-4 for each drive-model individually.
-For these comparisons, we calculate, pointwise, the posterior median of the proportion failing over a grid.
+For these comparisons we calculate pointwise the posterior median of the proportion failing over a grid.
 That is, for every model, and for a fixed set, $\tilde{t}=\{\tilde{t}_1,\ldots,\tilde{t}_M\}$, we compute
@@
 
+%stopped here
 The GLFP fit for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for similar plots for the other drive-models).
 The black step functions again correspond to the adjusted K-M estimates.
-As model complexity increases, the GLFP curves become more flexible and are better able to describe the observed failure data.
+As model complexity increases, the hierarchical-model GLFP curves become more flexible and are better able to describe the observed failure data.
 All of the GLFP models appear to produce higher estimates of the proportion failing, relative to K-M, for Drive-model 2.
-This behavior is not unexpected, however, since hierarchical models improve model stability through shrinkage; in this context, shrinkage toward an average cumulative time to failure.
+This behavior is not unexpected, however, because hierarchical models improve model stability through shrinkage; in this context, shrinkage is toward an average cdf.
 Model 4 is in closest agreement.
 
-Drive-model 9 shows a high proportion of  of early failures relative to Model 1.
+Drive-model 9 shows a high proportion of  of early failures, a fact which Model 1 fails to capture.
 Model 2 addresses this by estimating an earlier time to wearout, but still results in a poor fit to the data.
-Models 3 and 4 conform more closely to the K-M estimate.
-The disagreement between Models 3 and 4 for Drive-model 9, is interesting: Model 4 allows the proportion of defective units to vary by Drive-model, but Model 3 appears to fit just as well in this case, by increasing the scale parameter for the wearout failure mode.
+Models 3 and 4 conform more closely to the adjusted K-M estimate.
+The disagreement between Models 3 and 4 for Drive-model 9, is interesting: Model 4 allows the proportion of defective units to vary by drive-model, but Model 3 appears to fit just as well in this case, by increasing the scale parameter for the wearout failure mode.
 
 For Drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for Drive-model 40, they produce higher estimates.
-On the other hand, Model 4 follows the K-M quite closely across stages of lifetime.
+On the other hand, Model 4 follows the K-M closely across stages of lifetime.
 The differences between Models 3 and 4 and Models 1 and 2 are visually apparent.
@@
 
-\begin{equation}
-\mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] = \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
-\end{equation}
+\begin{align*}
+\mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
+&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
+\end{align*}
 \noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
@@
 % The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).  
-Results showing estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:2}.
+Results showing the estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:2}.
  We calculated the difference in expected predictive accuracy for Model 4 versus 3, Model 3 versus 2, and Model 2 versus 1, as well as the standard error of the difference (Table 1).
@@
 In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
-We choose to focus on the first of these.
+We choose to focus on the first of these and apply it to individual drive-models.
 There are many possible approaches to the posterior predictive check.
@@
 $$p(t_{rep,g}) = \int \prod_{i=1}^{n_g}f(t_{new}|t_{g,i}^L,c_{g,i},\theta)\,p(\theta|t)d\theta.$$
-Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing $\theta^{(s)}$ at random, then conditionally draw $t_{rep,g}$ from $f(t_{new}|t_{g,i}^L,c_{g,i},\theta^{(s)}).$
+Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing a draw, $\theta^{(s)}$, at random, then conditionally draw $t_{rep,g}$ from $f(t_{new}|t_{g,i}^L,c_{g,i},\theta^{(s)}).$
 
@@
 In particular, the GLFP predictions appear conservative, overpredicting the proportion failing after about 2.5 years.
-Due to the large number of failures for this Drive-model, the posterior for the hierarchical model is similar to the stand-alone model.
+Due to the large number of failures for this drive-model, the posterior for the hierarchical model is similar to the stand-alone model.
 Other drive-models showing a lack of fit are 4 and 34.
@@
 Further investigation reveals that, if we exclude the newest 75 drives, those with truncation times less than 800 hours, the K-M estimate falls within the posterior 90\% pointwise credible band.
-This situation suggests that, despite belonging to the same Drive-model, there may be something different about the newer drives leading to worse reliability.
+This situation suggests that, despite belonging to the same drive-model, there may be something different about the newer drives leading to worse reliability.
 
@@
 either improve reliability or to reduce cost. Such changes will
-sometimes have an important effect on the products failure-time
+sometimes have an important effect on the product's failure-time
 distribution, invalidating one of our important assumptions that
@@
 
-Figure 8 shows a similar pair of plots for Drive-model 43 illustrate.
+Figure 8 shows a similar pair of plots for Drive-model 43.
 The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.
-In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for Drive-model 43 to an average GLFP cdf for the entire population, which we call the `global average'; a description of the global average is given in Appendix \ref{global-avg}.
-We can see that the lack of data for this Drive-model during early life results in a diffuse posterior centered around the global average until about 10,000 hours when they diverge.
+In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for drive-model 43 to an average GLFP cdf for the entire population, which we call the ``global average;" a description of the global average is given in Appendix \ref{global-avg}.
+We can see that the lack of data for this drive-model during early life results in a diffuse posterior centered around the global average until about 10,000 hours when they diverge.
 Although the posterior is close to the adjusted K-M, it is shrunk toward the global average.
@@
 posterior uncertainty in the parameters, which largely depends on the
-sample size, and the uncertainty in future
+amount of data, and the uncertainty in future
 observations conditional on the parameters. The {\em posterior
-  predictive} distribution incorporates both sources.
+  predictive} distribution incorporates both sources of variability.
 \begin{equation*}
@@
 
-While MTTF is clearly important, due to the
+While MTTF is often an important metric, due to the
 anticipation of advancements in technology, we can expect that the
@@
 Posterior credible intervals for $L(t_{g,new})$ are shown in the left panel of Figure \ref{fig5}.
-A zoomed version of the top 8 drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel.
+A zoomed version of the top eight drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel.
 The panel above shows credible intervals for MTTF.
-We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit expected to have the last the longest, but possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric.
+We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit is expected to last the longest, but the possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric.
 The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
@@
 We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a measure of central tendency.
-The top panel of Figure 9 shows $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
+The top panel of Figure 9 shows estimates of $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
 In this application 50\% CIs are more meaningful as there is lots of uncertainty in some of the parameters: especially for those with few drives in operation.
-The top seven drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
+Interestingly, the top seven drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
 Another interesting feature to compare is $\pi$, the proportion defective.
@@
    \includegraphics[width=1\linewidth]{b10new.pdf}
-   \caption{50\% CI for $t_{0.10}$}
+   \caption{Point estimate and 50\% CI for $t_{0.10}$}
    \label{fig:Ng1} 
@@
    \includegraphics[width=1\linewidth]{pinew.pdf}
-   \caption{50\% CI for $\pi$}
+   \caption{Point estimate and 50\% CI for $\pi$}
    \label{fig:Ng2}
@@
 Products with long lifetimes often have few failures and when failures do occur the cause is frequently unknown to the analyst.
-The GLFP model provides a framework to accurately model real data with evidence of multiple failure modes.
+The GLFP model provides a framework to accurately model data with evidence of multiple failure modes.
 Moreover, when there are multiple product populations within the same product class, our hierarchical modeling approach borrows strength across brands with many observed failures to help make inference for those populations with little information.
@@
 \label{global-avg}
-In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a `global average' which represents an average GLFP cdf for the entire population, $\op{H}\left(\cdot|\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
-Since $\mu_1$ and $\sigma_1$ are shared across drive-models, these can already be interpreted as `global'.
+In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a ``global average" which represents an average GLFP cdf for the entire population, $\op{H}\left(\cdot|\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
+Since $\mu_1$ and $\sigma_1$ are shared across drive-models, these can already be interpreted as ``global".
 For the parameters that vary across drive models, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.

### 2017-10-31 paper/aosgflp.tex
Final round of edits.

@@
 Relatively high rates of failure occurring early (``infant mortality'') is due to manufacturing defects.
-After this ``burn-in'' period, failure rates stabilize after the majority of defective units have failed and are no longer in the population.
+After this ``burn-in'' period, failure rates stabilize after the majority of defective units have failed.
 Finally, after prolonged use, rates of failure increase due to wearout.
@@
 \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model.
-This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life.
+This model can provide a solution to the problem of unknown cause of failure in the special case where there is evidence of two modes of failure which impact different stages of product life.
 It avoids non-identifiability by introducing a parameter representing the population proportion defective.
@@
 This enables estimation of the lifetime distribution for all populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of population specific information.
-We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.
+We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via Markov Chain Monte Carlo (MCMC).
 
@@
 In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations.
-Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard disk drive data.
+Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard drive data.
 First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation.
@@
 Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
-Section~\ref{sec:Discussion} describes potential applications of our model and consider limitation for inference implied by the model assumptions.
+Section~\ref{sec:Discussion} describes potential applications of our model and limitations for inference implied by the model assumptions.
 
@@
 $\mu=\log(\alpha)$ and $\sigma=1/\beta$ are, respectively, location
-and scale parameters for the distribution of $\log(T)$. Thus the Weibull
+and scale parameters for the distribution of $\log(T)$. Therefore, the Weibull
 distribution is a member of the log-location-scale family of
@@
 distribution than it will be to elicit prior information about
-$\alpha$. There is, however, generally available
-information the shape parameter $\sigma$ for a given failure
+$\alpha$. In addition, there is generally available
+information about the shape parameter, $\sigma$, for a given failure
 mechanism (e.g., if the failure is due to a wearout mechanism, then
@@
 elicitation and specification of independent marginal prior
-distributions. For example, if a typical field data set has 10\%
+distributions. For example, if a typical field data set has 10\% of units
 failing, then choosing $t_{0.05}$ would 
@@
 The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$.
-$\delta_i$ is an indicator for whether or not a unit $i$ is defective.
+$\delta_i$ is an indicator for whether or not unit $i$ is defective.
 Expressed conditional on~$\delta_i$,
@@
 The likelihood for the GLFP is a function of the parameter $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.
-Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
+Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
 \begin{equation*}
@@
 Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, Accessed April 1, 2016)\nocite{backblaze}.
-In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.
+In addition, Backblaze publishes summary statistics of the different drive-models currently operating.
 No other analysis or modeling of the failure data is provided.
@@
 Some drive-models have been running since 2013 or before, while others were added at a later date.
-Data have been reported on 75,297 different hard drives that are or where in operation.
-The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.
+Data have been reported on 75,297 different hard drives that are or were in operation.
+The number of drives varies by drive-model; some drive-models only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.
 Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.
@@
 The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.
-The curve has a slow ramp up in failures till about 10,000 hours (early failures), increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).
+The curve has a slow ramp up in failures till about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).
 In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.
@@
 
-Probability plotting is a simple method to assess and compare the adequacy of members of the log-location-scale families of distributions.
+Probability plotting is a simple method to assess and compare the adequacy of members of the log-location-scale family of distributions.
 After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of a quantile and the shape parameter because the data we encounter are more informative about the lower tail of their respective lifetime distribution (note that if $T$ has a Weibull distribution, $\sigma$ is a shape parameter for the distribution of $T$ and a scale parameter for the distribution of $\log(T)$).
-We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (e.g., $\beta_{2g} = 1/\sigma_{2g}>1$).
+We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (i.e., $\beta_{2g} = 1/\sigma_{2g}>1$).
 %We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
@@
 Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).
-Also the failure-time distribution of the defective subpopulation is not of high interest (as opposed to the proportion defective.)
+Also the failure-time distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.
 
@@
 To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model.
-We select proper prior distributions to ensure a proper posterior distributions.
+We select proper prior distributions to ensure proper posterior distributions.
 
@@
 Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of the cdf for each drive-model.
-To make the plot comparable, we use an adjustment based on a parametric model (Model 4).
-The large left panel shows the adjusted K-M estimate of the same.
-The adjusted K-M estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit.
+The large left panel shows the adjusted K-M estimate.
 The smaller plots correspond to Models 1 through 4.
+To make the plots comparable, we use an adjustment based on the parametric model (Model 4).
 Note that while all the paths for the posterior estimates necessarily pass through the origin, the K-M estimates do not, because, for the K-M estimator, inference is entirely conditional on survival up to left-most left-trunctation time.
@@
 These plots suggest that Model 1 is insufficient to explain the observed data.
-Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in cdf, which may be too restrictive.
-Model 3, which removes this restriction on $\sigma_{2,g}$ suggests that there is evidence to support variation in this parameter among drive-models.
+Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2$ results in very similar progressions in the cdf, which may be too restrictive.
+Model 3, which allows $\sigma_2$ to vary by group, suggests there is evidence to support variation in this parameter among drive-models.
 While Model 3 and 4 do display differences, they are relatively subtle.
@@
 
-Drive-model 9 shows a high proportion of  of early failures, a fact which Model 1 fails to capture.
+Drive-model 9 shows a high proportion of early failures, a fact which Model 1 fails to capture.
 Model 2 addresses this by estimating an earlier time to wearout, but still results in a poor fit to the data.
@@
 To account
-for depreciation, rather than using MTTF as the metric for Drive-model comparison, we use the value at replacement. The US Internal Revenue Service considers
+for depreciation, rather than using MTTF as the metric for drive-model comparison, we use the value at replacement. The US Internal Revenue Service considers
 computer hardware to be ``5-year'' equipment \citep{f4562}. Using``the declining
@@
 The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
-However, given the discrepancy of the model fit for Drive-model 4, noted in the previous subsection, in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.
+However, given the discrepancy of the model fit for Drive-model 4 (noted in the previous subsection), in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.
 
@@
 We found that the use of weakly informative priors on the hyperparameters can increase the overall efficiency of the MCMC.
-%This makes standard parametric lifetime models inaccurate and difficult to estimate using classical estimation techniques such as ML.
-
 We now address some important model assumptions and their potential for impacting statistical inference.
@@
 We do not account for batch effects or varying conditions in the facility which have impacted the observed failure rates.
-These factors are confounded with Drive-model in our analysis.
+These factors are confounded with drive-model in our analysis.
 An example of a possible batch effect is noted in Section \ref{model-assessment} and illustrated by Figure \ref{fig:ex-mod-4}.

### 2017-11-02 paper/aosgflp.tex
Final round of edits from Meeker.

@@
 Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$).
-The likelihood for the GLFP is a function of the parameter $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.
+The likelihood for the GLFP is a function of the parameters $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.
 Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
@@
   $\beta_2 = 1/\sigma_2$ & 4.47 & 4.70 & 4.95 \\ 
-  $t_{0.5,1}$ & $1.03\times 10^3$  & $2.28 \times 10^3$ & $3.99 \times 10^3$\\ 
-  $t_{0.2,2}$ & $1.80 \times 10^4$ & $1.82 \times 10^4$ & $1.86 \times 10^4$ \\ 
+  $t_{0.5,1}$ & $1.03$  & $2.28$ & $3.99$\\ 
+  $t_{0.2,2}$ & $18.0$ & $18.2$ & $18.6$ \\ 
    \hline
 \end{tabular}
-\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.}
+\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.  Quantiles are in thousands of hours.}
 \label{table:1}
@@
 The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are shown in red.
-The solid line is an adjusted K-M estimate with pointwise 90\% credible band in blue.}
+The solid line is an adjusted K-M estimate with pointwise 90\% credible bounds in blue.}
   \label{ex1-overlay}
@@
 where $g=1,\ldots,G$ indexes the sub-populations.
-The likelihood for the hierarchical GLFP is a function of the parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.
+The likelihood for the hierarchical GLFP is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each group, $g$.
 Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
@@
 \label{eq:hier-model}
-t_{p_{1}g} \equiv \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
+t_{p_{1}g} = \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
 \end{equation}
 \begin{equation*}
-t_{p_{2}g} \equiv \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
+t_{p_{2}g} = \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
 \end{equation*}
@@
 Drive-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across drive-models in the right tails of the failure distribution.
-Going from a common model for all product brands and gradually increasing the complexity of the model, we conclude that Model 4 provides the best description of the data.
+Going from a common model for all drive-models and gradually increasing the complexity of the model, we conclude that Model 4 provides the best description of the data.
 Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by drive-model.
@@
 \caption{Left: Adjusted K-M estimates of the time to failure for each of the drive-models in the Backblaze data.
-Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
+Right: Pointwise posterior median cdf estimates for Models 1-4, ordered left to right, top to bottom.}
 \label{fig:fig2}

### 2017-11-06 paper/aosgflp.tex
Small edit to abstract

@@
 Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using draws from the joint posterior distribution of the model parameters.
+Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals is easy using draws from the joint posterior distribution of the model parameters.
 Potential applications include assessment and comparison of reliability to inform purchasing decisions.

### 2017-11-27 paper/m-lb-m-technometrics.tex
Change TeX as needed for Technometrics

--- 2017-11-06 paper/aosgflp.tex
@@
-% Template for the submission to:
-%   The Annals of Probability           [aop]
-%   The Annals of Applied Probability   [aap]
-%   The Annals of Statistics            [aos] 
-%   The Annals of Applied Statistics    [aoas]
-%   Stochastic Systems                  [ssy]
-%
-%Author: In this template, the places where you need to add information
-%        (or delete line) are indicated by {???}.  Mostly the information
-%        required is obvious, but some explanations are given in lines starting
-%Author:
-%All other lines should be ignored.  After editing, there should be
-%no instances of ??? after this line.
-
-% use option [preprint] to remove info line at bottom
-% journal options: aop,aap,aos,aoas,ssy
-% natbib option: authoryear
-\documentclass[aoas]{imsart}
-\RequirePackage{natbib}
-\usepackage{amsthm,amsmath,natbib,enumerate}
-\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url, imsart}
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url}
 \graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
-\bibliographystyle{imsart-nameyear}
-%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-
-% provide arXiv number if available:
-%\arxiv{arXiv:0000.0000}
+\usepackage{enumerate}
+\usepackage{natbib}
+\bibliographystyle{asa}
+\usepackage{url} % not crucial - just used below for the URL
 
@@
 \let\]\myendequation
-% put your definitions there:
-\startlocaldefs
-\endlocaldefs
+
+\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
 
@@
 
-\begin{frontmatter}
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
 
-% "Title of the paper"
-\title{A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{aug}
-  \author{\fnms{Eric}  \snm{Mittman}\corref{}\ead[label=e1]{first@somewhere.com}},
-  \author{\fnms{Colin} \snm{Lewis-Beck}\ead[label=e2]{second@somewhere.com}}
-  \and
-  \author{\fnms{William Q.}  \snm{Meeker}%
-  \ead[label=e3]{third@somewhere.com}%
-  \ead[label=u1,url]{http://www.foo.com}}
+\if0\blind
+{
+  \title{\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
+  \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
+    and \\
+    Colin Lewis-Beck\footnotemark[1]\\
+    and \\
+    William Q. Meeker\footnotemark[1]}
+  \maketitle
+} \fi
 
-  %\thankstext{t2}{Footnote to the first author with the `thankstext' command.}
-
-  %\runauthor{F. Author et al.}
-
-  \affiliation{Iowa State University}
-
-  %\address{Address of the First and Second authors,\\ 
-   %       \printead{e1,e2}}
-
-  %\address{Address of the Third author,\\
-   %       \printead{e3,u1}}
-
-\end{aug}
-
-% indicate corresponding author with \corref{}
-% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
-% \thankstext{t1}{Thanks to somebody} 
-% \address{line 1\\ line 2\\ printead{e1}}
-%\author{\fnms{Eric Mittman and Colin Lewis-Beck and William Q. Meeker}}
-%\affiliation{Iowa State University}
-%\address{\printead{e1}}
-%\and
-%\author{\fnms{???} \snm{???}\ead[label=e2]{???}}
-%\address{\printead{e2}}
-%\affiliation{???}
-
-%\runauthor{???}
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Title}
+\end{center}
+  \medskip
+} \fi
 
@@
 
-\begin{keyword}
-\kwd{GLFP, Bathtub hazard, Censored data, Hierarchical models, Bayesian estimation}
-%\kwd{}
-\end{keyword}
+\noindent%
+{\it Keywords:} GLFP, Bathtub hazard, Censored data, Hierarchical models, Bayesian estimation
+\vfill
 
-\end{frontmatter}
+\newpage
+% \tableofcontents
+% \newpage
+\spacingset{1.45} % DON'T change the spacing!
 
@@
 
-\begin{supplement}
-\sname{Supplement A}
-\stitle{Model code and plots}
-\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
-\sdescription{Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6.}
-\end{supplement}
-\begin{supplement}
-\sname{Supplement B}
-\stitle{Backblaze data set}
-\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
-\sdescription{Lifetime data set for all 44 drive-models used in our analysis.}
-\end{supplement}
+\section{Supplementary Material}
+\label{sec:supple}
+
+\begin{description}
+
+\item[Model code and plots] Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6 (supplementary.pdf)
+
+\item[Backblaze data set] Lifetime data set for all 44 drive-models used in our analysis. (backblaze\_hd\_data.csv)
+
+\end{description}
 

### 2017-12-14 paper/m-lb-m-technometrics.tex
Fix typo in likelihood
Edits to Section 5 and 6.

@@
 \section{Introduction}
+\subsection{Motivation}
 Consumers have an interest in accurate assessment of product reliability.
@@
 
+\subsection{Model considerations}
 Nonparametric methods require relatively few assumptions, but may not be suitable for prediction when extrapolation is required.
@@
 
+\subsection{Overview}
 The structure of the paper is as follows.
@@
 
-\section{Generalized Limited Failure Population Model}
+\section{Model for field data}
 \label{sec:GLFP model}
 
-\subsection{Weibull parameterization}
+\subsection{Weibull distribution parameterization}
 \label{sec:Weibull parameterization}
@@
 
-\subsection{GLFP model}
+\subsection{Generalized Limited Failure Population model}
 \label{subsec:GLFP model}
@@
 \begin{align*}
-h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
+h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right).
 \end{align*}
@@
 
-\subsubsection{Likelihood}
+\subsection{Likelihood}
 We now give the general form for the likelihood function, taking into account left truncation and right censoring.
@@
 
-\section{Motivating Example}
+\section{Motivating example}
 \label{sec:Data}
@@
 \label{sec:Hierarchical GLFP model}
+\subsection{Conditional likelihood}
 
-In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population-specific parameters hierarchically, borrowing strength across the sub-populations.
+In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across groups.
 Let
@@
 The likelihood for the hierarchical GLFP is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each group, $g$.
-Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
+Assuming the lifetimes of all units are independent within and across groups, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
 \begin{equation*}
@@
 All times are again in hours.
+\subsection{Parameter distributions}
 
-In the hierarchical model, the parameters are modeled as random variables, varying across the different sub-populations.  Let
+In a hierarchical model, the parameters are modeled as random variables, varying across different sub-populations.  For example
 \begin{equation*}
@@
 
-As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of a quantile and the shape parameter because the data we encounter are more informative about the lower tail of their respective lifetime distribution (note that if $T$ has a Weibull distribution, $\sigma$ is a shape parameter for the distribution of $T$ and a scale parameter for the distribution of $\log(T)$).
+As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of a quantile and the shape parameter because observed failure data are often more informative about the lower tail of their respective lifetime distribution (note that if $T$ has a Weibull distribution, $\sigma$ is a shape parameter for the distribution of $T$ and a scale parameter for the distribution of $\log(T)$).
 We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (i.e., $\beta_{2g} = 1/\sigma_{2g}>1$).
@@
 
-In principle, we prefer this full model because we tend to believe that every population has a distribution with a distinct set of parameters.
+\subsection{Priors}
+The complete GLFP model requires prior distributions on the hierarchical parameters.
+We suggest weakly informative proper distributions that allow the data to overwhelm the prior and provide information about the distribution of the population parameters.
+Prior knowledge about product lifetimes is useful for choosing the mean of the location parameter ($\eta$) distributions.
+However, the corresponding standard deviations should be diffuse to allow the location to vary depending on the observed data.
+For the scale parameters ($\tau$) we recommend half-Cauchy distributions as suggested by \citet{gelman2014bayesian}.
+
+\subsection{Nested models}
+In principle, we prefer the full model (\ref{eq:hier-model}) because we tend to believe that every population has a distribution with a distinct set of parameters.
 In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their inclusion.
-For example, we may consider a model with a common $\sigma_{1}$, letting $\sigma_{1g}=\sigma_1$ for all $g=1,\ldots,G$.
-A practical strategy is to begin with a simple model and gradually add complexity, reevaluating model adequacy at each iteration.
-In Section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.
+We considered four models, all based on (\ref{eq:hier-model}), differing by which model parameters are held constant across sub-populations.
+These are (from most to least restrictive):
 
-\section{Analysis of the Backblaze hard drive data}
-\label{sec:Data analysis}
-\subsection{Modeling}
-\subsubsection{Hierarchical models}
-We considered four models, all based on (\ref{eq:hier-model}), differing by which model parameters are held constant across drive-models. These are (from most to least restrictive):
+\paragraph{Model 1} $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad t_{p_{2}g} = t_{p2},\quad \sigma_{2g} = \sigma_2$
+\paragraph{Model 2} $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad \sigma_{2g} = \sigma_2$
+\paragraph{Model 3} $\pi_{g} = \pi,\quad  t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
+\paragraph{Model 4} $t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
 
-\begin{enumerate}
-\item $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad t_{p_{2}g} = t_{p2},\quad \sigma_{2g} = \sigma_2$
-\item $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad \sigma_{2g} = \sigma_2$
-\item $\pi_{g} = \pi,\quad  t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
-\item $t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
-\end{enumerate}
+The set of model specifications is chosen based data, interpretation of the model, as well as estimation considerations.
+Group-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.
+Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.
+Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by sub-population.
 
-The set of model specifications was chosen based data, interpretation of the model, as well as estimation considerations.
-Drive-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across drive-models in the right tails of the failure distribution.
-Going from a common model for all drive-models and gradually increasing the complexity of the model, we conclude that Model 4 provides the best description of the data.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by drive-model.
-
-For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.
+For all of the nested models, the parameters for the infant mortality failure mode are held in common across sub-populations.
 We found that there was often insufficient information to model these parameters hierarchically.
@@
 
-\subsubsection{Prior distributions}
+\subsection{Model selection}
+From the point of view of a consumer or purchaser, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
+In this case, we should choose the model that will provide the most accurate prediction for future units of previously observed products.
+One appropriate criterion for model selection, then, is the log pointwise predictive density (lpd),
+
+\begin{align*}
+\mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
+&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
+\end{align*}
+\noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
+Notationally suppressing the conditioning on $t^L$ and $c$,
+\begin{equation}
+\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
+\label{elpd}
+\end{equation}
+where $h$ is a density for the true distribution of $t_{new,g,i}$. Although $h$ is unknown, leave-one-out cross-validation provides a robust estimator:
+\begin{equation}
+\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
+\label{loo-est}
+\end{equation}
+The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
+It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\sum_{i=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$.
+In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.
+For details, refer to \cite{vehtari}.
+%When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
+% To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
+As outlined in Vehtari et.
+al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.
+For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.
+% The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
+
+\section{Model for the Backblaze hard drive data}
+\label{sec:Data analysis}
+We now apply this general methodology to the hard drive failure data from Backblaze.
+There are over 40 different drive-models in the data set, and significant heterogeneity in the number of observed failures and time observed (Figure 1).
+The hierarchical GLFP allows us to model the entire drive-model population, borrowing information across drive-models for those brands where little information exists.
+
+\subsection{Prior distributions}
 \label{sec:Prior distributions}
-To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model.
-We select proper prior distributions to ensure proper posterior distributions.
-
-For Models 1--4, different sets of restrictions required different prior distribution specifications, which were assigned as follows:
+Following the recommendations in Section 5.3, we specify the following prior distributions for each of the four nested GLFP models.
+Different sets of restrictions require different prior distribution specifications, which are assigned as follows:
 
@@
 
-Where applicable prior distributions on hyperparameters are as listed below.
-We follow the recommendation of \citet{gelman2014bayesian} to use half-Cauchy prior distributions on hierarchical scale parameters.
+Where applicable, prior distributions on hyperparameters are as listed below.
+We use half-Cauchy prior distributions on hierarchical scale parameters.
 As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
-% The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification?
-Should maybe be lower.) For $\eta_{\sigma, 2}$, $95\%$ of prior mass is on values greater than $0.037$.
-Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this corresponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$.
-The prior for $\eta_{t_{p_2},2}$ implies that the median 20th percentile for non-defective units is less than greater than 3 days and less than 24 years.
 
@@
 
-\subsection{Computation}
-\label{sec:Computation}
+\subsection{Model fitting}
+\label{sec:Model fitting}
 Each model was fit using the {\tt
@@
 Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.
-
-% HMC generates proposals by simulating paths along the posterior density.
-This is done to reduce autocorrelation between samples, but doing so also provides a useful diagnostic.
-\textit{Divergent transitions} will tend to occur when the posterior has regions of high curvature, which can present problems for MCMC by making it difficult to fully explore the posterior distribution.
-This feedback can be useful
 
@@
 
-From the point of view of a consumer of hard disk drives, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
-In this case, we should choose the model that will provide the most accurate prediction for future units of previously observed drive-models.
-One appropriate criterion for model selection, then, is the log pointwise predictive density (lpd),
-
-\begin{align*}
-\mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
-&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
-\end{align*}
-\noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$.
-Notationally suppressing the conditioning on $t^L$ and $c$,
-\begin{equation}
-\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
-\label{elpd}
-\end{equation}
-where $h$ is a density for the true distribution of $t_{new,g,i}$. Although $h$ is unknown, leave-one-out cross-validation provides a robust estimator:
-\begin{equation}
-\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
-\label{loo-est}
-\end{equation}
-The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
-It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\sum_{i=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$.
-In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.
-For details, refer to \cite{vehtari}.
-%When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
-% To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.
-As outlined in Vehtari et.
-al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.
-For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.
-% The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).  
 Results showing the estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:2}.

### 2018-01-09 paper/m-lb-m-technometrics.tex
Add/revise section headers

@@
 For inference, likelihood-based methods are able to deal with truncated and/or censored data.
-Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across groups, and incorporate prior information.
+Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across sub-populations, and incorporate prior information.
 
@@
 
-In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across groups.
+In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.
 Let
@@
 where $g=1,\ldots,G$ indexes the sub-populations.
-The likelihood for the hierarchical GLFP is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each group, $g$.
-Assuming the lifetimes of all units are independent within and across groups, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
+The likelihood for the hierarchical GLFP is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.
+Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
 \begin{equation*}
@@
 \end{equation*}
-where $t_{ig}$ is the observed failure or survival time of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.
+where $t_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored.
 All times are again in hours.
@@
 The set of model specifications is chosen based data, interpretation of the model, as well as estimation considerations.
-Group-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.
+sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.
 Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.
@@
 Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2$ results in very similar progressions in the cdf, which may be too restrictive.
-Model 3, which allows $\sigma_2$ to vary by group, suggests there is evidence to support variation in this parameter among drive-models.
+Model 3, which allows $\sigma_2$ to vary by sub-population, suggests there is evidence to support variation in this parameter among drive-models.
 While Model 3 and 4 do display differences, they are relatively subtle.
@@
 
-\subsection{Results}
-\subsubsection{Model assessment}
+\section{Model Assessment}
+\subsubsection{A simulation-based approach}
 \label{model-assessment}
@@
 
-\subsubsection{Drive-model comparisons}
-\label{subsubsec:Drive-model comparisons}
+\section{Comparing Sub-populations}
+\subsection{Comparisons based on model parameters}
+\subsection{Example: Drive-model comparison based on quantiles and proportion defective}
+\subsection{Comparisons based on functionals}
+\subsection{Example: Drive-model comparison based on predictive distribution of new unit}
+
 We consider the problem of ranking drive-models with respect to economical value. From a
@@
 Moreover, their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.
-The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product groups.
+The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product sub-populations.
 Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the data.

### 2018-01-10 paper/m-lb-m-technometrics.tex
Begin reworking model assessment section(s)

@@
 \section{Model Assessment}
-\subsubsection{A simulation-based approach}
+\subsection{A simulation-based approach}
 \label{model-assessment}
-Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate description of the observed data.
-One way to do this is to draw replicate data sets from the posterior predictive distribution.
-This is a general approach recommended by \cite{gelman1996postpred} for models where no classical goodness-of-fit test is available.
-In the hierarchical setting, there can be different formulations of the predictive distribution.
-In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
-We choose to focus on the first of these and apply it to individual drive-models.
+Having selected a model, it is good practice to check that the model provides an adequate representation of the data-generating mechanism.
+One way to do this is to draw replicate data sets from the posterior predictive distribution
+$$p(t_{rep,g}|t) = \int f(t_{\mbox{rep},g}|\theta_g)\; p(\theta_g|t)\;d\theta_g.$$
+This is a general approach recommended by \cite{gelman1996postpred}, which works when no classical goodness-of-fit test is available.
+When the model is adequate for the data, the predictive distribution of a finite set of data characteristics --- chosen to reflect features relevant to the goals of the analysis --- should not contradict the corresponding characteristics observed in the actualized data.
+
 There are many possible approaches to the posterior predictive check.
-For this problem, we chose to consider the posterior predictive distribution of the K-M estimate for each drive-model $g$.
-
-% Graphical comparisons can be highly informative in revealing limitations of the model.
-as in our cas a graphical posterior predictive check, wherein replicate data sets are generated from the posterior predictive distribution and compared with the observed data.
-Because of the rich information gleaned from plotting the data, this method can be very effective at identifying limitations of the model.
+We propose the comparison of K-M estimates for each sub-population derived from replicated data drawn from the posterior predictive distribution to the K-M estimate based on the actualized data.
+This graphical approach does not provide a test statistic, but provides rich feedback with respect to the fitted model's agreement with the data.
 
 Because the variability in the K-M estimator is primarily a function of the number of at-risk units, we want to replicate the pattern of censoring that was observed.
-That is, for unit $i$ of drive-model $g$,
+That is, for unit $i$ of sub-population $g$,
 $$p(t_{rep,g}) = \int \prod_{i=1}^{n_g}f(t_{new}|t_{g,i}^L,c_{g,i},\theta)\,p(\theta|t)d\theta.$$
@@
 
-An obstacle remains, in that the censoring mechanism itself is not fully observed.
+An obstacle remains, owing to the fact that the censoring time is unknown for observed failures.
 Our admittedly simple approach is to assume that failed units would have been the censored had they survived until the latest censoring or failure time for that drive model.
-We generated 19 replicates for each drive model.
-This gives a 1/20 chance under the hypothesis of no difference that the K-M estimate based on the observed data would be ``most extreme" relative to those from the replicated data, but not too many so as to produce an overly cluttered plot.
-Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots 19$, for each drive-model $g=1,\ldots,44$, is as follows:
+Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots R$, for each sub-population $g=1,\ldots,G$, is as follows:
 \begin{enumerate}
@@
 \end{enumerate}
+
+Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate description of the observed data.
+In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
+We choose to focus on the first of these and apply it to individual drive-models.
+There are many possible approaches to the posterior predictive check.
+For this problem, we chose to consider the posterior predictive distribution of the K-M estimate for each drive-model $g$.
+
+% Graphical comparisons can be highly informative in revealing limitations of the model.
+as in our cas a graphical posterior predictive check, wherein replicate data sets are generated from the posterior predictive distribution and compared with the observed data.
+Because of the rich information gleaned from plotting the data, this method can be very effective at identifying limitations of the model.
 

### 2018-01-10 paper/m-lb-m-technometrics.tex
Revised to acheive desired sectioning

@@
 \begin{equation}
-y_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
+T_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
 \end{equation}
@@
 \end{equation*}
-where $t_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored.
+where $T_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored.
 All times are again in hours.
@@
 \begin{equation*}
-\sigma_{2g} \ind \op{Log-normal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
+\sigma_{2g} \ind \op{Log-normal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{Tr}\left(0, 1\right)
 \end{equation*}
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of a quantile and the shape parameter because observed failure data are often more informative about the lower tail of their respective lifetime distribution (note that if $T$ has a Weibull distribution, $\sigma$ is a shape parameter for the distribution of $T$ and a scale parameter for the distribution of $\log(T)$).
-We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (i.e., $\beta_{2g} = 1/\sigma_{2g}>1$).
+We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $Tr(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (i.e., $\beta_{2g} = 1/\sigma_{2g}>1$).
 %We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
@@
 One way to do this is to draw replicate data sets from the posterior predictive distribution
-$$p(t_{rep,g}|t) = \int f(t_{\mbox{rep},g}|\theta_g)\; p(\theta_g|t)\;d\theta_g.$$
+$$p(t_{rep,g}|t) = \int f(t_{rep,g}|\theta_g)\; p(\theta_g|t)\;d\theta_g.$$
 This is a general approach recommended by \cite{gelman1996postpred}, which works when no classical goodness-of-fit test is available.
@@
 Our admittedly simple approach is to assume that failed units would have been the censored had they survived until the latest censoring or failure time for that drive model.
-Our method for generating replicate data sets, $y_{rep,g}^{(s)}$, $s = 1 \dots R$, for each sub-population $g=1,\ldots,G$, is as follows:
+Our method for generating replicate data sets, $t_{rep,g}^{(s)}$, $s = 1,\ldots, R$, for each sub-population $g=1,\ldots,G$, is as follows:
+
+For $s=1,\ldots, R$, do:
+
 \begin{enumerate}
-\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)}, \sigma_{2g}^{(s)})$ where each set of parameters, \\$\left( \pi_{g}^{(s)},t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)}, \sigma_{2g}^{(s)} \right)$, is a random draw from the full posterior distribution.
-\item Repeat step 1 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
+\item Sample $r$ uniformly from $\{1,\ldots,S\}$.\\ For $i=1,\ldots,n_g$, do:
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(r)},t_{p1}^{(r)},\sigma_1^{(r)},t_{p_{2}g}^{(r)}, \sigma_{2g}^{(r)})$.
+\item Repeat step 2 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
 \item Set $d_{g,i}^{(s)}= \begin{cases} \max \{t_{g,i}: i \ge 1\}, & \mbox{ if }c_{g,i}=0\\
@@
 \end{enumerate}
+Finally, plot K-M estimates based on the $R$ replicated data sets along with the one based on the actualized data on the same axes.
+Discrepancies between the replicated and actual data can be used to diagnose inadequacies of the model.
 
-Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate description of the observed data.
-In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model.
+\subsection{Application to the Backblaze data}
+For the hard-drive data, we used the posterior samples obtained in fitting Model 4 to generate $R=19$ replicate data sets as described above.
+The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
+
+% Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate description of the observed data.
+In the present context, one could  or a new drive-model.
 We choose to focus on the first of these and apply it to individual drive-models.
@@
 
-We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the K-M estimator.
-The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
+% We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the K-M estimator. 
 Only a representative sample of plots is shown here for brevity (see the Supplementary Material for similar plots for all of the drive-models).
@@
 \subsection{Comparisons based on model parameters}
+As discussed in Section \ref{sec:Weibull parameterization}, we use a parameterization of the Weibull distribution in terms of a chosen quantile, making prior elicitation straightforward.
+Posterior sampling immediately gives posterior distributions for these quantiles, which in reliability applications is typically more informative than a measure of central tendency.
+If the selected model includes varying $t_{p2g}$ across sub-poputations, side-by-side plots of credible intervals provides a way to compare lifetime of non-defective units.
+Depending on a user's application, a similar plot for the $\pi_g$'s as a comparison tool may be a relevant parameter of interest.
+
 \subsection{Example: Drive-model comparison based on quantiles and proportion defective}
-\subsection{Comparisons based on functionals}
-\subsection{Example: Drive-model comparison based on predictive distribution of new unit}
-
-We consider the problem of ranking drive-models with respect to economical value. From a
-business perspective, it is clear
-that we should prefer the drive-models which will provide more years
-of service. For ease of exposition, we will assume that the purchase
-price of a hard-drive is the same across drive-models.
-
-There are two sources of variation at play: the
-posterior uncertainty in the parameters, which largely depends on the
-amount of data, and the uncertainty in future
-observations conditional on the parameters. The {\em posterior
-  predictive} distribution incorporates both sources of variability.
-\begin{equation*}
-  p(t_{g,new}|t) = \int p(t_{g,new}|\theta_g) p(\theta_g|t_g) \, d\theta_g
-\end{equation*}
-
-We can sample from this distribution by drawing $t_{g,new}^{(s)}$ from
-$\operatorname{GLFP}(\pi_g^{(s)},
-t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)})$, for
-$s=1,...,S$, using the saved posterior draws for the model
-parameters. The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
-$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}.$$
-
-While MTTF is often an important metric, due to the
-anticipation of advancements in technology, we can expect that the
-relative value of computer hardware will depreciate rather quickly.
-In simple terms, given two hard drives, one would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all.
-To account
-for depreciation, rather than using MTTF as the metric for drive-model comparison, we use the value at replacement. The US Internal Revenue Service considers
-computer hardware to be ``5-year'' equipment \citep{f4562}. Using``the declining
-balance method'' the rate of depreciation is 40\% per year.
-
-Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
-relative to a new unit. We can rank the drive-models by 
-$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)}).$$
-
-Posterior credible intervals for $L(t_{g,new})$ are shown in the left panel of Figure \ref{fig5}.
-A zoomed version of the top eight drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel.
-The panel above shows credible intervals for MTTF.
-We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit is expected to last the longest, but the possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric.
-The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
-However, given the discrepancy of the model fit for Drive-model 4 (noted in the previous subsection), in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.
-
-We can also use the posterior draws to compute quantiles of interest, which in reliability applications is typically more informative than a measure of central tendency.
 The top panel of Figure 9 shows estimates of $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
@@
 While for many of the drive-models the ordinal ranking is the same as on the top, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
-Depending on a user's application or the hard drive warranty length, using the $\pi$'s as a comparison tool may be a relevant parameter of interest.
 
@@
 
+\subsection{Comparisons based on posterior predictive distribution}
+By utilizing samples from the full posterior distribution, in addition to estimation of the model parameters, we can also produce forecasts for new units or groups of units.
+In many situations, this may more directly address issues of importance.
+
+The {\em posterior predictive} distribution incorporates two sources of variability: the posterior uncertainty in the parameters, which largely depends on the
+amount of data collected, and the uncertainty in future
+observations conditional on those parameters.
+
+For example,
+\begin{equation*}
+  p(t_{g,new}|t) = \int p(t_{g,new}|\theta_g) p(\theta_g|t_g) \, d\theta_g
+\end{equation*}
+gives the posterior predictive density for a new unit from sub-population $g$.
+
+We can sample from this distribution by drawing $t_{g,new}^{(s)}$ from
+$\operatorname{GLFP}(\pi_g^{(s)},
+t_{p1}^{(s)},\sigma_1^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)})$, for
+$s=1,...,S$, using the saved posterior draws for the model
+parameters.
+
 \begin{figure}[H]
@@
   \includegraphics[width=.8\textwidth]{dm-eval}
-  \caption{Left: 90\% CI of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each drive-model.
+  \caption{Left: 90\% CI of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
   Right: A comparison (90\% CI) of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
@@
 \end{figure}
+
+\subsection{Example: Drive-model comparison based on predictive distribution of new unit}
+
+We consider the problem of ranking drive-models with respect to economical value. From a
+business perspective, it is clear
+that we should prefer the drive-models which will provide more years
+of service. For ease of exposition, we will assume that the purchase
+price of a hard-drive is the same across drive-models.
+
+The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
+$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}.$$
+
+While MTTF is often an important metric, due to the
+anticipation of advancements in technology, we can expect that the
+relative value of computer hardware will depreciate rather quickly.
+In simple terms, given two hard drives, one would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all.
+To account
+for depreciation, rather than using MTTF as the metric for drive-model comparison, we use the value at replacement. The US Internal Revenue Service considers
+computer hardware to be ``5-year'' equipment \citep{f4562}. Using``the declining
+balance method'' the rate of depreciation is 40\% per year.
+
+Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
+relative to a new unit. We can rank the drive-models by 
+$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)}).$$
+
+Posterior credible intervals for $L(t_{g,new})$ are shown in the left panel of Figure \ref{fig5}.
+A zoomed version of the top eight drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel.
+The panel above shows credible intervals for MTTF.
+We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit is expected to last the longest, but the possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric.
+The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
+However, given the discrepancy of the model fit for Drive-model 4 (noted in the previous subsection), in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.
 

### 2018-01-18 paper/m-lb-m-technometrics.tex
Make Meeker's suggested changes

@@
 {
-  \title{\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
+  \title{\bf A hierarchical failure-time model for heterogenous field data}
   \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
@@
 \begin{abstract}
-When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure.
+When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failures.
 The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions.
@@
 In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model.
-This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failures due to ``infant mortality'' and failures due to wearout.
-We fit the GLFP model using a hierarchical modeling approach.
-Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
-Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals is easy using draws from the joint posterior distribution of the model parameters.
+This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failures (sometimes referred to in the literature as ``infant mortality'') and failures due to wearout.
+We fit the GLFP model to a heterogenous population of devices using a hierarchical modeling approach.
+Borrowing strength across sub-populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.
+Moreover, using this Bayesian method, comparison of different product brands across the heterogenous population is straightforward because estimation of arbitrary functionals is easy using draws from the joint posterior distribution of the model parameters.
 Potential applications include assessment and comparison of reliability to inform purchasing decisions.
@@
 The reliability of many engineered products follows a similar pattern.
-Relatively high rates of failure occurring early (``infant mortality'') is due to manufacturing defects.
+Relatively high rates of early failure are due to manufacturing defects.
 After this ``burn-in'' period, failure rates stabilize after the majority of defective units have failed.
@@
 \begin{itemize}
-\item Making purchase decisions from among different suppliers of the components or subsystems
-\item Making predictions for the number of needed replacement components or subsystems for spare part provisioning
+\item Making purchase decisions from among different suppliers of the components or subsystems.
+\item Making predictions for the number of needed replacement components or subsystems for spare part provisioning.
 \end{itemize}
@@
 For inference, likelihood-based methods are able to deal with truncated and/or censored data.
-Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across sub-populations, and incorporate prior information.
+Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare sub-populations of interest, borrow strength across sub-populations, and incorporate prior information.
 
@@
 
-The GLFP model has a meaningful parameterization and accommodates lifetime data with both infant mortality and wearout.
+The GLFP model has a meaningful parameterization and accommodates lifetime data with both early and wearout failures.
 Unfortunately, it can require a lot of data to fit due to the model's complexity.
 When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates.
-When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
-As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters.
-This enables estimation of the lifetime distribution for all populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of population specific information.
-We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via Markov Chain Monte Carlo (MCMC).
+When multiple sub-populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data.
+As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across sub-populations which imposes ``soft'' constraints on model parameters.
+This enables estimation of the lifetime distribution for all sub-populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of sub-population specific information.
+We demonstrate the computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via Markov Chain Monte Carlo (MCMC).
 
@@
 In Section~\ref{sec:Data}, we describe the complete Backblaze Hard Drive data set, which is used throughout as a motivating example.
-In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations.
+In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple sub-populations.
 Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard drive data.
@@
 Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data.
-Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime.
+Finally, we present an evaluative comparison of sub-populations taking into account practical considerations related to product lifetime.
 Section~\ref{sec:Discussion} describes potential applications of our model and limitations for inference implied by the model assumptions.
@@
 The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.
-A common assumption in series systems is the time to failure for each risk is statistically independent.
-Thus, the overall reliability of a unit is the product of all the risks.
-Parameter estimation is straightforward if the cause of failure is known for each observation.
+A common assumption in series system modeling is the time to failure for the different risks are statistically independent.
+Thus, the overall reliability of a unit is the product of the reliability of each of the risks.
+Then parameter estimation is straightforward if the cause of failure is known for each observation.
 However, in many situations the cause of failure is unknown to the analyst.
-This is referred to in the literature as masking.\\
+This is referred to in the literature as masking.
 
@@
 When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.
-\cite{chan} labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.
+\cite{chan} labeled the cause of circuit board failures as early (due to defects), unknown, or wearout based on the time of observed failures.
 This helped identify parameters when using maximum likelihood (ML) estimation.
-Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure.
+Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the causes of failure.
 \cite{berger} introduced the Poly-Weibull distribution where the failure time is the minimum of a several Weibull distributions, corresponding to different failure modes.
-\cite{ranjan} considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions.
+\cite{ranjan} considered a competing risk model for early and wearout failures as a mixture of Weibull and exponential distributions.
 Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.
@@
 distribution than it will be to elicit prior information about
-$\alpha$. In addition, there is generally available
+$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
 information about the shape parameter, $\sigma$, for a given failure
@@
 work well.
+\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
 \end{itemize}
@@
 The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$.
-$\delta_i$ is an indicator for whether or not unit $i$ is defective.
+$\delta_i$ is an indicator for whether unit $i$ is defective or not (i.e., susceptible to an early failure).
 Expressed conditional on~$\delta_i$,
@@
 
-The parameter $\pi$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes.
+The parameter $\pi$ represents the proportion of units susceptible to early failure, and hence susceptible to both failure modes.
 Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.
 
-Taking the derivative of the (marginal) cdf, the density for the GLFP is
+Taking the derivative of the (marginal) cdf, the density for the GLFP model is
 \begin{align*}
@@
 In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.
-Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, Accessed April 1, 2016)\nocite{backblaze}.
+Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, accessed January 18, 2018)\nocite{backblaze}.
 In addition, Backblaze publishes summary statistics of the different drive-models currently operating.
-No other analysis or modeling of the failure data is provided.
+No other analysis or modeling of the failure data is provided, other than simple rates and proportions.
 Backblaze does, however, encourage others to further analyze their data.
@@
 Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.
-For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.
+For model identification, a minimum of three failures was the criterion for inclusion of drive-models into our analysis.
 
@@
   \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
-  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.  Each number labels a unique drive-model.   Axes are on the log scale.}
+  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.
+Each number labels a unique drive-model, numbered between 1 and 44.
+Axes are on the log scale.}
   \label{drive-scatter}
@@
 The curve has a slow ramp up in failures till about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).
-In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.
+Besides our empirical observation, it is well known that computer hard drives have a mixture of early and late failures \citep{chan}.
 Therefore, there is both an empirical and a theoretical justification to apply the GLFP model.
@@
 
-To estimate the parameters of the GLFP model we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution.
-We reparameterize the component Weibull distributions using the $0.50$ quantile for the infant mortality failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$).
+To estimate the parameters of the GLFP model we use a Bayesian approach, selecting proper, but generally diffuse, prior distributions to improve identification of the model parameters.
+Having a proper prior distribution also ensures a proper posterior distribution.
+We reparameterize the component Weibull distributions using the $0.50$ quantile for the early failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$).
 \\
@@
 %$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
-$$t_{0.5,1} \sim \op{Log-normal} \langle 1.7,7.6\times 10^6 \rangle,\quad
-t_{0.2,2} \sim \op{Log-normal}\langle 8.6, 5.6\times 10^7 \rangle$$
-$$\pi \sim \op{Logit-normal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Log-normal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
+$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
+t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
+$$\pi \sim \op{Logit-normal}\langle 1.0\times 10^{-3},\, 7.1\times 10^{-1} \rangle,\quad
+\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,\,1.3 \times 10^2 \rangle. $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 Table \ref{table:1} gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.
-The parameter $\pi$ is an estimate of the proportion of drives susceptible to infant mortality.
+The parameter $\pi$ is an estimate of the proportion of drives susceptible to early failure.
 As expected, this proportion is small with a median value of 0.05.
 The shape parameter estimates for the two Weibull distributions also match intuition.
-The infant mortality failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.
+The early failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.
 Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.
-The two quantiles are also well identified with the infant mortality failure mode having an much earlier time to reach the 0.50 quantile compared to the time to reach the 0.20 quantile for the wearout mode.
+The two quantiles are also well identified with the early failure mode having an much earlier time to reach the 0.50 quantile compared to the time to reach the 0.20 quantile for the wearout mode.
 
@@
 In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted K-M estimate with axes on the Weibull probability scale.
-The plot also contains 90\% pointwise uncertainty bands associated with each estimate.
+The plot also contains 90\% pointwise credible bands associated with each estimate.
 While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to adequately describe the rapid increase in the empirical cdf between 8000 and 20,000 hours.
@@
 where $g=1,\ldots,G$ indexes the sub-populations.
-The likelihood for the hierarchical GLFP is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.
+The likelihood for the hierarchical GLFP model is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.
 Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
@@
 \begin{equation*}
-\sigma_{1g} \ind \op{Log-normal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
+\sigma_{1g} \ind \op{Lognormal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
 \end{equation*}
 \begin{equation*}
-\sigma_{2g} \ind \op{Log-normal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{Tr}\left(0, 1\right)
+\sigma_{2g} \ind \op{Lognormal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{Tr}\left(0, 1\right)
 \end{equation*}
@@
 \label{eq:hier-model}
-t_{p_{1}g} = \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
+t_{p_{1}g} = \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
 \end{equation}
 \begin{equation*}
-t_{p_{2}g} = \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
+t_{p_{2}g} = \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
 \end{equation*}
@@
 
-The set of model specifications is chosen based data, interpretation of the model, as well as estimation considerations.
-sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.
+This set of model specifications is chosen based on the data, interpretation of the model, as well as estimation considerations.
+Sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.
 Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.
-Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by sub-population.
+Model 4 allows for the probability of early failure as well as the shape and scale parameters for the wearout failure mode to vary across sub-populations.
 
-For all of the nested models, the parameters for the infant mortality failure mode are held in common across sub-populations.
+For all of the nested models, the parameters for the early failure mode are common across sub-populations.
 We found that there was often insufficient information to model these parameters hierarchically.
-Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).
+Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).
 Also the failure-time distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.
@@
 We now apply this general methodology to the hard drive failure data from Backblaze.
-There are over 40 different drive-models in the data set, and significant heterogeneity in the number of observed failures and time observed (Figure 1).
-The hierarchical GLFP allows us to model the entire drive-model population, borrowing information across drive-models for those brands where little information exists.
+There are 44 different drive-models in the data set, and significant heterogeneity in the number of observed failures and time observed (Figure 1).
+The hierarchical GLFP allows us to model the entire drive-model population, borrowing information across drive-models, allowing estimation for those drive-models where little information exists.
 
@@
 
-\paragraph{Model 2} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
+\paragraph{Model 2} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
 $$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
-$$\sigma_1 \sim \op{Log-normal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
-$$t_{p1} \sim \op{Log-normal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
+$$\sigma_1 \sim \op{Lognormal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
+$$t_{p1} \sim \op{Lognormal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
 
@@
 
-\paragraph{Model 4} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.
+\paragraph{Model 4} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Prior distributions for constrained parameters are the same as for Model 2.
 
@@
 Model 3, which allows $\sigma_2$ to vary by sub-population, suggests there is evidence to support variation in this parameter among drive-models.
-While Model 3 and 4 do display differences, they are relatively subtle.
+While there are differences between Model 3 and 4, they are relatively subtle.
 
@@
 \end{equation}
-where $\theta^{(s)}$ is the $s^{th}$ posterior sample from $p(\theta|y)$.
+where $\theta^{(s)}$ is posterior sample $s$ from $p(\theta|y)$.
 %Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.
@@
 \includegraphics[width=\textwidth]{single-drive-4-Models-ex}}
-\caption{Posterior median of proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
+\caption{Posterior median of the proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
 The solid step function is the adjusted K-M estimate.}
@@
 \caption{\footnotesize Top: Event plot for all units of Drive-model 43.
-Data for all units are left-truncated, being observed conditional on survival beyond 12,000 hours.\hspace{\textwidth}Bottom: The dashed step function correponds to the adjusted K-M estimate.
-Due to the heavy left-truncation, the posterior median for Drive-model 43 (dashed line) coincides with the posterior median of the ``global model" (dotted line) until the first left-truncation time.}
+Data for all units are left-truncated, being observed conditional on survival beyond 12,000 hours.\hspace{\textwidth}Bottom: The dashed step function correponds to the adjusted K-M estimate (solid line).
+Due to the heavy left-truncation, the posterior median for Drive-model 43 (dotted line) coincides with the posterior median of the ``global model" (dashed line) until the first left-truncation time.}
 \label{fig:ex-mod-43}
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we use a parameterization of the Weibull distribution in terms of a chosen quantile, making prior elicitation straightforward.
-Posterior sampling immediately gives posterior distributions for these quantiles, which in reliability applications is typically more informative than a measure of central tendency.
+Posterior sampling immediately gives marginal posterior distributions for these quantiles, which in reliability applications is typically more informative than a measure of central tendency.
 If the selected model includes varying $t_{p2g}$ across sub-poputations, side-by-side plots of credible intervals provides a way to compare lifetime of non-defective units.
@@
 \subsection{Example: Drive-model comparison based on quantiles and proportion defective}
-The top panel of Figure 9 shows estimates of $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail), for all drive-models.
+The top panel of Figure 9 shows estimates of $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail, i.e.
+the quantile corresponding to the $H(t)$ distribution in Section \ref{subsec:GLFP model}), for all drive-models.
 In this application 50\% CIs are more meaningful as there is lots of uncertainty in some of the parameters: especially for those with few drives in operation.
@@
 
-\caption{(a) 50\% CI (in years) for $t_{0.10}$ (b) 50\% CI for $\pi$ plotted on the logit scale.  Both plots are sorted based on the median value of $t_{0.10}$.}
+\caption{(a) 50\% CI (in years) for $t_{0.10}$ (b) 50\% CI for $\pi$ plotted on the logit scale.  Both plots are sorted based on the median value of the marginal posterior distribution of $t_{0.10}$.}
 \end{figure}
@@
 relative value of computer hardware will depreciate rather quickly.
-In simple terms, given two hard drives, one would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all.
+In simple terms, given two hard drives, one would prefer to have both of them work for five years than for one to work for nine years and the other to work for just one year.
 To account
@@
 
-Second, we assume exchangability of units within a population.
+Second, we assume exchangability of units within a sub-population.
 This assumption is due in part to ignorance about which units are defective but also of potentially important covariates.
@@
 The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product sub-populations.
-Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the data.
+Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the actual application.
 
@@
 \section*{Acknowledgment}    
-The authors express their gratitude to Nick Clark for his comments and suggestions.
+The authors express their gratitude to Nick Clark for his helpful comments and suggestions.
 
@@
 \label{sec:trunc-adj}
-We first start with a nonparametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator.
+We first start with a nonparametric estimate of the empirical cdf for each sub-population using the Kaplan-Meier estimator.
 With left truncation, however, the standard Kaplan-Meier estimator for drive-model $g$, denoted by
 $\widehat{F_g(t)}_{KM}$, is conditional on survival up to
-$t_{g,\text{min}}^L$, the shortest reported running time of all units
-of drive-model $g$ for which records are available. To produce
-unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each drive-model we select
+$t_{g,\text{min}}^L$, the shortest reported running time of all units in
+of sub-population $g$ for which records are available. To produce
+unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each sub-population we select
 $t_{g,\text{min}}^L$, the smallest left truncated time in the sample.
-By sampling from the full posterior distribution, since
-$\Pr(T>t_{g,\text{min}}^L|\theta_g)$ (the probability a hard drive has
+By sampling from the full posterior distribution, because
+$\Pr(T>t_{g,\text{min}}^L|\theta_g)$ (the probability that a hard drive has
 survived up to $t_{g,\text{min}}^L$) is a function of the model
@@
 
-While this adjustment is negligible for the majority of drive-models, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
+While this adjustment is may be negligible for sub-populations with little truncation, in our Backblaze example, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
 2.3 years.
@@
 In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a ``global average" which represents an average GLFP cdf for the entire population, $\op{H}\left(\cdot|\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
-Since $\mu_1$ and $\sigma_1$ are shared across drive-models, these can already be interpreted as ``global".
-For the parameters that vary across drive models, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.
+Since $\mu_1$ and $\sigma_1$ are common across all sub-populations in our model, these can already be interpreted as ``global." For the parameters that vary across sub-populations, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.
 In particular, we set
 $$\bar{\pi}=\op{logit}^{-1}(\eta_{\pi}),\;\bar{\mu}_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2) \mbox{ and } \bar{\sigma}_2= m_{\sigma_2}.$$
-Let $J(\cdot|a,b),\;J^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a log-normal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
+Let $J(\cdot|a,b),\;J^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a ognormal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
 $$m_{\sigma_2}=J^{-1}[0.5 \cdot J(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}],$$
-which is the median of a log-normal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.
+which is the median of a ognormal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.
 
@@
 
-\item[Backblaze data set] Lifetime data set for all 44 drive-models used in our analysis. (backblaze\_hd\_data.csv)
+\item[Backblaze data set] Lifetime data set for all 44 Backblaze drive-models used in our analysis. (backblaze\_hd\_data.csv)
 

### 2018-01-19 paper/m-lb-m-technometrics.tex
Clean up small typos.

@@
 Nonparametric methods require relatively few assumptions, but may not be suitable for prediction when extrapolation is required.
-Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be useful.
+Moreover, nonparametric estimates made using small samples or data which are heavily censored may be too noisy to be useful.
 Appropriate parametric models can make better use of available data.
@@
 The structure of the paper is as follows.
-Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work.
+Section~\ref{sec:Background} introduces the proposed method, potential applications, and examines previous related work.
 Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of hard drive failure data from the company, Backblaze.
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of a quantile and the shape parameter because observed failure data are often more informative about the lower tail of their respective lifetime distribution (note that if $T$ has a Weibull distribution, $\sigma$ is a shape parameter for the distribution of $T$ and a scale parameter for the distribution of $\log(T)$).
-We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $Tr(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (i.e., $\beta_{2g} = 1/\sigma_{2g}>1$).
+We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $\text{Tr}(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (i.e., $\beta_{2g} = 1/\sigma_{2g}>1$).
 %We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.
@@
 Results showing the estimated elpd and associated standard errors for all 4 models are shown in Table \ref{table:2}.
-We calculated the difference in expected predictive accuracy for Model 4 versus 3, Model 3 versus 2, and Model 2 versus 1, as well as the standard error of the difference (Table 1).
+We calculated the difference in expected predictive accuracy for Model 4 versus 3, Model 3 versus 2, and Model 2 versus 1, as well as the standard error of the difference (Table 2).
 For the models we considered, each successive increase in model complexity increases elpd.

### 2018-01-21 paper/m-lb-m-technometrics.tex
Change to title, edits related to marginal conditional quantile distinction

@@
 {
-  \title{\bf A hierarchical failure-time model for heterogenous field data}
-  \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
-    and \\
-    Colin Lewis-Beck\footnotemark[1]\\
-    and \\
-    William Q. Meeker\footnotemark[1]}
+  \title{\bf A Hierarchical Model for Heterogenous Reliability Field Data}
+  \author{Eric Mittman, Colin Lewis-Beck, and William Q. Meeker\\
+  Department of Statistics\\
+  Iowa State University\\
+  Ames, Iowa}
   \maketitle
@@
 
+Note that the $p$ quantile may be found by setting $p=H(t_p)$ and solving numerically for $t_p$.
+
 \subsection{Censoring and truncation}
@@
 
-\paragraph{Model 1} $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad t_{p_{2}g} = t_{p2},\quad \sigma_{2g} = \sigma_2$
-\paragraph{Model 2} $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p1},\quad \sigma_{1g}=\sigma_1,\quad \sigma_{2g} = \sigma_2$
-\paragraph{Model 3} $\pi_{g} = \pi,\quad  t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
-\paragraph{Model 4} $t_{p_{1}g}= t_{p1},\quad \sigma_{1g}=\sigma_1$
+\paragraph{Model 1} $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p_1},\quad \sigma_{1g}=\sigma_1,\quad t_{p_{2}g} = t_{p_2},\quad \sigma_{2g} = \sigma_2$
+\paragraph{Model 2} $\pi_{g} = \pi,\quad t_{p_{1}g} = t_{p_1},\quad \sigma_{1g}=\sigma_1,\quad \sigma_{2g} = \sigma_2$
+\paragraph{Model 3} $\pi_{g} = \pi,\quad  t_{p_{1}g}= t_{p_1},\quad \sigma_{1g}=\sigma_1$
+\paragraph{Model 4} $t_{p_{1}g}= t_{p_1},\quad \sigma_{1g}=\sigma_1$
 
@@
 
-\paragraph{Model 1} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.
+\paragraph{Model 1} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Section \ref{subsec:ex1}.
 
-\paragraph{Model 2} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
+\paragraph{Model 2} $t_{p_{2}g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
 $$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
 $$\sigma_1 \sim \op{Lognormal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
-$$t_{p1} \sim \op{Lognormal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
+$$t_{p_1} \sim \op{Lognormal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
 
-\paragraph{Model 3} $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Prior distributions for the constrained parameters are the same as for Model 2.
+\paragraph{Model 3} $t_{p_{2}g}$ and $\sigma_{2g}$ vary by drive-model. Prior distributions for the constrained parameters are the same as for Model 2.
 
-\paragraph{Model 4} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Prior distributions for constrained parameters are the same as for Model 2.
+\paragraph{Model 4} $\pi_g$, $t_{p_{2}g}$ and $\sigma_{2g}$ vary by drive-model. Prior distributions for constrained parameters are the same as for Model 2.
 
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we use a parameterization of the Weibull distribution in terms of a chosen quantile, making prior elicitation straightforward.
-Posterior sampling immediately gives marginal posterior distributions for these quantiles, which in reliability applications is typically more informative than a measure of central tendency.
-If the selected model includes varying $t_{p2g}$ across sub-poputations, side-by-side plots of credible intervals provides a way to compare lifetime of non-defective units.
-Depending on a user's application, a similar plot for the $\pi_g$'s as a comparison tool may be a relevant parameter of interest.
+Posterior sampling immediately gives marginal posterior distributions for quantiles, which in reliability applications is typically more informative than a measure of central tendency.
+To compare the unconditional $p$ quantile for each sub-population, plots of posterior credible intervals may be used.
+
+Depending on the set of model restrictions used, simlar plots may be used to compare the $p$ quantiles for non-defective units ($t_{p_{2}g}$) or the proportion defective ($\pi_g$).
 
@@
 the quantile corresponding to the $H(t)$ distribution in Section \ref{subsec:GLFP model}), for all drive-models.
-In this application 50\% CIs are more meaningful as there is lots of uncertainty in some of the parameters: especially for those with few drives in operation.
-Interestingly, the top seven drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
-Another interesting feature to compare is $\pi$, the proportion defective.
-The bottom panel of Figure 9 shows the posterior credible intervals of $\pi$ for each drive-model.
+In this application 50\% CIs were used due to considerable posterior uncertainty in some of the parameters: especially for those with few drives in operation.
+%Interestingly, the top seven drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
+Another interesting feature to compare is $\pi_g$, the proportion defective.
+The bottom panel of Figure 9 shows the posterior credible intervals of $\pi_g$ for each drive-model.
 The plot is again sorted according to $t_{0.10}$ so it can be compared to the plot above.
-While for many of the drive-models the ordinal ranking is the same as on the top, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10}$ or $\pi$.
+While for many of the drive-models the ordinal ranking is the same as on the top, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10g}$ or $\pi_g$.
 

### 2018-01-22 paper/m-lb-m-technometrics.tex
Edits based on Skype call on Jan 21.

@@
 Consumers have an interest in accurate assessment of product reliability.
-Toward this end, there is a need for models that can accomodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
+Toward this end, there usually is a need for models that can accomodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
 Doing these things well enables good decision-making in matters related to product life, including purchase decisions and contingency planning.
@@
 
-Note that the $p$ quantile may be found by setting $p=H(t_p)$ and solving numerically for $t_p$.
+\noindent Note that the $p$ quantile of the GLFP model may be found by setting $p=H(t_p)$ and solving numerically for $t_p$.
 
@@
 A common feature of lifetime data is right-censoring.
-In the analysis of reliability field data, it is rare all units are observed until failure.
+In the analysis of reliability field data, it is rare that all units are observed until failure.
 If a unit has not yet failed when the data are analyzed it is considered right-censored.
@@
 The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.
-The curve has a slow ramp up in failures till about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).
+The curve has a slow ramp up in failures until about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).
 Besides our empirical observation, it is well known that computer hard drives have a mixture of early and late failures \citep{chan}.
@@
 We now apply this general methodology to the hard drive failure data from Backblaze.
-There are 44 different drive-models in the data set, and significant heterogeneity in the number of observed failures and time observed (Figure 1).
-The hierarchical GLFP allows us to model the entire drive-model population, borrowing information across drive-models, allowing estimation for those drive-models where little information exists.
+There are 44 different drive-models in the data set, and significant heterogeneity in the number of observed failures, time observed (Figure 1), and the parameter estimates of the underlying GLFP.
+The hierarchical GLFP allows us to model the entire drive-model population, borrowing information across drive-models, allowing estimation for those drive-models for which little information exists.
 
@@
 Data for all units are left-truncated, being observed conditional on survival beyond 12,000 hours.\hspace{\textwidth}Bottom: The dashed step function correponds to the adjusted K-M estimate (solid line).
-Due to the heavy left-truncation, the posterior median for Drive-model 43 (dotted line) coincides with the posterior median of the ``global model" (dashed line) until the first left-truncation time.}
+Due to the heavy left-truncation, the posterior median for Drive-model 43 (dotted line) coincides with the posterior median of the ``global average" (dashed line) until the first left-truncation time.}
 \label{fig:ex-mod-43}

### 2018-01-24 paper/m-lb-m-technometrics.tex
Enlarge legend for figure 5
small edits.

@@
 \subsection{Motivation}
-Consumers have an interest in accurate assessment of product reliability.
-Toward this end, there usually is a need for models that can accomodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
+Consumers have an interest in accurate assessment of product reliability}.
+Toward this end, there usually is a need for models that can accommodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
 Doing these things well enables good decision-making in matters related to product life, including purchase decisions and contingency planning.

### 2018-03-30 review/compare_posterior_ind_full.R
Scripts for plots to help address referee comments

@@ -1,0 +1,57 @@
+library(rstan)
+
+sfull <- readRDS("workflow/samples_lor_only3fails.rds")
+fullsamp <- extract(sfull)
+
+#3, 8, 14 correspond to 6, 14, 23 in full analys.
+s6 <- readRDS("workflow2/samples_nopool_data_dm3.rds")
+samp6 <- extract(s6)
+
+hist(samp6$log_tp2)
+hist(fullsamp$log_tp2_raw[,6] * fullsamp$tau_tp2 + fullsamp$eta_tp2, add=TRUE)
+plot(density(samp6$log_tp2))
+lines(density(fullsamp$log_tp2_raw[,6] * fullsamp$tau_tp2 + fullsamp$eta_tp2), lty=2)
+curve(dnorm(x, 10, 4), col=2, add=TRUE)
+
+s14 <- readRDS("workflow2/samples_nopool_data_dm8.rds")
+samp14 <- extract(s14)
+
+
+str(fullsamp)
+str(samp14)
+
+hist(samp14$log_tp2)
+hist(fullsamp$log_tp2_raw[,14] * fullsamp$tau_tp2 + fullsamp$eta_tp2, add=TRUE)
+plot(density(samp14$log_tp2))
+lines(density(fullsamp$log_tp2_raw[,14] * fullsamp$tau_tp2 + fullsamp$eta_tp2), lty=2)
+curve(dnorm(x, 10, 4), col=2, add=TRUE)
+
+s23 <- readRDS("workflow2/samples_nopool_data_dm14.rds")
+samp23 <- extract(s23)
+
+
+str(fullsamp)
+str(samp23)
+
+hist(samp23$log_tp2)
+hist(fullsamp$log_tp2_raw[,23] * fullsamp$tau_tp2 + fullsamp$eta_tp2, add=TRUE)
+plot(density(samp23$log_tp2))
+lines(density(fullsamp$log_tp2_raw[,23] * fullsamp$tau_tp2 + fullsamp$eta_tp2), lty=2)
+curve(dnorm(x, 10, 4), col=2, add=TRUE)
+
+
+plot(density(samp6$sigma2))
+lines(density(fullsamp$sigma2[,6]), lty=2)
+curve(dlnorm(x, 0, 2.5), col=2, add=TRUE)
+
+plot(density(samp6$logit_pi))
[…]

### 2018-03-30 review/look_at_doubling_var.R

@@ -1,0 +1,27 @@
+s <- readRDS("workflow/samples_lor_only3fails.rds")
+sdbl <- readRDS("workflow/samples_doublevar.rds")
+
+library(rstan)
+
+original <- extract(s)
+`inflated variance` <- extract(sdbl)
+
+plot(s, pars=c("eta_tp2","eta_s2","eta_pi"))
+plot(sdbl, pars=c("eta_tp2","eta_s2","eta_pi"))
+
+library(plyr)
+
+plot_dat <- ldply(c("original","inflated variance"), function(mod){
+  ldply(c("eta_tp2","eta_s2","eta_pi"), function(par){
+    data.frame(model = mod, param = par, t(quantile(get(mod)[[par]], c(.05,.5,.95))))
+  })
+})
+
+names(plot_dat)[3:5] <- c("lower","median","upper")
+
+
+library(ggplot2)
+
+ggplot(plot_dat, aes(x=model, ymin=lower, ymax=upper, y=median))+
+  geom_pointrange(aes(color=param)) + facet_wrap(~param, scales="free") +
+  theme_bw()

### 2018-03-30 paper/m-lb-m-technometrics.tex

@@
 \subsection{Motivation}
-Consumers have an interest in accurate assessment of product reliability}.
+Consumers have an interest in accurate assessment of product reliability.
 Toward this end, there usually is a need for models that can accommodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
@@
 \item Making purchase decisions from among different suppliers of the components or subsystems.
+\item Making inference about the lifetime distribution of products with few observed failures.
 \item Making predictions for the number of needed replacement components or subsystems for spare part provisioning.
@@
 Unfortunately, it can require a lot of data to fit due to the model's complexity.
+For example, clothing and electronic manufacturers, as well as pharmaceutical companies, often test products that rarely fail, even under accelerated testing.
+With few observed failures, fitting multi-parameter models can be unstable or impossible using standard estimation methods.
 When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates.
@@
 From the point of view of a consumer or purchaser, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
-In this case, we should choose the model that will provide the most accurate prediction for future units of previously observed products.
-One appropriate criterion for model selection, then, is the log pointwise predictive density (lpd),
+Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.
+Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
+posterior draws from a fitted model.  The log pointwise predictive density (lpd), is one method to quantify predictive accuracy,
 
@@
 The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent.
-It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\sum_{i=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$.
+It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $p(t_{g,i}|t_{-(g,i)})$ with $\sum_{s=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$.
 In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.
@@
 {\centering
-\includegraphics[width=\textwidth]{single-drive-4-Models-ex}}
+\includegraphics[width=\textwidth]{single-drive-4-Models-ex-v2}}
 \caption{Posterior median of the proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
@@
 \centering
-\begin{tabular}{rrrrrrr}
+\begin{tabular}{rrrrrrrr}
   \hline
- & $\widehat{\text{elpd}}$ \ & Difference in $\widehat{\text{elpd}}$ \ & SE of the Difference \\ 
+ & $\widehat{\text{elpd}}$ \ & Difference in $\widehat{\text{elpd}}$ \ & SE of the Difference & Effective No. Parameters\\ 
   \hline
-Model 4 & -13309.5 & 40.7 & 11.3 \\ 
-Model 3 & -13350.2 & 458.8 & 31.0  \\ 
-Model 2 & -13809.0 & 3674.6 & 96.2 \\ 
-Model 1 & -17483.6  \\ 
+Model 4 & -13309.5 & 40.7 & 11.3 & 92.5 \\ 
+Model 3 & -13350.2 & 458.8 & 31.0 & 68.8 \\ 
+Model 2 & -13809.0 & 3674.6 & 96.2 & 49.6 \\ 
+Model 1 & -17483.6 &  &  & 4.3 \\
    \hline
@@
 Each model is compared to the more parsimonious model below.
-The estimated difference of expected leave-one-out prediction errors between the two models, as well as the standard error of the difference, is also presented.}
+The estimated difference of expected leave-one-out prediction errors between the two models, the standard error of the difference, and the effective number of parameters in each model, is also presented.}
 \label{table:2}

### 2018-03-31 paper/supplement.tex
Adds density plots of the prior distributions to the Supplemental Materials document

@@
 \section{Posterior predictive graphical checks}
-\item Full set of Figures 5 and 6 for all 44 drive-models.
+Full set of Figures 5 and 6 for all 44 drive-models.
 Numbering in the top left indicates drive-model.
@@
 
+\section{Density Plots for Prior Distributions}
+To better visualize the prior distributions, below are plots of
+the prior densities for Models 1 - 4.
+\begin{figure}[H]
+\center
+\includegraphics[width=\textwidth]{priorsmod1.pdf}
+\caption{Prior Densities for Model 1}
+\end{figure}
+
+\begin{figure}[H]
+\center
+\includegraphics[width=.7\textwidth]{priorsmod2.pdf}
+\caption{Prior Densities for Model 2}
+\end{figure}
+
+\begin{figure}[H]
+\center
+\includegraphics[width=\textwidth]{priorsmod4.pdf}
+\caption{Prior Densities for Models 3 and 4}
+\end{figure}
+
 \end{document}

### 2018-04-06 plotting_fns/KM_plot2.R
Make greenwood confidence bands 'steps'

@@ -114,7 +114,8 @@ trimKMy <- function(fit, ylimits){
   fit
 }
 
-baseKMplot.multiple <- function(fit_list, xlimits=NULL, ylimits=NULL, color="black", linetype=1, logscale=NULL, prob=TRUE, label="nonparametric", alpha=1){
+baseKMplot.multiple <- function(fit_list, xlimits=NULL, ylimits=NULL, color="black",
+                                linetype=1, logscale=NULL, prob=TRUE, label="nonparametric", alpha=1){
   require(ggplot2)
   classes <- sapply(fit_list, function(li) "myKM" %in% class(li))
   if(!(all(classes))) stop("fits must be outputs from KM.survfit")
@@ -145,7 +146,8 @@ baseKMplot.multiple <- function(fit_list, xlimits=NULL, ylimits=NULL, color="bla
 }
 
 #Make base KMplot
-baseKMplot <- function(fit, xlimits=NULL, ylimits=NULL, color="black", linetype=1, logscale = FALSE, prob=TRUE, label="nonparametric", alpha=1){
+baseKMplot <- function(fit, xlimits=NULL, ylimits=NULL, color="black", linetype=1,
+                       logscale = FALSE, prob=TRUE, label="nonparametric", alpha=1){
   require(ggplot2)
   if(!("myKM" %in% class(fit))) stop("fit must be output from KM.survfit")
   if(!is.null(xlimits)){
@@ -289,7 +291,11 @@ plotFinally <- function(plotList, xbrks, ybrks, years=FALSE, greenwood=FALSE, al
   p <- plotList$base
   
   if(greenwood){
-    p <- p + geom_ribbon(inherit.aes = F, aes(x=t, ymin=lower, ymax=upper, fill=grp), alpha=alpha)
+    df_greenwood <- bind_rows(old=p$data,
+                             new=p$data %>% mutate(lower=lag(lower), upper=lag(upper)),
+                             .id = "source") %>%
+      arrange(t, source)
+    p <- p + geom_ribbon(data=df_greenwood, inherit.aes = F, aes(x=t, ymin=lower, ymax=upper, fill=grp), alpha=alpha)
   }
   
   

### 2018-04-06 paper/m-lb-m-technometrics.tex

@@
 $$\sigma_1 \sim \op{Lognormal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
-$$t_{p_1} \sim \op{Lognormal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
+$$t_{0.5} \sim \op{Lognormal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
 
@@
   \tau_{\pi} & \sim \op{Half-Cauchy}(0, 1)\\
-  \eta_{\sigma ,2} & \sim \op{Normal}(0, 2)\\
-  \tau_{\sigma ,2} & \sim \op{Half-Cauchy}(0, 1)\\
-  \eta_{t_{.2_2},2} & \sim \op{Normal}(9, 2)\\
-  \tau_{t_{.2_2},2} & \sim \op{Half-Cauchy}(0, 1).
+  \eta_{\sigma_2} & \sim \op{Normal}(0, 2)\\
+  \tau_{\sigma_2} & \sim \op{Half-Cauchy}(0, 1)\\
+  \eta_{t_{0.2}} & \sim \op{Normal}(9, 2)\\
+  \tau_{t_{0.2}} & \sim \op{Half-Cauchy}(0, 1).
  \end{align*}

### 2018-04-06 review/look_at_doubling_var.R
Make density plots comparing priors and posterior

@@ -3,8 +3,8 @@ sdbl <- readRDS("workflow/samples_doublevar.rds")
 
 library(rstan)
 
-original <- extract(s)
-`inflated variance` <- extract(sdbl)
+`original posterior`<- extract(s)
+`inflated variance posterior` <- extract(sdbl)
 
 plot(s, pars=c("eta_tp2","eta_s2","eta_pi"))
 plot(sdbl, pars=c("eta_tp2","eta_s2","eta_pi"))
@@ -25,3 +25,22 @@ library(ggplot2)
 ggplot(plot_dat, aes(x=model, ymin=lower, ymax=upper, y=median))+
   geom_pointrange(aes(color=param)) + facet_wrap(~param, scales="free") +
   theme_bw()
+
+normal.priors <- data.frame(param = rep(c("eta_tp2","eta_s2","eta_pi"),each=2),
+                            Distribution = rep(c("original prior","inflated variance prior"), 3),
+                            mu = rep(c(9,0,-3), each=2),
+                            sigma = c(2,4,2,4,1,2))
+
+df_plot_prior <- ddply(normal.priors, .(param,Distribution), function(x){
+  data.frame(param=x$param, Distribution = x$Distribution, value=rnorm(10000, x$mu,x$sigma))
+})
+
+df_plot_poster <- ldply(c("original posterior","inflated variance posterior"), function(m){
+  ldply(c("eta_tp2","eta_s2","eta_pi"), function(p){
+    data.frame(value=get(m)[[p]], Distribution=m, param=p)
+  })
+})
+df_plot = rbind(df_plot_prior,df_plot_poster)
+df_plot$Distribution <- factor(df_plot$Distribution, levels=c("original prior","inflated variance prior","original posterior","inflated variance posterior"))
+ggplot(df_plot, aes(value)) + geom_density(aes(fill=Distribution, color=Distribution), alpha=.5, lty=2) + facet_wrap(~param, scale="free",ncol=1)+
+  theme_bw(base_size=14) + xlab("")

### 2018-04-08 paper/paper.tex
Report consistent number of sigfigs in table
Edits bibtek and model selection section

@@
-\documentclass[12pt]{article}
+%\documentclass[12pt]{article}
+\documentclass[aap]{imsart}
 \usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url}
+\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url, imsart}
 \graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
@@
 {
-  \title{\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
-  \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
-    and \\
-    Colin Lewis-Beck\footnotemark[1]\\
-    and \\
-    William Q. Meeker\footnotemark[1]}
+  \title{\bf A HIERARCHICAL FAILURE-TIME MODEL FOR OBSERVATIONAL DATA EXHIBITING INFANT-MORTALITY AND WEAROUT FAILURE MODES}
+  \author{Eric Mittman and Colin Lewis-Beck and William Q. Meeker\\
+	\textit{Iowa State University}}
   \maketitle
@@
 \newpage
-\tableofcontents
+%\tableofcontents
 \newpage
@@
 
-\section{Supplementary Material}
+\section{Online Supplementary Material}
 \label{sec:supple}
+\begin{supplement}[id=suppA]
+  \sname{Supplement A}
+  \stitle{Stan Code and Plots}
+  \slink[doi]{COMPLETED BY THE TYPESETTER}
+  \sdatatype{.pdf}
+  \sdescription{Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6}
+\end{supplement}
 
-\begin{description}
-
-\item[Stan Code and Plots] Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6 (supplementary.pdf)
-
-\item[Backblaze hard-drive data] Lifetime data set for all 44 drive-models used in our analysis. (backblaze\_hd\_data.csv)
-
-\end{description}
+\begin{supplement}[id=suppB]
+  \sname{Supplement B}
+  \stitle{Backblaze Data}
+  \slink[doi]{COMPLETED BY THE TYPESETTER}
+  \sdatatype{.csv}
+  \sdescription{Lifetime data set for all 44 drive-models used in our analysis.}
+\end{supplement}
 

### 2018-04-10 paper/m-lb-m-technometrics.tex
Paper and bibtek edits after 4/8 meeting

@@
 \noindent%
-{\it Keywords:} GLFP, Bathtub hazard, Censored data, Hierarchical models, Bayesian estimation
+{\it Keywords:} GLFP, Bathtub hazard, Censored data, Bayesian estimation, Stan
 \vfill
@@
 $\mu=\log(\alpha)$ and $\sigma=1/\beta$ are, respectively, location
-and scale parameters for the distribution of $\log(T)$. Therefore, the Weibull
+and scale parameters for the distribution of $\log(T)$. In addition to showing the Weibull
 distribution is a member of the log-location-scale family of
-distributions.
+distributions, this step highlights the generalizability of our approach to other distributions 
+that fall into this class.
 
-We will use an alternative parameterization
+Going forward we use an alternative parameterization
 where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
@@
 \subsection{Model selection}
-From the point of view of a consumer or purchaser, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance.
+Bayes factors is a classical apporach to compare alternative model specifications.
+However, they can be problematic, especially in high dimensional settings where there is no concensus on how to properly compute the required marginal likelihood \citep{robert2016}.
+Moreover, from the point of view of a consumer or purchaser, the goal of an analysis such as this is to rate manufacturers and/or to inform future purchases based on predicted product performance.
 Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.
@@
 In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.
+A measure of the effective number of parameters is also available using the sum of the posterior variance of the log predictive density for each observation.
 For details, refer to \cite{vehtari}.

### 2018-04-22 paper/m-lb-m-technometrics.tex
fix the second typo

--- 2018-04-06 paper/m-lb-m-technometrics.tex
@@
 One way to do this is to draw replicate data sets from the posterior predictive distribution
-$$p(t_{rep,g}|t) = \int f(t_{rep,g}|\theta_g)\; p(\theta_g|t)\;d\theta_g.$$
+$$p(t_{rep,g}|t) = \int \prod_{i=1}^{n_g} f(t|\theta_g)\; p(\theta_g|t)\;d\theta_g,$$
+in the case that observations in subpopulation $g$ are exchangable.
+
 This is a general approach recommended by \cite{gelman1996postpred}, which works when no classical goodness-of-fit test is available.
@@
 Because the variability in the K-M estimator is primarily a function of the number of at-risk units, we want to replicate the pattern of censoring that was observed.
-That is, for unit $i$ of sub-population $g$,
-$$p(t_{rep,g}) = \int \prod_{i=1}^{n_g}f(t_{new}|t_{g,i}^L,c_{g,i},\theta)\,p(\theta|t)d\theta.$$
-Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing a draw, $\theta^{(s)}$, at random, then conditionally draw $t_{rep,g}$ from $f(t_{new}|t_{g,i}^L,c_{g,i},\theta^{(s)}).$
+To do this, we sample from a version of the posterior predictive that conditions on the observed truncation and censoring times, i.e.
+$$p(t_{rep,g}) = \int \prod_{i=1}^{n_g}f(t|t_{g,i}^L,c_{g,i},\theta)\,p(\theta|t)d\theta.$$
+Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing a draw, $\theta^{(s)}$, at random, then conditionally draw each $t_{rep,g,i}$ from $f(t|t_{g,i}^L,c_{g,i},\theta^{(s)}).$
 

### 2018-04-22 review/hier_accuracy_compare.R
Plot comparing unpooled vs hierarchical fit

@@ -1,0 +1,71 @@
+setwd("review")
+
+ov <- readRDS("../BB_data/overview.rds")
+full_id <- with(ov, which(f>=3))
+up_id <- with(ov, which(f>=5 & early_f>=1 & late_f>=1))
+ov$full_id <- NA
+ov$up_id <- NA
+ov$full_id[full_id] <- 1:length(full_id)
+ov$up_id[up_id] <- 1:length(up_id)
+
+s3 <- readRDS("../workflow2/samples_review_dm3.rds") #6
+s8 <- readRDS("../workflow2/samples_review_dm8.rds") #14
+s10 <- readRDS("../workflow2/samples_review_dm10.rds") #16
+s14 <- readRDS("../workflow2/samples_review_dm14.rds") #23
+s18 <- readRDS("../workflow2/samples_review_dm18.rds") #31
+
+
+full <- readRDS("../workflow/samples_lor_only3fails.rds")
+samp_full <- extract(full)
+
+psev = function(x, location = 0, scale = 1, lower.tail = TRUE, log.p = FALSE){
+  z = (x - location)/scale
+  upper <- exp(-exp(z))
+  if(lower.tail){
+    if(log.p)
+      return(log(1 - upper))
+    else
+      return(1-upper)
+  } else {
+    if(log.p)
+      return(log(upper))
+    else
+      return(upper)
+  }
+}
+solve4B10 <- Vectorize(function(log_pi, mu1, sigma1, mu2, sigma2){
+  #solve .1 = 1 - (1 - pi * pweibull(x, mu1, sigma1))(1 - pweibull(x, mu2, sigma2))
+  f <- function(x) {
+      .9 - (1 - exp(log_pi)*psev(x, location=mu1, scale=sigma1)) *
+                        (1 - psev(x, location=mu2, scale=sigma2))
+    }
+  uniroot(f, c(-100, 1000))$root
+})
+
+solve4B10(-4.4, 8.5, 1.2, 12.2, .4)
+
+unpooled_df <- ldply(c("s3","s8","s10","s14","s18"), function(i){
+  samp <- extract(get(i))
+  B10 <- with(samp, solve4B10(log_pi, mu1, sigma1, mu2, sigma2))
+  B10_interval <- quantile(B10, c(.05,.5,.95))
+  data.frame(set = i, lower_B10 = exp(B10_interval[1]), median_B10 = exp(B10_interval[2]),
[…]

### 2018-04-28 paper/m-lb-m-technometrics.tex
Clean up notation for predictive posterior intro section

@@
 One way to do this is to draw replicate data sets from the posterior predictive distribution
-$$p(t_{rep,g}|t) = \int \prod_{i=1}^{n_g} f(t|\theta_g)\; p(\theta_g|t)\;d\theta_g,$$
+$$p(t_{rep,g}|t) = \int \prod_{i=1}^{n_g} f(t_{rep,g,i}|\theta_g)\; p(\theta_g|t)\;d\theta_g,$$
 in the case that observations in subpopulation $g$ are exchangable.
 
-This is a general approach recommended by \cite{gelman1996postpred}, which works when no classical goodness-of-fit test is available.
+This is approach is recommended by \cite{gelman1996postpred} in part because it works well in cases when no classical goodness-of-fit test is available.
 When the model is adequate for the data, the predictive distribution of a finite set of data characteristics --- chosen to reflect features relevant to the goals of the analysis --- should not contradict the corresponding characteristics observed in the actualized data.
@@
 To do this, we sample from a version of the posterior predictive that conditions on the observed truncation and censoring times, i.e.
-$$p(t_{rep,g}) = \int \prod_{i=1}^{n_g}f(t|t_{g,i}^L,c_{g,i},\theta)\,p(\theta|t)d\theta.$$
-Because we are working with posterior draws, $\theta^{(s)}$, we can draw from this distribution by choosing a draw, $\theta^{(s)}$, at random, then conditionally draw each $t_{rep,g,i}$ from $f(t|t_{g,i}^L,c_{g,i},\theta^{(s)}).$
+$$p(t_{rep,g}|t) = \int \prod_{i=1}^{n_g}f(t_{rep,g,i}|t_{g,i}^L,c_{g,i},\theta_g)\;p(\theta_g|t)\;d\theta_g.$$
 
 An obstacle remains, owing to the fact that the censoring time is unknown for observed failures.
-Our admittedly simple approach is to assume that failed units would have been the censored had they survived until the latest censoring or failure time for that drive model.
+We resolve this by assuming that failed units would have been right censored had they survived until the latest censoring or failure time for that drive model.
+
 Our method for generating replicate data sets, $t_{rep,g}^{(s)}$, $s = 1,\ldots, R$, for each sub-population $g=1,\ldots,G$, is as follows:

### 2018-05-06 paper/supplement.tex
Edits prior densitites for supplemental material
Add plot to supplementary materials showing that posterior doesn't care if prior scale is doubled on hyperlocations

@@
 \includegraphics[width=\textwidth]{priorsmod1.pdf}
-\caption{Prior Densities for Model 1}
+\caption{Prior Densities for Model 1.  For $t_{p1 = 0.5}$, $t_{p2 = 0.2}$, $x$-axis is in thousands of hours.}
 \end{figure}
@@
 \center
-\includegraphics[width=.7\textwidth]{priorsmod2.pdf}
-\caption{Prior Densities for Model 2}
+\includegraphics[width=\textwidth]{priormod234.pdf}
+\caption{Prior Densities for Models 2 - 4 (top left to bottom right). For $t_{p1 = 0.5}$, $t_{p2 = 0.2g}$, $x$-axis is in thousands of hours.}
 \end{figure}
+
+\section{Sensitivity check}
+To check whether the posterior distribution of the location hyperparameters, $\eta_\pi$, $\eta_{t_{p_2}}$ and $\eta_{\sigma_2}$ in Model 4 was sensitive to our choice of prior distributions for these parameters, we refit Model 4 doubling the scale of the prior distributions, i.e.
+$$\eta_\pi \sim \text{N}(-3,2^2),\; \eta_{t_{p_2}} \sim \text{N}(9,4^2),\; \eta_{\sigma_2} \sim \text{N}(0, 4^2).$$
+Figure \ref{double_var} shows that the posterior was nearly identical with the wider prior distributions.
+This suggests that the posterior is overwhelmingly driven by the information from the data, rather than from the prior (for either choice of prior.)
 
@@
 \center
-\includegraphics[width=\textwidth]{priorsmod4.pdf}
-\caption{Prior Densities for Models 3 and 4}
+\includegraphics[width=\textwidth]{double_var.pdf}
+\caption{Comparison of prior and posterior for location hyperparameters, $\eta_\pi,\; \eta_{t_{p_2}}$ and $\eta_{\sigma_2}$ under the original priors from Section 6.1 and under a modified prior doubling the scale for these parameters.}
+\label{double_var}
 \end{figure}

### 2018-05-06 review/look_at_doubling_var.R

@@ -32,7 +32,7 @@ normal.priors <- data.frame(param = rep(c("eta_tp2","eta_s2","eta_pi"),each=2),
                             sigma = c(2,4,2,4,1,2))
 
 df_plot_prior <- ddply(normal.priors, .(param,Distribution), function(x){
-  data.frame(param=x$param, Distribution = x$Distribution, value=rnorm(10000, x$mu,x$sigma))
+  data.frame(param=x$param, Distribution = x$Distribution, value=rnorm(100000, x$mu,x$sigma))
 })
 
 df_plot_poster <- ldply(c("original posterior","inflated variance posterior"), function(m){
@@ -42,5 +42,43 @@ df_plot_poster <- ldply(c("original posterior","inflated variance posterior"), f
 })
 df_plot = rbind(df_plot_prior,df_plot_poster)
 df_plot$Distribution <- factor(df_plot$Distribution, levels=c("original prior","inflated variance prior","original posterior","inflated variance posterior"))
-ggplot(df_plot, aes(value)) + geom_density(aes(fill=Distribution, color=Distribution), alpha=.5, lty=2) + facet_wrap(~param, scale="free",ncol=1)+
-  theme_bw(base_size=14) + xlab("")
+
+
+p1 <- filter(df_plot, param == "eta_pi") %>%
+ggplot(aes(value)) + geom_density(aes(fill=Distribution, color=Distribution), alpha=.5, lty=2) + 
+  xlim(c(-6,1))+
+  # facet_wrap(~param, scale="free",ncol=1)+
+  theme_bw(base_size=14) + xlab("") + ylab("") + ggtitle(expression(eta[pi])) +
+  theme(axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.title = element_text(hjust = 0.5)) #+
+#  guides(color="none", fill="none")
+
+p2 <- filter(df_plot, param == "eta_tp2") %>%
+  ggplot(aes(value)) + geom_density(aes(fill=Distribution, color=Distribution), alpha=.5, lty=2) + 
+  xlim(c(6,13))+
+  # facet_wrap(~param, scale="free",ncol=1)+
+  theme_bw(base_size=14) + xlab("") + ylab("") + ggtitle(expression(eta[t[p[2]]])) +
+  theme(axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
[…]

### 2018-05-06 paper/m-lb-m-technometrics.tex
Add plot to supplementary materials showing that posterior doesn't care if prior scale is doubled on hyperlocations
Add to discussion a comment about relative accuracy of hierarchical fitting vs. separate fitting and supporting figure

@@
 \begin{align*}
-  \eta_{\pi} & \sim \op{Normal}(-3, 1)\\
+  \eta_{\pi} & \sim \op{Normal}(-3, 1^2)\\
   \tau_{\pi} & \sim \op{Half-Cauchy}(0, 1)\\
-  \eta_{\sigma_2} & \sim \op{Normal}(0, 2)\\
+  \eta_{\sigma_2} & \sim \op{Normal}(0, 2^2)\\
   \tau_{\sigma_2} & \sim \op{Half-Cauchy}(0, 1)\\
-  \eta_{t_{0.2}} & \sim \op{Normal}(9, 2)\\
+  \eta_{t_{0.2}} & \sim \op{Normal}(9, 2^2)\\
   \tau_{t_{0.2}} & \sim \op{Half-Cauchy}(0, 1).
@@
 
-\begin{figure}[H]
+\begin{figure}
 \includegraphics[width=\textwidth]{ppcheck-sample}
@@
 This can be important when trying to leverage multiple sources of data with various sample sizes and levels of censoring and truncation.
-We found that the use of weakly informative priors on the hyperparameters can increase the overall efficiency of the MCMC.
+
+In the case of the Backblaze data, we observed improved posterior precision for quantities of interest when using our hierarchical model when compared to separate models for lifetime by drive-model.
+This is illustrated in Figure \ref{improved_accuracy}.
+In addition, this leveraging information led to well-behaved posterior distributions for sub-populations with few observed failures.
+For many of these sub-populations, fitting separate GLFP models was unworkable.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\textwidth]{accuracy_compare}
+  \caption{Posterior distributions of 0.1 quantiles for a representative subset of drive-models where separate fitting of the GLFP model was successful (unimodal posteriors and passing MCMC convergence diagnostics.) $r$ indicates the number of observed failures.
+Separate fits use the same priors as Model 4, but have greater uncertainty because there is no borrowing of information across drive-models.}
+
+\label{improved_accuracy}
+\end{figure}
+
 We now address some important model assumptions and their potential for impacting statistical inference.

### 2018-05-06 review/hier_accuracy_compare.R

@@ -86,4 +86,6 @@ annotate_df <- ddply(comb, .(set), function(x){
   data.frame(y=max(x$upper_B10)+2500, text=paste(c("r = ",x$f[1]), collapse=""))
 })
 
+pdf(file="../paper/fig/accuracy_compare.pdf", width=11, height=6)
 p + geom_text(data=annotate_df, inherit.aes = FALSE, aes(x=set, y=y, label=text))
+dev.off()

### 2018-05-13 paper/m-lb-m-technometrics.tex
First pass at edits post 5/13 Sunday meeting

@@
 Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
-Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision.
+Furthermore, when the information carried by lifetime data are limited by sample size, censoring, and truncation, estimates can be unstable and suffer from imprecision.
 These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.
@@
 Physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive.
-This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution.
+This presents a dilemma for analysts: the data may indicate multiple failure modes, contradicting the use of a unimodal parametric distribution.
 However, without any knowledge of cause of failure, traditional competing risk models are not identifiable.
 \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model.
-This model can provide a solution to the problem of unknown cause of failure in the special case where there is evidence of two modes of failure which impact different stages of product life.
+This model can provide a solution to the problem of unknown causes of failure in the special case where there is evidence of two modes of failure which impact different stages of product life.
 It avoids non-identifiability by introducing a parameter representing the population proportion defective.
@@
 Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout of certain components.
-The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.
-A common assumption in series system modeling is the time to failure for the different risks are statistically independent.
+The general name for such a product is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.
+A common assumption in series system modeling is that the times to failure for the different risks are statistically independent.
 Thus, the overall reliability of a unit is the product of the reliability of each of the risks.
@@
 \cite{chan} labeled the cause of circuit board failures as early (due to defects), unknown, or wearout based on the time of observed failures.
-This helped identify parameters when using maximum likelihood (ML) estimation.
+Having knowledge of the failure mode for some of the failures helped identify parameters when using maximum likelihood (ML) estimation.
 Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the causes of failure.
@@
 \label{sec:GLFP model}
+The methods presented in this paper
+could employ any of the commonly-used lifetime distributions such as
+the Weibull or lognormal (see Chapters 4 and 5 of
+\cite{meeker} for others) to describe either the early
+(infant mortality) failures or the wearout failures. To simplify the
+presentation and because it adequately describes our example data,
+we will use the Weibull distribution to describe both the early and
+wearout failures. Changing to any other combination of distributions
+in the log-location-scale family of distributions is
+straightforward
 
-\subsection{Weibull distribution parameterization}
+\subsection{Weibull distribution and reparameterization}
 \label{sec:Weibull parameterization}
@@
 
-Going forward we use an alternative parameterization
-where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
-[-\log(1-p)\right ]^{\sigma}$. Replacing $\alpha$ in (1) with the
+Following the suggestions given in \cite{LiMeeker2014} and Section 15.2 of \cite{intervals} we use an alternative parameterization
+where the usual scale parameter $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
+[-\log(1-p)\right ]^{\sigma}$ (which is also a scale parameter). Replacing $\alpha$ in (1) with the
 $\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
@@
 \end{align*}
-There are two important reasons for using this parameterization.
+There are three important reasons for using this parameterization.
 \begin{itemize}
@@
 The parameter $\pi$ represents the proportion of units susceptible to early failure, and hence susceptible to both failure modes.
-Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.
+Here the cause of failure is not assumed to be known; thus units from the same population are exchangeable.
 
@@
 
-\noindent Note that the $p$ quantile of the GLFP model may be found by setting $p=H(t_p)$ and solving numerically for $t_p$.
+\noindent Note that the $p$ quantile of the GLFP model can be found by setting $p=H(t_p)$ and solving numerically for $t_p$.
 
@@
 \subsection{Likelihood}
+\label{subsec:likehood}
 We now give the general form for the likelihood function, taking into account left truncation and right censoring.
@@
 Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
-\begin{equation*}
+\begin{equation}
+\label{equation:likelihood}
 L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
-\end{equation*}
+\end{equation}
+
+\noindent This form of the likelihood assumes that the lifetimes are reported as exact
+failures. Most reliability and warranty databases have failure times
+reported to the nearest hour or day, where such a assumption would
+be warranted. In some applications, however, field failure data are
+summarized by reporting months is service (e.g., the Nevada format
+for warranty data). It is straightforward modify the likelihood to
+handle interval censored data by replacing $h(t_i;\bm{\theta})$ in
+(\ref{equation:likelihood}) with $H(t_{u i};\bm{\theta})-H(t_{\ell i};\bm{\theta})$
+where $t_{\ell i}$ and $t_{u i}$ are, respectively, the lower and
+upper endpoints of the interval in which unit $i$
+failed.
 
@@
 Backblaze is a company that offers cloud backup storage to protect against data loss.
-Since 2013 it has been collecting data on hard drives operating at its facility.
+Since 2013 it has been collecting daily operational data on all of the hard drives operating at its facility.
 The purpose is to provide consumers and businesses with reliability information on different hard drive-models.
@@
 In addition, Backblaze publishes summary statistics of the different drive-models currently operating.
-No other analysis or modeling of the failure data is provided, other than simple rates and proportions.
+No analysis or modeling of the failure data is provided, other than simple rates and proportions.
 Backblaze does, however, encourage others to further analyze their data.
@@
 The number of drives varies by drive-model; some drive-models only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.
-Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.
-For model identification, a minimum of three failures was the criterion for inclusion of drive-models into our analysis.
+Figure 1 shows a scatterplot of the total observed running time in hundreds of thousands hours versus the total number of failures for drive-models with at least 3 failures.
+For model identification, a minimum of three failures was the criterion for inclusion of a drive-model into our analysis.
 
@@
 Besides our empirical observation, it is well known that computer hard drives have a mixture of early and late failures \citep{chan}.
-Therefore, there is both an empirical and a theoretical justification to apply the GLFP model.
+Therefore, there is both an empirical and a theoretical justification to use the GLFP model in this application.
 
@@
 
-When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the parameters of the distribution.
-Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.
+When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the traditional parameters of the distribution.
+Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the priors).
 Using this convention the prior distributions used in our analyses are
@@
 t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
-$$\pi \sim \op{Logit-normal}\langle 1.0\times 10^{-3},\, 7.1\times 10^{-1} \rangle,\quad
-\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,\,1.3 \times 10^2 \rangle. $$
+$$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
+\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle. $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 \noindent These prior distributions put probability mass on a wide range of values for all model parameters --- much larger than we would expect for a typical Weibull distribution.
-Thus, we consider these prior distributions to be relatively uninformative.
+Thus, these prior distributions are relatively uninformative.
 
@@
   \hline
- & 2.5\% & 50\% & 97.5\% \\ 
+ & Lower & Estimate & Upper \\ 
   \hline
@@
 \end{tabular}
-\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.  Quantiles are in thousands of hours.}
+\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.  The quantiles are in thousands of hours.}
 \label{table:1}
@@
 Probability plotting is a simple method to assess and compare the adequacy of members of the log-location-scale family of distributions.
-After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.
+After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with confidence bands provides a visual assessment of distributional goodness of fit.
 We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.
@@
   \includegraphics[width=.8\textwidth]{km14-prob-plus}
-  \caption{The estimated GLFP for Drive-model 14 plotted on Weibull paper.
-The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are shown in red.
+  \caption{The estimated GLFP model for Drive-model 14 plotted on Weibull paper.
+The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are indicated by the smooth region.
 The solid line is an adjusted K-M estimate with pointwise 90\% credible bounds in blue.}
@@
 \label{sec:Hierarchical GLFP model}
+In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.
+
 \subsection{Conditional likelihood}
-
-In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.
 Let
@@
 where $g=1,\ldots,G$ indexes the sub-populations.
-The likelihood for the hierarchical GLFP model is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.
-Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
+The likelihood for the GLFP model for all subpopulations is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.
+Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on fixed values of the parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
 \begin{equation*}
@@
 where $T_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored.
-All times are again in hours.
+All times are in thousands of hours.
+Again, this likelihood
+is easy to modify for other situations like interval-censoring,
+as explained in Section~\ref{subsec:likehood}.
+
 \subsection{Parameter distributions}
@@
 
-\subsection{Priors}
+\subsection{Prior Distributions}
 The complete GLFP model requires prior distributions on the hierarchical parameters.
-We suggest weakly informative proper distributions that allow the data to overwhelm the prior and provide information about the distribution of the population parameters.
+We suggest weakly informative proper distributions that allow the data to overwhelm the prior distribution and provide information about the distribution of the population parameters.
 Prior knowledge about product lifetimes is useful for choosing the mean of the location parameter ($\eta$) distributions.
 However, the corresponding standard deviations should be diffuse to allow the location to vary depending on the observed data.
-For the scale parameters ($\tau$) we recommend half-Cauchy distributions as suggested by \citet{gelman2014bayesian}.
+For the scale parameters ($\tau$) we use half-Cauchy distributions as suggested by \citet{gelman2014bayesian}.
 
@@
 This set of model specifications is chosen based on the data, interpretation of the model, as well as estimation considerations.
-Sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.
+Sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the lifetime distribution.
 Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.
@@
 We found that there was often insufficient information to model these parameters hierarchically.
-Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).
-Also the failure-time distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.
+Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of the $\pi_g$ across the subpopulations.
+Also the lifetime distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.
 
 \subsection{Model selection}
-Bayes factors is a classical apporach to compare alternative model specifications.
+The use of Bayes factors is a classical apporach to compare alternative model specifications.
 However, they can be problematic, especially in high dimensional settings where there is no concensus on how to properly compute the required marginal likelihood \citep{robert2016}.
@@
 Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
-posterior draws from a fitted model.  The log pointwise predictive density (lpd), is one method to quantify predictive accuracy,
+posterior draws from a fitted model.  Following \cite{vehtari}, the log pointwise predictive density (lpd), is one method to quantify predictive accuracy,
 
@@
 A measure of the effective number of parameters is also available using the sum of the posterior variance of the log predictive density for each observation.
-For details, refer to \cite{vehtari}.
 %When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
@@
 We now apply this general methodology to the hard drive failure data from Backblaze.
-There are 44 different drive-models in the data set, and significant heterogeneity in the number of observed failures, time observed (Figure 1), and the parameter estimates of the underlying GLFP.
-The hierarchical GLFP allows us to model the entire drive-model population, borrowing information across drive-models, allowing estimation for those drive-models for which little information exists.
+There are 44 different drive-models in the data set, and significant heterogeneity in the number of observed failures, time observed (Figure 1), and the parameter estimates of the underlying GLFP model.
+The hierarchical GLFP model allows us to describe the entire drive-model population, borrowing information across drive-models, allowing estimation for those drive-models for which there is little information (i.e., few failures).
 
@@
 \paragraph{Model 2} $t_{p_{2}g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
-$$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
-$$\sigma_1 \sim \op{Lognormal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
-$$t_{0.5} \sim \op{Lognormal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$
+$$ \pi \sim \op{Logit-normal}\langle 0.007,\, 0.26 \rangle,$$
+$$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
+$$t_{0.5} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle.$$
 
@@
 As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
-
-\begin{align*}
-  \eta_{\pi} & \sim \op{Normal}(-3, 1^2)\\
-  \tau_{\pi} & \sim \op{Half-Cauchy}(0, 1)\\
-  \eta_{\sigma_2} & \sim \op{Normal}(0, 2^2)\\
-  \tau_{\sigma_2} & \sim \op{Half-Cauchy}(0, 1)\\
-  \eta_{t_{0.2}} & \sim \op{Normal}(9, 2^2)\\
-  \tau_{t_{0.2}} & \sim \op{Half-Cauchy}(0, 1).
- \end{align*}
+$$\eta_{\pi} \sim \op{Normal}(-3, 1^2),\quad \tau_{\pi} \sim \op{Half-Cauchy}(0, 1)$$
+$$\eta_{\sigma_2}  \sim \op{Normal}(0, 2^2),\quad
+\tau_{\sigma_2}  \sim \op{Half-Cauchy}(0, 1)$$
+$$\eta_{t_{0.2}}  \sim \op{Normal}(9, 2^2),\quad
+\tau_{t_{0.2}}  \sim \op{Half-Cauchy}(0, 1)$$
 
@@
 Other diagnostics provided by the software (tree-depth, and divergent transitions) did not indicate problems with sampling.
-Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.
+Plots of the posterior draws were inspected for parameters with the smallest effective sample size; these did not suggest features that were inadequately explored.
 
@@
 \label{sec:Model Comparisons}
-As we have discussed, while we prefer to allow all of the GLFP model parameters to be drive-model specific, there are practical issues with fitting the full model.
+As we have discussed, while we would prefer to allow all of the GLFP model parameters to be drive-model specific, there are practical issues with fitting the full model.
 For example, due to heavy left-truncation, there may be insufficient information for particular drive-models to estimate all of the parameters.
@@
 These plots suggest that Model 1 is insufficient to explain the observed data.
-Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2$ results in very similar progressions in the cdf, which may be too restrictive.
+Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2$ results in similar progressions in the cdf, which may be too restrictive.
 Model 3, which allows $\sigma_2$ to vary by sub-population, suggests there is evidence to support variation in this parameter among drive-models.
@@
 %stopped here
-The GLFP fit for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for similar plots for the other drive-models).
-The black step functions again correspond to the adjusted K-M estimates.
+The GLFP model fits for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for similar plots for the other drive-models).
+The step functions again correspond to the adjusted K-M estimates.
 As model complexity increases, the hierarchical-model GLFP curves become more flexible and are better able to describe the observed failure data.
@@
 For Drive-model 14, Models 1-3 produce lower estimates of the proportion of drives failing, and for Drive-model 40, they produce higher estimates.
-On the other hand, Model 4 follows the K-M closely across stages of lifetime.
+On the other hand, Model 4 follows the K-M closely across both stages of lifetime.
 The differences between Models 3 and 4 and Models 1 and 2 are visually apparent.
@@
 We calculated the difference in expected predictive accuracy for Model 4 versus 3, Model 3 versus 2, and Model 2 versus 1, as well as the standard error of the difference (Table 2).
-For the models we considered, each successive increase in model complexity increases elpd.
+For the models we considered, each successive increase in model complexity increases the elpd.
 Of all the models, Model 4 has significantly better elpd compared to the the other 3 models.
@@
 \end{tabular}
-\caption{Expected Log Pointwise Predictive Density (elpd) for each model specification.
-Each model is compared to the more parsimonious model below.
-The estimated difference of expected leave-one-out prediction errors between the two models, the standard error of the difference, and the effective number of parameters in each model, is also presented.}
+\caption{Expected log pointwise predictive density (elpd) and related statistics for each model.}
 \label{table:2}
@@
 in the case that observations in subpopulation $g$ are exchangable.
-
 This is approach is recommended by \cite{gelman1996postpred} in part because it works well in cases when no classical goodness-of-fit test is available.
@@
 
-For $s=1,\ldots, R$, do:
+For $s=1,\ldots, R$,
 
@@
 \end{enumerate}
-Finally, plot K-M estimates based on the $R$ replicated data sets along with the one based on the actualized data on the same axes.
+Finally, plot K-M estimates based on the $R$ replicated data sets along with the one based on the observed data on the same axes.
 Discrepancies between the replicated and actual data can be used to diagnose inadequacies of the model.
@@
 For the hard-drive data, we used the posterior samples obtained in fitting Model 4 to generate $R=19$ replicate data sets as described above.
-The resulting plots are displayed in Figure \ref{fig:post-pred-KM}.
-
-% Having chosen Model 4 as the best of the candidate models, we would like to assess whether our model provides an adequate description of the observed data.
-In the present context, one could  or a new drive-model.
-We choose to focus on the first of these and apply it to individual drive-models.
-There are many possible approaches to the posterior predictive check.
-For this problem, we chose to consider the posterior predictive distribution of the K-M estimate for each drive-model $g$.
-
-% Graphical comparisons can be highly informative in revealing limitations of the model.
-as in our cas a graphical posterior predictive check, wherein replicate data sets are generated from the posterior predictive distribution and compared with the observed data.
-Because of the rich information gleaned from plotting the data, this method can be very effective at identifying limitations of the model.
+The resulting plots for Drive-models 2, 9, 14, and 40 are displayed in Figure \ref{fig:post-pred-KM} (see the Supplementary Material for similar plots for all of the drive-models).
 
@@
 
-% We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to re-estimate the K-M estimator. 
-Only a representative sample of plots is shown here for brevity (see the Supplementary Material for similar plots for all of the drive-models).
 For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.
@@
 either improve reliability or to reduce cost. Such changes will
-sometimes have an important effect on the product's failure-time
+sometimes have an important effect on the product's lifetime
 distribution, invalidating one of our important assumptions that
-failure-time distribution for a drive-model does not depend on the
+lifetime distribution for a drive-model does not depend on the
 date of manufacture.
@@
 \caption{\footnotesize Top: Event plot for 60 randomly sampled drives of Drive-model 4.
-Note that all drives in the sample are left truncated well past 1000 hours.\hspace{\textwidth}Bottom: The adjusted K-M estimate (dashed step function) is substantially higher than the pointwise posterior median (dashed line; shaded region showing 90\% credible interval).
+Note that all drives in the sample are left truncated well past 1000 hours.\hspace{\textwidth}Bottom: The adjusted K-M estimate (dashed step function) is substantially higher than the pointwise posterior median (dashed line; shaded region showing pointwise 90\% credible intervals).
 However, this discrepancy is due to several early failures among the small set of drives with the earliest left-truncation times (the newest set).
@@
 The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.
-In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for drive-model 43 to an average GLFP cdf for the entire population, which we call the ``global average;" a description of the global average is given in Appendix \ref{global-avg}.
-We can see that the lack of data for this drive-model during early life results in a diffuse posterior centered around the global average until about 10,000 hours when they diverge.
-Although the posterior is close to the adjusted K-M, it is shrunk toward the global average.
+In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for drive-model 43 to an average GLFP cdf for the entire population, which we call the ``global average" cdf; a description of the global average cdf is given in Appendix \ref{global-avg}.
+We can see that the lack of data for this drive-model during early life results in a diffuse posterior centered around the global average cdf until about 10,000 hours when they diverge.
+Although the posterior is close to the adjusted K-M, it has been shrunk toward the global average cdf.
 
@@
 As discussed in Section \ref{sec:Weibull parameterization}, we use a parameterization of the Weibull distribution in terms of a chosen quantile, making prior elicitation straightforward.
-Posterior sampling immediately gives marginal posterior distributions for quantiles, which in reliability applications is typically more informative than a measure of central tendency.
+Posterior sampling immediately gives marginal posterior distributions for quantiles, which in reliability applications is typically more useful than a measure of central tendency.
 To compare the unconditional $p$ quantile for each sub-population, plots of posterior credible intervals may be used.
@@
 
-\subsection{Example: Drive-model comparison based on quantiles and proportion defective}
-The top panel of Figure 9 shows estimates of $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail, i.e.
-the quantile corresponding to the $H(t)$ distribution in Section \ref{subsec:GLFP model}), for all drive-models.
+\subsection{Example: Drive-model comparison based on lower-tail quantiles and proportion defective}
+The top panel of Figure 9 shows estimates of $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail, corresponding to the cdf $H(t)$ in Section \ref{subsec:GLFP model}), for all drive-models.
 In this application 50\% CIs were used due to considerable posterior uncertainty in some of the parameters: especially for those with few drives in operation.
 %Interestingly, the top seven drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
-Another interesting feature to compare is $\pi_g$, the proportion defective.
-The bottom panel of Figure 9 shows the posterior credible intervals of $\pi_g$ for each drive-model.
-The plot is again sorted according to $t_{0.10}$ so it can be compared to the plot above.
+The bottom panel of Figure 9 shows the posterior credible intervals for the proportion defective, $\pi_g$, for each drive-model.
+The plot is again sorted according to $t_{0.10}$ so it can be compared to the top plot.
 While for many of the drive-models the ordinal ranking is the same as on the top, there are some drive-model comparisons, for example 23 and 18, that differ in ranking if compared using $t_{0.10g}$ or $\pi_g$.
@@
   \caption{Left: 90\% CI of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
-  Right: A comparison (90\% CI) of the predictive distribution of TTF (top) and VAF (bottom) based on 40\% for best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
+  Right: A comparison (90\% CI) of the predictive distribution of TTF in hundreds of thousands of hours (top) and VAF (bottom) based on 40\% depreciation for the best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
   \label{fig5}
@@
 
-We consider the problem of ranking drive-models with respect to economical value. From a
+Here we show how to rank drive-models with respect to economical value. From a
 business perspective, it is clear
-that we should prefer the drive-models which will provide more years
+that we should prefer drive-models that will provide more years
 of service. For ease of exposition, we will assume that the purchase
@@
 Posterior credible intervals for $L(t_{g,new})$ are shown in the left panel of Figure \ref{fig5}.
-A zoomed version of the top eight drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel.
+A zoomed version for the top eight drive-models, as ranked by $\op{E}(L(t_{g,new})|y)$ are shown in the bottom right panel.
 The panel above shows credible intervals for MTTF.
 We can see that ranking by MTTF would lead to a different result -- a new Drive-model 8 unit is expected to last the longest, but the possibility of it failing early is higher than for the drive-models ranked highest by our depreciation-aware metric.
-The top four drive-models (4, 5, 2, 3) seem comparable, with a new Drive-model 4 unit expected to depreciate the most.
+The top four drive-models (4, 5, 2, 3) are comparable, with a new Drive-model 4 unit expected to depreciate the most.
 However, given the discrepancy of the model fit for Drive-model 4 (noted in the previous subsection), in the absence of more early life data, we might harbor some lingering concern about possible quality-control issues with this Drive-model.
@@
 Products with long lifetimes often have few failures and when failures do occur the cause is frequently unknown to the analyst.
-The GLFP model provides a framework to accurately model data with evidence of multiple failure modes.
-Moreover, when there are multiple product populations within the same product class, our hierarchical modeling approach borrows strength across brands with many observed failures to help make inference for those populations with little information.
+The GLFP model provides a framework to accurately model data with evidence of multiple failure modes that behave differently (e.g., early failures and wearout failures).
+Moreover, when there are multiple product populations within the same product class, our hierarchical modeling approach borrows strength across brands with many observed failures to help make inference for those brands with few failures.
 This can be important when trying to leverage multiple sources of data with various sample sizes and levels of censoring and truncation.
@@
 
-Consumers of manufactured goods usually lack the data and expertise to perform a failure analysis to determine the cause of failure.
-Moreover, their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.
+Consumers of manufactured products usually lack the data, expertise, or motivation to perform a failure analysis to determine the cause of failure.
+Their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.
 The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product sub-populations.
-Because the Bayesian approach provides full posteriors distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the actual application.
+Because the Bayesian approach provides full posterior distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the actual application.
 
@@
 \section*{Acknowledgment}    
-The authors express their gratitude to Nick Clark for his helpful comments and suggestions.
+The authors express their gratitude to Nick Clark for his helpful comments and suggestions.  We would also like to thank the anonymous referees who provided 
+many useful suggestions that helped improve this paper.
 
@@
 
-While this adjustment is may be negligible for sub-populations with little truncation, in our Backblaze example, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
+While this adjustment may be negligible for sub-populations with little truncation, in our Backblaze example, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
 2.3 years.
@@
 In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a ``global average" which represents an average GLFP cdf for the entire population, $\op{H}\left(\cdot|\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
-Since $\mu_1$ and $\sigma_1$ are common across all sub-populations in our model, these can already be interpreted as ``global." For the parameters that vary across sub-populations, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.
+Because $\mu_1$ and $\sigma_1$ are common across all sub-populations in our model, these can already be interpreted as ``global." For the parameters that vary across sub-populations, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.
 In particular, we set
 $$\bar{\pi}=\op{logit}^{-1}(\eta_{\pi}),\;\bar{\mu}_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2) \mbox{ and } \bar{\sigma}_2= m_{\sigma_2}.$$
-Let $J(\cdot|a,b),\;J^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a ognormal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
+Let $J(\cdot|a,b),\;J^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a lognormal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
 $$m_{\sigma_2}=J^{-1}[0.5 \cdot J(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}],$$
-which is the median of a ognormal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.
+which is the median of a lognormal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.
 
@@
 
-\item[Model code and plots] Stan code for Model 4 as well as the complete set of plots for Figures 5 and 6 (supplementary.pdf)
+\item[Model code and plots] Stan code for Model 4 as well as the complete set of plots similar to Figures 5 and 6 (supplementary.pdf)
 

### 2018-05-14 paper/m-lb-m-technometrics.tex
Fixes revised inserts

@@
 The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions.
-Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate.
-Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
+Because of this, standard lifetime models, such as using a single Weibull or
+lognormal distribution may be inadequate. Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
 Furthermore, when the information carried by lifetime data are limited by sample size, censoring, and truncation, estimates can be unstable and suffer from imprecision.
@@
 \cite{meeker} for others) to describe either the early
-(infant mortality) failures or the wearout failures. To simplify the
+(infant mortality) failures or the wearout failures. 
+In this paper
+we use the Weibull distribution to describe both the early and
+wearout failures for the following reasons.
+\begin{itemize}
+\item
+This combination of distributions provides a good fit to the data.
+\item
+Mechanistically, within both of the different kinds of failures there
+are almost certainly several, if not many, different competing
+causes. The actual cause for any particular failed drive
+would be the minimum time of all of these potential causes, suggesting a
+Weibull distribution because of extreme value theory (see page 86  of
+\cite{meeker}) for further discussion of this point.
+\end{itemize}
+
+To simplify the
 presentation and because it adequately describes our example data,
@@
 in the log-location-scale family of distributions is
-straightforward
+straightforward.
 
@@
 $\delta_i$ is an indicator for whether unit $i$ is defective or not (i.e., susceptible to an early failure).
-Expressed conditional on~$\delta_i$,
+Here $F_{1}(t)$ is the cdf for the early failures and $F_{2}(t)$ is
+the cdf for the wearout failures. Expressed conditional on~$\delta_i$,
 \begin{align*}

### 2018-05-14 paper/m-lb-m-technometrics.tex
Complete edits based on May 13 meeting, and add drive brand names to Supplemental Material

@@
 \usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url}
+\usepackage{graphicx,psfrag,epsf,float, bm, subcaption, url, makecell}
 \graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
@@
 Doing these things well enables good decision-making in matters related to product life, including purchase decisions and contingency planning.
-%While computationally convenient, simple lifetime models can fail at each of these.
 
@@
   \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
-  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.
+  \caption{\footnotesize Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.
 Each number labels a unique drive-model, numbered between 1 and 44.
@@
   \includegraphics[width=.6\textwidth]{km14-prob}
-  \caption{K-M estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
+  \caption{\footnotesize K-M estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
   \label{fig1}
@@
 \end{tabular}
-\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.  The quantiles are in thousands of hours.}
+\caption{\footnotesize Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.  The quantiles are in thousands of hours.}
 \label{table:1}
@@
   \includegraphics[width=.8\textwidth]{km14-prob-plus}
-  \caption{The estimated GLFP model for Drive-model 14 plotted on Weibull paper.
+  \caption{\footnotesize The estimated GLFP model for Drive-model 14 plotted on Weibull paper.
 The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are indicated by the smooth region.
@@
 \includegraphics[width=\textwidth]{heterogeneity-compare}
-\caption{Left: Adjusted K-M estimates of the time to failure for each of the drive-models in the Backblaze data.
+\caption{\footnotesize Left: Adjusted K-M estimates of the time to failure for each of the drive-models in the Backblaze data.
 Right: Pointwise posterior median cdf estimates for Models 1-4, ordered left to right, top to bottom.}
@@
 \includegraphics[width=\textwidth]{single-drive-4-Models-ex-v2}}
-\caption{Posterior median of the proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
+\caption{\footnotesize Posterior median of the proportion failing as a function of time under Models 1-4 for sample of drive-models, with axes scales as for Weibull probability plotting.
 The solid step function is the adjusted K-M estimate.}
@@
   \hline
- & $\widehat{\text{elpd}}$ \ & Difference in $\widehat{\text{elpd}}$ \ & SE of the Difference & Effective No. Parameters\\ 
+ & $\widehat{\text{elpd}}$ \ & \thead{Estimated Difference \\ in $\widehat{\text{elpd}}$} \ & \thead{Standard Error of \\ the Difference} & \thead{Effective Number of \\ Parameters}\\ 
   \hline
@@
 \end{tabular}
-\caption{Expected log pointwise predictive density (elpd) and related statistics for each model.}
+\caption{\footnotesize Expected log pointwise predictive density (elpd) and related statistics for each model.}
 \label{table:2}
@@
 \includegraphics[width=\textwidth]{ppcheck-sample}
-\caption{Adjusted K-M estimates of the proportion failing for a representative subset of drive-models.
+\caption{\footnotesize Adjusted K-M estimates of the proportion failing for a representative subset of drive-models.
 Both the original data (\textbf{bold} line) and 19 ``replicated" data sets from the posterior predictive distribution (\textit{dashed} lines) are shown.}
@@
    \includegraphics[width=1\linewidth]{b10new.pdf}
-   \caption{Point estimate and 50\% CI for $t_{0.10}$}
+   \caption{\footnotesize Point estimate and 50\% CI for $t_{0.10}$}
    \label{fig:Ng1} 
@@
    \includegraphics[width=1\linewidth]{pinew.pdf}
-   \caption{Point estimate and 50\% CI for $\pi$}
+   \caption{\footnotesize Point estimate and 50\% CI for $\pi$}
    \label{fig:Ng2}
@@
 
-\caption{(a) 50\% CI (in years) for $t_{0.10}$ (b) 50\% CI for $\pi$ plotted on the logit scale.  Both plots are sorted based on the median value of the marginal posterior distribution of $t_{0.10}$.}
+\caption{\footnotesize (a) 50\% CI (in years) for $t_{0.10}$ (b) 50\% CI for $\pi$ plotted on the logit scale.
+Both plots are sorted based on the median value of the marginal posterior distribution of $t_{0.10}$.}
 \end{figure}
@@
   \includegraphics[width=.8\textwidth]{dm-eval}
-  \caption{Left: 90\% CI of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
+  \caption{\footnotesize Left: 90\% CI of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
   Right: A comparison (90\% CI) of the predictive distribution of TTF in hundreds of thousands of hours (top) and VAF (bottom) based on 40\% depreciation for the best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
@@
 $$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}.$$
+where $t_{g,new}^{(s)}$ is a draw from the posterior predictive distribution for drive-model $g$.
 
@@
   \includegraphics[width=.7\textwidth]{accuracy_compare}
-  \caption{Posterior distributions of 0.1 quantiles for a representative subset of drive-models where separate fitting of the GLFP model was successful (unimodal posteriors and passing MCMC convergence diagnostics.) $r$ indicates the number of observed failures.
+  \caption{\footnotesize Posterior distributions of 0.1 quantiles for a representative subset of drive-models where separate fitting of the GLFP model was successful (unimodal posteriors and passing MCMC convergence diagnostics.) $r$ indicates the number of observed failures.
 Separate fits use the same priors as Model 4, but have greater uncertainty because there is no borrowing of information across drive-models.}

### 2018-05-14 paper/supplement.tex
Complete edits based on May 13 meeting, and add drive brand names to Supplemental Material

@@
 
+\section{Drive Model Brands}
+Below are the brand names (provided by Backblaze) corresponding to
+the model numbers used in the analysis.
+\begin{table}[H]
+\centering
+\begin{tabular}{lr}
+  \hline
+Brand Name & Model ID \\ 
+  \hline
+HGST HDS5C4040ALE630 &   1 \\ 
+  HGST HMS5C4040ALE640 &   2 \\ 
+  HGST HMS5C4040BLE640 &   3 \\ 
+  Hitachi HDS5C3030ALA630 &   4 \\ 
+  Hitachi HDS5C4040ALE630 &   5 \\ 
+  Hitachi HDS722020ALA330 &   6 \\ 
+  Hitachi HDS723020BLA642 &   7 \\ 
+  Hitachi HDS723030ALA640 &   8 \\ 
+  ST1500DL003 &   9 \\ 
+  ST2000DL001 &  10 \\ 
+  ST2000DL003 &  11 \\ 
+  ST250LM004 HN &  12 \\ 
+  ST250LT007 &  13 \\ 
+  ST3000DM001 &  14 \\ 
+  ST31500341AS &  15 \\ 
+  ST31500541AS &  16 \\ 
+  ST3160316AS &  17 \\ 
+  ST3160318AS &  18 \\ 
+  ST32000542AS &  19 \\ 
+  ST320005XXXX &  20 \\ 
+  ST320LT007 &  21 \\ 
+  ST33000651AS &  22 \\ 
+  ST4000DM000 &  23 \\ 
+  ST4000DX000 &  24 \\ 
+  ST500LM012 HN &  25 \\ 
+  ST6000DX000 &  26 \\ 
+  ST9250315AS &  27 \\ 
+  TOSHIBA DT01ACA300 &  28 \\ 
+  TOSHIBA MD04ABA400V &  29 \\ 
+  WDC WD10EACS &  30 \\ 
+  WDC WD10EADS &  31 \\ 
+  WDC WD10EADX &  32 \\ 
+  WDC WD1600AAJB &  33 \\ 
+  WDC WD1600AAJS &  34 \\ 
+  WDC WD1600BPVT &  35 \\ 
+  WDC WD20EFRX &  36 \\ 
+  WDC WD30EFRX &  37 \\ 
+  WDC WD30EZRX &  38 \\ 
+  WDC WD3200BEKX &  39 \\ 
+  WDC WD5000LPVX &  40 \\ 
+  WDC WD60EFRX &  41 \\ 
+  WDC WD800AAJB &  42 \\ 
+  WDC WD800BB &  43 \\ 
+  WDC WD800JB &  44 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
 \end{document}

### 2018-05-19 paper/m-lb-m-technometrics.tex
Fixes small equation typos

@@
 \label{equation:reparameterized.weibull.cdf}
-&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
+&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0
 \end{align*}
@@
 Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p1},\sigma_1)$ and $(t_{p2}, \sigma_2)$, respectively.
-The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$. Then
+The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$ Then
 $$\Pr(T \le t) = H(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
@@
 \begin{align*}
-P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
-P(T\le t | \delta_i=0) =& F_2(t).
+P(T\le t | \delta_i=1) &=  1 -(1-F_1(t))(1-F_2(t))\\
+P(T\le t | \delta_i=0) &=  F_2(t)
 \end{align*}
@@
 \begin{align*}
-h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right).
+h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right)
 \end{align*}
@@
 $$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
-\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle. $$
+\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle $$
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
 \begin{equation*}
-\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi).
+\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi)
 \end{equation*}
@@
 \mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
-&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
+&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right]
 \end{align*}
@@
 \begin{equation}
-\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
+\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i}
 \label{elpd}
@@
 \begin{equation}
-\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
+\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})]
 \label{loo-est}
@@
 $$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
-$$t_{0.5} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle.$$
+$$t_{0.5} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle$$
 
@@
 \item Sample $r$ uniformly from $\{1,\ldots,S\}$.\\ For $i=1,\ldots,n_g$, do:
-\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(r)},t_{p1}^{(r)},\sigma_1^{(r)},t_{p_{2}g}^{(r)}, \sigma_{2g}^{(r)})$.
-\item Repeat step 2 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(r)},t_{p1}^{(r)},\sigma_1^{(r)},t_{p_{2}g}^{(r)}, \sigma_{2g}^{(r)})$
+\item Repeat step 2 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$
 \item Set $d_{g,i}^{(s)}= \begin{cases} \max \{t_{g,i}: i \ge 1\}, & \mbox{ if }c_{g,i}=0\\
@@
   (d_{g,i}^{(s)},1), & \mbox{ if } \tilde{t}_{rep,g,i}^{(s)}>d_{g,i}^{(s)}\\
-  (\tilde{t}_{rep,g,i}^{(s)},0), & \mbox{ if }\tilde{t}_{rep,g,i}^{(s)} \le d_{g,i}^{(s)}.
+  (\tilde{t}_{rep,g,i}^{(s)},0), & \mbox{ if }\tilde{t}_{rep,g,i}^{(s)} \le d_{g,i}^{(s)}
 \end{cases}$
@@
 The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
-$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}.$$
+$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}$$
 where $t_{g,new}^{(s)}$ is a draw from the posterior predictive distribution for drive-model $g$.
@@
 relative to a new unit. We can rank the drive-models by 
-$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)}).$$
+$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)})$$
 

### 2018-05-21 paper/m-lb-m-technometrics.tex
Change font size
Fixes periods at end of equations

--- 2018-05-14 paper/m-lb-m-technometrics.tex
@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 \usepackage{amsmath}
@@
 \begin{align*}
-P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
-P(T\le t | \delta_i=0) =& F_2(t).
+P(T\le t | \delta_i=1) &=  1 -(1-F_1(t))(1-F_2(t))\\
+P(T\le t | \delta_i=0) &=  F_2(t).
 \end{align*}
@@
 $$\eta_{t_{0.2}}  \sim \op{Normal}(9, 2^2),\quad
-\tau_{t_{0.2}}  \sim \op{Half-Cauchy}(0, 1)$$
+\tau_{t_{0.2}}  \sim \op{Half-Cauchy}(0, 1).$$
 
@@
 \item Sample $r$ uniformly from $\{1,\ldots,S\}$.\\ For $i=1,\ldots,n_g$, do:
-\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(r)},t_{p1}^{(r)},\sigma_1^{(r)},t_{p_{2}g}^{(r)}, \sigma_{2g}^{(r)})$.
-\item Repeat step 2 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
+\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(r)},t_{p1}^{(r)},\sigma_1^{(r)},t_{p_{2}g}^{(r)}, \sigma_{2g}^{(r)}).$
+\item Repeat step 2 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$
 \item Set $d_{g,i}^{(s)}= \begin{cases} \max \{t_{g,i}: i \ge 1\}, & \mbox{ if }c_{g,i}=0\\

### 2018-05-30 paper/m-lb-m-technometrics.tex
Complete small list of edits..
- change notation t_{p_i} to be consistent throughout
- add "Drive model " to plot names for clarity
- fix typos/small grammar/wording things
- add reference (Gelman 2004)
- add acknowledgment (Backblaze)
- move Appendix to Supplementary materials

@@
 Because of this, standard lifetime models, such as using a single Weibull or
-lognormal distribution may be inadequate. Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
+lognormal distribution, may be inadequate. Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
 Furthermore, when the information carried by lifetime data are limited by sample size, censoring, and truncation, estimates can be unstable and suffer from imprecision.
@@
 Weibull distribution because of extreme value theory (see page 86  of
-\cite{meeker}) for further discussion of this point.
+\cite{meeker} for further discussion of this point.)
 \end{itemize}
 
-To simplify the
-presentation and because it adequately describes our example data,
-we will use the Weibull distribution to describe both the early and
-wearout failures. Changing to any other combination of distributions
+%To simplify the presentation and because it adequately describes our example data, we will use the Weibull distribution to describe both the early and wearout failures.
+
+Changing to any other combination of distributions
 in the log-location-scale family of distributions is
@@
 \end{itemize}
+ \cite{gelman2004parameterization} provides a more general discussion of reparameterization and Bayesian inference.
 
@@
 \label{subsec:GLFP model}
-Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p1},\sigma_1)$ and $(t_{p2}, \sigma_2)$, respectively.
-The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$ Then
+Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p_1},\sigma_1)$ and $(t_{p_2}, \sigma_2)$, respectively.
+The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p_1},\sigma_1,t_{p_2},\sigma_2)$ Then
 $$\Pr(T \le t) = H(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$
@@
 \begin{align*}
-h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right).
+h(t|\pi, t_{p_1},\sigma_1, t_{p_2}, \sigma_2) = \pi & f_1(t|t_{p_1},\sigma_1)\left(1-F_2(t|t_{p_2},\sigma_2)\right) + \\ & f_2(t|t_{p_2},\sigma_2)\left(1-\pi F_1(t|t_{p_1},\sigma_1\right).
 \end{align*}
@@
 Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$).
-The likelihood for the GLFP is a function of the parameters $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.
+The likelihood for the GLFP is a function of the parameters $\bm{\theta} = (\pi,t_{p_1}, \sigma_1, t_{p_2}, \sigma_2)$.
 Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
@@
 \label{subsec:ex1}
-To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.
+To illustrate the GLFP model, we will first present an analysis of the drive-model with the most observed failures, Drive-model 14.
 The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.
@@
 Having a proper prior distribution also ensures a proper posterior distribution.
-We reparameterize the component Weibull distributions using the $0.50$ quantile for the early failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$).
+We reparameterize the component Weibull distributions using the $0.50$ quantile for the early failure mode ($t_{p_1=0.5}$) and the $0.20$ quantile for the wearout failure mode ($t_{p_2=0.2}$).
 \\
@@
 %$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
-$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
-t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
+$$t_{p_1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
+t_{p_2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
 $$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
-\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle. $$
+\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle$$
+where $p_1=0.2$ and $p_2=0.5$.
 %\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$
@@
   $\beta_2 = 1/\sigma_2$ & 4.47 & 4.70 & 4.95 \\ 
-  $t_{0.5,1}$ & $1.03$  & $2.28$ & $3.99$\\ 
-  $t_{0.2,2}$ & $18.0$ & $18.2$ & $18.6$ \\ 
+  $t_{p_1=0.5}$ & $1.03$  & $2.28$ & $3.99$\\ 
+  $t_{p_2=0.2}$ & $18.0$ & $18.2$ & $18.6$ \\ 
    \hline
@@
 We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.
-With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (see Appendix for a detailed description).
+With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (details are given in Appendix A in the Supplementary Materials).
 
@@
 The plot also contains 90\% pointwise credible bands associated with each estimate.
-While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to adequately describe the rapid increase in the empirical cdf between 8000 and 20,000 hours.
+While a single Weibull distribution is inadequate to describe these data, the GLFP model fits quite well, as it is able to adequately describe the rapid increase in the empirical cdf between 8000 and 20,000 hours.
 %The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.
@@
 \label{sec:Hierarchical GLFP model}
-In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.
+In order to describe an entire product population consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.
 
@@
 
-In a hierarchical model, the parameters are modeled as random variables, varying across different sub-populations.  For example
+In a hierarchical model, the parameters are modeled as random variables, varying across different sub-populations.
+Although other distributions might be specified in general, for the Backblaze example, we used
 \begin{equation*}
@@
 \begin{equation}
-\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i}
+\mbox{elpd} = \op{E}_h (\mbox{lpd}) = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i}
 \label{elpd}
@@
 $$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
-$$t_{0.5} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle.$$
+$$t_{p_1} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle.$$
+where $p_{1}=0.5$.
 
@@
 \tau_{\sigma_2}  \sim \op{Half-Cauchy}(0, 1)$$
-$$\eta_{t_{0.2}}  \sim \op{Normal}(9, 2^2),\quad
-\tau_{t_{0.2}}  \sim \op{Half-Cauchy}(0, 1).$$
+$$\eta_{t_{p_2}}  \sim \op{Normal}(9, 2^2),\quad
+\tau_{t_{p_2}}  \sim \op{Half-Cauchy}(0, 1)$$
+where$p_{2}=0.2$.
 
@@
 \centering
-\begin{tabular}{rrrrrrrr}
+\begin{tabular}{rrccc}
   \hline
@@
   \hline
-Model 4 & -13309.5 & 40.7 & 11.3 & 92.5 \\ 
-Model 3 & -13350.2 & 458.8 & 31.0 & 68.8 \\ 
-Model 2 & -13809.0 & 3674.6 & 96.2 & 49.6 \\ 
-Model 1 & -17483.6 &  &  & 4.3 \\
+Model 4 & $-13{,}309.5$ & 40.7 & 11.3 & 92.5 \\ 
+Model 3 & -13{,}350.2 & 458.8 & 31.0 & 68.8 \\ 
+Model 2 & -13{,}809.0 & 3674.6 & 96.2 & 49.6 \\ 
+Model 1 & -17{,}483.6 &  &  & 4.3 \\
    \hline
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{ppcheck-sample}
+\includegraphics[width=\textwidth]{ppcheck-sample-v2}
 \caption{\footnotesize Adjusted K-M estimates of the proportion failing for a representative subset of drive-models.
@@
 The data for this Drive-model are heavily truncated; the observed units had run for quite some time before data were made available with the earliest left-truncation time at 12,189 hours.
-In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for drive-model 43 to an average GLFP cdf for the entire population, which we call the ``global average" cdf; a description of the global average cdf is given in Appendix \ref{global-avg}.
+In the bottom half of Figure \ref{fig:ex-mod-43} we contrast the posterior median for drive-model 43 to an average GLFP cdf for the entire population, which we call the ``global average" cdf; a description of the global average cdf is given in Appendix B in the Supplementary Materials.
 We can see that the lack of data for this drive-model during early life results in a diffuse posterior centered around the global average cdf until about 10,000 hours when they diverge.
@@
 The top panel of Figure 9 shows estimates of $t_{0.10}$ (i.e., the amount of time it takes for 10\% of hard drives to fail, corresponding to the cdf $H(t)$ in Section \ref{subsec:GLFP model}), for all drive-models.
-In this application 50\% CIs were used due to considerable posterior uncertainty in some of the parameters: especially for those with few drives in operation.
+In this application 50\% CIs were used due to considerable posterior uncertainty in some of the parameters, especially for those with few drives in operation.
 %Interestingly, the top seven drive-models ranked by MTTF are also the drive-models with the largest $t_{0.10}$.
@@
   \caption{\footnotesize Left: 90\% CI of the posterior predictive distributions for $e^{-0.4 t_{g,new}}$ the depreciated value at failure (VAF) for a new unit for each Drive-model.
-  Right: A comparison (90\% CI) of the predictive distribution of TTF in hundreds of thousands of hours (top) and VAF (bottom) based on 40\% depreciation for the best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
+  Right: A comparison (90\% CI) of the predictive distribution of TTF in thousands of hours (top) and VAF (bottom) based on 40\% depreciation for the best 11 drive-models, shows that rankings produced by the two measures are not equivalent.}
   \label{fig5}
@@
 The authors express their gratitude to Nick Clark for his helpful comments and suggestions.  We would also like to thank the anonymous referees who provided 
-many useful suggestions that helped improve this paper.
+many useful suggestions that helped improve this paper. Finally, we would like to thank Andy Klein of Backblaze and the Backblaze company for making their hard drive data publicly available.
 
@@
 
-\appendix
-\section{Truncation adjusted Kaplan-Meier estimate of lifetime}
-\label{sec:trunc-adj}
-We first start with a nonparametric estimate of the empirical cdf for each sub-population using the Kaplan-Meier estimator.
-With left truncation, however, the standard Kaplan-Meier estimator for drive-model $g$, denoted by
-$\widehat{F_g(t)}_{KM}$, is conditional on survival up to
-$t_{g,\text{min}}^L$, the shortest reported running time of all units in
-of sub-population $g$ for which records are available. To produce
-unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each sub-population we select
-$t_{g,\text{min}}^L$, the smallest left truncated time in the sample.
-By sampling from the full posterior distribution, because
-$\Pr(T>t_{g,\text{min}}^L|\theta_g)$ (the probability that a hard drive has
-survived up to $t_{g,\text{min}}^L$) is a function of the model
-parameters, we can easily compute its posterior median,
-$\widehat{A}_{\text{med}} = \widehat{\Pr}(T>t_{g,\text{min}}^L|\theta_g)$. We compute the adjusted estimate by
-
-$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_g(t)}_{KM},\; t>t_{g,\text{min}}^L.$$
-
-While this adjustment may be negligible for sub-populations with little truncation, in our Backblaze example, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
-2.3 years.
-
-\section{Definition of global average for Model 4}
-\label{global-avg}
-In Section \ref{model-assessment}, to illustrate the concept of shrinkage in our hierarchical model, we refer to a ``global average" which represents an average GLFP cdf for the entire population, $\op{H}\left(\cdot|\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
-Because $\mu_1$ and $\sigma_1$ are common across all sub-populations in our model, these can already be interpreted as ``global." For the parameters that vary across sub-populations, we select values corresponding to the medians of the hierarchical distributions (\ref{eq:hier-model}) conditional on the hyperparameters.
-In particular, we set
-$$\bar{\pi}=\op{logit}^{-1}(\eta_{\pi}),\;\bar{\mu}_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2) \mbox{ and } \bar{\sigma}_2= m_{\sigma_2}.$$
-Let $J(\cdot|a,b),\;J^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a lognormal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
-$$m_{\sigma_2}=J^{-1}[0.5 \cdot J(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}],$$
-which is the median of a lognormal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.
-
-We estimate the global average pointwise, using draws from the joint posterior distribution, $H\left(\tilde{t}|\eta_\pi^{(s)}, \mu_1^{(s)}, \sigma_1^{(s)}, \eta_{t_2}^{(s)},\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$, $s=1,2,\ldots,S$.
-The computation is thus similar to that shown in (\ref{pointwise-medians}).
  % We get this by setting the $\op{GLFP}$ parameters equal to the median of their hierarchical distribution, i.e., $\pi=\op{logit}^{-1}(\eta_{\pi}^{(s)}),\, \mu_1=\mu_1^{(s)},\, \sigma_1=\sigma_1^{(s)},\, \mu_2=\eta_{t_2}^{(s)} - m_{\sigma_2}^{(s)} \Phi_{SEV}^{-1}(.2)$ and $\sigma_2= m_{\sigma_2}^{(s)}$, where $m_{\sigma_2}^{(s)}=F^{-1}[0.5 \cdot F(1;\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)});\eta_{\sigma_2}^{(s)}, \tau_{\sigma_2}^{(s)}]$ is the population median of the $\sigma_{2g}$, conditioning on $\left(\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$.
@@
 
-\section{Supplementary Material}
+\section*{Supplementary Material}
 \label{sec:supple}
@@
 
-\item[Model code and plots] Stan code for Model 4 as well as the complete set of plots similar to Figures 5 and 6 (supplementary.pdf)
+\item[Appendices, stan model code and plots] Description of truncation adjustment to Kaplan Meier estimator, global average, Stan code for Model 4 as well as the complete set of plots similar to Figures 5 and 6 (supplementary.pdf)
 

### 2018-05-30 paper/supplement.tex
Complete small list of edits..
- change notation t_{p_i} to be consistent throughout
- add "Drive model " to plot names for clarity
- fix typos/small grammar/wording things
- add reference (Gelman 2004)
- add acknowledgment (Backblaze)
- move Appendix to Supplementary materials

@@
 
-\begin{description}
+{\appendix
+\section{Truncation adjusted Kaplan-Meier estimate of lifetime}
+\label{sec:trunc-adj}
+We first start with a nonparametric estimate of the empirical cdf for each sub-population using the Kaplan-Meier estimator.
+With left truncation, however, the standard Kaplan-Meier estimator for drive-model $g$, denoted by
+$\widehat{F_g(t)}_{KM}$, is conditional on survival up to
+$t_{g,\text{min}}^L$, the shortest reported running time of all units in
+of sub-population $g$ for which records are available. To produce
+unconditional estimates, we adapt the adjustment method outlined by \citet[Chapter 11]{meeker}.  For each sub-population we select
+$t_{g,\text{min}}^L$, the smallest left truncated time in the sample.
+By sampling from the full posterior distribution, because
+$\Pr(T>t_{g,\text{min}}^L|\theta_g)$ (the probability that a hard drive has
+survived up to $t_{g,\text{min}}^L$) is a function of the model
+parameters, we can easily compute its posterior median,
+$\widehat{A}_{\text{med}} = \widehat{\Pr}(T>t_{g,\text{min}}^L|\theta_g)$. We compute the adjusted estimate by
 
-\item Stan code for Model 4, estimated using the Backblaze data set (also available as a supplementary .csv file).
+$$\widehat{F(t)}_{KMadj} = \widehat{A}_{\text{med}} + \left(1 - \widehat{A}_{\text{med}}\right)\widehat{F_g(t)}_{KM},\; t>t_{g,\text{min}}^L.$$
 
-\end{description}
+While this adjustment may be negligible for sub-populations with little truncation, in our Backblaze example, five drive-models receive upward adjustments of greater than 5 percent and the estimated time to failure distribution of one drive-model (30) was adjusted by nearly 16 percent, in part because the shortest truncation time for all observed units was approx.
+2.3 years.
+
+\section{Definition of global average for Model 4}
+\label{global-avg}
+In Section 7, to illustrate the concept of shrinkage in our hierarchical model, we refer to a ``global average" which represents an average GLFP cdf for the entire population, $\op{H}\left(\cdot|\bar{\pi},\mu_1,\sigma_1,\bar{\mu}_2,\bar{\sigma}_2\right)$.
+Because $\mu_1$ and $\sigma_1$ are common across all sub-populations in our model, these can already be interpreted as ``global." For the parameters that vary across sub-populations, we select values corresponding to the medians of the hierarchical distributions (4) conditional on the hyperparameters.
+In particular, we set
+$$\bar{\pi}=\op{logit}^{-1}(\eta_{\pi}),\;\bar{\mu}_2=\eta_{t_2} - m_{\sigma_2}\Phi^{-1}(.2) \mbox{ and } \bar{\sigma}_2= m_{\sigma_2}.$$
+Let $J(\cdot|a,b),\;J^{-1}(\cdot|a,b)$ denote the cdf and inverse cdf, respectively, for a lognormal distribution with log-location parameter $a$ and log-scale parameter $b$. Then
+$$m_{\sigma_2}=J^{-1}[0.5 \cdot J(1|\eta_{\sigma_2},\tau_{\sigma_2})|\eta_{\sigma_2}, \tau_{\sigma_2}],$$
+which is the median of a lognormal distribution with parameters $\eta_{\sigma_2}$, and $\tau_{\sigma_2}$, truncated to the interval $(0, 1)$.
+
+We estimate the global average pointwise, using draws from the joint posterior distribution, $H\left(\tilde{t}|\eta_\pi^{(s)}, \mu_1^{(s)}, \sigma_1^{(s)}, \eta_{t_2}^{(s)},\eta_{\sigma_2}^{(s)},\tau_{\sigma_2}^{(s)}\right)$, $s=1,2,\ldots,S$.
+The computation is thus similar to that shown in (7).
+}
+
+%\begin{description}
+
+%\item Stan code for Model 4, estimated using the Backblaze data set (also available as a supplementary .csv file).
+
+%\end{description}
 
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-1.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-1.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-2}
+\includegraphics[width=\textwidth]{ppcheck-v3-2}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-3.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-3.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-4.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-4.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-5.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-5.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-6.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-6.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-7.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-7.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-8.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-8.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-9.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-9.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-10.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-10.pdf}
 \end{figure}
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{ppcheck-v2-11.pdf}
+\includegraphics[width=\textwidth]{ppcheck-v3-11.pdf}
 \end{figure}
@@
 
+\bibliography{./sample}
+
 \end{document}

### 2018-05-30 workflow/predKM1.R
Add "Drive-model " to plot title

@@ -69,7 +69,7 @@ y_reps <- lapply(1:44, function(dm){
 })
 
 saveRDS(y_reps,"../plots/y_rep.rds")
-
+y_reps <- readRDS("../plots/y_rep.rds")
 get_plotlist <- function(y_reps, dat, xlimits=c(100,100000)){
   plist <- list()
   xaxis <- NULL
@@ -92,7 +92,7 @@ get_plotlist <- function(y_reps, dat, xlimits=c(100,100000)){
       theme_bw(base_size=14) + #theme(#axis.title.y=element_blank(),
                                      #plot.margin = unit(c(0,0,0,0), "cm")
         #) + 
-      ggtitle(as.character(i)) + ylab("Proportion failing")
+      ggtitle(paste(c("Drive-model ", as.character(i)), collapse="")) + ylab("Proportion failing")
   
     if(i==1) legend <- cowplot::get_legend(pnew)
     
@@ -117,7 +117,7 @@ plist <- get_plotlist(y_reps, dat, xlimits=c(100,100000))
 
 # plist <- get_plotlist(all, firstcol = 1+(0:8)*5, lastrow = 41:44)
 library(cowplot)
-pdf("../paper/fig/post-pred-KM-all2.pdf", height=14, width=9)
+pdf("../paper/fig/post-pred-KM-all3.pdf", height=14, width=9)
 plot_grid(plotlist = plist$plots, ncol=5, align="hv")
 dev.off()
 
@@ -140,7 +140,7 @@ for(j in 1:11){
     # plot_grid(plot_grid(plotlist=plist$plots[1:4 + (j-1)*8], ncol=2, align="h"), plist$legend,
                      # nrow=1, rel_widths = c(1,.2))
   # }
-  ggsave(paste("../paper/fig/ppcheck-v2-",j,".pdf",sep=""), width = 8.5, height=11)
+  ggsave(paste("../paper/fig/ppcheck-v3-",j,".pdf",sep=""), width = 8.5*1.7, height=11*1.7)
 }
 
 # set.seed(12122017)
@@ -148,9 +148,9 @@ for(j in 1:11){
 smplID <- c(2,9,14,40)
 smp_plist <- get_plotlist(y_reps, dat = dat, xlimits = c(100,30000))
 plot_grid(plot_grid(plotlist = smp_plist$plots[smplID], ncol=2, align="hv"), plist$legend,
-          nrow=1, rel_widths = c(1,.2))
[…]

### 2018-06-09 paper/m-lb-m-technometrics.tex
Fixes 2 small typos in May 31 version
Incorporates Meeker edits from June 8

@@
 \label{equation:reparameterized.weibull.cdf}
-&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0
+&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
 \end{align*}
@@
 \label{equation:likelihood}
-L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
+L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}.
 \end{equation}
@@
 When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the traditional parameters of the distribution.
-Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the priors).
+Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the prior distributions).
 Using this convention the prior distributions used in our analyses are
@@
 Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of the $\pi_g$ across the subpopulations.
-Also the lifetime distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.
+Also the lifetime distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest in practical applications.
 
@@
 Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
-posterior draws from a fitted model.  Following \cite{vehtari}, the log pointwise predictive density (lpd), is one method to quantify predictive accuracy,
+posterior draws from a fitted model.  Following \cite{vehtari}, the log pointwise predictive density (lpd) is one method to quantify predictive accuracy,
 
@@
 $$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
-$$t_{p_1} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle.$$
+$$t_{p_1} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle,$$
 where $p_{1}=0.5$.
@@
 \tau_{t_{p_2}}  \sim \op{Half-Cauchy}(0, 1)$$
-where$p_{2}=0.2$.
+where $p_{2}=0.2$.
 
@@
 \end{equation}
-where $\theta^{(s)}$ is posterior sample $s$ from $p(\theta|y)$.
+where $\theta^{(s)}$ is a posterior sample $s$ from $p(\theta|y)$.
 %Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.
@@
 %stopped here
-The GLFP model fits for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for similar plots for the other drive-models).
+The GLFP model fits for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (see the Supplementary Material for similar plots for the other drive-models).
 The step functions again correspond to the adjusted K-M estimates.
@@
 Drive-model 14 shows a discrepancy between the GFLP fit and the K-M fit in the right tail.
-In particular, the GLFP predictions appear conservative, overpredicting the proportion failing after about 2.5 years.
+In particular, the GLFP predictions overpredict the proportion failing after about 2.5 years.
 Due to the large number of failures for this drive-model, the posterior for the hierarchical model is similar to the stand-alone model.
@@
 Their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.
-The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product sub-populations.
+The ability to fit marginal models, such as the GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product sub-populations.
 Because the Bayesian approach provides full posterior distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the actual application.
@@
 
-\item[Appendices, stan model code and plots] Description of truncation adjustment to Kaplan Meier estimator, global average, Stan code for Model 4 as well as the complete set of plots similar to Figures 5 and 6 (supplementary.pdf)
+\item[Appendices, stan model code, plots, and tables] Description of truncation adjustment to Kaplan Meier estimator global average; Stan code for Model 4; complete set of plots similar to Figures 5 and 6; plots of prior distributions for Models 1-4; sensitivity analysis of the prior distributions for Model 4;
+and a complete list of brand names corresponding to the model numbers used in the analysis (supplementary.pdf)
 
